### 2023-01-31 scripts/01 Filter Bioeconomy.qmd
add scripts to upload tags to db

---
title: "01 Filter Forest Bioeconomy"
format: html
---
```{r}
here::i_am("scripts/01 Filter Bioeconomy.qmd")

library(tidyverse)
library(here)
library(DataEditR)

data_path <- here("data")

```

```{r}
#| warning: false

df <- readxl::read_xlsx(here(data_path, "raw-data", "20201221 SWINNO 1970-2019.xlsx"), sheet="SWINNO 1970-2019")

additional_info <- df %>% select(starts_with("Additional")) %>% colnames()
use_sectors <- df %>% select(starts_with("Use")) %>% colnames()

df <- df %>% 
  unite(info, additional_info, sep =" ", remove=TRUE, na.rm=TRUE) %>%
  unite(sectors, use_sectors, sep=",", remove= TRUE, na.rm=TRUE)

df <- df %>%
  select(c(`SINNO ID`, `Literal name of innovating firm ; SINNO's common name ; SINNO ID`  ,`Year of commercialization`, `Description in Swedish`, info, sectors))

filtered_df  <- df %>%
  filter(sectors == 2) %>%
  rename(year = `Year of commercialization`, description = `Description in Swedish`)

write_excel_csv(filtered_df, here(data_path, "modified-data", "classification-forest.csv"))

```

```{r}
data\raw-data\

df <- read_csv(here(data_path, "raw-data", "chaminade-bayou-bioeconomy.csv"))

additional_info <- df %>% select(starts_with("Additional")) %>% colnames()

df <- df %>% 
  unite(info, additional_info, sep =" ", remove=TRUE, na.rm=TRUE)

df <- df %>%
  select(c( SINNO.ID,
        Name.Swe,
        Descri.Swe,
        Literal.name.of.innovating.firm...SINNO.s.common.name...SINNO.ID,
       info))

write_excel_csv(df, here(data_path, "modified-data", "classification-bioeconomy.csv"))

```

```{python}

import pandas as pd

df = pd.read_csv(
    "../data/raw-data/chaminade-bayou-bioeconomy.csv",
    usecols=[
        "SINNO.ID",
        "Name.Swe",
        "Descri.Swe",
        "Literal.name.of.innovating.firm...SINNO.s.common.name...SINNO.ID",
        "Additional.information.if.Origin...1a.New.scientific.discovery",
        "Additional.information.if.Origin...1b.New.technologies.or.materials",
        "Additional.info.if.Origin...2a.Official.regulation..legislation.and.standards.",
        "Additional.information.if.Origin...5b.Solution.for.a.problem",
        "Additional.information.if.Origin...7.Performance",
        "Additional.information.if.Origin...20.Other",
    ],
)

add_info = list(df.iloc[:, 4:].columns)

df["info"] = df[add_info].apply(lambda x: x.str.cat(sep=''), axis=1)

df.to_csv("../data/modified-data/bioeonomy-classification.csv")

```

### 2023-01-31 scripts/combine-categories.py

@@ -1,0 +1,52 @@
+import pandas as pd
+from pathlib import Path
+from src.utils import get_project_root
+from src.swinno_helpers import *
+
+ROOT = get_project_root()
+
+def clean_codes(input_df, code_digits, col):
+    old = "(\d{code_digits})(?!\s+$)"
+    new = "$1,"
+    input_df[col] = input_df[col].replace(old, new, regex=True)
+    return input_df
+
+def check_duplicates(df, output_name, subset='sinno_id'):
+    duplicates = df.loc[df.duplicated(subset=subset, keep=False),:]
+    duplicates = duplicates.sort_values(by=subset)
+
+    if duplicates.empty:
+        print('No duplicates found.')
+    else:
+        print(f'{len(duplicates)} duplicates found.')
+        df.to_csv(Path(ROOT, "data", "modified-data", f'{output_name}-duplicates.csv'))
+        print(f'{output_name} written to csv.')
+
+engine = connect_swinno_db()
+
+for csv in Path(ROOT, "data", "raw-data").glob("*.csv"):
+
+    file_name = Path(csv).stem
+
+    print(f'****************** \n {file_name}')
+   
+    df = clean_import('csv', csv)
+    df["sinno_id"] = df["sinno_id"].astype(int)
+   
+    cleaned_innovation_types = clean_codes(df, code_digits=3, col='innovation_type')
+    cleaned_tags = clean_codes(df, code_digits=1, col='bioeconomy_vision')
+
+    check_duplicates(cleaned_tags, output_name=file_name)
+
+    split_innovation_types = split_cols(cleaned_tags, col_to_split='innovation_type', sep=',')
+    melted_innovation_types = melt_table(split_innovation_types, id_vars='sinno_id', col_start='innovation', value_name='innovation_type')
+    melted_innovation_types.to_sql(name = 'eco_innovations', con = engine, if_exists='append', echo=False)
+
+    split_visions = split_cols(cleaned_tags, col_to_split='bioeconomy_visions', sep=',')
+    melted_visions = melt_table(split_innovation_types, id_vars='sinno_id', col_start='bio', value_name='bioeconomy_vision')
+    melted_visions.to_sql(name = 'bioeconomy_visions', con = engine, if_exists='append', echo=False)
+
+
+
+
+

### 2023-01-31 scripts/test.py

@@ -1,0 +1,3 @@
+from src.utils import get_project_root
+
+print(get_project_root())

### 2023-01-31 src/swinno_helpers.py
add swinno tools for parsing tags into db

@@ -1,0 +1,86 @@
+import pandas as pd
+import numpy as np
+from sqlalchemy import create_engine
+from src.utils import get_project_root
+
+
+def split_cols(df, col_to_split: str, sep: str):
+    """Splits columns that have joined information (e.g., SNI codes)
+
+    Args:
+        df (pd.DataFrame): The data frame in which the split should occur
+        col_to_split (str): The name of the column to split
+
+    Returns:
+        pd.DataFrame: A new data frame with the original column content split into new columns and the original column removed.
+    """
+    temp_df = df
+    split_cols = (
+        temp_df.loc[:, col_to_split]
+        .astype(str)
+        .str.strip()
+        .str.split(sep, expand=True)
+        .fillna(np.nan)
+    )
+
+    for col in split_cols:
+        temp_df[f"{col_to_split}_{col}"] = split_cols.loc[:, col]
+
+    temp_df.drop(labels=col_to_split, axis=1, inplace=True)
+
+    return temp_df
+
+
+def melt_table(df, id_vars: str, col_start: str, value_name: str, **kwargs):
+    """Melts tables into lookup tables
+
+    Args:
+        df (pd.DataFrame): The dataframe to melt
+        id_vars (str): Variables to keep as ids
+        col_start (str): The start of the columns to select
+        value_name (str): The name of the value column in the melted table
+
+    Returns:
+        df (pd.DataFrame): Melted dataframe
+    """
+    temp_df = df.melt(
+        id_vars=id_vars,
+        value_vars=df.columns[df.columns.str.startswith(col_start)],
+        value_name=value_name,
+    )
+
+    temp_df.drop("variable", axis=1, inplace=True)
+    temp_df.dropna(inplace=True)
+
+    temp_df = temp_df.sort_values(by=id_vars).reset_index(drop=True)
+
+    return temp_df
+
+
+
+def clean_column_names(df):
+    df.columns = df.columns.str.strip().str.lower().str.replace(" ", "_")
+    return df
+
+
+def clean_import(file_type, file_path, **kwargs):
+
+    if file_type == "csv":
+        df = pd.read_csv(file_path, **kwargs)
+
+    elif file_type == "xlsx":
+        df = pd.read_excel(file_path, **kwargs)
+
+    else:
+        raise ValueError("Unsupported file type")
+
+    df = clean_column_names(df)
+    return df
+
+
+def connect_swinno_db():
+
[…]

### 2023-01-31 src/utils.py

@@ -1,0 +1,4 @@
+from pathlib import Path
+
+def get_project_root():
+    return Path(__file__).absolute().parent.parent

### 2023-01-31 tests/test_melt.py
add tests for swinnotools

@@ -1,0 +1,27 @@
+import pandas as pd
+import numpy as np
+from pandas.testing import assert_frame_equal
+from src.swinno_helpers import melt_table
+
+
+test_input_df = pd.DataFrame(
+    {
+        "ID": [1, 2, 3, 4],
+        "A_0": ["1", np.nan, "0", "1"],
+        "A_1": ["2", np.nan, np.nan, "2"],
+    }
+)
+
+expected_result = pd.DataFrame(
+    {
+        "ID": [1, 1, 2, 2, 3, 3, 4, 4],
+        "B": ["1", "2", np.nan, np.nan, "0", np.nan, "1", "2"],
+    }
+)
+
+
+def test_melt_tables():
+    function_result = melt_table(
+        test_input_df, id_vars="ID", col_start="A", value_name="B"
+    )
+    assert_frame_equal(expected_result, function_result)

### 2023-01-31 tests/test_split_cols.py

@@ -1,0 +1,15 @@
+import pandas as pd
+import numpy as np
+from pandas.testing import assert_frame_equal
+from src.swinno_helpers import split_cols
+
+test_input_df = pd.DataFrame({"A": ["1 2", np.nan, "0", "1 2"]})
+
+expected_result_df = pd.DataFrame(
+    {"A_0": ["1", np.nan, "0", "1"], "A_1": ["2", np.nan, np.nan, "2"]}
+)
+
+
+def test_df_equal():
+    function_result = split_cols(test_input_df, "A", sep=" ")
+    assert_frame_equal(expected_result_df, function_result, check_dtype=False)

### 2023-02-01 scripts/combine-categories.py
update to v0.1 of combine-categories program

@@ -31,6 +31,7 @@ for csv in Path(ROOT, "data", "raw-data").glob("*.csv"):
     print(f'****************** \n {file_name}')
    
     df = clean_import('csv', csv)
+    df.dropna(subset="sinno_id", inplace=True)
     df["sinno_id"] = df["sinno_id"].astype(int)
    
     cleaned_innovation_types = clean_codes(df, code_digits=3, col='innovation_type')
@@ -40,13 +41,19 @@ for csv in Path(ROOT, "data", "raw-data").glob("*.csv"):
 
     split_innovation_types = split_cols(cleaned_tags, col_to_split='innovation_type', sep=',')
     melted_innovation_types = melt_table(split_innovation_types, id_vars='sinno_id', col_start='innovation', value_name='innovation_type')
-    melted_innovation_types.to_sql(name = 'eco_innovations', con = engine, if_exists='append', echo=False)
+    melted_innovation_types.to_sql(name = 'eco_innovations', con = engine, if_exists='append', index=False)
 
-    split_visions = split_cols(cleaned_tags, col_to_split='bioeconomy_visions', sep=',')
+    split_visions = split_cols(cleaned_tags, col_to_split='bioeconomy_vision', sep=',')
     melted_visions = melt_table(split_innovation_types, id_vars='sinno_id', col_start='bio', value_name='bioeconomy_vision')
-    melted_visions.to_sql(name = 'bioeconomy_visions', con = engine, if_exists='append', echo=False)
+    melted_visions.to_sql(name = 'bioeconomy_visions', con = engine, if_exists='append', index=False)
+
+    notes = df[["sinno_id", "notes"]]
+    notes.to_sql(name =  'classification_notes', con=engine, index=False, if_exists='append')
 
 
 
 
+# TODO: sanitze nans before entering in db
+# TODO: find error which enters company / innovation names instead of sinno ids
+# Maybe should instead combine all dfs together before running duplicate check
 

### 2023-02-01 scripts/combine-categories.py
add helpers tests and script to enter tags to db

@@ -1,59 +1,75 @@
 import pandas as pd
-from pathlib import Path
 from src.utils import get_project_root
+from src.categorization_helpers import *
 from src.swinno_helpers import *
 
 ROOT = get_project_root()
-
-def clean_codes(input_df, code_digits, col):
-    old = "(\d{code_digits})(?!\s+$)"
-    new = "$1,"
-    input_df[col] = input_df[col].replace(old, new, regex=True)
-    return input_df
-
-def check_duplicates(df, output_name, subset='sinno_id'):
-    duplicates = df.loc[df.duplicated(subset=subset, keep=False),:]
-    duplicates = duplicates.sort_values(by=subset)
-
-    if duplicates.empty:
-        print('No duplicates found.')
-    else:
-        print(f'{len(duplicates)} duplicates found.')
-        df.to_csv(Path(ROOT, "data", "modified-data", f'{output_name}-duplicates.csv'))
-        print(f'{output_name} written to csv.')
+ 
 
 engine = connect_swinno_db()
 
-for csv in Path(ROOT, "data", "raw-data").glob("*.csv"):
+list_innovation_type_df = []
+list_visions_df = []
+list_notes_df = []
+
+for csv in Path(ROOT, "data", "raw-data", "tags").glob("*.csv"):
 
     file_name = Path(csv).stem
 
-    print(f'****************** \n {file_name}')
-   
-    df = clean_import('csv', csv)
+    print(f"****************** \n {file_name}")
+
+    df = clean_import("csv", csv)
     df.dropna(subset="sinno_id", inplace=True)
     df["sinno_id"] = df["sinno_id"].astype(int)
-   
-    cleaned_innovation_types = clean_codes(df, code_digits=3, col='innovation_type')
-    cleaned_tags = clean_codes(df, code_digits=1, col='bioeconomy_vision')
+    replaced_letters = replace_letters_codes(df, "bioeconomy_vision")
 
-    check_duplicates(cleaned_tags, output_name=file_name)
+    cleaned_innovation_types = clean_codes(replaced_letters, code_digits=3, col="innovation_type")
+    list_innovation_type_df.append(
+        cleaned_innovation_types.loc[
+            :,
+            ["sinno_id", "innovation_type"],
+        ]
+    )
 
-    split_innovation_types = split_cols(cleaned_tags, col_to_split='innovation_type', sep=',')
-    melted_innovation_types = melt_table(split_innovation_types, id_vars='sinno_id', col_start='innovation', value_name='innovation_type')
[…]

### 2023-02-01 src/categorization_helpers.py

@@ -1,0 +1,29 @@
+import pandas as pd
+from pathlib import Path
+
+
+def clean_codes(input_df, code_digits, col):
+    old = r"(\d{{}})(?!\s+$)".format(code_digits)
+    new = "$1,"
+    input_df[col] = input_df[col].replace(old, new, regex=True).str.strip()
+    return input_df
+
+
+def replace_letters_codes(input_df, col):
+    d = {"A": "1", "B": "2", "C": "3"}
+    input_df[col] = input_df[col].str.replace(
+        "(A|B|C)", lambda x: d[x.group()], regex=True
+    )
+    return input_df
+
+
+def check_duplicates(df, output_name, output_path, subset="sinno_id"):
+    duplicates = df.loc[df.duplicated(subset=subset, keep=False), :]
+    duplicates = duplicates.sort_values(by=subset)
+
+    if duplicates.empty:
+        print("No duplicates found.")
+    else:
+        print(f"{len(duplicates)} duplicates found.")
+        df.to_csv(output_path)
+        print(f"{output_name} written to csv.")

### 2023-02-01 src/swinno_helpers.py

@@ -77,10 +77,9 @@ def clean_import(file_type, file_path, **kwargs):
     df = clean_column_names(df)
     return df
 
-
 def connect_swinno_db():
 
     database_dir = get_project_root().parent.absolute()
     database_uri = f"sqlite:///{database_dir}/swinno.db"
-    engine = create_engine(database_uri, echo=True)
+    engine = create_engine(database_uri)
     return engine

### 2023-02-01 tests/test_clean_codes.py

@@ -1,0 +1,30 @@
+import pandas as pd
+from pandas.testing import assert_frame_equal
+from  src.categorization_helpers import *
+
+input_df = pd.DataFrame(
+    {"id": [1, 2, 3, 4, 5], "codes": ["A 0", "   1,23", "A,1", "ABC", " 1, 2, 0"]}
+)
+
+def test_letter_to_code():
+
+    expected_df = pd.DataFrame(
+        {"id": [1, 2, 3, 4, 5], "codes": ["1 0", "   1,23", "1,1", "123", " 1, 2, 0"]}
+    )
+
+    output_df = replace_letters_codes(input_df, "codes")
+
+    assert_frame_equal(expected_df, output_df, check_names=False)
+
+
+def test_clean_codes():
+
+    expected_df = pd.DataFrame(
+        {"id": [1, 2, 3, 4, 5], "codes": ["1,0", "1,2,3", "1,1", "1,2,3", "1,2,0"]}
+    )
+
+    replaced_letters = replace_letters_codes(input_df, "codes")
+
+    output_df = clean_codes(replaced_letters, code_digits=1, col="codes")
+
+    assert expected_df["codes"].all() == output_df["codes"].all()

### 2023-02-02 src/categorization_helpers.py
fix clean codes has problem with some str / conv

@@ -5,7 +5,7 @@ from pathlib import Path
 def clean_codes(input_df, code_digits, col):
     old = r"(\d{{}})(?!\s+$)".format(code_digits)
     new = "$1,"
-    input_df[col] = input_df[col].replace(old, new, regex=True).str.strip()
+    input_df[col] = input_df[col].astype('string').replace(old, new, regex=True).str.strip()
     return input_df
 
 

### 2023-02-02 tests/test_check_duplicates.py

@@ -1,0 +1,13 @@
+# TODO check dupliactes function
+# TODO check duplicate detection
+# TODO check duplicate df creation
+# TODO check writing to csv
+
+
+def test_find_duplicates():
+
+
+def test_create_duplicate_df():
+
+
+def test_write_csv():

### 2023-02-02 tests/test_clean_codes.py

@@ -17,7 +17,7 @@ def test_letter_to_code():
     assert_frame_equal(expected_df, output_df, check_names=False)
 
 
-def test_clean_codes():
+def test_clean_codes_1_digit():
 
     expected_df = pd.DataFrame(
         {"id": [1, 2, 3, 4, 5], "codes": ["1,0", "1,2,3", "1,1", "1,2,3", "1,2,0"]}
@@ -26,5 +26,36 @@ def test_clean_codes():
     replaced_letters = replace_letters_codes(input_df, "codes")
 
     output_df = clean_codes(replaced_letters, code_digits=1, col="codes")
+
+    assert expected_df["codes"].all() == output_df["codes"].all()
+
+
+def test_clean_codes_3_digit():
+
+    input_df = pd.DataFrame(
+        {"id": [1, 2, 3], "codes": ["112322,0", "1230", "123123"]}
+    )
+    expected_df = pd.DataFrame(
+        {"id": [1, 2, 3], "codes": ["112,322,0", "123,0", "123, 123"]}
+    )
+
+    replaced_letters = replace_letters_codes(input_df, "codes")
+
+    output_df = clean_codes(replaced_letters, code_digits=3, col="codes")
+
+    assert expected_df["codes"].all() == output_df["codes"].all()
+
+def test_clean_codes_int_handling():
+
+    input_df = pd.DataFrame(
+        {"id": [1, 2, 3], "codes": ["112322,0", "1230", 123123]}
+    )
+    expected_df = pd.DataFrame(
+        {"id": [1, 2, 3], "codes": ["112,322,0", "123,0", "123, 123"]}
+    )
+
+    replaced_letters = replace_letters_codes(input_df, "codes")
+
+    output_df = clean_codes(replaced_letters, code_digits=3, col="codes")
 
     assert expected_df["codes"].all() == output_df["codes"].all()

### 2023-03-22 docs/paper.qmd
add paper draft

---
title: Classification of Swedish Forest Sector Innovation

author: Jonas Kreutzer
date: last-modified
abstract: |

bibliography: swinno-shades-green.bib
csl: https://zotero.org/styles/apa?source=1
link-citations: true
link-bibliography: true
# colorlinks: true
execute: 
  echo: false
  warning: false
format: 
  pdf:
    fig-width: 10
    fig-height: 10
---

```{python}
from pathlib import Path
import pandas as pd
import seaborn as sns

import matplotlib.pyplot as plt
from IPython.display import display, Markdown

from src.utils import get_project_root
from src.swinno_helpers import connect_swinno_db

ROOT = get_project_root()
conn = connect_swinno_db()

sns.set_theme(style="whitegrid")

def format_table(df):
    # create a Styler object and set the wrapping style
    styler = df.style.set_properties(**{"width": "50em", "text-wrap": "break-word"})

    return styler

def plot_counts(df1, df2, category_col, y, show_legend=True, legend_cols=None, y_label="Count"):
    categories = df1[category_col].astype(str).unique()
    categories.sort()

    legend_ncols = len(categories) if legend_cols is None else legend_cols

    colors = sns.color_palette("colorblind", len(categories))
    palette = dict(zip(categories, colors))

    fig, axes = plt.subplots(
        ncols=1, nrows=2, sharex=False, sharey=True, layout="tight"
    )

    marker_size = 100
    df1 = df1.sort_values(by=category_col)
    df2 = df2.sort_values(by=category_col)

    sns.scatterplot(
        x=df1.index,
        y=y,
        hue=category_col,
        data=df1,
        ax=axes[0],
        palette=palette,
        s=marker_size,
    )

    sns.scatterplot(
        x=df2.index,
        y=y,
        hue=category_col,
        data=df2,
        ax=axes[1],
        palette=palette,
        s=marker_size,
    )

    for axs in axes:
        axs.set_ylabel(y_label)
        axs.set_xlabel("Year")
        axs.legend().remove()

    axes[0].set_title("Including Uncertains")
    axes[0].title.set_position([0.5, 1.05])  # adjust the position of the title

    axes[1].set_title("Excluding Uncertains")
    axes[1].title.set_position([0.5, 1.05])  # adjust the position of the title

    if not show_legend:
        fig.legend().remove()
    else:
        handles, labels = axes[0].get_legend_handles_labels()
        fig.legend(
            handles,
            labels,
            loc="lower center",
            ncol=legend_ncols,
            frameon=False,
            bbox_to_anchor=(0.5, -0.1),
        )  # adjust the position of the legend

    plt.subplots_adjust(bottom=0.2)

    plt.show()

def plot_heatmap(data, title):
    fig, ax = plt.subplots(1, 1)
    ax = sns.heatmap(
        data, annot=True, cmap="Blues", fmt="g", cbar=False, annot_kws={"fontsize": 10}
    )
    ax.set(title=title, ylabel=None, xlabel=None)
    ax.set_xticklabels(ax.get_xticklabels(), rotation=45, ha="right")

    plt.show()

def prepare_heatmap_data(df, index_col, cols_col, values_col, ascending=False):
    """
    Prepare data for a heatmap visualization by pivoting and sorting a dataframe.

    Args:
    - df: input dataframe
    - index_col: column name for the index of the pivoted dataframe
    - cols_col: column name for the columns of the pivoted dataframe
    - values_col: column name for the values of the pivoted dataframe
    - ascending: boolean to specify whether to sort in ascending or descending order (default: False)

    Returns:
    - pivoted and sorted dataframe
    """
    # Pivot the data to create a matrix
    matrix = df.pivot(index=index_col, columns=cols_col, values=values_col)

    # Sort the matrix by descending order of values in the column with the highest value
    highest_col = matrix.idxmax(axis=1).values[0]
    matrix = matrix.sort_values(highest_col, ascending=ascending)

    return matrix

```

# Classification Categories

The classification categories for eco-innovation types were sourced from a single paper and augmented during the classification process.
In addition to the 30 firm performance indicators identified by García-Granero's [-@garcia-granero2018EcoinnovationMeasurementReview] review of management literature, XX categories were added.
Some are mere extensions of their categories (e.g., specifying if an innovation prolongs the life cylce of the product itself or downstream products), others are new types of existing categories (e.e.g., reduced other pollution such as noise), yet others are entirely new categories (human health and safety).
Clearly not all new categories are eco-innovations.
The newly created pesticide category, for example, can be an eco-innovation if it reduces chemical pollution, but it could also increase chemical pollution if pesticides are introduced in novel ways.
These codes are therefore not a final categorization of eco-innovation, but rather steps towards a taxonomy, or at least usable dataset.

The bioeconomy vision categories were combined from
@bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy.
In addition to the three visions discussed in these articles, the classified data includes two additional categories: one to keep track of doubt cases and one for innovations which do not belong to any bioeconomy vision.

## Decision Process

For the eco-innovations the decision process is straightforward.
Based on the information contained in the text, the innovation is labeled with one or more categories.
In case of uncertainty, the doubt label is added.
Because the list of eco-innvation types was relatively exhaustive from the beginning, few adjustments had to be made.
These include, as alluded to above, expanding the labeling for other sustaianbility criteria, such as human health and safety, and specific forest related aspects / tasks to gain a more detailed understanding of an innovations aim.
For example, XX innnovations aim to protect harvested timber, either at harvest, during transport or storage.
Since these could be tagged in different ways e.g., longer lifecycle of (downstream) product, optimized use of raw materials, or even reduced waste, it seemed more appropriate to create a specific category for these cases.
Depending on subsequent analysis needs this could then be mapped to other categories.

The bioeconomy vision classification remains more challenging.
Identifying the larger aim, or correspondence to a vision, of an innovation can be unclear both from SWINNO descriptions and from sources.
An innovation is tagged as belonging to a bioeconomy vision if it clearly fullfils at least one of the criteria unique to the vision discussed in Bugge et al.[-@bugge2016WhatBioeconomyReview, @tbl-bv-bugge] and Vivien et al. [-@vivien2019HijackingBioeconomy, @tbl-bv-vivien].
This means that an innovation which references economic performance improvements is not tagged as belonging to both the bioresource and the biotechnology vision. An innovation would be tagged as belonging to the bioecology vision if it states that the innovation aims to protect biodiversity.
If an innovation only aims to upgrade a bioresource, it is tagged as belonging to the bioresource vision, wheras if these products also feature a territorial idenity, it would be tagged as both bioresource and bioecology vision.
Of course there are innovations belonging to the forestry sector which fit into no vision.
<!-- TODO: Include Examples -->

For both classficiations, uncertain cases were tagged to allow for later review, with intentions to use consensus between experts for final decision making.

|  | Biotechnology | Bioresource | Bioecology |
|---|---|---|---|
| Aim &   Objectives | Economic growth & job   creation | Economic growth &   sustainability | Sustainability, biodiversity,   conservation of ecosystems, avoiding soil degradation |
| Value   Creation | Application of biotechnology,   commercialisation of research & technology | Conversion and upgrading of   bio-resources (process oriented) | Development of integrated   production systems and high-quality products with territorial identity |
| Drivers   & mediators of innovation | R & D, patents, TTOs,   Research councils and funders (Science push, linear model) | Interdisciplinary, optimisation   of land use, include degraded land in the production of biofuels, use and   availability of bio-resources, waste management, engineering, science &   market (Interactive & networked production mode) | Identification of favourable   organic agro-ecological practices, ethics, risk, transdisciplinary   sustainability, ecological interactions, re-use & recycling of waste,   land use, (Circular and self-sustained production mode) |

: Visions from @bugge2016WhatBioeconomyReview {#tbl-bv-bugge}

| | Type II | Type III | Type I |
|---|---|---|---|
|  | A science-based economy driven   by industrial biotechnology<br>     <br>     The cell is a factory | Biomass replaces fossil fuels   and mining to produce energy and materials<br>     <br>     Biorefining at the heart of ecological transition (multilevel perspective). | An ecological economy, that is   compatible with the biosphere<br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |

: Visions from @vivien2019HijackingBioeconomy {#tbl-bv-vivien}

<!-- TODO: Read up on theoretical drawbacks of post-hoc tagging of visions -->

# Data

```{python}

bv = pd.read_sql(
    "SELECT sinno_id, bioeconomy_vision as bv FROM bioeconomy_visions", conn
)
bva = pd.read_sql(
    "SELECT sinno_id, bioeconomy_vision as bv FROM bioeconomy_visions_articles", conn
)

ei = pd.read_sql("SELECT sinno_id, innovation_type as it FROM eco_innovations", conn)
eia = pd.read_sql(
    "SELECT sinno_id, innovation_type as it FROM eco_innovations_articles", conn
)
swinno = pd.read_sql(
    "SELECT [SINNO ID] as sinno_id, [Year of commercialization] as year, [Innovation name in Swedish] as innovation_name, [Literal name of innovating firm ; SINNO's common name ; SINNO ID] as firm, Location as location FROM innovation",
    conn,
)

display(
    Markdown(
        f"There are {len(bv['sinno_id'].unique())} unique innovations classified for bioeconomy visions using only swinno descriptions and  {len(bva['sinno_id'].unique())} innovations classified using the source articles. For the eco-innovation classification without and with articles the totals are {len(ei['sinno_id'].unique())} and {len(eia['sinno_id'].unique())}, respectively."
    )
)
# Why are the unique ids for eco-innovation and visions different for the non-article chekes.
```

## Intra Rater Reliability of Labels

Because the classifications using articles were done on innovations already tagged using only SWINNO descriptions, the reliability of the rating between these two approaches can be calculated.
The reliability can also be calculated for some of the innovations classified only on SWINNO data, as during export some innovations were duplicated.
In total 15 innovations were duplicated; they are not included in the reported figures on SWINNO only classifications.
<!-- TODO: Check if duplicates are reported and remove if they are included -->
<!-- TODO: make sure number is correct -->

Different ways to assess the reliability of repeated ratings exist.
Treating the intra-rater reliability of repeated codings like inter-rater reliabilities suggests use of different suggested reliability measures.
However, most of them only consider the case of a single rating on different items between more than one observation.
Considering that one rating can include multiple labels, an extension of the Kappa-coefficient as proposed by @kraemer1980ExtensionKappaCoefficient was tried.
The statistic measures the proportion of agreement between ratings corrected for chance agreements and can be calculated by

$$ \kappa_0 = \frac{\bar{P} - P_e}{1-P_e} + \frac{(1-\bar{P})}{Nm_0(1-P_e)}, $$

where $\bar{P}$ is the average proportion of agreement pairs, 
$P_e = \sum p_i^2$ are the "overall proportions of observations in which the several response categories were selected" [@kraemer1980ExtensionKappaCoefficient 210], and $N$, $m_0$ the number of subjects and the number of of observations per subject, respectively.

The difficulty of calculating this measure lies in deciding on the correct value for $P_e$, as any combination of the 19 given eco-innovation labels could be the possible.

<!-- TODO fix this and calculate new -->

## Classifications Using Only SWINNO

```{python}

bv_count = pd.read_sql(
    """SELECT
  codes.Category AS Vision,
  COUNT(bioeconomy_vision) AS Count
FROM
  bioeconomy_visions
  JOIN [classification-codes] AS codes ON codes.code = bioeconomy_visions.bioeconomy_vision
GROUP BY
  Vision
ORDER BY
  Count DESC """,
    conn,
)

bv_count_zero = pd.read_sql(
    """
SELECT
  codes.Category AS Vision,
  COUNT(bioeconomy_vision) AS Count
FROM
  bioeconomy_visions
  JOIN [classification-codes] AS codes ON codes.code = bioeconomy_visions.bioeconomy_vision
WHERE
  NOT EXISTS (
    SELECT
      *
    FROM
      bioeconomy_visions AS bv2
    WHERE
      bv2.sinno_id = bioeconomy_visions.sinno_id
      AND bv2.bioeconomy_vision = 0
  )
GROUP BY
  Vision
ORDER BY
  Count DESC;
""",
    conn,
)

ei_count = pd.read_sql(
    """
SELECT
  codes.Category AS Type,
  COUNT(ei.innovation_type) AS Count
FROM
  eco_innovations as ei
  JOIN [classification-codes] AS codes ON codes.code = ei.innovation_type
GROUP BY
  innovation_type
ORDER BY
  Count DESC;
""",
    conn,
)

ei_count_0 = pd.read_sql(
    """
SELECT
  codes.Category AS Type,
  COUNT(ei.innovation_type) AS Count
FROM
  eco_innovations as ei
  JOIN [classification-codes] AS codes ON codes.code = ei.innovation_type
WHERE
  NOT EXISTS(
    SELECT * 
    FROM
      eco_innovations AS ei2
    WHERE
      ei2.sinno_id = ei.sinno_id
      AND ei2.innovation_type = "000"
  )
GROUP BY
  Type
ORDER BY
  Count DESC;
""",
    conn,
)
```

It took approximately one minute to assign labels for bioeconomy visions and eco-innovations using only SWINNO data.
However, this speed came at the cost of a high number of uncertain cases, which required follow up checking with articles, as can be seen from @tbl-bv-swinno-count and @tbl-ei-swinno-count.
When examining these tables it is important to keep in mind that one innovation can have been labeled as belonging to more than one category.

The large number of innovations not belonging to either a bioeconomy vision or an eco-innovation type is not very surprising, as it is the result of an overrepresentation of the mining sector in this data.
During the selection of data from SWINNO, at first an error led to the inclusion of many mining areas.
While this does inflate the number of non-applicable labels, it is perhaps a good sign that the labels are indeed useful and accurate when differentiating between eco-innovations / bioeconomy related innovations and those that do not belonge to these categories.

No additional steps were taken at this point to remedy the error, instead this sector was simply excluded before moving on the classificaiton using source articles.

This sections main take away is that the classification using predefined labels from literature has worked relatively well.
After some initial tweaks and few new categories (mostly more specific under categories of eco-innovation types @tbl-full-codes) a complete taxonomy can be applied to work with the sources.

```{python}

def get_max_categories(dataframes):
    max_categories = {}
    for name, df in dataframes.items():
        df = df.loc[df.iloc[:,0] != "Unsure", :]
        max_category = df.iloc[0, 0]
        max_categories[name] = max_category
    return max_categories

count_dfs = {"ei_count": ei_count, "ei_count_0": ei_count_0, "bv_count": bv_count, "bv_count_zero":bv_count_zero}

max_counts = get_max_categories(count_dfs)

# display(Markdown(f"Ignoring the Unsure category, the biggest group of innovations belonged to the vision {max_counts['bv_count']}, and {max_counts['bv_count_zero']}. For the eco-innovation types, the biggest categories are {max_counts['ei_count']} if uncertains are included and {max_counts['ei_count_0']} if they are excluded."))
```

```{python}

#| label: tbl-bv-swinno-count
#| tbl-cap: Count of Bioeconomy Visions Using Swinno Data

bv_counts = pd.merge(bv_count, bv_count_zero, on="Vision", how="left")

bv_counts.columns = ["Vision", "Count", "Count Excl. Unsure"]
bv_counts.loc[:, "Count Excl. Unsure"] = (
    bv_counts["Count Excl. Unsure"].fillna(0).astype(int)
)

bv_counts.style.set_properties(**{"text-align": "right"}).hide_index()

```

```{python}
#| label: tbl-ei-swinno-count
#| tbl-cap: Count of Eco-Innovations Using Swinno Data

ei_counts = pd.merge(ei_count, ei_count_0, on="Type", how="left")

ei_counts.columns = ["Type", "Count", "Count Excl. Unsure"]
ei_counts.loc[:, "Count Excl. Unsure"] = (
    ei_counts["Count Excl. Unsure"].fillna(0).astype(int)
)

ei_counts.style.set_properties(**{"text-align": "right"}).hide_index()

```

## Classificaitions Using Source Articles

```{python}
bv_count_articles = pd.read_sql(
    """
SELECT
  codes.Category AS Vision,
  COUNT(bioeconomy_vision) AS Count
FROM
  bioeconomy_visions_articles
  JOIN [classification-codes] AS codes ON codes.code = bioeconomy_visions_articles.bioeconomy_vision
GROUP BY
  Vision
ORDER BY
  Count DESC """,
    conn,
)

bv_count_zero_articles = pd.read_sql(
    """
SELECT
  codes.Category AS Vision,
  COUNT(bioeconomy_vision) AS Count
FROM
  bioeconomy_visions_articles
  JOIN [classification-codes] AS codes ON codes.code = bioeconomy_visions_articles.bioeconomy_vision
WHERE
  NOT EXISTS (
    SELECT
      *
    FROM
      bioeconomy_visions_articles AS bv2
    WHERE
      bv2.sinno_id = bioeconomy_visions_articles.sinno_id
      AND bv2.bioeconomy_vision = 0
  )
GROUP BY
  Vision
ORDER BY
  Count DESC;
""",
    conn,
)

```

```{python}
# |output: asis
bv_counts_articles = pd.merge(
    bv_count_articles, bv_count_zero_articles, on="Vision", how="left"
)

bv_counts_articles.columns = ["Vision", "Count", "Count Excl. Unsure"]
bv_counts_articles.loc[:, "Count Excl. Unsure"] = (
    bv_counts_articles["Count Excl. Unsure"].fillna(0).astype(int)
)

bv_counts_articles.style.set_properties(**{"text-align": "right"}).hide_index()

```

```{python}

ei_count_articles = pd.read_sql(
    """
SELECT
  codes.Category AS Type,
  COUNT(ei.innovation_type) AS Count
FROM
  eco_innovations_articles as ei
  JOIN [classification-codes] AS codes ON codes.code = ei.innovation_type
GROUP BY
  innovation_type
ORDER BY
  Count DESC;
""",
    conn,
)

ei_count_articles_0 = pd.read_sql(
    """
SELECT
  codes.Category AS Type,
  COUNT(ei.innovation_type) AS Count
FROM
  eco_innovations_articles as ei
  JOIN [classification-codes] AS codes ON codes.code = ei.innovation_type
WHERE
  NOT EXISTS(
    SELECT * 
    FROM
      eco_innovations_articles AS ei2
    WHERE
      ei2.sinno_id = ei.sinno_id
      AND ei2.innovation_type = "000"
  )
GROUP BY
  Type
ORDER BY
  Count DESC;
""",
    conn,
)
```

```{python}
# |output: asis
ei_counts_articles = pd.merge(
    ei_count_articles, ei_count_articles_0, on="Type", how="left"
)

ei_counts_articles.columns = ["Type", "Count", "Count Excl. Unsure"]
ei_counts_articles.loc[:, "Count Excl. Unsure"] = (
    ei_counts_articles["Count Excl. Unsure"].fillna(0).astype(int)
)

ei_counts_articles.style.set_properties(**{"text-align": "right"}).hide_index()
```

# Results

## Trends

### By Sector

#### Bioeconomy Vision

```{python}

bv_sectors = pd.read_sql(
    """
SELECT
  c.Category as vision,
  sni_codes.label as sector,
  count(bv.bioeconomy_vision) as count
FROM
  bioeconomy_visions as bv
  join use_sectors ON bv.sinno_id = use_sectors.[SINNO ID]
  join sni_codes ON SUBSTR(use_sectors.use_sectors, 1, 2) = sni_codes.code
  join [classification-codes] as c ON bv.bioeconomy_vision = c.Code
WHERE 
  NOT EXISTS(
   SELECT
      *
    FROM
      bioeconomy_visions AS bv2
    WHERE
      bv2.sinno_id = bv.sinno_id
      AND bv2.bioeconomy_vision = 0
)
GROUP BY
  sector,
  bioeconomy_vision;
""",
    conn,
)

bv_matrix = prepare_heatmap_data(
    bv_sectors, index_col="sector", cols_col="vision", values_col="count"
)

plot_heatmap(bv_matrix, "Count Bioeconomy Vision by Sector Excluding Uncertains")
```

#### Eco-Innovations

```{python}
ei_sectors = pd.read_sql(
    """
SELECT
  c.Category as innovation_type,
  sni_codes.label as sector,
  count(ei.innovation_type) as count
FROM
  eco_innovations as ei
  join use_sectors ON ei.sinno_id = use_sectors.[SINNO ID]
  join sni_codes ON SUBSTR(use_sectors.use_sectors, 1, 2) = sni_codes.code
  join [classification-codes] as c ON ei.innovation_type = c.Code
WHERE 
  NOT EXISTS(
   SELECT
      *
    FROM
      eco_innovations AS ei2
    WHERE
      ei2.sinno_id = ei.sinno_id
      AND ei2.innovation_type = "000"
)
GROUP BY
  sector,
  innovation_type;
""",
    conn,
)

ei_sectors_matrix = prepare_heatmap_data(
    ei_sectors, index_col="sector", cols_col="innovation_type", values_col="count"
)

plot_heatmap(ei_sectors_matrix, title="Eco-Innovation Type by Sector")

```
### Over Time

#### Bioeconomy Visions Using SWINNO

```{python}

bv_years = pd.read_sql(
    """
SELECT
  codes.Category AS Vision,
  count(bv.bioeconomy_vision) AS vision_count,
  innovation.[Year of commercialization] AS year
FROM
  bioeconomy_visions AS bv
  JOIN innovation ON bv.sinno_id = innovation.[SINNO ID]
  JOIN [classification-codes] AS codes ON bv.bioeconomy_vision = codes.Code
WHERE
  TRIM(innovation.[Year of commercialization]) IS NOT ""
GROUP BY
  year,
  bv.bioeconomy_vision;
""",
    conn,
)

bvc_years = pd.read_sql(
    """
SELECT
  codes.Category AS Vision,
  count(bv.bioeconomy_vision) AS vision_count,
  innovation.[Year of commercialization] AS year
FROM
  bioeconomy_visions AS bv
  JOIN innovation ON bv.sinno_id = innovation.[SINNO ID]
  JOIN [classification-codes] AS codes ON bv.bioeconomy_vision = codes.Code
WHERE
  NOT EXISTS (
    SELECT
      *
    FROM
      bioeconomy_visions AS bv2
    WHERE
      bv2.sinno_id = bv.sinno_id
      AND bv2.bioeconomy_vision = 0
  )
  AND TRIM(innovation.[Year of commercialization]) IS NOT ""
GROUP BY
  year,
  bv.bioeconomy_vision;
""",
    conn,
)

bv_years = bv_years.set_index("year")
bvc_years = bvc_years.set_index("year")

plot_counts(
    bv_years,
    bvc_years,
    category_col="Vision",
    y="vision_count",
    y_label="Count of Vision",
)
```

#### Bioeconomy Visions Using Articles

```{python}

bva_years = pd.read_sql(
    """
SELECT
  codes.Category AS Vision,
  count(bv.bioeconomy_vision) AS vision_count,
  innovation.[Year of commercialization] AS year
FROM
  bioeconomy_visions_articles AS bv
  JOIN innovation ON bv.sinno_id = innovation.[SINNO ID]
  JOIN [classification-codes] AS codes ON bv.bioeconomy_vision = codes.Code
WHERE
  TRIM(innovation.[Year of commercialization]) IS NOT ""
GROUP BY
  year,
  bv.bioeconomy_vision;
""",
    conn,
)

bvac_years = pd.read_sql(
    """
SELECT
  codes.Category AS Vision,
  count(bv.bioeconomy_vision) AS vision_count,
  innovation.[Year of commercialization] AS year
FROM
  bioeconomy_visions_articles AS bv
  JOIN innovation ON bv.sinno_id = innovation.[SINNO ID]
  JOIN [classification-codes] AS codes ON bv.bioeconomy_vision = codes.Code
WHERE
  NOT EXISTS (
    SELECT
      *
    FROM
      bioeconomy_visions_articles AS bv2
    WHERE
      bv2.sinno_id = bv.sinno_id
      AND bv2.bioeconomy_vision = 0
  )
  AND TRIM(innovation.[Year of commercialization]) IS NOT ""
GROUP BY
  year,
  bv.bioeconomy_vision;
""",
    conn,
)

bva_years = bva_years.set_index("year")
bvac_years = bvac_years.set_index("year")

plot_counts(
    bva_years,
    bvac_years,
    category_col="Vision",
    y="vision_count",
    y_label="Count of Vision",
)

```

At most we record

# Outlook

## Open Questions and Doubts

## Thoughts

At times it can be difficult to decide wether an innovation should be counted as optimization of raw material, reduction of waste, or both.
Generally it appears sensible that an optimization of waste can, but does not have to, lead to a reduction of waste.
Keeping both labels seems sensible here.
To avoid double counting eco-innovation in statistics, a prudent solution is to simple count the unique SWINNO IDs tagged with at least one eco-innovation type

Considering the difference between innovations labeled as unsure between SWINNO data only and articles, there is no way around using articles exclusively.
Utilizing articles drastically reduced the amount of uncertainty of a classification.
However, even with articles some doubt cases remain.
This is in so far expected as articles vary in length and detail as does my understanding of the likely impact of an innovation.
Considering the number of doubt cases is rather low, it might be valuable to ask a second rater (ideally an expert on environmental evaluation) to rate a subset of articles deemed clear in addition to the doubt cases.
An important question is who to approach for this task and with what instructions.
As the level of expertise increases, perhaps the willingness to make decisive decisions on such limited information decreases.
That, on the other hand, would in itself be valuable feedback regarding the overall approach.

Code | Use Sector
|:-|:---------|
01 | Jordbruk, jakt och service i anslutning härtill
02 | Skogsbruk och service till skogsbruk
13 | Utvinning av metallmalmer
16 | Tobaksvarutillverkning
17 | Textilvarutillverkning
18 | Tillverkning av kläder; pälsberedning
20 | Tillverkning av trä och varor av trä, kork, rotting o.d. utom möbler
41 | Vattenförsörjning
: Included Sectors {#tbl-included-sectors tbl-colwidths="[10,90]"}

<!-- Some of these are not clearly related to either the forest of bioeconomy. Mistake on my part. Category 16 needs to be removed.-->

There are a number of  additional use sectors that need to be looked at

Code | Use Sector
|:---|:---|
21 | Massa-, pappers- och pappersvarutillverkning
24 | Tillverkning av kemikalier och kemiska produkter
25 | Tillverkning av gummi- och plastvaror
34 | Tillverkning av motorfordon, släpfordon och påhängsvagnar
35 | Tillverkning av andra transportmedel
36 | Tillverkning av möbler; annan tillverkning
37 | Återvinning
40 | El-, gas-, ång- och hetvattenförsörjning
41 | Vattenförsörjning
45 | Byggverksamhet
50 | Handel med och service av motorfordon; detaljhandel med drivmedel
90 | Avloppsrening, avfallshantering, renhållning o.d.
: Missing Sectors {#tbl-missing-sectors tbl-colwidths="[10,90]"}

<!-- This query yields 1678 innovation. Filtering with '%skog%', 'trä%', '%papper%' and '%cellulose' yields 1598 innovation. This likely has many overlaps with sni codes however.
Filtering on the other sectors is a bit trickier than I thought.
Might be that this only results in additional 119 innovation. -->

<!-- 
A cursory glance at sector code 70 (Fastighetsverksamhet) reveals many innovations aimed at reducing moisture in rooms.
Could there be an argument to be made about development blocks for budilding with wood?
TODO: Ask Ronnie about difficulties of building with wood from a building management perspective
8596001

It would perhaps also be a smart idea to go through all agricultural sectors, at least to the extent that they relate to the forest.
Perhaps with a keyword search?
Possible keywords: skog, trä, cellulose, papper
What about the other machine sectors (29-30)? The same? -->

## Appendix {.appendix}

#### Bioeconomy Visions by Sector

```{python}
bvu_sectors = pd.read_sql(
    """
SELECT
  c.Category AS vision,
  sni_codes.label AS sector,
  count(bv.bioeconomy_vision) AS count
FROM
  bioeconomy_visions AS bv
  join use_sectors ON bv.sinno_id = use_sectors.[SINNO ID]
  join sni_codes ON SUBSTR(use_sectors.use_sectors, 1, 2) = sni_codes.code
  join [classification-codes] AS c ON bv.bioeconomy_vision = c.Code
GROUP BY
  sector,
  bioeconomy_vision;
""",
    conn,
)

bvu_matrix = prepare_heatmap_data(
    bvu_sectors, index_col="sector", cols_col="vision", values_col="count"
)

plot_heatmap(bvu_matrix, "Count Bioeconomy Vision by Sector Including Uncertains")
```

#### Eco-Innovations Over Time

##### SWINNO

##### Articles
```{python}

#| fig-caption: Count of Eco-Innovations by Year.

eia_years = pd.read_sql(
    """
SELECT
  codes.Category AS Type,
  count(ei.innovation_type) AS innovation_count,
  innovation.[Year of commercialization] AS Year,
  SUBSTR(ei.innovation_type, 1,1) as top_cat
FROM
  eco_innovations_articles AS ei
  JOIN innovation ON ei.sinno_id = innovation.[SINNO ID]
  JOIN [classification-codes] AS codes ON ei.innovation_type = codes.Code
WHERE
  TRIM(innovation.[Year of commercialization]) IS NOT ""
GROUP BY
  Year,
  ei.innovation_type;
""",
    conn,
)

eiac_years = pd.read_sql(
    """
SELECT
  codes.Category AS Type,
  count(ei.innovation_type) AS innovation_count,
  innovation.[Year of commercialization] AS year,
  SUBSTR(ei.innovation_type, 1,1)
FROM
  eco_innovations_articles AS ei
  JOIN innovation ON ei.sinno_id = innovation.[SINNO ID]
  JOIN [classification-codes] AS codes ON ei.innovation_type = codes.Code
WHERE
  NOT EXISTS (
    SELECT
      *
    FROM
      eco_innovations_articles AS ei2
    WHERE
      ei2.sinno_id = ei.sinno_id
      AND ei2.innovation_type = "000"
  )
  AND TRIM(innovation.[Year of commercialization]) IS NOT ""
GROUP BY
  year,
  ei.innovation_type;
""",
    conn,
)

eia_years = eia_years.set_index("Year")
eiac_years = eiac_years.set_index("year")

plot_counts(
    eia_years,
    eiac_years,
    category_col="Type",
    y="innovation_count",
    legend_cols = 8,
    y_label="Count of Type",
    show_legend=False
)

```

```{python}
for i, df in enumerate([eia_years, eiac_years]):
  df = df.iloc[:, :2]
  df = df.sort_values(by="innovation_count", ascending=False)
  df.columns = ["Type", "Count"]
  df.index.rename("Year")
  if i == 0:
    eia_years = df
  else:
    eiac_years = df

```

```{python}
#| tbl-cap: Top 5 Most Frequent Categories by Year Including Uncertains
format_table(eia_years.head())
```

```{python}
#| tbl-cap: Top 5 Most Frequent Categories by Year Excluding Uncertains
format_table(eiac_years.head())
```

```{python}
#| label: tbl-full-codes
# | tbl-cap: List of Labels

codes_df = pd.read_sql("SELECT * FROM [classification-codes]", conn, index_col="Code")

format_table(codes_df)
```

## References

### 2023-03-23 docs/paper.qmd
update paper with certainty analysis

@@
 In addition to the 30 firm performance indicators identified by García-Granero's [-@garcia-granero2018EcoinnovationMeasurementReview] review of management literature, XX categories were added.
-Some are mere extensions of their categories (e.g., specifying if an innovation prolongs the life cylce of the product itself or downstream products), others are new types of existing categories (e.e.g., reduced other pollution such as noise), yet others are entirely new categories (human health and safety).
+Some are mere extensions of their categories (e.g., specifying if an innovation prolongs the life cycle of the product itself or downstream products), others are new types of existing categories (e.e.g., reduced other pollution such as noise), yet others are entirely new categories (human health and safety).
 Clearly not all new categories are eco-innovations.
@@
 In case of uncertainty, the doubt label is added.
-Because the list of eco-innvation types was relatively exhaustive from the beginning, few adjustments had to be made.
-These include, as alluded to above, expanding the labeling for other sustaianbility criteria, such as human health and safety, and specific forest related aspects / tasks to gain a more detailed understanding of an innovations aim.
-For example, XX innnovations aim to protect harvested timber, either at harvest, during transport or storage.
+Because the list of eco-innovation types was relatively exhaustive from the beginning, few adjustments had to be made.
+These include, as alluded to above, expanding the labeling for other sustainability criteria, such as human health and safety, and specific forest related aspects / tasks to gain a more detailed understanding of an innovations aim.
+For example, XX innovations aim to protect harvested timber, either at harvest, during transport or storage.
 Since these could be tagged in different ways e.g., longer lifecycle of (downstream) product, optimized use of raw materials, or even reduced waste, it seemed more appropriate to create a specific category for these cases.
@@
 Identifying the larger aim, or correspondence to a vision, of an innovation can be unclear both from SWINNO descriptions and from sources.
-An innovation is tagged as belonging to a bioeconomy vision if it clearly fullfils at least one of the criteria unique to the vision discussed in Bugge et al.[-@bugge2016WhatBioeconomyReview, @tbl-bv-bugge] and Vivien et al.
+An innovation is tagged as belonging to a bioeconomy vision if it clearly fulfills at least one of the criteria unique to the vision discussed in Bugge et al.[-@bugge2016WhatBioeconomyReview, @tbl-bv-bugge] and Vivien et al.
 [-@vivien2019HijackingBioeconomy, @tbl-bv-vivien].
@@
 An innovation would be tagged as belonging to the bioecology vision if it states that the innovation aims to protect biodiversity.
-If an innovation only aims to upgrade a bioresource, it is tagged as belonging to the bioresource vision, wheras if these products also feature a territorial idenity, it would be tagged as both bioresource and bioecology vision.
+If an innovation only aims to upgrade a bioresource, it is tagged as belonging to the bioresource vision, whereas if these products also feature a territorial identity, it would be tagged as both bioresource and bioecology vision.
 Of course there are innovations belonging to the forestry sector which fit into no vision.
@@
 
-For both classficiations, uncertain cases were tagged to allow for later review, with intentions to use consensus between experts for final decision making.
+For both classifications, uncertain cases were tagged to allow for later review, with intentions to use consensus between experts for final decision-making.
 
@@
 | Aim & Objectives | Economic growth & job creation | Economic growth &   sustainability | Sustainability, biodiversity,   conservation of ecosystems, avoiding soil degradation |
-| Value   Creation | Application of biotechnology,   commercialisation of research & technology | Conversion and upgrading of   bio-resources (process oriented) | Development of integrated   production systems and high-quality products with territorial identity |
-| Drivers   & mediators of innovation | R & D, patents, TTOs,   Research councils and funders (Science push, linear model) | Interdisciplinary, optimisation   of land use, include degraded land in the production of biofuels, use and   availability of bio-resources, waste management, engineering, science &   market (Interactive & networked production mode) | Identification of favourable   organic agro-ecological practices, ethics, risk, transdisciplinary   sustainability, ecological interactions, re-use & recycling of waste,   land use, (Circular and self-sustained production mode) |
+| Value   Creation | Application of biotechnology,   commercialization of research & technology | Conversion and upgrading of   bio-resources (process oriented) | Development of integrated   production systems and high-quality products with territorial identity |
+| Drivers   & mediators of innovation | R & D, patents, TTOs,   Research councils and funders (Science push, linear model) | Interdisciplinary, optimization   of land use, include degraded land in the production of biofuels, use and   availability of bio-resources, waste management, engineering, science &   market (Interactive & networked production mode) | Identification of favorable   organic agro-ecological practices, ethics, risk, transdisciplinary   sustainability, ecological interactions, re-use & recycling of waste,   land use, (Circular and self-sustained production mode) |
 
@@
 |  | A science-based economy driven   by industrial biotechnology<br>     <br>     The cell is a factory | Biomass replaces fossil fuels   and mining to produce energy and materials<br>     <br>     Biorefining at the heart of ecological transition (multilevel perspective).
-| An ecological economy, that is   compatible with the biosphere<br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
+| An ecological economy, that is   compatible with the biosphere<br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decision-making centers |
 
@@
 ```
-
+<!-- 
 ## Intra Rater Reliability of Labels
@@
 In total 15 innovations were duplicated; they are not included in the reported figures on SWINNO only classifications.
-<!-- TODO: Check if duplicates are reported and remove if they are included -->
-<!-- TODO: make sure number is correct -->
+
+TODO: Check if duplicates are reported and remove if they are included 
+TODO: make sure number is correct
 
@@
 
-<!-- TODO fix this and calculate new -->
+TODO fix this and calculate new -->
 
@@
 It took approximately one minute to assign labels for bioeconomy visions and eco-innovations using only SWINNO data.
-However, this speed came at the cost of a high number of uncertain cases, which required follow up checking with articles, as can be seen from @tbl-bv-swinno-count and @tbl-ei-swinno-count.
+However, this speed came at the cost of a high number of uncertain cases, which required follow-up checking with articles, as can be seen from @tbl-bv-swinno-count and @tbl-ei-swinno-count.
 When examining these tables it is important to keep in mind that one innovation can have been labeled as belonging to more than one category.
@@
 During the selection of data from SWINNO, at first an error led to the inclusion of many mining areas.
-While this does inflate the number of non-applicable labels, it is perhaps a good sign that the labels are indeed useful and accurate when differentiating between eco-innovations / bioeconomy related innovations and those that do not belonge to these categories.
+While this does inflate the number of non-applicable labels, it is perhaps a good sign that the labels are indeed useful and accurate when differentiating between eco-innovations / bioeconomy related innovations and those that do not belong to these categories.
 
-No additional steps were taken at this point to remedy the error, instead this sector was simply excluded before moving on the classificaiton using source articles.
+No additional steps were taken at this point to remedy the error, instead this sector was simply excluded before moving on the classification using source articles.
 
@@
 
-## Classificaitions Using Source Articles
+## Classifications Using Source Articles
 
@@
 ```{python}
-# |output: asis
-bv_counts_articles = pd.merge(
-    bv_count_articles, bv_count_zero_articles, on="Vision", how="left"
-)
-
-bv_counts_articles.columns = ["Vision", "Count", "Count Excl. Unsure"]
-bv_counts_articles.loc[:, "Count Excl. Unsure"] = (
-    bv_counts_articles["Count Excl. Unsure"].fillna(0).astype(int)
-)
-
-bv_counts_articles.style.set_properties(**{"text-align": "right"}).hide_index()
-
-```
-
-```{python}
 
@@
 )
+
+```
+
+Beginning with the innovations tagged as Unsure in the SWINNO only classification, the source articles were read and tags for bioeconomy visions and eco-innovations given on this basis.
+@tbl-bv-article-count and @tbl-ei-article-count show the results thus far.
+
+```{python}
+
+def label_confidence(df):
+  count_unsure = df.loc[df.iloc[:, 0] == "Unsure", "Count"]
+  total_labels = df.iloc[:, 1].sum()
+  return float(1 - (count_unsure / total_labels).iloc[0]) #to return the value instead of series
+
+count_dfs_confidence = {"ei_count": ei_count, "ei_count_articles":ei_count_articles, "bv_count":bv_count, "bv_count_articles":bv_count_articles}
+
+label_confidences = {k: label_confidence(v) for  k, v in count_dfs_confidence.items()}
+
+display(Markdown(f"The confidence of labels for each approch, measured as the share of Unsure to total classifications is present in each data set shows that relying on source articles drastically reduces the uncertainty of labelling decisions(Eco-Innovation: {label_confidences['ei_count']:.2%} to {label_confidences['ei_count_articles']:.2%} and Bioeconomy: {label_confidences['bv_count']:.2%} to {label_confidences['bv_count_articles']:.2%}, for SWINNO and articles, respectively)."))
+```
+
+The only down side to the articles approach is a quintupling of labeling time (from 1 min on average to 5 min on average per innovation).
+
+```{python}
+#| label: tbl-bv-article-count
+#| tbl-cap: Count of Bioeconomy Visions Using Articles
+bv_counts_articles = pd.merge(
+    bv_count_articles, bv_count_zero_articles, on="Vision", how="left"
+)
+
+bv_counts_articles.columns = ["Vision", "Count", "Count Excl. Unsure"]
+bv_counts_articles.loc[:, "Count Excl. Unsure"] = (
+    bv_counts_articles["Count Excl. Unsure"].fillna(0).astype(int)
+)
+
+bv_counts_articles.style.set_properties(**{"text-align": "right"}).hide_index()
+
 ```
@@
 ```{python}
-# |output: asis
+#| label: tbl-ei-article-count
+#| tbl-cap: Count of Eco-Innovations Using Swinno Data
+
 ei_counts_articles = pd.merge(
@@
 # Results
+
+While still early, a first analysis of the labeling results in regards to trends over time and across sectors follows.
+In this section only data based on source articles is presented.
+The figures and tables for SWINNO articles are included in the appendix.
 
@@
 
-At times it can be difficult to decide wether an innovation should be counted as optimization of raw material, reduction of waste, or both.
+At times, it can be difficult to decide whether an innovation should be counted as optimization of raw material, reduction of waste, or both.
 Generally it appears sensible that an optimization of waste can, but does not have to, lead to a reduction of waste.

### 2023-03-23 docs/EHCS-session-abstract.qmd
add to analysis and abstract

---
title: EHCS Session
subtitle: Abstracts
author: Jonas Kreutzer
date: last-modified
abstract: |

csl: https://zotero.org/styles/apa?source=1
bibliography: swinno-shades-green.bib
link-citations: true
link-bibliography: true
colorlinks: true

---

<!-- From Chat GPT -->

Abstract 1:
The Swedish forestry sector has undergone a significant transformation towards a bioeconomy in recent years. This paper presents an economic history analysis of the sector's innovation activity since the 1970s in regards to the three bioeconomy visions: biotechnology, bioresource and bioecology vision. By examining innovation trends and policy developments, we shed light on which vision the Swedish innovation system is pursuing in the forestry sector. Our findings reveal that while the biotechnology vision dominated innovation activity in the past, there has been a significant shift towards the bioresource and bioecology visions in recent years.

Abstract 2:
The bioeconomy has emerged as a new economic paradigm that prioritizes sustainable and renewable resources. The Swedish forestry sector, which accounts for a significant share of the country's economy, has been at the forefront of this transition. This paper presents an economic history analysis of the sector's innovation activity since the 1970s in regards to the three bioeconomy visions: biotechnology, bioresource and bioecology vision.

Abstract 3:
The Swedish forestry sector has been undergoing a significant transformation towards a bioeconomy in recent years. This paper presents an economic history analysis of the sector's innovation activity since the 1970s in regards to the three bioeconomy visions: biotechnology, bioresource and bioecology vision.

---

Abstract 4:
Proponents of a bioeconomy argue it will combine prioritized sustainability with new avenues for economic development.
The Swedish forestry sector is positioning itself at the forefront of this transition nationally and internationally.
However, the aims of this transition are contested in public debate, with three possibilities competing: economic visions centered around biotechnology, bioresources, and bioecology. 
By examining the innovation activities of the sector, the directionality of the sector can be assessed beyond the levels of discourse and funding decisions.
This raises questions to the nuance embedded in sustainability transitions beyond material structure alone and highlights the potential of the Swedish forestry sector as a case study for understanding the challenges and opportunities of sustainability transitions.

### 2023-03-23 docs/paper.qmd
add to analysis and abstract

@@
 
-## Trends
-
-### By Sector
-
-#### Bioeconomy Vision
-
-```{python}
-
-bv_sectors = pd.read_sql(
+## Bioeconomy Vision
+
+Examining the split of bioeconomy visions across sectors in @fig-bva-sectors, reveals a strong concentration of bioresource related innovations in the forestry sector.
+
+```{python}
+#| label: fig-bva-sectors
+#| fig-cap: Heatmap of Bioeconomy Visions by Sector
+
+bva_sectors = pd.read_sql(
     """
@@
   sni_codes.label as sector,
-  count(bv.bioeconomy_vision) as count
-FROM
-  bioeconomy_visions as bv
-  join use_sectors ON bv.sinno_id = use_sectors.[SINNO ID]
+  count(bva.bioeconomy_vision) as count
+FROM
+  bioeconomy_visions_articles as bva
+  join use_sectors ON bva.sinno_id = use_sectors.[SINNO ID]
   join sni_codes ON SUBSTR(use_sectors.use_sectors, 1, 2) = sni_codes.code
-  join [classification-codes] as c ON bv.bioeconomy_vision = c.Code
+  join [classification-codes] as c ON bva.bioeconomy_vision = c.Code
 WHERE 
@@
     FROM
-      bioeconomy_visions AS bv2
-    WHERE
-      bv2.sinno_id = bv.sinno_id
-      AND bv2.bioeconomy_vision = 0
+      bioeconomy_visions_articles AS bva2
+    WHERE
+      bva2.sinno_id = bva.sinno_id
+      AND bva2.bioeconomy_vision = 0
 )
@@
 
-bv_matrix = prepare_heatmap_data(
-    bv_sectors, index_col="sector", cols_col="vision", values_col="count"
-)
-
-plot_heatmap(bv_matrix, "Count Bioeconomy Vision by Sector Excluding Uncertains")
-```
-
-#### Eco-Innovations
-
-```{python}
-ei_sectors = pd.read_sql(
-    """
-SELECT
-  c.Category as innovation_type,
-  sni_codes.label as sector,
-  count(ei.innovation_type) as count
-FROM
-  eco_innovations as ei
-  join use_sectors ON ei.sinno_id = use_sectors.[SINNO ID]
-  join sni_codes ON SUBSTR(use_sectors.use_sectors, 1, 2) = sni_codes.code
-  join [classification-codes] as c ON ei.innovation_type = c.Code
-WHERE 
-  NOT EXISTS(
-   SELECT
-      *
-    FROM
-      eco_innovations AS ei2
-    WHERE
-      ei2.sinno_id = ei.sinno_id
-      AND ei2.innovation_type = "000"
-)
-GROUP BY
-  sector,
-  innovation_type;
-""",
-    conn,
-)
-
-ei_sectors_matrix = prepare_heatmap_data(
-    ei_sectors, index_col="sector", cols_col="innovation_type", values_col="count"
-)
-
-plot_heatmap(ei_sectors_matrix, title="Eco-Innovation Type by Sector")
-
-```
-### Over Time
-
-#### Bioeconomy Visions Using SWINNO
-
-```{python}
-
-bv_years = pd.read_sql(
-    """
-SELECT
-  codes.Category AS Vision,
-  count(bv.bioeconomy_vision) AS vision_count,
-  innovation.[Year of commercialization] AS year
-FROM
-  bioeconomy_visions AS bv
-  JOIN innovation ON bv.sinno_id = innovation.[SINNO ID]
-  JOIN [classification-codes] AS codes ON bv.bioeconomy_vision = codes.Code
-WHERE
-  TRIM(innovation.[Year of commercialization]) IS NOT ""
-GROUP BY
-  year,
-  bv.bioeconomy_vision;
-""",
-    conn,
-)
-
-bvc_years = pd.read_sql(
-    """
-SELECT
-  codes.Category AS Vision,
-  count(bv.bioeconomy_vision) AS vision_count,
-  innovation.[Year of commercialization] AS year
-FROM
-  bioeconomy_visions AS bv
-  JOIN innovation ON bv.sinno_id = innovation.[SINNO ID]
-  JOIN [classification-codes] AS codes ON bv.bioeconomy_vision = codes.Code
-WHERE
-  NOT EXISTS (
-    SELECT
-      *
-    FROM
-      bioeconomy_visions AS bv2
-    WHERE
-      bv2.sinno_id = bv.sinno_id
-      AND bv2.bioeconomy_vision = 0
-  )
-  AND TRIM(innovation.[Year of commercialization]) IS NOT ""
-GROUP BY
-  year,
-  bv.bioeconomy_vision;
-""",
-    conn,
-)
-
-bv_years = bv_years.set_index("year")
-bvc_years = bvc_years.set_index("year")
-
-plot_counts(
-    bv_years,
-    bvc_years,
-    category_col="Vision",
-    y="vision_count",
-    y_label="Count of Vision",
-)
-```
-
-#### Bioeconomy Visions Using Articles
-
-```{python}
+bva_matrix = prepare_heatmap_data(
+    bva_sectors, index_col="sector", cols_col="vision", values_col="count"
+)
+
+plot_heatmap(bva_matrix, "Count Bioeconomy Vision by Sector Excluding Uncertains")
+```
+
+@fig-bva-years suggests no clear pattern in the absolute count of visions assigned to innovations.
+Partially this might stem from a small number of innovations labeled so far, however, no clearer pattern emerges in the larger dataset using only SWINNO either (@fig-bv-years).
+The small number of biotechnology related visions is due to the chemistry sector not yet being included in the labelling.
+
+```{python}
+#| label: fig-bva-years
+#| fig-cap: Count of Bioeconomy Visions by Year
 
@@
 
-At most we record
+## Eco-Innovations
+
+Looking at the eco-innovation types by sector reveals that the biggest category and most heavily used in the forest and timber sector are innovations aimed to prolong the lifecylce of products used therin (@fig-ei-sectors)
+```{python}
+#| label: fig-eia-sectors
+#| fig-cap: Heatmap of Eco-Innovations by Sector
+
+eia_sectors = pd.read_sql(
+    """
+SELECT
+  c.Category as innovation_type,
+  sni_codes.label as sector,
+  count(eia.innovation_type) as count
+FROM
+  eco_innovations_articles as eia
+  join use_sectors ON eia.sinno_id = use_sectors.[SINNO ID]
+  join sni_codes ON SUBSTR(use_sectors.use_sectors, 1, 2) = sni_codes.code
+  join [classification-codes] as c ON eia.innovation_type = c.Code
+WHERE 
+  NOT EXISTS(
+   SELECT
+      *
+    FROM
+      eco_innovations_articles AS eia2
+    WHERE
+      eia2.sinno_id = eia.sinno_id
+      AND eia2.innovation_type = "000"
+)
+GROUP BY
+  sector,
+  innovation_type;
+""",
+    conn,
+)
+
+eia_sectors_matrix = prepare_heatmap_data(
+    eia_sectors, index_col="sector", cols_col="innovation_type", values_col="count"
+)
+
+plot_heatmap(eia_sectors_matrix, title="Eco-Innovation Type by Sector")
+
+```
+
+For the count of eco-innovation types by year, the legend had to be surpressed, as the amount of categories made the plots unreadable.
+Just as for bioeconomy visions @fig-eia-year shows no clear pattern yet.
+@fig-ei-year-swinno does little to improve this.
+The biggest categories in that dataset are "Not Eco-Innovation" (dark blue), and "Automatization" (dark pink).
+Visualizing these data is challenging and suggestions warmly welcome.
+
+```{python}
+#| label: fig-eia-year
+#| fig-cap: Count of Eco-Innovations by Year
+
+eia_years = pd.read_sql(
+    """
+SELECT
+  codes.Category AS Type,
+  count(eia.innovation_type) AS innovation_count,
+  innovation.[Year of commercialization] AS Year,
+  SUBSTR(eia.innovation_type, 1,1) as top_cat
+FROM
+  eco_innovations_articles AS eia
+  JOIN innovation ON eia.sinno_id = innovation.[SINNO ID]
+  JOIN [classification-codes] AS codes ON eia.innovation_type = codes.Code
+WHERE
+  TRIM(innovation.[Year of commercialization]) IS NOT ""
+GROUP BY
+  Year,
+  eia.innovation_type;
+""",
+    conn,
+)
+
+eiac_years = pd.read_sql(
+    """
+SELECT
+  codes.Category AS Type,
+  count(eia.innovation_type) AS innovation_count,
+  innovation.[Year of commercialization] AS year,
+  SUBSTR(eia.innovation_type, 1,1)
+FROM
+  eco_innovations_articles AS eia
+  JOIN innovation ON eia.sinno_id = innovation.[SINNO ID]
+  JOIN [classification-codes] AS codes ON eia.innovation_type = codes.Code
+WHERE
+  NOT EXISTS (
+    SELECT
+      *
+    FROM
+      eco_innovations_articles AS eia2
+    WHERE
+      eia2.sinno_id = eia.sinno_id
+      AND eia2.innovation_type = "000"
+  )
+  AND TRIM(innovation.[Year of commercialization]) IS NOT ""
+GROUP BY
+  year,
+  eia.innovation_type;
+""",
+    conn,
+)
+
+eia_years = eia_years.set_index("Year")
+eiac_years = eiac_years.set_index("year")
+
+plot_counts(
+    eia_years,
+    eiac_years,
+    category_col="Type",
+    y="innovation_count",
+    legend_cols = 8,
+    y_label="Count of Type",
+    show_legend=False
+)
+```
+
+## Trends
+
+So far, there is not much to make of these figures.
+The subset of innovations in SWINNO analyzed is too small to descern any real trends.
+However, there is probable cause to believe that the tendency of dominating bioresource vision related innovations will persist.
+It is unsurprising that in the core forestry sector (logging, timber) these type of innovations are more frequent than e.g., biotechnology related visions.
+Perhpas surprising is that, despite the small number of innovation labelled so far and the methodological disadvantage of LBIO towards bioecology innovations, three were possibly already identified.
+There occurance in the 1980s might suggest taht they were the outcome of a temporary increased awareness of environmental issues in the forestry sector at that time.
 
@@
 
+### SWINNO
+
 #### Bioeconomy Visions by Sector
@@
 ```{python}
+#| label: fig-bv-sectors
+#| fig-cap: Heatmap of Bioeconomy Visions by Sector (SWINNO)
+
 bvu_sectors = pd.read_sql(
@@
 
+#### Bioeconomy Visions Over Time
+
+```{python}
+#| label: fig-bv-years
+#| fig-cap: Count of Bioeconomy Visions by Year (SWINNO)
+
+bv_years = pd.read_sql(
+    """
+SELECT
+  codes.Category AS Vision,
+  count(bv.bioeconomy_vision) AS vision_count,
+  innovation.[Year of commercialization] AS year
+FROM
+  bioeconomy_visions AS bv
+  JOIN innovation ON bv.sinno_id = innovation.[SINNO ID]
+  JOIN [classification-codes] AS codes ON bv.bioeconomy_vision = codes.Code
+WHERE
+  TRIM(innovation.[Year of commercialization]) IS NOT ""
+GROUP BY
+  year,
+  bv.bioeconomy_vision;
+""",
+    conn,
+)
+
+bvc_years = pd.read_sql(
+    """
+SELECT
+  codes.Category AS Vision,
+  count(bv.bioeconomy_vision) AS vision_count,
+  innovation.[Year of commercialization] AS year
+FROM
+  bioeconomy_visions AS bv
+  JOIN innovation ON bv.sinno_id = innovation.[SINNO ID]
+  JOIN [classification-codes] AS codes ON bv.bioeconomy_vision = codes.Code
+WHERE
+  NOT EXISTS (
+    SELECT
+      *
+    FROM
+      bioeconomy_visions AS bv2
+    WHERE
+      bv2.sinno_id = bv.sinno_id
+      AND bv2.bioeconomy_vision = 0
+  )
+  AND TRIM(innovation.[Year of commercialization]) IS NOT ""
+GROUP BY
+  year,
+  bv.bioeconomy_vision;
+""",
+    conn,
+)
+
+bv_years = bv_years.set_index("year")
+bvc_years = bvc_years.set_index("year")
+
+plot_counts(
+    bv_years,
+    bvc_years,
+    category_col="Vision",
+    y="vision_count",
+    y_label="Count of Vision",
+)
+```
+
+#### Eco-Innovations by Sector
+
+```{python}
+#| label: fig-ei-sectors
+#| fig-cap: Heatmap of Eco-Innovations by Sector (SWINNO)
+
+ei_sectors = pd.read_sql(
+    """
+SELECT
+  c.Category as innovation_type,
+  sni_codes.label as sector,
+  count(ei.innovation_type) as count
+FROM
+  eco_innovations as ei
+  join use_sectors ON ei.sinno_id = use_sectors.[SINNO ID]
+  join sni_codes ON SUBSTR(use_sectors.use_sectors, 1, 2) = sni_codes.code
+  join [classification-codes] as c ON ei.innovation_type = c.Code
+WHERE 
+  NOT EXISTS(
+   SELECT
+      *
+    FROM
+      eco_innovations AS ei2
+    WHERE
+      ei2.sinno_id = ei.sinno_id
+      AND ei2.innovation_type = "000"
+)
+GROUP BY
+  sector,
+  innovation_type;
+""",
+    conn,
+)
+
+ei_sectors_matrix = prepare_heatmap_data(
+    ei_sectors, index_col="sector", cols_col="innovation_type", values_col="count"
+)
+
+plot_heatmap(ei_sectors_matrix, title="Eco-Innovation Type by Sector")
+
+```
+
 #### Eco-Innovations Over Time
-
-##### SWINNO
-
-##### Articles
-```{python}
-
-#| fig-caption: Count of Eco-Innovations by Year.
-
-eia_years = pd.read_sql(
+```{python}
+#| label: fig-ei-year-swinno
+#| fig-cap: Count of Eco-Innovations by Year (SWINNO).
+
+ei_years = pd.read_sql(
     """
@@
 FROM
-  eco_innovations_articles AS ei
+  eco_innovations AS ei
   JOIN innovation ON ei.sinno_id = innovation.[SINNO ID]
@@
 
-eiac_years = pd.read_sql(
+eic_years = pd.read_sql(
     """
@@
     FROM
-      eco_innovations_articles AS ei2
+      eco_innovations AS ei2
     WHERE
@@
 
-eia_years = eia_years.set_index("Year")
-eiac_years = eiac_years.set_index("year")
+ei_years = ei_years.set_index("Year")
+eic_years = eic_years.set_index("year")
 
 plot_counts(
-    eia_years,
-    eiac_years,
+    ei_years,
+    eic_years,
     category_col="Type",

### 2023-03-24 docs/paper.qmd
update version sent to astrid and josef

@@
 ```{python}
+#| label: import-functions
 from pathlib import Path
@@
 The classification categories for eco-innovation types were sourced from a single paper and augmented during the classification process.
-In addition to the 30 firm performance indicators identified by García-Granero's [-@garcia-granero2018EcoinnovationMeasurementReview] review of management literature, XX categories were added.
-Some are mere extensions of their categories (e.g., specifying if an innovation prolongs the life cycle of the product itself or downstream products), others are new types of existing categories (e.e.g., reduced other pollution such as noise), yet others are entirely new categories (human health and safety).
+In addition to the 30 firm performance indicators identified by García-Granero's [-@garcia-granero2018EcoinnovationMeasurementReview] review of management literature, 15 categories were added.
+Some are mere extensions of their categories (e.g., specifying if an innovation prolongs the life cycle of the product itself or downstream products), others are new types of existing categories (e.e.g., reduced other pollution such as noise), yet others are entirely new categories (human health and safety), the categories of automatization and precision by themselves are not eco-innovations, but can be used to gain a deeper understanding of what makes an innovation an eco-innovation.
 Clearly not all new categories are eco-innovations.
@@
 These include, as alluded to above, expanding the labeling for other sustainability criteria, such as human health and safety, and specific forest related aspects / tasks to gain a more detailed understanding of an innovations aim.
-For example, XX innovations aim to protect harvested timber, either at harvest, during transport or storage.
+For example, some innovations aim to protect harvested timber, either at harvest, during transport or storage.
 Since these could be tagged in different ways e.g., longer lifecycle of (downstream) product, optimized use of raw materials, or even reduced waste, it seemed more appropriate to create a specific category for these cases.
@@
 
-The only down side to the articles approach is a quintupling of labeling time (from 1 min on average to 5 min on average per innovation).
+The only downside to the articles approach is a quintupling of labeling time (from 1 min on average to 5 min on average per innovation).
 
@@
 
-While still early, a first analysis of the labeling results in regards to trends over time and across sectors follows.
+While still early, a first analysis of the labeling results regarding trends over time and across sectors follows.
 In this section only data based on source articles is presented.
@@
 @fig-bva-years suggests no clear pattern in the absolute count of visions assigned to innovations.
-Partially this might stem from a small number of innovations labeled so far, however, no clearer pattern emerges in the larger dataset using only SWINNO either (@fig-bv-years).
+Partially this might stem from the small number of innovations labeled so far, however, no clearer pattern emerges in the larger dataset using only SWINNO either (@fig-bv-years).
 The small number of biotechnology related visions is due to the chemistry sector not yet being included in the labelling.
@@
 
-Looking at the eco-innovation types by sector reveals that the biggest category and most heavily used in the forest and timber sector are innovations aimed to prolong the lifecylce of products used therin (@fig-ei-sectors)
+Looking at the eco-innovation types by sector reveals that the biggest category and most heavily used in the forest and timber sector are innovations aimed to prolong the lifecycle of products used therein (@fig-ei-sectors)
 ```{python}
@@
 
-For the count of eco-innovation types by year, the legend had to be surpressed, as the amount of categories made the plots unreadable.
+For the count of eco-innovation types by year, the legend had to be suppressed, as the amount of categories made the plots unreadable.
 Just as for bioeconomy visions @fig-eia-year shows no clear pattern yet.
@@
 So far, there is not much to make of these figures.
-The subset of innovations in SWINNO analyzed is too small to descern any real trends.
+The subset of innovations in SWINNO analyzed is too small to discern any real trends.
 However, there is probable cause to believe that the tendency of dominating bioresource vision related innovations will persist.
 It is unsurprising that in the core forestry sector (logging, timber) these type of innovations are more frequent than e.g., biotechnology related visions.
-Perhpas surprising is that, despite the small number of innovation labelled so far and the methodological disadvantage of LBIO towards bioecology innovations, three were possibly already identified.
-There occurance in the 1980s might suggest taht they were the outcome of a temporary increased awareness of environmental issues in the forestry sector at that time.
+Perhaps surprising is that, despite the small number of innovation labelled so far and the methodological disadvantage of LBIO towards bioecology innovations, three were possibly already identified.
+There occurrence in the 1980s might suggest that they were the outcome of a temporary increased awareness of environmental issues in the forestry sector at that time.
 
@@
 ## Open Questions and Doubts
+
+<!-- ```{python}
+#| label: tbl-bioeco-examples
+#| tbl-cap: Examples of Bioecology Vision
+
+bv3_examples = pd.read_sql(
+  """
+  SELECT
+  bv.sinno_id AS [SINNO ID],
+  innovation.[Description in Swedish]
+FROM
+  bioeconomy_visions_articles AS bv
+JOIN innovation ON bv.sinno_id = innovation.[SINNO ID]
+WHERE
+bv.bioeconomy_vision = 3 
+AND
+  NOT EXISTS (
+    SELECT
+      *
+    FROM
+      bioeconomy_visions_articles AS bv2
+    WHERE
+      bv2.sinno_id = bv.sinno_id
+      AND bv2.bioeconomy_vision = 0
+  );
+  """
+  , conn
+)
+
+def wrap_text(s):
+    return 'white-space: pre-wrap'
+
+styled_df = bv3_examples.style.set_properties(subset=[bv3_examples.columns[-1]], **{'width': '70%', 'max-width': '80%', 'text-align': 'justify', 'word-wrap': 'break-word', 'white-space': 'pre-wrap'})
+
+# display the styled dataframe
+display(Markdown(styled_df.hide_index().render()))
+
+```-->
+
+| SINNO ID | Description in Swedish |
+|:---|:---|
+| 7440001  | Trämelass.
+Trämelassen består av de mest lättlösliga substanserna i vedråvaran för skivframställning  vilka går i lösning i tillverkningsprocessen och inblandas i processvattnet.
+Sedan mitten av 1960-talet återvinnes substanserna koncentreras och torkas samt ges en viss efterbehandling som gör dem lämpliga som bl a kreatursfoder.
+Produkten är alltså dels rent kreatursfoder och dels en ingående ingredients i kreatursfoder.
+|
+| 7441001  | Maskinsystem för småskalig avverkning av energiskog (ESO).                                                          
+| 7446001  | Maskinsystem för storskalig avverkning av energiskog (ESO) i företrädesvis våtmark. I systemet ingår både skördare och buntare. |                                                      
+| 7508001  | Elsåg för motorsågsdrivna sågverk. Elsågen håller en jämnare hastighet jämfört med en dieseldriven såg det medför lägre bränslekostnader.
+| 7555001  | Patenterad lösning för stegdämpning i trähus. Innovationen består av en extremt kraftig gummilist som tar upp vibrationer. |
+: Examples of Bioecology Vision {#tbl-bioeco-examples tbl-colwidths="[15,85]"}
+
+@tbl-bioeco-examples shows the five innovations tagged without uncertainty as belonging to the bioecology vision.
+While the first is perhaps the clearest example, the other three are not as obvious.
+After some deliberation, I decided that they ought to belong to the bioecology vision, as they all include strong territorial identity.
+But I have doubts if that is true enough.
+
+Further, doubts regarding the overall approach come from the fact that it feels very subjective at times, and I struggle with designing a more rigorous approach to classification of bioeconomy visions.
 

### 2023-04-11 synopsis.qmd
big git add-mess

---
title: Synopsis for Umeå
---
With the first paper of my thesis, I contribute to the bioeconomy transition debate in Sweden.
Previous research has mapped out the discourse, its central actors and their key aims, finding three competing visions for a bioeconomy.
The biotechnology vision is most dominant outside the EU, focusing on the scientific exploitation of biological material for new technologies.
The bioresource vision, heavily promoted by the Swedish forstry industry, focuses on processess and upgrading of product value chains.
A counter vision centering on ecological aspects draws these first visions into question, proposing instead a reoriented economy embedded within the ecological limits.

By classifiying significant Swedish innovations from 1970 onwards identified by the SWINNO project, accroding to their corresponding visions, my paper moves beyond intent and analyses the concrete direction taken by the transition.
At Umeå I look forward to presenting preliminary results and the method used for classifying the innovations.
Input on the validity of the approach as well as discussions about strategies to mitigate data biases favoring the biotechnology and bioresource vision are much appreciated.
I will circulate the data together with a detailed method description beforehand.

For more information about my PhD project and how this paper fits in, please visit [pjkreutzer.github.io/research](pjkreutzer.github.io/research), where you will also find a research plan.

### 2023-04-11 rename_files.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+# specify the directory to perform the replacement in
+dir="scripts"
+
+# loop through all files and directories in the specified directory
+for file in "$dir"/*; do
+    # check if the file name contains "-"
+    if [[ "$file" == *-* ]]; then
+        # replace "-" with "_"
+        new_file="${file//-/_}"
+        # rename the file with the new name
+        mv "$file" "$new_file"
+    fi
+done

### 2023-04-11 scripts/find_sources.py

@@ -5,13 +5,36 @@ from src.utils import get_project_root
 from collections import defaultdict
 
 parser = argparse.ArgumentParser()
-parser.add_argument("-i", "--input", required=True, help="Text file containing input ids.")
+parser.add_argument(
+    "-i", "--input", required=True, help="Text file containing input ids."
+)
 
-parser.add_argument("-l", "--lookup", required=True, help="Text file containing id lookup pairs.")
+parser.add_argument(
+    "-l",
+    "--lookup",
+    required=False,
+    type=Path,
+    default="notes/innovation_id_to_source_id.txt",
+    help="Text file containing id lookup pairs.",
+)
 
-parser.add_argument("-s", "--source", type=Path, default=Path("/mnt/c/Users/ph8148kr/Lund University/Mathias Johansson - non_missing_source_materials/images").absolute(), help="Path for source of files.")
+parser.add_argument(
+    "-s",
+    "--source",
+    type=Path,
+    default=Path(
+        "/mnt/c/Users/ph8148kr/Lund University/Mathias Johansson - non_missing_source_materials/images"
+    ).absolute(),
+    help="Path for source of files.",
+)
 
-parser.add_argument("-d", "--destination", required=False, type=Path, help="Destination path in which to create destination dir")
+parser.add_argument(
+    "-d",
+    "--destination",
+    required=False,
+    type=Path,
+    help="Destination path in which to create destination dir",
+)
 
 args = vars((parser.parse_args()))
 
@@ -19,20 +42,34 @@ ROOT = get_project_root()
 
 
 def copy_sources():
-    input_path = Path(ROOT, args["input"])
-    lookup_path = Path(ROOT, args["lookup"])
+
+    if not Path(ROOT, args["input"]).exists():
+        raise FileNotFoundError(f"Input file {args['input']} does not exist.")
+    else:
+        input_path = Path(ROOT, args["input"])
+
+    if not Path(ROOT, args["lookup"]).exists():
+        raise FileNotFoundError(f"Lookup file {args['lookup']} does not exist.")
+    else:
+        lookup_path = Path(ROOT, args["lookup"])
     source_path = args["source"]
-    destination_path = args["destination"]
-    
+
+    if not Path(ROOT, args["destination"]).is_dir():
+        Path(ROOT, args["destination"]).mkdir(parents=True)
+
+    destination_path = Path(ROOT, args["destination"])
+
     input_ids = get_input_ids(input_path)
     source_names = get_source_names(lookup_path, input_ids)
[…]

### 2023-09-07 docs/reports/paper/paper1.qmd
big add

---
title: The Direction of Search for a Forest Based Bioeconomy
author: 
    name: Jonas Kreutzer
    affiliation: Department of Economic History, Lund University
date: last-modified
abstract: |

csl: https://zotero.org/styles/apa?source=1
bibliography: [../../swinno-shades-green.bib]
link-citations: true
link-bibliography: true
colorlinks: true
format:
    pdf
execute: 
    include: false
---

```{python}
#| label: Imports

from IPython.display import display, Markdown
import pandas as pd
```

# Debate around Bioeconomies

The development of a bioeconomy firmly features at the top of many national and international economic agendas.
In Europe, 2012 saw the publication of a EU wide bioeconomy strategy, which outlined core tenenats sourrounding the aims and methods by which economic inputs with biological origin are to replace those based on fossile materials.

The concept and idea of a bioeconomy is hence not novel.
With progressing concern about environmental degradation the discussion around a bioeconomy has intensified in recent years.

## Advanced Research Fronts

Public, scientific and political debate of bioeconomies has brought forth a number of discourse analyses.
Both @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy conclude that the concept of a bioeconomy conceals different, possibly conflicting visions of transfromation outcomes.

# Context of Swedish Forest Based Bioeconomy

## Size of Forestry Sector

##

# Data and Methods

To capture the actual innovation output associated with the Swedish forest based bioeconomy the SWINNO database was searched for innovation associated with core forest sectors and keywords describing forest products which may be used in sectors along the bioeconomy value chain.
The list of keywords and sectors is presented in the appendix.
Due to the query matching keywords whenever they appear as substrings in the description of the innovation, the inital search result is rather susceptible to false positives.
In later manual review of each innovation for classification these false positives were removed from the search results.

The SWINNO database is constructed according to  literature-based innovation output (LBIO) method.
This method relies on the judgment of editiors and journalists at trade journals and hence captures what can be described as significant innovation.
Measuring innovation with LBIO has substantial advantages.
For one, it ensures that actual innovation, meaning commericialized products and processes, are captured.
Second, because the source articles are preserved, it provides a rich source of data, giving frequently providing insights into the aims, motivations, advantages and challenges of the covered innovation.

An important limitation of LBIO measures, especially for research into forest sector innovation, is a bias against non-commercialized innovation.
New processes, or new organizational apporaches for example, will only be caputred if they are sold to another company.
This is especially sensitive for bioeconomy research as, for example, new methods or plans for managing forest stands are less likely to be included.
Relevant product innovation which faciliates these plans, for example products related to the planting of trees, on the other hand, are captured.

A likely consequence of this limitation is that the number of innovation related to the bioecology vision are under counted, as this vision highlights "Counter-expertise rather than concrete technical solutions" [@vivien2019HijackingBioeconomy 194].
Nonetheless, a small number of innovation associated with this vision were henceforth identified in the database.

## Classification Methodology

Innovations were classified as belonging to a bioeconomy vision if they fulfilled one, or more criteria outlined in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy.
@tbl-tbl-bioeconomy_vision_summary shows the summary of bioeconomy visions presented in the cited papers.

Economic growth is cited as the aim & objective of biotechnology and bioresource visions by @bugge2016WhatBioeconomyReview.

It can be argued that all innovation can contribute to economic growth, which is why these aspects were not given hightened importance in the classification decision.

<!-- Use examples of clear cases here -->

# Results

```{python}
#| label: Queries

```

```{python}
#| label: Number of Results

display
```

# References

:::{#refs}
:::

# Appendix {.appendix}

| @bugge2016WhatBioeconomyReview [10] | Biotechnology | Bioresource | Bioecology |
|---|---|---|---|
| Aim &   Objectives | Economic growth & job   creation | Economic growth &   sustainability | Sustainability, biodiversity,   conservation of ecosystems, avoiding soil degradation |
| Value   Creation | Application of biotechnology,   commercialisation of research & technology | Conversion and upgrading of   bio-resources (process oriented) | Development of integrated   production systems and high-quality products with territorial identity |
| Drivers   & mediators of innovation | R & D, patents, TTOs,   Research councils and funders (Science push, linear model) | Interdisciplinary, optimisation   of land use, include degraded land in the production of biofuels, use and   availability of bio-resources, waste management, engineering, science &   market (Interactive & networked production mode) | Identification of favourable   organic agro-ecological practices, ethics, risk, transdisciplinary   sustainability, ecological interactions, re-use & recycling of waste,   land use, (Circular and self-sustained production mode) |

| @vivien2019HijackingBioeconomy [194] | Type II | Type III | Type I |
|---|---|---|---|
|  | A science-based economy driven   by industrial biotechnology<br>     <br>     The cell is a factory | Biomass replaces fossil fuels   and mining to produce energy and materials<br>     <br>     Biorefining at the heart of ecological transition (multilevel perspective). | An ecological economy, that is   compatible with the biosphere<br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
| Classificaiton Code | 1 | 2 | 3 |

<!-- Definition An ecological economy, that is compatible with the biosphere A science-based economy driven by industrial biotechnology A biomass-based economy Nature/economy relations Struggle against entropy and coevolution with the biosphere. Economic development in line with biological evolution The cell is a factory Technology has the power to “correct God's mistakes” Biomass replaces fossil fuels and mining to produce energy and materials Science, technology and society Megatrends with “promethean technologies” Pending the advent of Prometheus III, the economics of prudence and sharing Prospect of a fifth Kondratiev wave based on the rise of biotechnology. The “economy of techno-scientific promises” Biorefining at the heart of ecological transition (multilevel perspective). The economy of learning Sustainability “Strong sustainability” approach and degrowth perspective Very “weak sustainability” approach “Weak sustainability” approach -->

<!-- TODO merge tables -->

: Bioeconomy Vision Summary {#tbl-bioeconomy_vision_summary}

### 2023-09-07 scripts/filter_sectors.py

@@ -1,0 +1,43 @@
+import sqlite3
+import pandas as pd
+
+
+def unite(df, old_col_start: str, new_col_name: str, position: str):
+    temp_df = df
+    col_list = [col for col in temp_df if col.startswith(old_col_start)]
+    temp_col = temp_df[col_list].astype(str).apply(lambda x: x.str.cat(sep=""), axis=1)
+    temp_df = temp_df.drop(columns=col_list)
+    temp_df.insert(loc=position, column=new_col_name, value=temp_col)
+
+    return temp_df
+
+
+db = sqlite3.connect("C:\\Users\ph8148kr\OneDrive - Lund University/research/swinno.db")
+
+df = pd.read_sql(
+    "SELECT [SINNO ID], [Use sectors],  [Use sector 2], [Use sector 3],[Use sector 4], [Use sector 5], [Use sector 6], [Use sector 7], [Use sector 8], [Use sector 9], [Literal name of innovating firm ; SINNO's common name ; SINNO ID], [Innovation name in Swedish], [Description in Swedish], [Additional information if Origin = New scientific discovery],[Additional information if Origin = New technologies or materials], [Additional info if Origin = Official regulation, legislation and standards.], [Additional information if Origin = Solution for a problem.], [Additional information if Origin = Performance], [Additional information if Origin = Other] FROM innovation",
+    db,
+)
+
+df.columns = df.columns.str.strip().str.lower().str.replace(" ", "_")
+
+sni_codes = [
+    "01",
+    "02",
+    "13",
+    "16",
+    "17",
+    "18",
+    "20",
+    "38",
+    "41",
+]
+
+df = unite(df, "use", "use_sectors", position=1)
+df = unite(df, "add", "info", position=5)
+
+filtered_df = df[df["use_sectors"].isin(sni_codes)]
+
+filtered_df.to_excel(
+    "data/modified-data/filtered_swinno_1.xlsx", encoding="utf-8-sig", index=False
+)

### 2023-09-07 scripts/test.py

@@ -1,0 +1,4 @@
+import  sys
+
+# Print command-line arguments
+print(sys.argv)

### 2023-09-07 src/categorization_helpers.py

@@ -5,9 +5,11 @@ import re
 
 def clean_codes(df, column, code_digits, remove_whitespace=True, drop_na=True):
     if drop_na:
-        df = df.dropna(subset=[column])
+        df = df.dropna(subset=[column]).copy()
 
     df.loc[:, column] = df[column].astype(str)
+
+    df.loc[:, column] = df[column].apply(lambda x: re.sub(r";", r",", x))
 
     df.loc[:, column] = df[column].apply(
         lambda x: re.sub(r"\b(0)\b", r"0" * code_digits, x)
@@ -19,6 +21,8 @@ def clean_codes(df, column, code_digits, remove_whitespace=True, drop_na=True):
     df.loc[:, column] = df[column].apply(
         lambda x: re.sub(rf"(\d{ {code_digits} })(?!$)", r"\1,", str(x))
     )
+
+    df.loc[:, column] = df[column].apply(lambda x: re.sub(r",,", r",", str(x)))
 
     if remove_whitespace:
         df.loc[:, column] = df[column].apply(lambda x: re.sub(r"\s", "", x))

### 2023-09-07 src/file_utils.py

@@ -60,10 +60,6 @@ def get_source_names(lookup_path, input_ids):
 
     return {k: lookup[k] for k in set(lookup).intersection(input_ids)}
 
-
-from pathlib import Path
-from shutil import copy
-
 def copy_files(source_links: dict, source_dir: Path, destination_dir: Path):
     """
     Copies files from the source directory to the destination directory based on the provided source links.

### 2023-09-07 src/swinno_helpers.py

@@ -24,7 +24,7 @@ def split_cols(df, col_to_split: str, sep: str):
     )
 
     for col in split_cols:
-        temp_df[f"{col_to_split}_{col}"] = split_cols.loc[:, col]
+        temp_df.loc[:, f"{col_to_split}_{col}"] = split_cols.loc[:, col]
 
     temp_df.drop(labels=col_to_split, axis=1, inplace=True)
 
@@ -57,13 +57,13 @@ def melt_table(df, id_vars: str, col_start: str, value_name: str, **kwargs):
 
     return temp_df
 
+
 def clean_column_names(df):
     df.columns = df.columns.str.strip().str.lower().str.replace(" ", "_")
     return df
 
 
 def clean_import(file_type, file_path, **kwargs):
-
     if file_type == "csv":
         df = pd.read_csv(file_path, **kwargs)
 
@@ -76,8 +76,8 @@ def clean_import(file_type, file_path, **kwargs):
     df = clean_column_names(df)
     return df
 
-def connect_swinno_db():
 
+def connect_swinno_db():
     database_dir = get_project_root().parent.absolute()
     database_uri = f"sqlite:///{database_dir}/swinno-db/data/swinno.db"
     engine = create_engine(database_uri)

### 2023-09-18 docs/reports/paper/paper1.qmd
add text from phd plan & sql query

@@
 abstract: |
+    The transition to a bioeconomy is high on the agenda for research, industry and policy.
+Beyond conceptual uncertainties of what is meant by a bioeconomy, three different visions have emerged, differing in their aims and objectives, drivers and mediators of innovation and relation to the biosphere.
+The use of detailed innovation data from the SWINNO data base allows to capture the actual innovation output of the Swedish forest based bioeconomy innovation system.
+Qualitative coding of the underlying source material associates innovation from 1970 to 2020 with vision of a bioresource, biotechnology and / or bioecology bioeconomy.
+Preliminary results suggest that the overall innovation output of the forest based bioeconomy has followed overall trends in the SWINNO database, failing to gain in prominence.
+Of the three visions, the bioresource vision has been the most dominant, especially in the 1970s and after 2005.
+At the same time, innovation which fit into multiple visions exist, suggesting that the opportunity space for future development is not yet locked in.
 
@@
 format:
-    pdf
+    wordcount-pdf: default
 execute: 
-    include: false
+    echo: false
 ---
@@
 ```{python}
-#| label: Imports
+# | label: Imports
+# | eval: true
 
+import pandas as pd
 from IPython.display import display, Markdown
-import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+from src.figures import *
+from src.utils import get_project_root, format_table
+from src.swinno_helpers import connect_swinno_db
+
+ROOT = get_project_root()
+conn = connect_swinno_db()
+
+# set all text on plots to be monospace with rcParams
+plt.rcParams["font.family"] = "monospace"
+
+sns.set_theme(style="whitegrid")
+
 ```
-
 # Debate around Bioeconomies
 
-The development of a bioeconomy firmly features at the top of many national and international economic agendas.
-In Europe, 2012 saw the publication of a EU wide bioeconomy strategy, which outlined core tenenats sourrounding the aims and methods by which economic inputs with biological origin are to replace those based on fossile materials.
+An idea which is bound to increase the pressure and trade-offs between different ecosystem services is the bioeconomy.
+In recent years the concept has gained relevance among policymakers, practitioners and researchers.
+The term bioeconomy refers broadly to an economy which uses bio-based inputs instead of fossil based ones.
+Connected to this are hopes of increased resource efficiency and renewability.
+In a report commissioned by the Swedish Government, the @2012SwedishResearchInnovation defines a vision of a Swedish bioeconomy based on two central pillars.
+It is to be an economy based on,
 
-The concept and idea of a bioeconomy is hence not novel.
-With progressing concern about environmental degradation the discussion around a bioeconomy has intensified in recent years.
+> \[a\] sustainable production of biomass to enable increased use within a number of different sectors of society. The objective is to reduce climate effects and the use of fossil-based raw materials.
 
-## Advanced Research Fronts
+With,
+
+> \[a\]n increased added value for biomass materials, concomitant with a reduction in energy consumption and recovery of nutrients and energy as additional end products.
+The objective is to optimize the value and contribution of ecosystem services to the economy.
+(p.
+9).
+
+These pillars reflect the two most influential visions the bioeconomy concept.
+Notions of a bio-technology and bio-resource oriented bioeconomy are widespread and dominate discourse at various levels [@dieken2021MultitudesBioeconomiesSystematic].
+Despite -- or perhaps because of -- the bioeconomy lacking a strong definition, it is used to refer to different concepts, with different aims and objectives, foci, normative assumptions and hence implications for society and environment [@bugge2016WhatBioeconomyReview].
+
+The bio-technology vision focuses on economic growth and job provision through the application of bio-technology [@bugge2016WhatBioeconomyReview].
+Science driven innovation features prominently in this vision, with collaborations between research institutions in academia and industry taking on an important role [@bugge2016WhatBioeconomyReview].
+@vivien2019HijackingBioeconomy present a review of the concept showing that the focus on science as an absolute motor of change also shaped policy recommendations, driven especially by the OECD, towards increasing the speed and diffusion of innovation through partnerships between new start-ups and establish pharmaceutical companies.
+As a policy narrative this vision was particularly potent and widespread at the turn of the 21st century.
+According to them, the intellectual heritage of Schumpeter and Kontradieff play an important role in this vision.
+Technological breakthroughs are poised to solve economic and ecological challenges, for example through genetic engineering.
+Consequently, this vision has a weak sustainability notion, viewing sustainability challenges as temporary and nature as substitutable given sufficient research and knowledge creation [@vivien2019HijackingBioeconomy].
+
+In more recent years, visions of a bio-resource bioeconomy have taken over as the leading narrative.
+Unlike the bio-technology vision, this conceptualization does not only focus on economic growth and job creation, but attempts to include ecological sustainability challenges [@bugge2016WhatBioeconomyReview].
+Biological resources, not technologies, form the basis of a new economy.
+In essence, there are two ways in which bio-resources are poised to replace fossil resources: as substitution, e.g., bio-fuels replacing fossil fuels, and as high value added products, e.g., Cross-laminated Timber used as a concrete alternative in construction [@vivien2019HijackingBioeconomy].
+To substitute fossil resources requires significant efforts in innovation.
+In contrast to the bio-technology vision, however, innovation is not confined to scientific advances from a narrow biochemical and pharmaceutical field.
+Instead, new and old actors are required to collaborate between and beyond established sectors [@bugge2016WhatBioeconomyReview].
+Although sustainability plays a more important role in this vision of the bioeconomy, this vision is considered an example of weak sustainability [@damato2017GreenCircularBio; @vivien2019HijackingBioeconomy].
+The reason is the focus of substitution and high biomass demands.
+The consumption of raw materials may even increase, not only for bio-based, but also for fossil-based resources [@asada2020BioeconomicTransitionProjecting].
+
+The exact role played by forests in a bioeconomy transition depends on the desired vision [@kleinschmit2014ShadesGreenSocial].
+There is no doubt that the Swedish vision places high demands on forests as primary providers of the required biomass.
+Both in the strategy formulated in @2012SwedishResearchInnovation and in public discourse, a small, well-connected group of actors center the bio-resource vision as legitimate and desirable [@holmgren2022WhoseTransformationThis].
+The result is a forest sector which operates under a "more of everything" paradigm [@belandlindahl2015SwedishForestryModel].
+
+#### Innovation for a Bioeconomy
+
+Apart from the substitution of fossil resources through bio-based resources, a central narrative used in the Swedish transition is that, "\[t\]echnological innovation is key to a greener future" [@holmgren2022WhoseTransformationThis 42].
+According to @jankovsky2021InnovationsForestBioeconomy four types of innovation are central for a forest based bioeconomy:
+
+1.  Substitute Products,
+2.  New (bio-based) Processes,
+3.  New (bio-based) Products,
+4.  New Behavior.
+
+Currently, most research into bioeconomy innovation relies on patent data or qualitative case studies [e.g., @bennich2021BiobasedEconomy2030; @jander2019MonitoringTransitionBioeconomy].
+Unfortunately, these approaches are not without their flaws.
+Focusing on patents could overestimate the importance of the bio-technology vision, whose science oriented innovation approach is driven by patents [@bugge2016WhatBioeconomyReview].
+Qualitative case studies, on the other hand, may miss broader trends and features of a bioeconomy transition.
+
+@hausknost2017TransitionWhichBioeconomy
 
@@
 
-# Context of Swedish Forest Based Bioeconomy
+Recent scholarship has called for advancing research fronts instead of remaining entrenched in contested discourse positions @allain2022BioeconomicsVsBioeconomy.
 
-## Size of Forestry Sector
+@jankovsky2021InnovationForestBioeconomy
+
+@bauer 
+refineries
+
+Research Question:
+
+While much research has been done on the research, less is known over the actual direction of technical change in the Swedish forest bioeconomy.
+How innovative, in terms of output, is the system?
+And what vision does the innovation system pursue?
 
@@
 
-The SWINNO database is constructed according to  literature-based innovation output (LBIO) method.
+The SWINNO database is constructed according to  literature-based innovation output (LBIO) method [@sjoo2014DatabaseSwedishInnovations].
 This method relies on the judgment of editiors and journalists at trade journals and hence captures what can be described as significant innovation.
-Measuring innovation with LBIO has substantial advantages.
-For one, it ensures that actual innovation, meaning commericialized products and processes, are captured.
-Second, because the source articles are preserved, it provides a rich source of data, giving frequently providing insights into the aims, motivations, advantages and challenges of the covered innovation.
-
 An important limitation of LBIO measures, especially for research into forest sector innovation, is a bias against non-commercialized innovation.
@@
 A likely consequence of this limitation is that the number of innovation related to the bioecology vision are under counted, as this vision highlights "Counter-expertise rather than concrete technical solutions" [@vivien2019HijackingBioeconomy 194].
-Nonetheless, a small number of innovation associated with this vision were henceforth identified in the database.
+Nonetheless, a small number of innovation associated with this vision were already identified in the database.
+
+But, measuring innovation with LBIO has also substantial advantages.
+For one, it ensures that actual innovation, meaning commericialized products and processes, are captured.
+Second, because the source articles are preserved, it provides a rich source of data, giving frequently providing insights into the aims, motivations, advantages and challenges of the covered innovation.
 
@@
 
+These source articles were carefully read and their associated unique identifiers in the database tagged with two types of labels.
+The first label is the main interest of this study: the bioeconomy vision an innovation can be associated with.
+The second label provides a deeper understanding if and how the innovation reduces environmental impacts.
+As @broring2020InnovationTypesBioeconomy highlight, not all bioeconomy innovations need to be sustainability oriented innovation.
+
+### Bioeconomy Visions
+
 Innovations were classified as belonging to a bioeconomy vision if they fulfilled one, or more criteria outlined in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy.
-@tbl-tbl-bioeconomy_vision_summary shows the summary of bioeconomy visions presented in the cited papers.
+@tbl-bioeconomy_vision_summary shows the summary of bioeconomy visions presented in the cited papers.
 
-Economic growth is cited as the aim & objective of biotechnology and bioresource visions by @bugge2016WhatBioeconomyReview.
+While following these criteria, a number of decisions were made concerning content of the summary presented in @tbl-bioeconomy_vision_summary and the nature of the corresponding vision.
+For one, economic growth is cited as the aim & objective of biotechnology and bioresource visions by @bugge2016WhatBioeconomyReview.
+It can be argued that all innovation can contribute to economic growth, which is why these aspects were not given hightened importance in the classification decision.
+A second topic of judgment has to be the inclusion of innovation which increases the recyclability of a product, or uses recycled waste, water and material in the process.
+Based on the definition of a bioecology vision presented in @bugge2016WhatBioeconomyReview, these innovations should be included as contributing to this vision.
+The extent to which such innovation contributes in reality to a "[c]ircular and self-sustained production mode" [@bugge2016WhatBioeconomyReview 10], however, can be debated.
 
-It can be argued that all innovation can contribute to economic growth, which is why these aspects were not given hightened importance in the classification decision.
+```{python}
+# | label: Count Recycling Innovation
+
+recycle_innovation = pd.read_sql(
+    """
+SELECT *
+FROM classification_codes
+LEFT JOIN eco_innovations ON classification_codes.Code = eco_innovations.innovation_type
+WHERE eco_innovations.innovation_type = 107 OR eco_innovations.innovation_type = 206 AND sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes LIKE '%not forest%')
+""",
+    conn,
+)
+
+recycle_innovation_count = recycle_innovation["sinno_id"].nunique()
+
+display(
+    Markdown(
+        f"Currently, {recycle_innovation_count} innovations were identified which involve either a product or process innovation related to recyling."
+    )
+)
+```
+
+<!-- TODO count all innovation with eco-innovation type 107 and 206 as bv 3 -->
+
+@tbl-vision_examples shows five examples of classfied innovation.
+This selection also highlights that multible labels can be attached to one innovation.
+The last innovation, describing a waterproof paper, is classified as a bioresource vission innovaition because it describes a process by which biomass is converted and upgraded.
+It is also classified as a biotechnology innovation since the process mimics trees' microfiber formation, accroding to the innovation source article.
+
+```{python}
+# | label: tbl-vision_examples
+# | tbl-cap: "List of Eco-Innovation Labels"
+
+classification_examples = pd.read_sql(
+    """SELECT
+  n.sinno_id,
+  i.description_in_swedish,
+  GROUP_CONCAT(c.Category, ' & ') AS Categories,
+  n.notes
+FROM
+  categorization_notes n
+  JOIN innovation i ON i.sinno_id = n.sinno_id
+  JOIN bioeconomy_visions bv
+  ON n.sinno_id = bv.sinno_id
+  JOIN classification_codes c ON bv.bioeconomy_vision = c.Code
+WHERE
+  n.notes LIKE "%good example%"
+GROUP BY
+  n.sinno_id,
+  i.description_in_swedish,
+  n.notes""",
+    conn,
+)
+
+classification_examples_out = (
+    classification_examples.copy()
+    .loc[:, ["Categories", "description_in_swedish"]]
+    .iloc[2:]
+)
+
+rowIDs = range(1, len(classification_examples_out) + 1)
+display(
+    Markdown(
+        classification_examples_out.rename(
+            columns={"description_in_swedish": "Short Description"}
+        ).to_markdown(index=rowIDs)
+    )
+)
+
+```
+
+### Eco-Innovation Types
+
+Understanding the bioeconomy vision that an innovation is contributing to is the focus of this paper.
+In order to gain a deeper understanding of the nature of innovation within each bioeconomy vision, however, the innovation were also tagged with labels of eco-innovation types.
+This classification uses eco-innovation types identified in a recent literature review by @garcia-granero2018EcoinnovationMeasurementReview.
+They review 104 publications and produce a list 30 key Eco-Innovation Perfromance Indicators of product, process, organizational and marketing innovation.
+
+This initial list was tested on a subset of innovation in SWINNO and augmented with 15 categories.
+Some are mere extensions of their categories (for example, specifying if an innovation prolongs the life cycle of the product itself or downstream products), others are new types of existing categories (for example, reduced other pollution such as noise), yet others are entirely new categories (human health and safety), the categories of automatization and precision by themselves are not eco-innovations, but can be used to gain a deeper understanding of what makes an innovation an eco-innovation.
+Clearly not all new categories are eco-innovations.
+The newly created pesticide category, for example, can be an eco-innovation if it reduces chemical pollution, but it could also increase chemical pollution if pesticides are introduced in novel ways.
+
+For the eco-innovations the decision process is straightforward.
+Based on the information contained in the text, the innovation is labeled with one or more categories.
+In case of uncertainty, the doubt label is added.
+Because the list of eco-innovation types was relatively exhaustive from the beginning, few adjustments had to be made.
+These include, as alluded to above, expanding the labeling for other sustainability criteria, such as human health and safety, and specific forest related aspects / tasks to gain a more detailed understanding of an innovations aim.
+For example, some innovations aim to protect harvested timber, either at harvest, during transport or storage.
+Since these could be tagged in different ways, for example, longer lifecycle of (downstream) product, optimized use of raw materials, or even reduced waste, it seemed more appropriate to create a specific category for these cases.
+Depending on subsequent analysis needs this could then be mapped to other categories.
 
@@
 
-# Results
+# Results {#sec-results}
 
 ```{python}
-#| label: Queries
+# | label: Queries
+
+bioeconomy = pd.read_sql(
+    """
+select
+  i.sinno_id,
+  i.innovation_name_in_swedish AS name,
+  i.description_in_swedish AS description,
+  i.additional_information_if_origin_new_scientific_discovery || i.additional_information_if_origin_new_technologies_or_materials || i.additional_info_if_origin_official_regulation_legislation_and_standards || i.additional_information_if_origin_solution_for_a_problem || i.additional_information_if_origin_performance || i.additional_information_if_origin_other AS info,
+  i.year_of_commercialization AS year,
+  us.use_sector
+from
+  innovation i
+  join use_sectors us on i.sinno_id = us.sinno_id
+where
+  (
+    us.use_sector like '02%'
+    or us.use_sector like '20%'
+    or us.use_sector like '21%'
+    or us.use_sector like '36%'
+    or product_code like '02%'
+    or product_code like '20%'
+    or product_code like '21%'
+    or product_code like '36%'
+  )
+  or (
+    description like '%virke%'
+    or description like '%cellulos%'
+    or description like '%lignin%'
+    or description like '%spån%'
+    or description like '%bark%'
+    or description like '%levulinsyra%'
+    or description like '%furfural%'
+    or description like '%svarttjära%'
+    or description like '%svartlut%'
+    or description like '%växtbas%'
+    or description like '%ved%'
+    or description like '%trä%'
+    or description like '%skog%'
+    or description like '%biobränsle%'
+    or description like '%biologisk%'
+    or description like '%nedbrytbar%'
+    or description like '%papper%'
+    or description like '%pappret%'
+    or description like '%karton%'
+    or description like '%tencel%'
+  );
+""",
+    conn,
+)
+
+sni_codes = pd.read_sql_query(
+    """
+select
+*
+from 
+sni_codes
+""",
+    conn,
+)
+
+sni_codes = sni_codes.rename(columns={"code": "use_sector"})
+
+swinno = pd.read_sql_query(
+    """
+select sinno_id, year_of_commercialization as year, innovation_name_in_swedish as name
+from innovation;
+""",
+    conn,
+)
+
+swinno["bioeconomy"] = swinno["sinno_id"].isin(bioeconomy["sinno_id"].unique())
 
@@
 
+:::{#tbl-bioeconomy_vision_summary layout-ncol=1}
 | @bugge2016WhatBioeconomyReview [10] | Biotechnology | Bioresource | Bioecology |
@@
 | Classificaiton Code | 1 | 2 | 3 | 
+:
+
+Bioeconomy Vision Summary 
+:::
 
@@
 
-<!-- TODO merge tables -->
+```{python}
+# | label: tbl-full_codes
+# | tbl-cap: "List of Eco-Innovation Labels"
 
-: Bioeconomy Vision Summary {#tbl-bioeconomy_vision_summary}
+codes_df = pd.read_sql(
+    "SELECT * FROM classification_codes", conn, index_col="Code"
+).iloc[:-5, :]
+
+format_table(codes_df)
+```
+
+```sql
+SELECT *
+FROM table
+WHERE
+    (
+        us.use_sectors LIKE '02%' -- Forestry and related services
+-- Furniture manufacturing; other manufacturing
+        OR us.use_sectors LIKE '20%' -- Wood and wood product manufacturing except furniture
+        OR us.use_sectors LIKE '21%' -- Pulp, paper and paper product manufacturing
+        OR us.use_sectors LIKE '36%' -- Furniture manufacturing; other manufacturing
+        us.use_sectors LIKE '02%' -- Forestry and related services
+-- Furniture manufacturing; other manufacturing
+        OR product_code LIKE '20%' -- Wood and wood product manufacturing except furniture
+        OR product_code LIKE '21%' -- Pulp, paper and paper product manufacturing
+        OR product_code LIKE '36%' -- Furniture manufacturing; other manufacturing
+    )
+    OR (
+        description LIKE '%virke%'
+        OR description LIKE '%cellulos%'
+        OR description LIKE '%lignin%'
+        OR description LIKE '%spån%'
+        OR description LIKE '%bark%'
+        OR description LIKE '%levulinsyra%'
+        OR description LIKE '%furfural%'
+        OR description LIKE '%svarttjära%'
+        OR description LIKE '%svartlut%'
+        OR description LIKE '%växtbas%'
+        OR description LIKE '%ved%'
+        OR description LIKE '%trä%'
+        OR description LIKE '%skog%'
+        OR description LIKE '%biobränsle%'
+        OR description LIKE '%biologisk%'
+        OR description LIKE '%nedbrytbar%'
+        OR description LIKE '%papper%'
+        OR description LIKE '%pappret%'
+        OR description LIKE '%karton%'
+        OR description LIKE '%tencel%'
+    );
+```
+
+::::{.columns}
+
+:::{.column width="50%"}
+**Using producer / user codes
+**
+
+|SNI Code | Description                            |
+|--------:|:---------------------------------------|
+|02       | Forestry and related services
+|20       | Wood and wood product manufacturing except furniture
+|21       | Pulp, paper and paper product manufacturing
+|36       | Furniture manufacturing; other manufacturing
+:::
+
+:::{.column width="50%"}
+**Adding keywords of value chain**
+
+    - "virke",
+    - "cellulos",
+    - "lignin",
+    - "spån", 
+    - "bark",
+    - "levulinsyra" (Levulinic acid),
+    - "furfural" (Furfural),
+    - "svarttjära",
+    - "svartlut",
+    - "växtbas",
+    - "ved",
+    - "trä",
+    - "skog",
+    - "papper",
+    - "biobränsle",
+    - "biologiskt",
+    - "nedbrytbar",
+    - "papper",
+    - "pappret",
+    - "karton",
+    - "tencel",
+:::
+
+::::

### 2023-09-18 src/figures.py
add figures, references, update utils

@@ -1,0 +1,109 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+import plotly.graph_objects as po
+import plotly.express as px
+
+
+def plot_counts(
+    df1, df2, category_col, y, show_legend=True, legend_cols=None, y_label="Count"
+):
+    categories = df1[category_col].astype(str).unique()
+    categories.sort()
+
+    legend_ncols = len(categories) if legend_cols is None else legend_cols
+
+    colors = sns.color_palette("colorblind", len(categories))
+    palette = dict(zip(categories, colors))
+
+    fig, axes = plt.subplots(
+        ncols=1, nrows=2, sharex=False, sharey=True, layout="tight"
+    )
+
+    marker_size = 100
+    df1 = df1.sort_values(by=category_col)
+    df2 = df2.sort_values(by=category_col)
+
+    sns.scatterplot(
+        x=df1.index,
+        y=y,
+        hue=category_col,
+        data=df1,
+        ax=axes[0],
+        palette=palette,
+        s=marker_size,
+    )
+
+    sns.scatterplot(
+        x=df2.index,
+        y=y,
+        hue=category_col,
+        data=df2,
+        ax=axes[1],
+        palette=palette,
+        s=marker_size,
+    )
+
+    for axs in axes:
+        axs.set_ylabel(y_label)
+        axs.set_xlabel("Year")
+        axs.legend().remove()
+
+    axes[0].set_title("Including Uncertains")
+    axes[0].title.set_position([0.5, 1.05])  # adjust the position of the title
+
+    axes[1].set_title("Excluding Uncertains")
+    axes[1].title.set_position([0.5, 1.05])  # adjust the position of the title
+
+    if not show_legend:
+        fig.legend().remove()
+    else:
+        handles, labels = axes[0].get_legend_handles_labels()
+        fig.legend(
+            handles,
+            labels,
+            loc="lower center",
+            ncol=legend_ncols,
+            frameon=False,
+            bbox_to_anchor=(0.5, -0.1),
+        )  # adjust the position of the legend
+
+    plt.subplots_adjust(bottom=0.2)
+
+    plt.show()
+
+
+def plot_heatmap(data, title):
+    fig, ax = plt.subplots(1, 1)
+    ax = sns.heatmap(
+        data, annot=True, cmap="Blues", fmt="g", cbar=False, annot_kws={"fontsize": 10}
+    )
+    ax.set(title=title, ylabel=None, xlabel=None)
[…]

### 2023-09-18 src/utils.py

@@ -2,3 +2,10 @@ from pathlib import Path
 
 def get_project_root():
     return Path(__file__).absolute().parent.parent
+
+def format_table(df):
+    # create a Styler object and set the wrapping style
+    styler = df.style.set_properties(**{"width": "50em", "text-wrap": "break-word"})
+
+    return styler
+

### 2023-09-19 docs/reports/paper/paper1.qmd
add trend and heatmap figure

@@
 # set all text on plots to be monospace with rcParams
-plt.rcParams["font.family"] = "monospace"
+# plt.rcParams["font.family"] = "Ubunto Mono"
 
-sns.set_theme(style="whitegrid")
+sns.set_theme(style="white")
 
 ```
+
 # Debate around Bioeconomies
@@
 And what vision does the innovation system pursue?
-
-##
 
@@
 ```{python}
-# | label: Queries
+# | label: Query bioeconomy
 
@@
 ```{python}
-#| label: Number of Results
+#| label: Query eco-innovation
 
-display
+eco_innovations = pd.read_sql(
+    """
+SELECT 
+    sinno_id
+    , innovation_type
+FROM
+  eco_innovations;
+""",
+    conn,
+)
+
 ```
+
+```{python}
+# | label: Query bioeconomy visions
+
+bioeconomy_visions = pd.read_sql(
+    """
+SELECT bv.sinno_id
+, codes.Category AS bioeconomy_vision
+, i.year_of_commercialization AS "year"
+FROM bioeconomy_visions AS bv
+JOIN [classification_codes] AS codes ON codes.code = bv.bioeconomy_vision
+JOIN innovation AS i ON i.sinno_id = bv.sinno_id
+WHERE bv.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
+UNION
+SELECT DISTINCT(ei.sinno_id)
+, 'Bioecology Vision' AS bioeconomy_vision
+, i.year_of_commercialization AS "i.year"
+FROM eco_innovations AS ei
+JOIN innovation AS i ON i.sinno_id = ei.sinno_id
+WHERE ei.innovation_type IN (206, 107)
+AND ei.sinno_id NOT IN (
+  SELECT ei2.sinno_id
+  FROM eco_innovations as ei2
+  WHERE ei2.innovation_type IN (999, '000')
+  );
+-- adds recycling innovation to bioecology
+""",
+    conn,
+)
+
+bioeconomy_visions.loc[
+    bioeconomy_visions["bioeconomy_vision"] == "Not Bioeconomy",
+    ["bioeconomy_vision"],
+] = "No Bioeconomy Vision"
+
+uncertain_bioeconomy_visions = bioeconomy_visions.query("bioeconomy_vision == 'Unsure'")
+
+certain_bioeconomy_visions = bioeconomy_visions.loc[
+    ~bioeconomy_visions["sinno_id"].isin(uncertain_bioeconomy_visions["sinno_id"]), :
+]
+
+```
+
+```{python}
+# | label: Number of Results
+# | output: asis
+
+display(
+    Markdown(
+        f"So far, {certain_bioeconomy_visions['sinno_id'].nunique()} innovation have been classified for the period between {certain_bioeconomy_visions['year'].min()} and {certain_bioeconomy_visions['year'].max()}."
+    ))
+```
+
+```{python}
+# | label: fig-bioeconomy_trends
+# | fig-cap: Bioeconomy Vision Trends
+
+count_certain_bioeconomy_visions = (
+    certain_bioeconomy_visions.groupby(["year", "bioeconomy_vision"])
+    .size()
+    .reset_index(name="count")
+)
+
+def plot_bv_trends(df, title):
+    palette = sns.color_palette("colorblind")[:5]
+
+    # Create a custom palette with a default light gray color for "not bio" hue
+    custom_palette = {
+        hue: color if bioeconomy_vision != "No Bioeconomy Vision" else (0.8, 0.8, 0.8)
+        for hue, color, bioeconomy_vision in zip(
+            df["bioeconomy_vision"], palette, df["bioeconomy_vision"]
+        )
+    }
+
+    chart = sns.lineplot(
+        data=df,
+        x="year",
+        y="count",
+        hue="bioeconomy_vision",
+        palette=custom_palette,
+        linewidth=2.5,
+    )
+
+    chart.set_title(title, fontdict={"fontsize": 16, "fontweight": "normal"})
+    chart.set_xlabel("Year", fontdict={"fontsize": 12, "fontweight": "normal"})
+    chart.set_ylabel("Count", fontdict={"fontsize": 12, "fontweight": "normal"})
+    chart.tick_params(labelsize=10)
+
+    # Customize x-axis tick labels
+    years = df["year"].unique()
+    chart.set_xticks(years[::5])
+    chart.set_xticklabels(years[::5], rotation=45, ha="right")
+
+    # Create a custom legend with colored labels (no lines)
+    handles, labels = chart.get_legend_handles_labels()
+
+    # Set font color for legend labels based on the hue
+    legend = chart.legend(
+        bbox_to_anchor=(0.5, -0.35), loc="upper center", ncol=2, frameon=False
+    )
+    for text, label in zip(legend.texts, labels):
+        text.set_color(custom_palette[label])
+
+    plt.tight_layout()
+
+plot_bv_trends(count_certain_bioeconomy_visions, title="")
+
+plt.show()
+```
+
+```{python}
+# | label: fig-vision_heatmap
+# | fig-caption: Sectors Contribution to Bioeconomy Visions
+
+sectors = pd.read_sql(
+    """
+SELECT
+  bv.sinno_id as sinno_id,
+  c.Category AS vision,
+  sni_codes.label AS sector
+FROM
+  bioeconomy_visions AS bv
+  JOIN use_sectors ON bv.sinno_id = use_sectors.[sinno_id]
+  JOIN sni_codes ON SUBSTR(use_sectors.use_sector, 1, 2) = sni_codes.code
+  JOIN classification_codes AS c ON bv.bioeconomy_vision = c.Code;
+""",
+    conn,
+)
+
+sectors = sectors.loc[
+    sectors["sinno_id"].isin(set(certain_bioeconomy_visions["sinno_id"])), :
+]
+
+sectors.loc[sectors["vision"] == "Not Bioeconomy", "vision"] = "No Bioeconomy Vision"
+
+sector_counts = (
+    sectors.groupby(["sector", "vision"]).size().to_frame("count").reset_index()
+)
+# Question is, should I weight this following the way in the env-innovation paper?
+
+sector_counts = sector_counts.sort_values("count", ascending=False)
+
+heatmap_data = prepare_heatmap_data(sector_counts, "sector", "vision", "count")
+
+plot_heatmap(heatmap_data, "")
+
+# TODO: Better y axis labels
+
+```
+
+### How many eco-innovations are there?
+#### How many by bioeconomy vision?
 
@@
 format_table(codes_df)
-```
-
-```sql
-SELECT *
-FROM table
-WHERE
-    (
-        us.use_sectors LIKE '02%' -- Forestry and related services
--- Furniture manufacturing; other manufacturing
-        OR us.use_sectors LIKE '20%' -- Wood and wood product manufacturing except furniture
-        OR us.use_sectors LIKE '21%' -- Pulp, paper and paper product manufacturing
-        OR us.use_sectors LIKE '36%' -- Furniture manufacturing; other manufacturing
-        us.use_sectors LIKE '02%' -- Forestry and related services
--- Furniture manufacturing; other manufacturing
-        OR product_code LIKE '20%' -- Wood and wood product manufacturing except furniture
-        OR product_code LIKE '21%' -- Pulp, paper and paper product manufacturing
-        OR product_code LIKE '36%' -- Furniture manufacturing; other manufacturing
-    )
-    OR (
-        description LIKE '%virke%'
-        OR description LIKE '%cellulos%'
-        OR description LIKE '%lignin%'
-        OR description LIKE '%spån%'
-        OR description LIKE '%bark%'
-        OR description LIKE '%levulinsyra%'
-        OR description LIKE '%furfural%'
-        OR description LIKE '%svarttjära%'
-        OR description LIKE '%svartlut%'
-        OR description LIKE '%växtbas%'
-        OR description LIKE '%ved%'
-        OR description LIKE '%trä%'
-        OR description LIKE '%skog%'
-        OR description LIKE '%biobränsle%'
-        OR description LIKE '%biologisk%'
-        OR description LIKE '%nedbrytbar%'
-        OR description LIKE '%papper%'
-        OR description LIKE '%pappret%'
-        OR description LIKE '%karton%'
-        OR description LIKE '%tencel%'
-    );
 ```

### 2023-09-19 src/figures.py

@@ -4,73 +4,28 @@ import seaborn as sns
 import plotly.graph_objects as po
 import plotly.express as px
 
-
-def plot_counts(
-    df1, df2, category_col, y, show_legend=True, legend_cols=None, y_label="Count"
-):
-    categories = df1[category_col].astype(str).unique()
-    categories.sort()
-
-    legend_ncols = len(categories) if legend_cols is None else legend_cols
-
-    colors = sns.color_palette("colorblind", len(categories))
-    palette = dict(zip(categories, colors))
-
-    fig, axes = plt.subplots(
-        ncols=1, nrows=2, sharex=False, sharey=True, layout="tight"
+def plot_bv_trends(df, title):
+    chart = sns.lineplot(
+        data=df,
+        x="year",
+        y="count",
+        hue="bioeconomy_vision",
+        palette="colorblind",
+        linewidth=2.5,
     )
+    chart.set_title(title, fontdict={"fontsize": 16, "fontweight": "bold"})
+    chart.set_xlabel("Year", fontdict={"fontsize": 12, "fontweight": "bold"})
+    chart.set_ylabel("Count", fontdict={"fontsize": 12, "fontweight": "bold"})
+    chart.tick_params(labelsize=10)
+    plt.legend(title="Vision", fontsize=10, title_fontsize=12)
+    plt.tight_layout()
 
-    marker_size = 100
-    df1 = df1.sort_values(by=category_col)
-    df2 = df2.sort_values(by=category_col)
+    # Customize x-axis tick labels
+    years = df["year"].unique()
+    chart.set_xticks(years[::5])
+    chart.set_xticklabels(years[::5], rotation=45, ha="right")
 
-    sns.scatterplot(
-        x=df1.index,
-        y=y,
-        hue=category_col,
-        data=df1,
-        ax=axes[0],
-        palette=palette,
-        s=marker_size,
-    )
-
-    sns.scatterplot(
-        x=df2.index,
-        y=y,
-        hue=category_col,
-        data=df2,
-        ax=axes[1],
-        palette=palette,
-        s=marker_size,
-    )
-
-    for axs in axes:
-        axs.set_ylabel(y_label)
-        axs.set_xlabel("Year")
-        axs.legend().remove()
-
-    axes[0].set_title("Including Uncertains")
-    axes[0].title.set_position([0.5, 1.05])  # adjust the position of the title
-
-    axes[1].set_title("Excluding Uncertains")
[…]

### 2023-09-21 docs/reports/paper/paper1.qmd
paper draft send to Josef

@@
 execute:
+    include: true
     echo: false
+    warning: false
 ---
@@
 # | label: Imports
-# | eval: true
 
 import pandas as pd
+import numpy as np
 from IPython.display import display, Markdown
 import matplotlib.pyplot as plt
+import matplotlib.ticker as mtick
 import seaborn as sns
@@
 # set all text on plots to be monospace with rcParams
-# plt.rcParams["font.family"] = "Ubunto Mono"
+plt.rcParams["font.family"] = "Space Grotesk"
 
@@
 
-```
+# full code labels for merging
 
-# Debate around Bioeconomies
-
-An idea which is bound to increase the pressure and trade-offs between different ecosystem services is the bioeconomy.
-In recent years the concept has gained relevance among policymakers, practitioners and researchers.
-The term bioeconomy refers broadly to an economy which uses bio-based inputs instead of fossil based ones.
-Connected to this are hopes of increased resource efficiency and renewability.
-In a report commissioned by the Swedish Government, the @2012SwedishResearchInnovation defines a vision of a Swedish bioeconomy based on two central pillars.
-It is to be an economy based on,
-
-> \[a\] sustainable production of biomass to enable increased use within a number of different sectors of society. The objective is to reduce climate effects and the use of fossil-based raw materials.
-
-With,
-
-> \[a\]n increased added value for biomass materials, concomitant with a reduction in energy consumption and recovery of nutrients and energy as additional end products.
-The objective is to optimize the value and contribution of ecosystem services to the economy.
-(p.
-9).
-
-These pillars reflect the two most influential visions the bioeconomy concept.
-Notions of a bio-technology and bio-resource oriented bioeconomy are widespread and dominate discourse at various levels [@dieken2021MultitudesBioeconomiesSystematic].
-Despite -- or perhaps because of -- the bioeconomy lacking a strong definition, it is used to refer to different concepts, with different aims and objectives, foci, normative assumptions and hence implications for society and environment [@bugge2016WhatBioeconomyReview].
-
-The bio-technology vision focuses on economic growth and job provision through the application of bio-technology [@bugge2016WhatBioeconomyReview].
-Science driven innovation features prominently in this vision, with collaborations between research institutions in academia and industry taking on an important role [@bugge2016WhatBioeconomyReview].
-@vivien2019HijackingBioeconomy present a review of the concept showing that the focus on science as an absolute motor of change also shaped policy recommendations, driven especially by the OECD, towards increasing the speed and diffusion of innovation through partnerships between new start-ups and establish pharmaceutical companies.
-As a policy narrative this vision was particularly potent and widespread at the turn of the 21st century.
-According to them, the intellectual heritage of Schumpeter and Kontradieff play an important role in this vision.
-Technological breakthroughs are poised to solve economic and ecological challenges, for example through genetic engineering.
-Consequently, this vision has a weak sustainability notion, viewing sustainability challenges as temporary and nature as substitutable given sufficient research and knowledge creation [@vivien2019HijackingBioeconomy].
-
-In more recent years, visions of a bio-resource bioeconomy have taken over as the leading narrative.
-Unlike the bio-technology vision, this conceptualization does not only focus on economic growth and job creation, but attempts to include ecological sustainability challenges [@bugge2016WhatBioeconomyReview].
-Biological resources, not technologies, form the basis of a new economy.
-In essence, there are two ways in which bio-resources are poised to replace fossil resources: as substitution, e.g., bio-fuels replacing fossil fuels, and as high value added products, e.g., Cross-laminated Timber used as a concrete alternative in construction [@vivien2019HijackingBioeconomy].
-To substitute fossil resources requires significant efforts in innovation.
-In contrast to the bio-technology vision, however, innovation is not confined to scientific advances from a narrow biochemical and pharmaceutical field.
-Instead, new and old actors are required to collaborate between and beyond established sectors [@bugge2016WhatBioeconomyReview].
-Although sustainability plays a more important role in this vision of the bioeconomy, this vision is considered an example of weak sustainability [@damato2017GreenCircularBio; @vivien2019HijackingBioeconomy].
-The reason is the focus of substitution and high biomass demands.
-The consumption of raw materials may even increase, not only for bio-based, but also for fossil-based resources [@asada2020BioeconomicTransitionProjecting].
-
-The exact role played by forests in a bioeconomy transition depends on the desired vision [@kleinschmit2014ShadesGreenSocial].
-There is no doubt that the Swedish vision places high demands on forests as primary providers of the required biomass.
-Both in the strategy formulated in @2012SwedishResearchInnovation and in public discourse, a small, well-connected group of actors center the bio-resource vision as legitimate and desirable [@holmgren2022WhoseTransformationThis].
-The result is a forest sector which operates under a "more of everything" paradigm [@belandlindahl2015SwedishForestryModel].
-
-#### Innovation for a Bioeconomy
-
-Apart from the substitution of fossil resources through bio-based resources, a central narrative used in the Swedish transition is that, "\[t\]echnological innovation is key to a greener future" [@holmgren2022WhoseTransformationThis 42].
-According to @jankovsky2021InnovationsForestBioeconomy four types of innovation are central for a forest based bioeconomy:
-
-1.  Substitute Products,
-2.  New (bio-based) Processes,
-3.  New (bio-based) Products,
-4.  New Behavior.
-
-Currently, most research into bioeconomy innovation relies on patent data or qualitative case studies [e.g., @bennich2021BiobasedEconomy2030; @jander2019MonitoringTransitionBioeconomy].
-Unfortunately, these approaches are not without their flaws.
-Focusing on patents could overestimate the importance of the bio-technology vision, whose science oriented innovation approach is driven by patents [@bugge2016WhatBioeconomyReview].
-Qualitative case studies, on the other hand, may miss broader trends and features of a bioeconomy transition.
-
-@hausknost2017TransitionWhichBioeconomy
-
-Public, scientific and political debate of bioeconomies has brought forth a number of discourse analyses.
-Both @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy conclude that the concept of a bioeconomy conceals different, possibly conflicting visions of transfromation outcomes.
-
-Recent scholarship has called for advancing research fronts instead of remaining entrenched in contested discourse positions @allain2022BioeconomicsVsBioeconomy.
-
-@jankovsky2021InnovationForestBioeconomy
-
-@bauer 
-refineries
-
-Research Question:
-
-While much research has been done on the research, less is known over the actual direction of technical change in the Swedish forest bioeconomy.
-How innovative, in terms of output, is the system?
-And what vision does the innovation system pursue?
-
-# Data and Methods
-
-To capture the actual innovation output associated with the Swedish forest based bioeconomy the SWINNO database was searched for innovation associated with core forest sectors and keywords describing forest products which may be used in sectors along the bioeconomy value chain.
-The list of keywords and sectors is presented in the appendix.
-Due to the query matching keywords whenever they appear as substrings in the description of the innovation, the inital search result is rather susceptible to false positives.
-In later manual review of each innovation for classification these false positives were removed from the search results.
-
-The SWINNO database is constructed according to  literature-based innovation output (LBIO) method [@sjoo2014DatabaseSwedishInnovations].
-This method relies on the judgment of editiors and journalists at trade journals and hence captures what can be described as significant innovation.
-An important limitation of LBIO measures, especially for research into forest sector innovation, is a bias against non-commercialized innovation.
-New processes, or new organizational apporaches for example, will only be caputred if they are sold to another company.
-This is especially sensitive for bioeconomy research as, for example, new methods or plans for managing forest stands are less likely to be included.
-Relevant product innovation which faciliates these plans, for example products related to the planting of trees, on the other hand, are captured.
-
-A likely consequence of this limitation is that the number of innovation related to the bioecology vision are under counted, as this vision highlights "Counter-expertise rather than concrete technical solutions" [@vivien2019HijackingBioeconomy 194].
-Nonetheless, a small number of innovation associated with this vision were already identified in the database.
-
-But, measuring innovation with LBIO has also substantial advantages.
-For one, it ensures that actual innovation, meaning commericialized products and processes, are captured.
-Second, because the source articles are preserved, it provides a rich source of data, giving frequently providing insights into the aims, motivations, advantages and challenges of the covered innovation.
-
-## Classification Methodology
-
-These source articles were carefully read and their associated unique identifiers in the database tagged with two types of labels.
-The first label is the main interest of this study: the bioeconomy vision an innovation can be associated with.
-The second label provides a deeper understanding if and how the innovation reduces environmental impacts.
-As @broring2020InnovationTypesBioeconomy highlight, not all bioeconomy innovations need to be sustainability oriented innovation.
-
-### Bioeconomy Visions
-
-Innovations were classified as belonging to a bioeconomy vision if they fulfilled one, or more criteria outlined in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy.
-@tbl-bioeconomy_vision_summary shows the summary of bioeconomy visions presented in the cited papers.
-
-While following these criteria, a number of decisions were made concerning content of the summary presented in @tbl-bioeconomy_vision_summary and the nature of the corresponding vision.
-For one, economic growth is cited as the aim & objective of biotechnology and bioresource visions by @bugge2016WhatBioeconomyReview.
-It can be argued that all innovation can contribute to economic growth, which is why these aspects were not given hightened importance in the classification decision.
-A second topic of judgment has to be the inclusion of innovation which increases the recyclability of a product, or uses recycled waste, water and material in the process.
-Based on the definition of a bioecology vision presented in @bugge2016WhatBioeconomyReview, these innovations should be included as contributing to this vision.
-The extent to which such innovation contributes in reality to a "[c]ircular and self-sustained production mode" [@bugge2016WhatBioeconomyReview 10], however, can be debated.
-
-```{python}
-# | label: Count Recycling Innovation
-
-recycle_innovation = pd.read_sql(
-    """
-SELECT *
-FROM classification_codes
-LEFT JOIN eco_innovations ON classification_codes.Code = eco_innovations.innovation_type
-WHERE eco_innovations.innovation_type = 107 OR eco_innovations.innovation_type = 206 AND sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes LIKE '%not forest%')
-""",
-    conn,
+codes_df = pd.read_sql(
+    "SELECT * FROM classification_codes", conn
 )
 
-recycle_innovation_count = recycle_innovation["sinno_id"].nunique()
+codes_df["Category"] = codes_df["Category"].str.strip()
+uncertain_eco = {"Code":"000", "Category":"Unsure"}
+codes_df = codes_df.append(uncertain_eco, ignore_index=True)
 
-display(
-    Markdown(
-        f"Currently, {recycle_innovation_count} innovations were identified which involve either a product or process innovation related to recyling."
-    )
-)
+codes_df = codes_df[~codes_df["Code"].duplicated(keep='last')
+]
+
+codes_df =codes_df.set_index("Code")
 ```
-
-<!-- TODO count all innovation with eco-innovation type 107 and 206 as bv 3 -->
-
-@tbl-vision_examples shows five examples of classfied innovation.
-This selection also highlights that multible labels can be attached to one innovation.
-The last innovation, describing a waterproof paper, is classified as a bioresource vission innovaition because it describes a process by which biomass is converted and upgraded.
-It is also classified as a biotechnology innovation since the process mimics trees' microfiber formation, accroding to the innovation source article.
-
-```{python}
-# | label: tbl-vision_examples
-# | tbl-cap: "List of Eco-Innovation Labels"
-
-classification_examples = pd.read_sql(
-    """SELECT
-  n.sinno_id,
-  i.description_in_swedish,
-  GROUP_CONCAT(c.Category, ' & ') AS Categories,
-  n.notes
-FROM
-  categorization_notes n
-  JOIN innovation i ON i.sinno_id = n.sinno_id
-  JOIN bioeconomy_visions bv
-  ON n.sinno_id = bv.sinno_id
-  JOIN classification_codes c ON bv.bioeconomy_vision = c.Code
-WHERE
-  n.notes LIKE "%good example%"
-GROUP BY
-  n.sinno_id,
-  i.description_in_swedish,
-  n.notes""",
-    conn,
-)
-
-classification_examples_out = (
-    classification_examples.copy()
-    .loc[:, ["Categories", "description_in_swedish"]]
-    .iloc[2:]
-)
-
-rowIDs = range(1, len(classification_examples_out) + 1)
-display(
-    Markdown(
-        classification_examples_out.rename(
-            columns={"description_in_swedish": "Short Description"}
-        ).to_markdown(index=rowIDs)
-    )
-)
-
-```
-
-### Eco-Innovation Types
-
-Understanding the bioeconomy vision that an innovation is contributing to is the focus of this paper.
-In order to gain a deeper understanding of the nature of innovation within each bioeconomy vision, however, the innovation were also tagged with labels of eco-innovation types.
-This classification uses eco-innovation types identified in a recent literature review by @garcia-granero2018EcoinnovationMeasurementReview.
-They review 104 publications and produce a list 30 key Eco-Innovation Perfromance Indicators of product, process, organizational and marketing innovation.
-
-This initial list was tested on a subset of innovation in SWINNO and augmented with 15 categories.
-Some are mere extensions of their categories (for example, specifying if an innovation prolongs the life cycle of the product itself or downstream products), others are new types of existing categories (for example, reduced other pollution such as noise), yet others are entirely new categories (human health and safety), the categories of automatization and precision by themselves are not eco-innovations, but can be used to gain a deeper understanding of what makes an innovation an eco-innovation.
-Clearly not all new categories are eco-innovations.
-The newly created pesticide category, for example, can be an eco-innovation if it reduces chemical pollution, but it could also increase chemical pollution if pesticides are introduced in novel ways.
-
-For the eco-innovations the decision process is straightforward.
-Based on the information contained in the text, the innovation is labeled with one or more categories.
-In case of uncertainty, the doubt label is added.
-Because the list of eco-innovation types was relatively exhaustive from the beginning, few adjustments had to be made.
-These include, as alluded to above, expanding the labeling for other sustainability criteria, such as human health and safety, and specific forest related aspects / tasks to gain a more detailed understanding of an innovations aim.
-For example, some innovations aim to protect harvested timber, either at harvest, during transport or storage.
-Since these could be tagged in different ways, for example, longer lifecycle of (downstream) product, optimized use of raw materials, or even reduced waste, it seemed more appropriate to create a specific category for these cases.
-Depending on subsequent analysis needs this could then be mapped to other categories.
-
-<!-- Use examples of clear cases here -->
-
-# Results {#sec-results}
 
@@
 
+# Debate around Bioeconomies
+
+An idea which is bound to increase the pressure and trade-offs between different ecosystem services is the bioeconomy.
+In recent years the concept has gained relevance among policymakers, practitioners and researchers.
+The term bioeconomy refers broadly to an economy which uses bio-based inputs instead of fossil based ones.
+Connected to this are hopes of increased resource efficiency and renewability.
+In a report commissioned by the Swedish Government, the @2012SwedishResearchInnovation defines a vision of a Swedish bioeconomy based on two central pillars.
+It is to be an economy based on,
+
+> \[a\] sustainable production of biomass to enable increased use within a number of different sectors of society. The objective is to reduce climate effects and the use of fossil-based raw materials.
+
+With,
+
+> \[a\]n increased added value for biomass materials, concomitant with a reduction in energy consumption and recovery of nutrients and energy as additional end products.
+The objective is to optimize the value and contribution of ecosystem services to the economy.
+(p.
+9).
+
+These pillars reflect the two most influential visions the bioeconomy concept.
+Notions of a bio-technology and bio-resource oriented bioeconomy are widespread and dominate discourse at various levels [@dieken2021MultitudesBioeconomiesSystematic].
+Despite -- or perhaps because of -- the bioeconomy lacking a strong definition, it is used to refer to different concepts, with different aims and objectives, foci, normative assumptions and hence implications for society and environment [@bugge2016WhatBioeconomyReview].
+
+The bio-technology vision focuses on economic growth and job provision through the application of bio-technology [@bugge2016WhatBioeconomyReview].
+Science driven innovation features prominently in this vision, with collaborations between research institutions in academia and industry taking on an important role [@bugge2016WhatBioeconomyReview].
+@vivien2019HijackingBioeconomy present a review of the concept showing that the focus on science as an absolute motor of change also shaped policy recommendations, driven especially by the OECD, towards increasing the speed and diffusion of innovation through partnerships between new start-ups and establish pharmaceutical companies.
+As a policy narrative this vision was particularly potent and widespread at the turn of the 21st century.
+According to them, the intellectual heritage of Schumpeter and Kontradieff play an important role in this vision.
+Technological breakthroughs are poised to solve economic and ecological challenges, for example through genetic engineering.
+Consequently, this vision has a weak sustainability notion, viewing sustainability challenges as temporary and nature as substitutable given sufficient research and knowledge creation [@vivien2019HijackingBioeconomy].
+
+In more recent years, visions of a bio-resource bioeconomy have taken over as the leading narrative.
+Unlike the bio-technology vision, this conceptualization does not only focus on economic growth and job creation, but attempts to include ecological sustainability challenges [@bugge2016WhatBioeconomyReview].
+Biological resources, not technologies, form the basis of a new economy.
+In essence, there are two ways in which bio-resources are poised to replace fossil resources: as substitution, e.g., bio-fuels replacing fossil fuels, and as high value added products, e.g., Cross-laminated Timber used as a concrete alternative in construction [@vivien2019HijackingBioeconomy].
+To substitute fossil resources requires significant efforts in innovation.
+In contrast to the bio-technology vision, however, innovation is not confined to scientific advances from a narrow biochemical and pharmaceutical field.
+Instead, new and old actors are required to collaborate between and beyond established sectors [@bugge2016WhatBioeconomyReview].
+Although sustainability plays a more important role in this vision of the bioeconomy, this vision is considered an example of weak sustainability [@damato2017GreenCircularBio; @vivien2019HijackingBioeconomy].
+The reason is the focus of substitution and high biomass demands.
+The consumption of raw materials may even increase, not only for bio-based, but also for fossil-based resources [@asada2020BioeconomicTransitionProjecting].
+
+The exact role played by forests in a bioeconomy transition depends on the desired vision [@kleinschmit2014ShadesGreenSocial].
+There is no doubt that the Swedish vision places high demands on forests as primary providers of the required biomass.
+Both in the strategy formulated in @2012SwedishResearchInnovation and in public discourse, a small, well-connected group of actors center the bio-resource vision as legitimate and desirable [@holmgren2022WhoseTransformationThis].
+The result is a forest sector which operates under a "more of everything" paradigm [@belandlindahl2015SwedishForestryModel].
+
+#### Innovation for a Bioeconomy
+
+Apart from the substitution of fossil resources through bio-based resources, a central narrative used in the Swedish transition is that, "\[t\]echnological innovation is key to a greener future" [@holmgren2022WhoseTransformationThis 42].
+According to @jankovsky2021InnovationsForestBioeconomy four types of innovation are central for a forest based bioeconomy:
+
+1.  Substitute Products,
+2.  New (bio-based) Processes,
+3.  New (bio-based) Products,
+4.  New Behavior.
+
+Currently, most research into bioeconomy innovation relies on patent data or qualitative case studies [e.g., @bennich2021BiobasedEconomy2030; @jander2019MonitoringTransitionBioeconomy].
+Unfortunately, these approaches are not without their flaws.
+Focusing on patents could overestimate the importance of the bio-technology vision, whose science oriented innovation approach is driven by patents [@bugge2016WhatBioeconomyReview].
+Qualitative case studies, on the other hand, may miss broader trends and features of a bioeconomy transition.
+
+@hausknost2017TransitionWhichBioeconomy
+
+Public, scientific and political debate of bioeconomies has brought forth a number of discourse analyses.
+Both @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy conclude that the concept of a bioeconomy conceals different, possibly conflicting visions of transfromation outcomes.
+
+Recent scholarship has called for advancing research fronts instead of remaining entrenched in contested discourse positions @allain2022BioeconomicsVsBioeconomy.
+
+Research Question:
+
+While much research has been done on the research, less is known over the actual direction of technical change in the Swedish forest bioeconomy.
+How innovative, in terms of output, is the system?
+What vision does the innovation system pursue?
+And how do bioeconomy visions differ in their eco-innovation orientation?
+
+# Data and Methods
+
+To capture the actual innovation output associated with the Swedish forest based bioeconomy the SWINNO database was searched for innovation associated with core forest sectors and keywords describing forest products which may be used in sectors along the bioeconomy value chain.
+The list of keywords and sectors is presented in the appendix.
+Due to the query matching keywords whenever they appear as substrings in the description of the innovation, the inital search result is rather susceptible to false positives.
+In later manual review of each innovation for classification these false positives were removed from the search results.
+
+```{python}
+#| label: Total SWINNO Innovation
+#| output: asis
+
+n_swinno = swinno["sinno_id"].nunique()
+year_min = swinno["year"].min()
+year_max = swinno["year"].max()
+
+display(Markdown(
+    f"The SWINNO database used here contains {n_swinno} innovations commercialized between {year_min} and {year_max}.
+It is constructed according to  literature-based innovation output (LBIO) method [@sjoo2014DatabaseSwedishInnovations].
+This method relies on the judgment of editiors and journalists at trade journals and hence captures what can be described as significant innovation."
+))
+```
+
+An important limitation of LBIO measures, especially for research into forest sector innovation, is a bias against non-commercialized innovation.
+New processes, or new organizational apporaches for example, will only be caputred if they are sold to another company.
+This is especially sensitive for bioeconomy research as, for example, new methods or plans for managing forest stands are less likely to be included.
+Relevant product innovation which faciliates these plans, for example products related to the planting of trees, on the other hand, are captured.
+A likely consequence of this limitation is that the number of innovation related to the bioecology vision are under counted, as this vision highlights "Counter-expertise rather than concrete technical solutions" [@vivien2019HijackingBioeconomy 194].
+Nonetheless, a small number of innovation associated with this vision were already identified in the database.
+
+But, measuring innovation with LBIO has also substantial advantages.
+For one, it ensures that actual innovation, meaning commericialized products and processes, are captured.
+Second, because the source articles are preserved, it provides a rich source of data, giving frequently providing insights into the aims, motivations, advantages and challenges of the covered innovation.
+
+## Classification Methodology
+
+These source articles were carefully read and their associated unique identifiers in the database tagged with two types of labels.
+The first label is the main interest of this study: the bioeconomy vision an innovation can be associated with.
+The second label provides a deeper understanding if and how the innovation reduces environmental impacts.
+As @broring2020InnovationTypesBioeconomy highlight, not all bioeconomy innovations need to be sustainability oriented innovation.
+
+### Bioeconomy Visions
+
+Innovations were classified as belonging to a bioeconomy vision if they fulfilled one, or more criteria outlined in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy.
+@tbl-bioeconomy_vision_summary shows the summary of bioeconomy visions presented in the cited papers.
+
+While following these criteria, a number of decisions were made concerning content of the summary presented in @tbl-bioeconomy_vision_summary and the nature of the corresponding vision.
+For one, economic growth is cited as the aim & objective of biotechnology and bioresource visions by @bugge2016WhatBioeconomyReview.
+It can be argued that all innovation can contribute to economic growth, which is why these aspects were not given hightened importance in the classification decision.
+A second topic of judgment has to be the inclusion of innovation which increases the recyclability of a product, or uses recycled waste, water and material in the process.
+Based on the definition of a bioecology vision presented in @bugge2016WhatBioeconomyReview, these innovations should be included as contributing to this vision.
+The extent to which such innovation contributes in reality to a "[c]ircular and self-sustained production mode" [@bugge2016WhatBioeconomyReview 10], however, can be debated.
+
+@tbl-vision_examples shows five examples of classfied innovation.
+This selection also highlights that multiple labels can be attached to one innovation.
+The last innovation, describing a waterproof paper, is classified as a bioresource vission innovaition because it describes a process by which biomass is converted and upgraded.
+It is also classified as a biotechnology innovation since the process mimics trees' microfiber formation, accroding to the innovation source article.
+
+```{python}
+# | label: tbl-vision_examples
+# | tbl-cap: "List of Eco-Innovation Labels"
+
+classification_examples = pd.read_sql(
+    """SELECT
+  n.sinno_id,
+  i.description_in_swedish,
+  GROUP_CONCAT(c.Category, ' & ') AS Categories,
+  n.notes
+FROM
+  categorization_notes n
+  JOIN innovation i ON i.sinno_id = n.sinno_id
+  JOIN bioeconomy_visions bv
+  ON n.sinno_id = bv.sinno_id
+  JOIN classification_codes c ON bv.bioeconomy_vision = c.Code
+WHERE
+  n.notes LIKE "%good example%"
+GROUP BY
+  n.sinno_id,
+  i.description_in_swedish,
+  n.notes""",
+    conn,
+)
+
+classification_examples_out = (
+    classification_examples.copy()
+    .loc[:, ["Categories", "description_in_swedish"]]
+    .iloc[2:]
+)
+
+rowIDs = range(1, len(classification_examples_out) + 1)
+display(
+    Markdown(
+        classification_examples_out.rename(
+            columns={"description_in_swedish": "Short Description"}
+        ).to_markdown(index=rowIDs)
+    )
+)
+
+```
+
+### Eco-Innovation Types
+
+The focus of this paper is understanding what bioeconomy visions dominate the Swedish innovation system's output.
+In order to gain a deeper understanding of the nature of innovation within each bioeconomy vision, however, the innovation were also tagged with labels of eco-innovation types.
+This classification uses eco-innovation types identified in a recent literature review by @garcia-granero2018EcoinnovationMeasurementReview.
+They review 104 publications and produce a list 30 key Eco-Innovation Perfromance Indicators of product, process, organizational and marketing innovation.
+
+This initial list was tested on a subset of innovation in SWINNO and augmented with 15 categories.
+Because the list of eco-innovation types was relatively exhaustive from the beginning, few adjustments had to be made.
+These include: expanding the labeling for other sustainability criteria, such as human health and safety, and specific forest related aspects / tasks to gain a more detailed understanding of an innovations aim.
+For example, some innovations aim to protect harvested timber, either at harvest, during transport or storage.
+While @tbl-full_codes shows all new added labels, in the following text those that do not clearly represent an eco-innovation -- following @garcia-granero2018EcoinnovationMeasurementReview "…fewer adverse effects on the environment and more efficient use of resources" (p.
+305) -- are excluded.
+These excluded labels are Gen-manipulated plant, Increased plant yield, Gen-manipulated animal (incl new breed), Human Health and Safety, Automatization, Precision, and Biocide.
+
+The decision process for labelling eco-innovation types is also based on the source texts for each innovation.
+Based on the information contained therein, the innovation is labeled with one or more categories.
+In case of uncertainty, the uncertain label is added, in case where the text does not mention any eco-innovation aspect, the innovation is labeled as not an eco-innovation.
+Where the source text only mentions that the innovation is less harmful to the environment, the label "Environmental-friendly technologies" is used.
+
+<!-- Use examples of clear cases here -->
+
+# Results {#sec-results}
+
+```{python}
+#| label: Number Bioeconomy Innovation
+#| output: asis
+
+sum_bioeconomy_visions = swinno["bioeconomy"].sum()
+
+display(
+    Markdown(
+        f"Querying the SWINNO database for forest based bioeconomy innovations initally yields {sum_bioeconomy_visions} innovations.
+This number includes false positives from substrings matching keywords included in the search.
+While there is no indication that these false positives are non-randomly distributed across time and sectors, the following analysis for bioeconomy innovation trends at large needs to be viewed in the light of this being unlceaned data.
+Where the analysis turns to specific bioeconomy visions, only clean and certain cases are discussed."
+    )
+)
+```
+
 ```{python}
@@
 SELECT
-    sinno_id
-    , innovation_type
+  sinno_id,
+  innovation_type as innovation_type_code,
+  c.Category as innovation_type
 FROM
-  eco_innovations;
+  eco_innovations
+JOIN classification_codes AS c ON c.Code = eco_innovations.innovation_type
+WHERE 
+innovation_type NOT IN (701, 602, 601, 501, 111, 109, 108)
+;
 """,
@@
 
+eco_innovations["innovation_type"] = eco_innovations["innovation_type"].str.strip()
+
 ```
+
+```{python}
+#| label: Calculate Bioeconomy Share
+
+def calculate_bioeconomy_share(swinno):
+    df = swinno.copy()
+    df = df[["sinno_id", "year", "bioeconomy"]]
+    df = (
+        df.groupby("year")
+        .agg(
+            total_count=("sinno_id", "nunique"), bioeconomy_count=("bioeconomy", "sum")
+        )
+        .reset_index()
+    )
+    df["bioeconomy_share"] = df["bioeconomy_count"] / df["total_count"]
+
+    return df
+
+bioeconomy_share = calculate_bioeconomy_share(swinno)
+
+```
+
+```{python}
+#| label: Bioeconomy Trend Description
+#| output: asis
+
+share_max = bioeconomy_share["bioeconomy_share"].max()
+share_max_year = bioeconomy_share.loc[bioeconomy_share["bioeconomy_share"] == share_max, "year"].values[0]
+share_last_year = bioeconomy_share["year"].max()
+share_latest = bioeconomy_share.loc[bioeconomy_share["year"] == share_last_year, "bioeconomy_share"].values[0]
+
+display(
+    Markdown(
+        f"@fig-bioeconomy_share shows the count (1) and share relative to all innovation contained in SWINNO (2) of bioeconomy innovation by year and as five year moving averages.
+From visual inspection alone, two features of the forest based bioeconomy innovation system emerge.
+First, two periods of high output can be identified from 1970-1990 and from 2005 to approximately 2015.
+The period in between 1990 and 2005 is characterized by low average output with high volatility.
+Second, the forest based bioeconomy has failed to gain in importance relative to other innovation captures by SWINNO.
+From a high of {share_max:.2%} in {share_max_year} to {share_latest:.2%} in {share_last_year}."
+    )
+)
+
+```
+
+```{python}
+# | label: fig-bioeconomy_share
+# | fig-cap: "Bioeconomy Innovation Trends in SWINNO (Uncleaned)"
+# | fig-subcap:
+# |   - "Bioeconomy Innovation Count"
+# |   - "Bioeconomy Innovation Share in %"
+# | layout-ncol: 1
+
+def plot_count(data):
+    fig, ax = plt.subplots()
+    ax.plot(
+        data["year"],
+        data["bioeconomy_count"],
+        label="Yearly",
+    )
+
+        # Calculate rolling mean and plot on same axis
+    rolling_mean = data["bioeconomy_count"].rolling(window=5).mean()
+    ax.plot(
+        data["year"],
+        rolling_mean,
+        label="5-year moving average",
+        linestyle="dashed"
+    )
+
+    ax.set_xticks(data["year"][::10])
+    ax.tick_params(labelsize=10)
+
+    ax.spines["top"].set_visible(False)
+    ax.spines["right"].set_visible(False)
+    ax.legend(frameon=False)
+
+    plt.show()
+
+def plot_share(data):
+    fig, ax = plt.subplots()
+    ax.plot(data["year"], data["bioeconomy_share"], label="Yearly")
+    ax.set_xticks(data["year"][::10])
+    ax.yaxis.set_major_formatter(mtick.PercentFormatter(xmax=1.0))
+    ax.tick_params(which="major", labelsize=10)
+
+    # Calculate rolling mean and plot on same axis
+    rolling_mean = data["bioeconomy_share"].rolling(window=5).mean()
+    ax.plot(
+        data["year"],
+        rolling_mean,
+        label="5-year moving average",
+        linestyle="dashed"
+    )
+    ax.set_ylim((0, .5))
+    ax.legend(frameon=False)
+    ax.spines["top"].set_visible(False)
+    ax.spines["right"].set_visible(False)
+    plt.show()
+
+plot_count(bioeconomy_share)
+plot_share(bioeconomy_share)
+```
+
+To find plausible explanations for the decline in innovation output, as well as explaining the variation in high and low output periods will be the focus of future versions of this paper.
+For now, it seems likely that the 1970s oil crisis and 1969 Environmental Protection Act provided strong pressures and windows of opportunity for the forest bioeconomy to innovate.
+Aspects of this have been discussed by many authors, for example @bergquist2011GreenInnovationSystems who illustrate how several government projects between 1960 and 1980 supported green innovation in Sweden's pulp and paper industry, and @taalbi2017WhatDrivesInnovation who argues that innovation in biomass harvest and use for energy was driven by the oil crisis and simultaneous wood shortage.
+Perhaps more interesting is to explain the slowdown in innovation output, despite the increasing role of the bioeconomy transition in political discourse.
+
+##  Visions of the Swedish Forest Bioeconomy Innnovation System
 
@@
 
+count_certain_bioeconomy_visions = (
+    certain_bioeconomy_visions
+    .groupby(["year", "bioeconomy_vision"])
+    .size()
+    .reset_index(name="count")
+)
+
+total_certain_bioeconomy_visions = (count_certain_bioeconomy_visions
+        .groupby("bioeconomy_vision")
+        .agg("sum", numeric_only=True)
+        .sort_values("count", ascending=False)
+        .reset_index()
+)
+
+```
+
+```{python}
+#| label: Number of Uncertains
+#| output: asis
+
+n_uncertain_bioeconomy_visions = uncertain_bioeconomy_visions["sinno_id"].nunique()
+
+display(
+    Markdown(
+        f"For the following part, innovations which do not relate to the forest are excluded.
+Excluded are also the {n_uncertain_bioeconomy_visions} innovations whose source text did not allow labeling without significant uncertaintiy.
+These innovations are planned to undergo a revision by subject matter experts at a later point."
+    )
+)
 ```
@@
 
+n_certain_bioeconomy_visions = certain_bioeconomy_visions['sinno_id'].nunique()
+
+certain_bioeconomy_visions_min_year = certain_bioeconomy_visions['year'].min()
+
+certain_bioeconomy_visions_max_year = certain_bioeconomy_visions['year'].max()
+
+n_most_bioeconomy_vision = total_certain_bioeconomy_visions.iloc[0, 1]
+
 display(
     Markdown(
-        f"So far, {certain_bioeconomy_visions['sinno_id'].nunique()} innovation have been classified for the period between {certain_bioeconomy_visions['year'].min()} and {certain_bioeconomy_visions['year'].max()}."
+        f"This leaves {n_certain_bioeconomy_visions} innovations have been classified for the period between {certain_bioeconomy_visions_min_year} and {certain_bioeconomy_visions_max_year}.
+With {n_most_bioeconomy_vision} innovations, the {total_certain_bioeconomy_visions.iloc[0, 0]} has the most associated innovations.
+Next the {total_certain_bioeconomy_visions.iloc[1, 0]} was associated with {total_certain_bioeconomy_visions.iloc[1,1]}, while {total_certain_bioeconomy_visions.iloc[2,1]} and  {total_certain_bioeconomy_visions.iloc[3,1]} innovation were classified as {total_certain_bioeconomy_visions.iloc[2,0]} and {total_certain_bioeconomy_visions.iloc[3,0]}, respectively.
+Note that the sum by category will not yield the total number of classified innovations, as one innovation can be associated with more than one bioeconomy vision."
     ))
+
+```
+
+Perhaps surprisingly, the majority of innovation related to the forest based bioeconomy does not involve one of the visions discussed in policy, industry and research.
+Working with the source articles suggests that one possible reason is that many innovation used or produced by sectors in the forest bioeconomy broadly aim to improve tools.
+For example to increase the lifespan of a milling machine metal tip.
+<!-- SINNO ID: 10022001 -->
+The second suprising thing, is the relative absence of biotechnology vision related innovation.
+Considering that this vision was promoted heavily at an international level by the OECD, it could be expected to play a more significant role in the Swedish innovation system.
+
+```{python}
+# | label: Count Recycling Innovation
+
+recycle_innovation = pd.read_sql(
+    """
+SELECT *
+FROM classification_codes
+LEFT JOIN eco_innovations ON classification_codes.Code = eco_innovations.innovation_type
+WHERE eco_innovations.innovation_type = 107 OR eco_innovations.innovation_type = 206 AND sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes LIKE '%not forest%')
+""",
+    conn,
+)
+
+recycle_innovation_count = recycle_innovation["sinno_id"].nunique()
+
+n_bioecology_vision = count_certain_bioeconomy_visions.loc[count_certain_bioeconomy_visions["bioeconomy_vision"] == "Bioecology Vision", "count"].sum()
+
+display(
+    Markdown(
+        f'For the Bioecology Vision it is important to note that a majority, but not all, innovation associated with this vision earned their label through the mention of recycling in the source.
+Currently, this applies to {recycle_innovation_count} innovations involving either a product or process innovation related to recyling.
+Hence, {n_bioecology_vision - recycle_innovation_count} innovation are associated with the bioecology vision, without using or facilitating recycling.
+These innovations require special focus, as they illustrate that the innovation system does already possess some "[c]ounter-expertise rather than concrete technical solutions" [@vivien2019HijackingBioeconomy 194], to embedd economic activity within the biosphere.
+This focus, however, is not within the scope of the current paper.'
+    )
+)
+```
+
+```{python}
+# | label: tbl-count_bioeconomy_visions
+# | tbl-cap: "Count of Bioeconomy Visions"
+# | output: asis
+# | include: false
+
+display(
+    Markdown(
+        count_certain_bioeconomy_visions.groupby("bioeconomy_vision")
+        .agg("sum", numeric_only=True)
+        .sort_values("count", ascending=False)
+        .reset_index()
+        .rename(columns={"bioeconomy_vision": "Vision", "count": "Count"})
+        .to_markdown(index=False)
+    )
+)
 ```
@@
 
-count_certain_bioeconomy_visions = (
-    certain_bioeconomy_visions.groupby(["year", "bioeconomy_vision"])
-    .size()
-    .reset_index(name="count")
-)
+plot_bv_trends(count_certain_bioeconomy_visions)
 
-def plot_bv_trends(df, title):
-    palette = sns.color_palette("colorblind")[:5]
+```
 
-    # Create a custom palette with a default light gray color for "not bio" hue
-    custom_palette = {
-        hue: color if bioeconomy_vision != "No Bioeconomy Vision" else (0.8, 0.8, 0.8)
-        for hue, color, bioeconomy_vision in zip(
-            df["bioeconomy_vision"], palette, df["bioeconomy_vision"]
-        )
-    }
-
-    chart = sns.lineplot(
-        data=df,
-        x="year",
-        y="count",
-        hue="bioeconomy_vision",
-        palette=custom_palette,
-        linewidth=2.5,
-    )
-
-    chart.set_title(title, fontdict={"fontsize": 16, "fontweight": "normal"})
-    chart.set_xlabel("Year", fontdict={"fontsize": 12, "fontweight": "normal"})
-    chart.set_ylabel("Count", fontdict={"fontsize": 12, "fontweight": "normal"})
-    chart.tick_params(labelsize=10)
-
-    # Customize x-axis tick labels
-    years = df["year"].unique()
-    chart.set_xticks(years[::5])
-    chart.set_xticklabels(years[::5], rotation=45, ha="right")
-
-    # Create a custom legend with colored labels (no lines)
-    handles, labels = chart.get_legend_handles_labels()
-
-    # Set font color for legend labels based on the hue
-    legend = chart.legend(
-        bbox_to_anchor=(0.5, -0.35), loc="upper center", ncol=2, frameon=False
-    )
-    for text, label in zip(legend.texts, labels):
-        text.set_color(custom_palette[label])
-
-    plt.tight_layout()
-
-plot_bv_trends(count_certain_bioeconomy_visions, title="")
-
-plt.show()
-```
+@fig-bioeconomy_trends illustrates that the peaks in innovation output, identified in @fig-bioeconomy_share and discussed above, were driven by innovation related to the Bioresource Vision as well as No Bioeconomy Vision.
+While the Bioresource Visions did strongly contribute to the first peak, the second appears to have been primarliy driven by innovation which cannot be associated as belonging to any vision of a bioeconomy.
+This raises a couple of questions regarding the effectiveness of past and current policy.
+It does add, for example, additional heft to the argument put forth by @holmgren2022WhoseTransformationThis that industry interest dominate the transformation to a bioeconomy.
+Further, the persistent dominance of the bioresource vision over the full study period suggests that there might not be much effort, at least in terms of production of innovations, put into creating a transformation towards a bioeconomy.
 
@@
 # | label: fig-vision_heatmap
-# | fig-caption: Sectors Contribution to Bioeconomy Visions
+# | fig-caption: "Sector's Contribution to Bioeconomy Visions"
+# | include: false
 
@@
 )
-# Question is, should I weight this following the way in the env-innovation paper?
+
+short_sector_labels = {
+    "Manufacture of other transport equipment": "Transport Equipment",
+    "Manufacture of motor vehicles, trailers and semi-trailers": "Motor Vehicle Manufacturing",
+    "Manufacture of furniture; manufacturing n.e.c.": "Furniture Manufacturing",
+    "Manufacture of fabricated metal products, except machinery and equipment": "Metal Products Manufacturing",
+    "Manufacture of machinery and equipment n.e.c.": "Machinery and Equipment",
+    "Construction": "Construction",
+    "Manufacture of electrical machinery and apparatus n.e.c.": "Electrical Machinery Manufacturing",
+    "Manufacture of radio, television and communication equipment and apparatus": "Communication Equipment",
+    "Manufacture of rubber and plastic products": "Rubber and Plastic Products",
+    "Forestry, logging and related service activities": "Forestry and Logging",
+    "Manufacture of wood and of products of wood and cork, except furniture; manufacture of articles of straw and plaiting materials": "Wood and Cork Products",
+    "Agriculture, hunting and related service activities": "Agriculture and Hunting",
+    "Land transport; transport via pipelines": "Land Transport",
+    "Supporting and auxiliary transport activities; activities of travel agencies": "Transport Services",
+    "Water transport": "Water Transport",
+    "Air transport": "Air Transport",
+    "Manufacture of medical, precision and optical instruments, watches and clocks": "Precision Instruments",
+    "Recycling": "Recycling",
+    "Manufacture of basic metals": "Basic Metals Manufacturing",
+    "Public administration and defence; compulsory social security": "Public Administration",
+    "Electricity, gas, steam and hot water supply": "Utilities",
+    "Sewage and refuse disposal, sanitation and similar activities": "Waste Management",
+    "Manufacture of chemicals and chemical products": "Chemical Products",
+    "Manufacture of pulp, paper and paper products": "Paper Products",
+    "Manufacture of coke, refined petroleum products and nuclear fuel": "Petroleum Products",
+    "Collection, purification and distribution of water": "Water Services",
+    "Publishing, printing and reproduction of recorded media": "Publishing and Media",
+    "Other mining and quarrying": "Mining and Quarrying",
+    "Mining of metal ores": "Metal Ore Mining",
+    "Tanning and dressing of leather; manufacture of luggage, handbags, saddlery, harness and footwear": "Leather and Footwear",
+    "Manufacture of food products and beverages": "Food and Beverages",
+    "Manufacture of other non-metallic mineral products": "Non-metallic Products",
+    "Wholesale trade and commission trade, except of motor vehicles and motorcycles": "Wholesale Trade",
+    "Other business activities": "Business Services",
+    "Research and development": "R&D",
+    "Manufacture of textiles": "Textile Manufacturing",
+    "Financial intermediation, except insurance and pension funding": "Financial Services",
+    "Education": "Education",
+    "Health and social work": "Healthcare",
+    "Fishing, operation of fish hatcheries and fish farms; service activities incidental to fishing": "Fishing and Aquaculture",
+    "Recreational, cultural and sporting activities": "Recreation and Culture",
+    "Extraction of crude petroleum and natural gas; service activities incidental to oil and gas extraction excluding surveying": "Oil and Gas Extraction",
+    "Post and telecommunications": "Post and Telecom",
+    "Sale, maintenance and repair of motor vehicles and motorcycles; retail sale of automotive fuel": "Automotive Services",
+    "Manufacture of wearing apparel; dressing and dyeing of fur": "Apparel Manufacturing",
+}
 
@@
 
+sector_counts["sector"] = sector_counts["sector"].map(short_sector_labels)
+
 heatmap_data = prepare_heatmap_data(sector_counts, "sector", "vision", "count")
 
-plot_heatmap(heatmap_data, "")
-
-# TODO: Better y axis labels
+plot_heatmap(heatmap_data)
 
@@
 ### How many eco-innovations are there?
-#### How many by bioeconomy vision?
+
+How does each vision contribute to lessening the environmental impact of the forest bioeconomy?
+While a quantiative assessment of technological impact on the environment comes with a host of challenges, the labelling of eco-innovations at least allows a glimpse at what could be expected from each vision.
+@fig-eco_innovation_by_bioeconomy_vision shows the count of eco-innovation associated with each bioeconomy vision by type.
+As before, there is no weighting, which is why the total sum of all eco-innovation across all bioeconomy visions outnumbers the sum of innovations in the bioeconomy.
+A single innovation can be more than one eco-innovation type, for example, if it reduces the amount of raw material needed by recylcing waste.
+The picture of innovation in the forest bioeconomy which @fig-eco_innovation_by_bioeconomy_vision suggests fits previous literature and what can be expected from LBIO data.
+Product and Process innovation are by far the most frequent across all visions.
+The Bioresource Vision is heavily involved with improving raw-material utilization, while the Bioecology Vision is most frequently associated with recycling.
+At the very least, this suggests that the definitions for each Vision and the classification scheme are in agreement.
+
+A noteworthy feature of innovation belonging to No Bioresource Vision is that this category has by far the most Non-Eco innovations.
+For the sustainability transition of the forest bioeconomy this could imply that having some Vision of a bioeconomy creates more eco-innovation than pursuing other innovation paths.
+
+```{python}
+# | label: merge eco-innovation with bioeconomy visions
+
+eco_innovation_merged_bv = certain_bioeconomy_visions.merge(
+    eco_innovations, on="sinno_id", validate="many_to_many"
+)
+
+eco_innovation_by_bioeconomy_visions = (
+    eco_innovation_merged_bv.groupby(["bioeconomy_vision", "innovation_type_code"])
+    .size()
+    .to_frame("count")
+    .reset_index()
+)
+
+eco_innovation_by_bioeconomy_visions["innovation_type_group"] = np.where(
+    eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0] == "1",
+    "Product",
+    np.where(
+        eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0] == "2",
+        "Process",
+        np.where(
+            eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0] == "3",
+            "Organizational",
+            np.where(
+                eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0]
+                == "4",
+                "Marketing",
+                "No Eco-Innovation",
+            ),
+        ),
+    ),
+)
+```
+
+```{python}
+#| label: fig-eco_innovation_by_bioeconomy_vision
+#| fig-cap: Count of Eco-Innovation Types by Bioeconomy Vision
+p = plot_eco_innovation_by_bioeconomy_vision(eco_innovation_by_bioeconomy_visions, codes_df)
+
+sns.move_legend(p, "lower center", ncol=5)
+# TODO: perhaps a per cent display by group size would be interesting too?
+# TODO: order y by label groups, color bars
+
+# def barplot_eco_innovation_by_bioeconomy_vision(data):
+
+```
+
+<!-- To answer this, I need to decide what is an eco innovation in the first place.
+should take the original values from @garcia-granero2018EcoinnovationMeasurementReview and remove codes 
+501 -- health and saftey
+701 -- Biocide
+108 -- Gen manipulated plant
+109 -- increased plant yield
+111 -- Gen manipulated animal
+  -->
+
+<!-- From the review of innovation two clusters of highly related innovations emerge.
+One revolves around optimizing harvest of biomass from forest as a response to the 1970s oil crisis.
+The second is a response to the envrionmental pollution problems faced by the pulp and paper industry.
+
+While in the first cluster environmental considerations played only a secondary role, they were crucial in the cluster around recyclability and water pollution prevention of the pulp and paper industry.
+This is not to say that the environmental impacts of harvest intensification were not at least to some extent on the agenda of the innovation system.
+One innovation source cites possible concerns about the environemntal effects of a root harvester, pointing to the need of future studies to establish the extent to which this demages forests.
+<!-- TODO: find innovation that discusses forest floor issues -->
+<!-- One source, after describing a new root harvester from  -->
+<!-- TODO: find company -->
+<!-- , also included a hand drawn charicature of a hiker, who, taking a break on the last stump on a hill and watching the harvester at work, exclaims "Thankfully they left at least one standing". -->
+
+## Discussion
 
@@
 | An ecological economy, that is   compatible with the biosphere<br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
-| Classificaiton Code | 1 | 2 | 3 | 
+| Classification Code | 1 | 2 | 3 | 
 :
@@
 # | tbl-cap: "List of Eco-Innovation Labels"
+format_table(codes_df.iloc[:5, :])
 
-codes_df = pd.read_sql(
-    "SELECT * FROM classification_codes", conn, index_col="Code"
-).iloc[:-5, :]
-
-format_table(codes_df)
 ```

### 2023-09-21 src/figures.py

@@ -4,37 +4,60 @@ import seaborn as sns
 import plotly.graph_objects as po
 import plotly.express as px
 
-def plot_bv_trends(df, title):
+def plot_bv_trends(df):
+    
+    palette = sns.color_palette("colorblind")[:5]
+
+    # Create a custom palette with a default light gray color for "not bio" hue
+    custom_palette = {
+        hue: color if bioeconomy_vision != "No Bioeconomy Vision" else (0.8, 0.8, 0.8, 0.8)
+        for hue, color, bioeconomy_vision in zip(
+            df["bioeconomy_vision"], palette, df["bioeconomy_vision"]
+        )
+    }
+
     chart = sns.lineplot(
         data=df,
         x="year",
         y="count",
         hue="bioeconomy_vision",
-        palette="colorblind",
-        linewidth=2.5,
+        palette=custom_palette,
+        linewidth=2
     )
-    chart.set_title(title, fontdict={"fontsize": 16, "fontweight": "bold"})
-    chart.set_xlabel("Year", fontdict={"fontsize": 12, "fontweight": "bold"})
-    chart.set_ylabel("Count", fontdict={"fontsize": 12, "fontweight": "bold"})
+
+    chart.set_xlabel("Year", fontdict={"fontsize": 12, "fontweight": "normal"})
+    chart.set_ylabel("Count", fontdict={"fontsize": 12, "fontweight": "normal"})
     chart.tick_params(labelsize=10)
-    plt.legend(title="Vision", fontsize=10, title_fontsize=12)
-    plt.tight_layout()
 
     # Customize x-axis tick labels
     years = df["year"].unique()
-    chart.set_xticks(years[::5])
-    chart.set_xticklabels(years[::5], rotation=45, ha="right")
+    chart.set_xticks(years[::10])
+    chart.set_xticklabels(years[::10])
 
-    return chart
+    # Create a custom legend with colored labels (no lines)
+    handles, labels = chart.get_legend_handles_labels()
+
+    # Set font color for legend labels based on the hue
+    legend = chart.legend(
+        bbox_to_anchor=(0.5, -0.35), loc="upper center", ncol=2, frameon=False
+    )
+    for text, label in zip(legend.texts, labels):
+        text.set_color(custom_palette[label])
+
+    sns.despine()
+
+    plt.show()
 
 
-def plot_heatmap(data, title):
+def plot_heatmap(data):
     fig, ax = plt.subplots(1, 1)
     ax = sns.heatmap(
[…]

### 2023-09-21 docs/reports/paper/paper1.qmd
add split table bioeconomy visions

@@
 Beyond conceptual uncertainties of what is meant by a bioeconomy, three different visions have emerged, differing in their aims and objectives, drivers and mediators of innovation and relation to the biosphere.
-The use of detailed innovation data from the SWINNO data base allows to capture the actual innovation output of the Swedish forest based bioeconomy innovation system.
+The use of detailed innovation data from a Swedish innovation database, SWINNO, allows to capture the actual innovation output of the Swedish forest based bioeconomy innovation system.
 Qualitative coding of the underlying source material associates innovation from 1970 to 2020 with vision of a bioresource, biotechnology and / or bioecology bioeconomy.
-Preliminary results suggest that the overall innovation output of the forest based bioeconomy has followed overall trends in the SWINNO database, failing to gain in prominence.
+Preliminary results suggest that the innovation output of the forest based bioeconomy has failed to gain in prominence.
 Of the three visions, the bioresource vision has been the most dominant, especially in the 1970s and after 2005.
@@
 colorlinks: true
+number-sections: true
 format:
@@
 
-# Debate around Bioeconomies
+# Introduction {#sec-introduction}
 
-An idea which is bound to increase the pressure and trade-offs between different ecosystem services is the bioeconomy.
-In recent years the concept has gained relevance among policymakers, practitioners and researchers.
-The term bioeconomy refers broadly to an economy which uses bio-based inputs instead of fossil based ones.
+The Baltic forests are facing a trilemma of expectations to mitigate climate change, preserve biodiversity, render other ecosystem services and provide the resources for a future bioeconomy [@hogbom2021TrilemmaNordicBaltic].
+Balancing these expectations includes managing conflicting goals and hence difficult trade-offs between ecosystem services. 
+Transitioning to a bioeconomy, meaning an economy based on bio-based rather than fossil-based inputs is set to increase the pressure on Baltic forests in the coming years [@hogbom2021TrilemmaNordicBaltic].
+In recent years transitioning to a bioeconomy has gained relevance among policymakers, practitioners and researchers. 
 Connected to this are hopes of increased resource efficiency and renewability.
-In a report commissioned by the Swedish Government, the @2012SwedishResearchInnovation defines a vision of a Swedish bioeconomy based on two central pillars.
+In a report commissioned by the Swedish Government, the @research2012StrategyBiobasedEconomy defines a vision of a Swedish bioeconomy based on two central pillars.
 It is to be an economy based on,
@@
 These pillars reflect the two most influential visions the bioeconomy concept.
-Notions of a bio-technology and bio-resource oriented bioeconomy are widespread and dominate discourse at various levels [@dieken2021MultitudesBioeconomiesSystematic].
+Notions of a biotechnology and bio-resource oriented bioeconomy are widespread and dominate discourse at various levels [@dieken2021MultitudesBioeconomiesSystematic].
 Despite -- or perhaps because of -- the bioeconomy lacking a strong definition, it is used to refer to different concepts, with different aims and objectives, foci, normative assumptions and hence implications for society and environment [@bugge2016WhatBioeconomyReview].
 
-The bio-technology vision focuses on economic growth and job provision through the application of bio-technology [@bugge2016WhatBioeconomyReview].
+The application of biotechnology to promote economic growth and provide jobs lends the biotechnology vision its name[@bugge2016WhatBioeconomyReview].
 Science driven innovation features prominently in this vision, with collaborations between research institutions in academia and industry taking on an important role [@bugge2016WhatBioeconomyReview].
@@
 As a policy narrative this vision was particularly potent and widespread at the turn of the 21st century.
-According to them, the intellectual heritage of Schumpeter and Kontradieff play an important role in this vision.
+According to @vivien2019HijackingBioeconomy, the intellectual heritage of Schumpeter and Kontradieff play an important role in this vision.
 Technological breakthroughs are poised to solve economic and ecological challenges, for example through genetic engineering.
@@
 In more recent years, visions of a bio-resource bioeconomy have taken over as the leading narrative.
-Unlike the bio-technology vision, this conceptualization does not only focus on economic growth and job creation, but attempts to include ecological sustainability challenges [@bugge2016WhatBioeconomyReview].
+Unlike the biotechnology vision, this conceptualization does not only focus on economic growth and job creation, but attempts to include ecological sustainability challenges [@bugge2016WhatBioeconomyReview].
 Biological resources, not technologies, form the basis of a new economy.
-In essence, there are two ways in which bio-resources are poised to replace fossil resources: as substitution, e.g., bio-fuels replacing fossil fuels, and as high value added products, e.g., Cross-laminated Timber used as a concrete alternative in construction [@vivien2019HijackingBioeconomy].
+In essence, there are two ways in which bio-resources are poised to replace fossil resources: as substitution, e.g., biofuels replacing fossil fuels, and as high value added products, e.g., Cross-laminated timber used as an alternative to concrete in construction [@vivien2019HijackingBioeconomy].
 To substitute fossil resources requires significant efforts in innovation.
-In contrast to the bio-technology vision, however, innovation is not confined to scientific advances from a narrow biochemical and pharmaceutical field.
+In contrast to the biotechnology vision, however, innovation is not confined to scientific advances from a narrow biochemical and pharmaceutical field.
 Instead, new and old actors are required to collaborate between and beyond established sectors [@bugge2016WhatBioeconomyReview].
@@
 There is no doubt that the Swedish vision places high demands on forests as primary providers of the required biomass.
-Both in the strategy formulated in @2012SwedishResearchInnovation and in public discourse, a small, well-connected group of actors center the bio-resource vision as legitimate and desirable [@holmgren2022WhoseTransformationThis].
+Both in the strategy formulated in @research2012StrategyBiobasedEconomy and in public discourse, a small, well-connected group of actors center the bio-resource vision as legitimate and desirable [@holmgren2022WhoseTransformationThis].
 The result is a forest sector which operates under a "more of everything" paradigm [@belandlindahl2015SwedishForestryModel].
@@
 
-Apart from the substitution of fossil resources through bio-based resources, a central narrative used in the Swedish transition is that, "\[t\]echnological innovation is key to a greener future" [@holmgren2022WhoseTransformationThis 42].
-According to @jankovsky2021InnovationsForestBioeconomy four types of innovation are central for a forest based bioeconomy:
+Apart from the substitution of fossil resources with bio-based resources, a central narrative used in the Swedish transition is that, "\[t\]echnological innovation is key to a greener future" [@holmgren2022WhoseTransformationThis 42].
+According to @jankovsky2021InnovationsForestBioeconomy there are four types of innovation crucial for a forest-based bioeconomy:
 
@@
 Unfortunately, these approaches are not without their flaws.
-Focusing on patents could overestimate the importance of the bio-technology vision, whose science oriented innovation approach is driven by patents [@bugge2016WhatBioeconomyReview].
-Qualitative case studies, on the other hand, may miss broader trends and features of a bioeconomy transition.
-
-@hausknost2017TransitionWhichBioeconomy
+Concentrating on patents might overemphasize the significance of the biotechnology vision, which predominantly relies on a science-oriented innovation approach driven by patents [@bugge2016WhatBioeconomyReview].
+On the other hand, qualitative case studies, may miss broader trends and features of a bioeconomy transition.
 
 Public, scientific and political debate of bioeconomies has brought forth a number of discourse analyses.
-Both @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy conclude that the concept of a bioeconomy conceals different, possibly conflicting visions of transfromation outcomes.
+Both @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy conclude that the concept of a bioeconomy conceals different, possibly conflicting visions of transformation outcomes.
 
@@
 
-Research Question:
-
-While much research has been done on the research, less is known over the actual direction of technical change in the Swedish forest bioeconomy.
-How innovative, in terms of output, is the system?
-What vision does the innovation system pursue?
-And how do bioeconomy visions differ in their eco-innovation orientation?
+This paper takes up this call and asks what bioeconomy the Swedish innovation system pursues.
+A unique database of actual innovation output is first used to answer how innovative the system has been between 1970 and 2020.
+The database sources are then used to provide a qualitative insight into the extent to which specific bioeconomy visions appear in commercialized innovation and the extent to which they include eco-innovations which could lessen environmental impact.
 
@@
 The list of keywords and sectors is presented in the appendix.
-Due to the query matching keywords whenever they appear as substrings in the description of the innovation, the inital search result is rather susceptible to false positives.
+Due to the query matching keywords whenever they appear as substrings in the description of the innovation, the initial search result is rather susceptible to false positives.
 In later manual review of each innovation for classification these false positives were removed from the search results.
@@
 An important limitation of LBIO measures, especially for research into forest sector innovation, is a bias against non-commercialized innovation.
-New processes, or new organizational apporaches for example, will only be caputred if they are sold to another company.
+New processes, or new organizational approaches for example, will only be captured if they are sold to another company.
 This is especially sensitive for bioeconomy research as, for example, new methods or plans for managing forest stands are less likely to be included.
-Relevant product innovation which faciliates these plans, for example products related to the planting of trees, on the other hand, are captured.
+Relevant product innovation which facilitates these plans, for example products related to the planting of trees, on the other hand, are captured.
 A likely consequence of this limitation is that the number of innovation related to the bioecology vision are under counted, as this vision highlights "Counter-expertise rather than concrete technical solutions" [@vivien2019HijackingBioeconomy 194].
-Nonetheless, a small number of innovation associated with this vision were already identified in the database.
+Nonetheless, a few innovations associated with this vision were already identified in the database which will be discussed later.
 
 But, measuring innovation with LBIO has also substantial advantages.
-For one, it ensures that actual innovation, meaning commericialized products and processes, are captured.
+For one, it ensures that actual innovation, meaning commercialized products and processes, are captured.
 Second, because the source articles are preserved, it provides a rich source of data, giving frequently providing insights into the aims, motivations, advantages and challenges of the covered innovation.
@@
 @tbl-bioeconomy_vision_summary shows the summary of bioeconomy visions presented in the cited papers.
+The analysis of bioeconomy discourse which revealed different visions, highlights that the visions have emerged at different times and with different actors at their centers [@vivien2019HijackingBioeconomy, @bugge2016WhatBioeconomyReview].
+As discourse elements the bioresource vision developed in
 
@@
 For one, economic growth is cited as the aim & objective of biotechnology and bioresource visions by @bugge2016WhatBioeconomyReview.
-It can be argued that all innovation can contribute to economic growth, which is why these aspects were not given hightened importance in the classification decision.
+It can be argued that all innovation can contribute to economic growth, which is why these aspects were not given heightened importance in the classification decision.
 A second topic of judgment has to be the inclusion of innovation which increases the recyclability of a product, or uses recycled waste, water and material in the process.
@@
 
-@tbl-vision_examples shows five examples of classfied innovation.
+:::{#tbl-bioeconomy_vision_summary layout-ncol=1}
++-------------------------------------+------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
+|                                     | Biotechnology                                                                      | Bioresource                                                                                                                                                                                                                              | Bioecology                                                                                                                                                                                                                          |
++=====================================+====================================================================================+==========================================================================================================================================================================================================================================+=====================================================================================================================================================================================================================================+
+| Aim &   Objectives                  | Economic growth & job   creation                                                   | Economic growth &   sustainability                                                                                                                                                                                                       | Sustainability, biodiversity,   conservation of ecosystems, avoiding soil degradation                                                                                                                                               |
++-------------------------------------+------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
+| Value   Creation                    | Application of biotechnology,   commercialisation of research & technology         | Conversion and upgrading of   bio-resources (process oriented)                                                                                                                                                                           | Development of integrated   production systems and high-quality products with territorial identity                                                                                                                                  |
++-------------------------------------+------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
+| Drivers   & mediators of innovation | R & D, patents, TTOs,   Research councils and funders (Science push, linear model) | Interdisciplinary, optimisation   of land use, include degraded land in the production of biofuels, use and   availability of bio-resources, waste management, engineering, science &   market (Interactive & networked production mode) | Identification of favourable   organic agro-ecological practices, ethics, risk, transdisciplinary   sustainability, ecological interactions, re-use & recycling of waste,   land use, (Circular and self-sustained production mode) |
++-------------------------------------+------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
+: Adapted from @bugge2016WhatBioeconomyReview [10] {tbl-colwidths="[25,25,25,25]"}
+
+|  | Type II | Type III | Type I |
+|:---|:---|:---|:---|
+|  | A science-based economy driven   by industrial biotechnology<br>     <br>     The cell is a factory | Biomass replaces fossil fuels   and mining to produce energy and materials<br>     <br>     Biorefining at the heart of ecological transition (multilevel perspective).
+| An ecological economy, that is   compatible with the biosphere<br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
+: Adapted from @vivien2019HijackingBioeconomy [194] {tbl-colwidths="[25,25,25,25]"}
+
+Bioeconomy Vision Summary 
+:::
+
+@tbl-vision_examples shows five examples of classified innovation.
 This selection also highlights that multiple labels can be attached to one innovation.
-The last innovation, describing a waterproof paper, is classified as a bioresource vission innovaition because it describes a process by which biomass is converted and upgraded.
-It is also classified as a biotechnology innovation since the process mimics trees' microfiber formation, accroding to the innovation source article.
+The last innovation, describing a waterproof paper, is classified as a bioresource vision innovation because it describes a process by which biomass is converted and upgraded.
+It is also classified as a biotechnology innovation since the process mimics trees' microfiber formation, according to the innovation source article.
 
@@
 This classification uses eco-innovation types identified in a recent literature review by @garcia-granero2018EcoinnovationMeasurementReview.
-They review 104 publications and produce a list 30 key Eco-Innovation Perfromance Indicators of product, process, organizational and marketing innovation.
+They review 104 publications and produce a list 30 key Eco-Innovation Performance Indicators of product, process, organizational and marketing innovation.
 
@@
 
-##  Visions of the Swedish Forest Bioeconomy Innnovation System
+##  Visions of the Swedish Forest Bioeconomy Innovation System
 
@@
 Perhaps surprisingly, the majority of innovation related to the forest based bioeconomy does not involve one of the visions discussed in policy, industry and research.
-Working with the source articles suggests that one possible reason is that many innovation used or produced by sectors in the forest bioeconomy broadly aim to improve tools.
+Working with the source articles suggests that one possible reason is that many innovations used or produced by sectors in the forest bioeconomy broadly aim to improve tools.
 For example to increase the lifespan of a milling machine metal tip.
 <!-- SINNO ID: 10022001 -->
-The second suprising thing, is the relative absence of biotechnology vision related innovation.
+The second surprising thing, is the relative absence of biotechnology vision related innovation.
 Considering that this vision was promoted heavily at an international level by the OECD, it could be expected to play a more significant role in the Swedish innovation system.
@@
 @fig-bioeconomy_trends illustrates that the peaks in innovation output, identified in @fig-bioeconomy_share and discussed above, were driven by innovation related to the Bioresource Vision as well as No Bioeconomy Vision.
-While the Bioresource Visions did strongly contribute to the first peak, the second appears to have been primarliy driven by innovation which cannot be associated as belonging to any vision of a bioeconomy.
+While the Bioresource Visions did strongly contribute to the first peak, the second appears to have been primarily driven by innovation which cannot be associated as belonging to any vision of a bioeconomy.
 This raises a couple of questions regarding the effectiveness of past and current policy.
@@
 How does each vision contribute to lessening the environmental impact of the forest bioeconomy?
-While a quantiative assessment of technological impact on the environment comes with a host of challenges, the labelling of eco-innovations at least allows a glimpse at what could be expected from each vision.
+While a quantitative assessment of technological impact on the environment comes with a host of challenges, the labelling of eco-innovations at least allows a glimpse at what could be expected from each vision.
 @fig-eco_innovation_by_bioeconomy_vision shows the count of eco-innovation associated with each bioeconomy vision by type.
 As before, there is no weighting, which is why the total sum of all eco-innovation across all bioeconomy visions outnumbers the sum of innovations in the bioeconomy.
-A single innovation can be more than one eco-innovation type, for example, if it reduces the amount of raw material needed by recylcing waste.
+A single innovation can be more than one eco-innovation type, for example, if it reduces the amount of raw material needed by recycling waste.
 The picture of innovation in the forest bioeconomy which @fig-eco_innovation_by_bioeconomy_vision suggests fits previous literature and what can be expected from LBIO data.
@@
 
-:::{#tbl-bioeconomy_vision_summary layout-ncol=1}
-| @bugge2016WhatBioeconomyReview [10] | Biotechnology | Bioresource | Bioecology |
-|---|---|---|---|
-| Aim &   Objectives | Economic growth & job   creation | Economic growth &   sustainability | Sustainability, biodiversity,   conservation of ecosystems, avoiding soil degradation |
-| Value   Creation | Application of biotechnology,   commercialisation of research & technology | Conversion and upgrading of   bio-resources (process oriented) | Development of integrated   production systems and high-quality products with territorial identity |
-| Drivers   & mediators of innovation | R & D, patents, TTOs,   Research councils and funders (Science push, linear model) | Interdisciplinary, optimisation   of land use, include degraded land in the production of biofuels, use and   availability of bio-resources, waste management, engineering, science &   market (Interactive & networked production mode) | Identification of favourable   organic agro-ecological practices, ethics, risk, transdisciplinary   sustainability, ecological interactions, re-use & recycling of waste,   land use, (Circular and self-sustained production mode) |
-
-| @vivien2019HijackingBioeconomy [194] | Type II | Type III | Type I |
-|---|---|---|---|
-|  | A science-based economy driven   by industrial biotechnology<br>     <br>     The cell is a factory | Biomass replaces fossil fuels   and mining to produce energy and materials<br>     <br>     Biorefining at the heart of ecological transition (multilevel perspective).
-| An ecological economy, that is   compatible with the biosphere<br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
-| Classification Code | 1 | 2 | 3 | 
-:
-
-Bioeconomy Vision Summary 
-:::
-
 <!-- Definition An ecological economy, that is compatible with the biosphere A science-based economy driven by industrial biotechnology A biomass-based economy Nature/economy relations Struggle against entropy and coevolution with the biosphere.

### 2023-09-21 docs/reports/paper/paper1.qmd
update classification description

@@
 
-#### Innovation for a Bioeconomy
+## Innovation for a Bioeconomy {#sec-innovation_bioeconomy}
 
@@
 
-# Data and Methods
+# Data and Methods {#sec-data_methods}
 
@@
 
-## Classification Methodology
+## Classification Methodology {#sec-classification_method}
 
@@
 
-### Bioeconomy Visions
+### Bioeconomy Visions {#sec-bioeconomy_classification_method}
 
 Innovations were classified as belonging to a bioeconomy vision if they fulfilled one, or more criteria outlined in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy.
-@tbl-bioeconomy_vision_summary shows the summary of bioeconomy visions presented in the cited papers.
+@tbl-bugge and @tbl-vivien show the summary of bioeconomy visions presented in the cited papers.
+For the remainder of the paper the nomenclature of @bugge2016WhatBioeconomyReview is adopted.
+Vivien et al.'s [-@vivien2019HijackingBioeconomy] Type I bioeconomy is taken to correspond with the bioecology vision, Type II and Type III to correspond with the bioresource and biotechnology vision, respectively.
 The analysis of bioeconomy discourse which revealed different visions, highlights that the visions have emerged at different times and with different actors at their centers [@vivien2019HijackingBioeconomy, @bugge2016WhatBioeconomyReview].
-As discourse elements the bioresource vision developed in
+@vivien2019HijackingBioeconomy trace the introduction of the term bioeconomics to the 1920s and a work on fishery by the Russian biologist Baranoff.
+The term bioeoconomics which they label as Type I bioeconomy, was introduced by Georgescu-Roegen,  in -@roegen1975BioeconomicAspectsEntropy.
+@vivien2019HijackingBioeconomy argue that the term biotechnology gradually emerged between 1990 and 2000, in the wake of enthusiasm about biotechnological adavances, such as the discovery of DNA, in preceeding years.
+It gained in popularity and international importance when the OECD began increasingly supporting and reshaping its meaning at the turn of the millenium [@vivien2019HijackingBioeconomy].
+According to them the last concept to enter the discourse was the bioresource oriented bioeconomy.
+While biomass has been used throughout history in various ways, the renewed interest with the intention to replace fossil inputs, for example in European Comission communication since 2010.
+However, @vivien2019HijackingBioeconomy point to the long history of biorefinery as a core aspect of this vision, which was discussed as early as the 1930s.
+Hence, while the organization of the discourse into the labels biotechnology, bioecology and bioresource vision is a very recent research outcome, the underlying concepts have a richer history.
+This alleviates concerns that using these labels retroactively confines the innovations to current policy terms.
 
-While following these criteria, a number of decisions were made concerning content of the summary presented in @tbl-bioeconomy_vision_summary and the nature of the corresponding vision.
-For one, economic growth is cited as the aim & objective of biotechnology and bioresource visions by @bugge2016WhatBioeconomyReview.
+While following the criteria of @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy, a number of decisions were made concerning content of the summary presented in @tbl-bugge and @tbl-vivien, and the nature of the corresponding vision.
+For one, economic growth is cited as the aim and objective of biotechnology and bioresource visions by @bugge2016WhatBioeconomyReview.
 It can be argued that all innovation can contribute to economic growth, which is why these aspects were not given heightened importance in the classification decision.
@@
 
-:::{#tbl-bioeconomy_vision_summary layout-ncol=1}
-+-------------------------------------+------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
 | | Biotechnology | Bioresource | Bioecology |
-+=====================================+====================================================================================+==========================================================================================================================================================================================================================================+=====================================================================================================================================================================================================================================+
+|---|---|---|---|
 | Aim &   Objectives | Economic growth & job   creation | Economic growth &   sustainability | Sustainability, biodiversity,   conservation of ecosystems, avoiding soil degradation |
-+-------------------------------------+------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
 | Value   Creation | Application of biotechnology,   commercialisation of research & technology | Conversion and upgrading of   bio-resources (process oriented) | Development of integrated   production systems and high-quality products with territorial identity |
-+-------------------------------------+------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
 | Drivers   & mediators of innovation | R & D, patents, TTOs,   Research councils and funders (Science push, linear model) | Interdisciplinary, optimisation   of land use, include degraded land in the production of biofuels, use and   availability of bio-resources, waste management, engineering, science &   market (Interactive & networked production mode) | Identification of favourable   organic agro-ecological practices, ethics, risk, transdisciplinary   sustainability, ecological interactions, re-use & recycling of waste,   land use, (Circular and self-sustained production mode) |
-+-------------------------------------+------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
-: Adapted from @bugge2016WhatBioeconomyReview [10] {tbl-colwidths="[25,25,25,25]"}
+: Table adapted from @bugge2016WhatBioeconomyReview [10] {#tbl-bugge}
 
 |  | Type II | Type III | Type I |
-|:---|:---|:---|:---|
+|---|---|---|---|
 |  | A science-based economy driven   by industrial biotechnology<br>     <br>     The cell is a factory | Biomass replaces fossil fuels   and mining to produce energy and materials<br>     <br>     Biorefining at the heart of ecological transition (multilevel perspective).
 | An ecological economy, that is   compatible with the biosphere<br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
-: Adapted from @vivien2019HijackingBioeconomy [194] {tbl-colwidths="[25,25,25,25]"}
-
-Bioeconomy Vision Summary 
-:::
+: Table adapted from @vivien2019HijackingBioeconomy [194] {#tbl-vivien}
 
@@
 It is also classified as a biotechnology innovation since the process mimics trees' microfiber formation, according to the innovation source article.
+Lastly, there is also the possibility that an innovation cannot be placed into at least one of the visions.
+This is the case for innovation which brings general technological improvements, purely advances economic performance, or relates to a general purpose technology such as appliance of computers.
+These innovations are labeled as No Bioeconomy Vision, which does not imply that the innovation is not used, or originates in the forest based bioeconomy as defined above.
+An illustrative example of such an innovation are improved work boots, which promote safety of forest workers.
 
@@
 
-### Eco-Innovation Types
+### Eco-Innovation Types {#sec-eco_innovatoin_types}
 

### 2023-09-21 docs/reports/paper/paper1.qmd
update papersize to a4

@@
 format:
-    wordcount-pdf: default
+    wordcount-pdf: 
+        papersize: A4
 execute:
@@
 
-## Discussion
+# Discussion
 
-# References
+{{< pagebreak >}}
+
+# References {.unnumbered}
 

### 2023-09-22 docs/reports/paper/paper1.qmd
update figure2 bv count

@@
 ---
-title: The Direction of Search for a Forest Based Bioeconomy
+title: Shades of Green in the Swedish Forest-Based Bioeconomy Transition 
+subtitle: Draft
 author: 
@@
 @vivien2019HijackingBioeconomy trace the introduction of the term bioeconomics to the 1920s and a work on fishery by the Russian biologist Baranoff.
-The term bioeoconomics which they label as Type I bioeconomy, was introduced by Georgescu-Roegen,  in -@roegen1975BioeconomicAspectsEntropy.
+The term bioeoconomics which they label as Type I bioeconomy, was introduced by Georgescu-Roegen,  in -@georgescu-roegen1975BioeconomicAspectsEntropy.
 @vivien2019HijackingBioeconomy argue that the term biotechnology gradually emerged between 1990 and 2000, in the wake of enthusiasm about biotechnological adavances, such as the discovery of DNA, in preceeding years.
@@
 # | label: fig-bioeconomy_trends
-# | fig-cap: Bioeconomy Vision Trends
+# | fig-cap: "Count of Bioeconomy Visions by Year"
 
@@
 
-# Discussion
-
 {{< pagebreak >}}

### 2023-09-22 src/figures.py

@@ -10,7 +10,7 @@ def plot_bv_trends(df):
 
     # Create a custom palette with a default light gray color for "not bio" hue
     custom_palette = {
-        hue: color if bioeconomy_vision != "No Bioeconomy Vision" else (0.8, 0.8, 0.8, 0.8)
+        hue: color if bioeconomy_vision != "No Bioeconomy Vision" else (0.7, 0.7, 0.7, 0.6)
         for hue, color, bioeconomy_vision in zip(
             df["bioeconomy_vision"], palette, df["bioeconomy_vision"]
         )
@@ -25,24 +25,27 @@ def plot_bv_trends(df):
         linewidth=2
     )
 
-    chart.set_xlabel("Year", fontdict={"fontsize": 12, "fontweight": "normal"})
-    chart.set_ylabel("Count", fontdict={"fontsize": 12, "fontweight": "normal"})
     chart.tick_params(labelsize=10)
 
     # Customize x-axis tick labels
     years = df["year"].unique()
     chart.set_xticks(years[::10])
     chart.set_xticklabels(years[::10])
+    chart.set_xlabel(None)
+    chart.set_ylabel(None)
 
     # Create a custom legend with colored labels (no lines)
     handles, labels = chart.get_legend_handles_labels()
 
     # Set font color for legend labels based on the hue
     legend = chart.legend(
-        bbox_to_anchor=(0.5, -0.35), loc="upper center", ncol=2, frameon=False
+        bbox_to_anchor=(0.5, -0.1), loc="upper center", ncol=2, frameon=False
     )
     for text, label in zip(legend.texts, labels):
-        text.set_color(custom_palette[label])
+        if label == "No Bioeconomy Vision":
+            text.set_color((0.4,0.4,0.4))
+        else:
+            text.set_color(custom_palette[label])
 
     sns.despine()
 

### 2023-09-22 docs/reports/paper/paper1.qmd
add doublespaced lines

@@
         papersize: A4
+        linestretch: 1.5
 execute:

### 2023-09-25 docs/reports/paper/paper1.qmd
update draft sehm handin

@@
 The use of detailed innovation data from a Swedish innovation database, SWINNO, allows to capture the actual innovation output of the Swedish forest based bioeconomy innovation system.
-Qualitative coding of the underlying source material associates innovation from 1970 to 2020 with vision of a bioresource, biotechnology and / or bioecology bioeconomy.
-Preliminary results suggest that the innovation output of the forest based bioeconomy has failed to gain in prominence.
+Qualitative coding of the underlying source material associates innovation from 1970 to 2010 with visions of a bioresource, biotechnology and / or bioecology bioeconomy.
+Preliminary results suggest that the innovation output of the forest based bioeconomy has decreased relative to the total output.
 Of the three visions, the bioresource vision has been the most dominant, especially in the 1970s and after 2005.
@@
 Connected to this are hopes of increased resource efficiency and renewability.
-In a report commissioned by the Swedish Government, the @research2012StrategyBiobasedEconomy defines a vision of a Swedish bioeconomy based on two central pillars.
+In a report commissioned by the Swedish Government, the @2012StrategyBiobasedEconomy defines a vision of a Swedish bioeconomy based on two central pillars.
 It is to be an economy based on,
@@
 
-These pillars reflect the two most influential visions the bioeconomy concept.
+These pillars reflect the two most influential visions of the bioeconomy concept.
 Notions of a biotechnology and bio-resource oriented bioeconomy are widespread and dominate discourse at various levels [@dieken2021MultitudesBioeconomiesSystematic].
-Despite -- or perhaps because of -- the bioeconomy lacking a strong definition, it is used to refer to different concepts, with different aims and objectives, foci, normative assumptions and hence implications for society and environment [@bugge2016WhatBioeconomyReview].
+Despite -- or perhaps because of -- bioeconomy lacking a strong definition, the term is used to refer to different concepts, with different aims and objectives, foci, normative assumptions and hence implications for society and environment [@bugge2016WhatBioeconomyReview].
 
-The application of biotechnology to promote economic growth and provide jobs lends the biotechnology vision its name[@bugge2016WhatBioeconomyReview].
+The application of biotechnology to promote economic growth and provide jobs lends the biotechnology vision its name [@bugge2016WhatBioeconomyReview].
 Science driven innovation features prominently in this vision, with collaborations between research institutions in academia and industry taking on an important role [@bugge2016WhatBioeconomyReview].
@@
 Technological breakthroughs are poised to solve economic and ecological challenges, for example through genetic engineering.
-Consequently, this vision has a weak sustainability notion, viewing sustainability challenges as temporary and nature as substitutable given sufficient research and knowledge creation [@vivien2019HijackingBioeconomy].
+Consequently, this vision has a weak sustainability notion, viewing sustainability challenges as temporary, and nature as substitutable given sufficient research and knowledge creation [@vivien2019HijackingBioeconomy].
 
@@
 Biological resources, not technologies, form the basis of a new economy.
-In essence, there are two ways in which bio-resources are poised to replace fossil resources: as substitution, e.g., biofuels replacing fossil fuels, and as high value added products, e.g., Cross-laminated timber used as an alternative to concrete in construction [@vivien2019HijackingBioeconomy].
+In essence, there are two ways in which bio-resources are poised to replace fossil resources: as substitution, e.g., biofuels replacing fossil fuels, and as high value added products, e.g., cross-laminated timber used as an alternative to concrete in construction [@vivien2019HijackingBioeconomy].
 To substitute fossil resources requires significant efforts in innovation.
@@
 Both @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy conclude that the concept of a bioeconomy conceals different, possibly conflicting visions of transformation outcomes.
-
-Recent scholarship has called for advancing research fronts instead of remaining entrenched in contested discourse positions @allain2022BioeconomicsVsBioeconomy.
+Recent scholarship has called for advancing research fronts instead of remaining entrenched in contested discourse positions [@allain2022BioeconomicsVsBioeconomy].
 
@@
     f"The SWINNO database used here contains {n_swinno} innovations commercialized between {year_min} and {year_max}.
-It is constructed according to  literature-based innovation output (LBIO) method [@sjoo2014DatabaseSwedishInnovations].
+It is constructed according to  literature-based innovation output (LBIO) method using Swedish source material [@sjoo2014DatabaseSwedishInnovations].
 This method relies on the judgment of editiors and journalists at trade journals and hence captures what can be described as significant innovation."
@@
 @vivien2019HijackingBioeconomy trace the introduction of the term bioeconomics to the 1920s and a work on fishery by the Russian biologist Baranoff.
-The term bioeoconomics which they label as Type I bioeconomy, was introduced by Georgescu-Roegen,  in -@georgescu-roegen1975BioeconomicAspectsEntropy.
-@vivien2019HijackingBioeconomy argue that the term biotechnology gradually emerged between 1990 and 2000, in the wake of enthusiasm about biotechnological adavances, such as the discovery of DNA, in preceeding years.
-It gained in popularity and international importance when the OECD began increasingly supporting and reshaping its meaning at the turn of the millenium [@vivien2019HijackingBioeconomy].
-According to them the last concept to enter the discourse was the bioresource oriented bioeconomy.
-While biomass has been used throughout history in various ways, the renewed interest with the intention to replace fossil inputs, for example in European Comission communication since 2010.
+The term bioeoconomics which they label as Type I bioeconomy, was introduced by Georgescu-Roegen, in [-@georgescu-roegen1975BioeconomicAspectsEntropy].
+@vivien2019HijackingBioeconomy argue that the term biotechnology gradually emerged between 1990 and 2000, in the wake of enthusiasm about biotechnological advances, such as the discovery of DNA.
+It gained popularity and international importance when the OECD began increasingly supporting and reshaping its meaning at the turn of the millennium [@vivien2019HijackingBioeconomy].
+According to them, the last concept to enter the discourse was the bioresource oriented bioeconomy.
+While biomass has been used throughout history in various ways, the renewed interest with the intention to replace fossil inputs, for example in European Commission communication since 2010.
 However, @vivien2019HijackingBioeconomy point to the long history of biorefinery as a core aspect of this vision, which was discussed as early as the 1930s.
@@
 The last innovation, describing a waterproof paper, is classified as a bioresource vision innovation because it describes a process by which biomass is converted and upgraded.
-It is also classified as a biotechnology innovation since the process mimics trees' microfiber formation, according to the innovation source article.
+It is also classified as a biotechnology innovation since, according to the innovation source article, the process mimics trees' microfiber formation.
 Lastly, there is also the possibility that an innovation cannot be placed into at least one of the visions.
@@
 These innovations are labeled as No Bioeconomy Vision, which does not imply that the innovation is not used, or originates in the forest based bioeconomy as defined above.
-An illustrative example of such an innovation are improved work boots, which promote safety of forest workers.
+An illustrative example of such an innovation are improved work boots, which promote safety of forest workers but do not correspond with any of the criteria in @tbl-bugge or @tbl-vivien.
 
@@
 ```{python}
-#| label: Bioeconomy Trend Description
-#| output: asis
+# | label: Bioeconomy Trend Description
+# | output: asis
 
 share_max = bioeconomy_share["bioeconomy_share"].max()
-share_max_year = bioeconomy_share.loc[bioeconomy_share["bioeconomy_share"] == share_max, "year"].values[0]
+share_max_year = bioeconomy_share.loc[
+    bioeconomy_share["bioeconomy_share"] == share_max, "year"
+].values[0]
 share_last_year = bioeconomy_share["year"].max()
-share_latest = bioeconomy_share.loc[bioeconomy_share["year"] == share_last_year, "bioeconomy_share"].values[0]
+share_latest = bioeconomy_share.loc[
+    bioeconomy_share["year"] == share_last_year, "bioeconomy_share"
+].values[0]
 
@@
     Markdown(
-        f"@fig-bioeconomy_share shows the count (1) and share relative to all innovation contained in SWINNO (2) of bioeconomy innovation by year and as five year moving averages.
+        f"@fig-bioeconomy_share shows the count (a) and share relative to all innovation contained in SWINNO (b) of bioeconomy innovation by year and as five year moving averages.
 From visual inspection alone, two features of the forest based bioeconomy innovation system emerge.

### 2023-09-25 docs/presentations/20230829-Umea_workshop.qmd
add presentations

---
title: Shades of Green
subtitle: Direction of the Swedish Forest Innovation System 1970-2020
author: Jonas Kreutzer
date: last-modified
abstract: |

csl: https://zotero.org/styles/apa?source=1
bibliography: [../../swinno-shades-green.bib]

format: revealjs
# format:
#   revealjs:
#     theme: [../assets/revealme.scss]
#     transition: none
#     title-slide-attributes:
#       data-background-image: '../assets/ground-fog-valley-black-forest-sunset.jpg'
#       data-background-size: fit
#       data-background-opacity: '0.3'
#     embed-resources: true

execute:
    freeze: auto
    echo: false
jupyter: python3
---

```{python}

from IPython.display import display, Markdown

import pandas as pd
import numpy as np

import matplotlib.pyplot as plt
import seaborn as sns
import altair as alt

from src.swinno_helpers import connect_swinno_db

swinno_db = connect_swinno_db()

# set all text on plots to be monospace with rcParams
plt.rcParams["font.family"] = "monospace"

sns.set_theme(style="whitegrid")

def format_table(df):
    # create a Styler object and set the wrapping style
    styler = df.style.set_properties(**{"width": "50em", "text-wrap": "break-word"})

    return styler

def plot_counts(
    df1, df2, category_col, y, show_legend=True, legend_cols=None, y_label="Count"
):
    categories = df1[category_col].astype(str).unique()
    categories.sort()

    legend_ncols = len(categories) if legend_cols is None else legend_cols

    colors = sns.color_palette("colorblind", len(categories))
    palette = dict(zip(categories, colors))

    fig, axes = plt.subplots(
        ncols=1, nrows=2, sharex=False, sharey=True, layout="tight"
    )

    marker_size = 100
    df1 = df1.sort_values(by=category_col)
    df2 = df2.sort_values(by=category_col)

    sns.scatterplot(
        x=df1.index,
        y=y,
        hue=category_col,
        data=df1,
        ax=axes[0],
        palette=palette,
        s=marker_size,
    )

    sns.scatterplot(
        x=df2.index,
        y=y,
        hue=category_col,
        data=df2,
        ax=axes[1],
        palette=palette,
        s=marker_size,
    )

    for axs in axes:
        axs.set_ylabel(y_label)
        axs.set_xlabel("Year")
        axs.legend().remove()

    axes[0].set_title("Including Uncertains")
    axes[0].title.set_position([0.5, 1.05])  # adjust the position of the title

    axes[1].set_title("Excluding Uncertains")
    axes[1].title.set_position([0.5, 1.05])  # adjust the position of the title

    if not show_legend:
        fig.legend().remove()
    else:
        handles, labels = axes[0].get_legend_handles_labels()
        fig.legend(
            handles,
            labels,
            loc="lower center",
            ncol=legend_ncols,
            frameon=False,
            bbox_to_anchor=(0.5, -0.1),
        )  # adjust the position of the legend

    plt.subplots_adjust(bottom=0.2)

    plt.show()

def plot_heatmap(data, title):
    fig, ax = plt.subplots(1, 1)
    ax = sns.heatmap(
        data, annot=True, cmap="Blues", fmt="g", cbar=False, annot_kws={"fontsize": 10}
    )
    ax.set(title=title, ylabel=None, xlabel=None)
    ax.set_xticklabels(ax.get_xticklabels(), rotation=45, ha="right")

    plt.show()

def prepare_heatmap_data(df, index_col, cols_col, values_col, ascending=False):
    """
    Prepare data for a heatmap visualization by pivoting and sorting a dataframe.

    Args:
    - df: input dataframe
    - index_col: column name for the index of the pivoted dataframe
    - cols_col: column name for the columns of the pivoted dataframe
    - values_col: column name for the values of the pivoted dataframe
    - ascending: boolean to specify whether to sort in ascending or descending order (default: False)

    Returns:
    - pivoted and sorted dataframe
    """
    # Pivot the data to create a matrix
    matrix = df.pivot(index=index_col, columns=cols_col, values=values_col)

    # Sort the matrix by descending order of values in the column with the highest value
    highest_col = matrix.idxmax(axis=1).values[0]
    matrix = matrix.sort_values(highest_col, ascending=ascending)

    return matrix

```

## Bioeconomy Demarkations {.smaller}

::::{.columns}

:::{.column width="50%"}
**Using producer / user codes
**

|SNI Code | Description                            |
|--------:|:---------------------------------------|
|02       | Forestry and related services
|20       | Wood and wood product manufacturing except furniture
|21       | Pulp, paper and paper product manufacturing
|36       | Furniture manufacturing; other manufacturing
:::

:::{.column width="50%"}
**Adding keywords of value chain**

    - "virke",
    - "cellulos",
    - "lignin",
    - "spån", 
    - "bark",
    - "levulinsyra" (Levulinic acid),
    - "furfural" (Furfural),
    - "svarttjära",
    - "svartlut",
    - "växtbas",
    - "ved",
    - "trä",
    - "skog",
    - "papper",
    - "biobränsle",
    - "biologiskt",
    - "nedbrytbar",
    - "papper",
    - "pappret",
    - "karton",
    - "tencel",
:::

::::

## First Estimation of Forest Bioeconomy Innovativeness

```{python}

bioeconomy = pd.read_sql(
    """
select
  i.sinno_id,
  i.innovation_name_in_swedish AS name,
  i.description_in_swedish AS description,
  i.additional_information_if_origin_new_scientific_discovery || i.additional_information_if_origin_new_technologies_or_materials || i.additional_info_if_origin_official_regulation_legislation_and_standards || i.additional_information_if_origin_solution_for_a_problem || i.additional_information_if_origin_performance || i.additional_information_if_origin_other AS info,
  i.year_of_commercialization AS year,
  us.use_sector
from
  innovation i
  join use_sectors us on i.sinno_id = us.sinno_id
where
  (
    us.use_sector like '02%'
    or us.use_sector like '20%'
    or us.use_sector like '21%'
    or us.use_sector like '36%'
    or product_code like '02%'
    or product_code like '20%'
    or product_code like '21%'
    or product_code like '36%'
  )
  or (
    description like '%virke%'
    or description like '%cellulos%'
    or description like '%lignin%'
    or description like '%spån%'
    or description like '%bark%'
    or description like '%levulinsyra%'
    or description like '%furfural%'
    or description like '%svarttjära%'
    or description like '%svartlut%'
    or description like '%växtbas%'
    or description like '%ved%'
    or description like '%trä%'
    or description like '%skog%'
    or description like '%biobränsle%'
    or description like '%biologisk%'
    or description like '%nedbrytbar%'
    or description like '%papper%'
    or description like '%pappret%'
    or description like '%karton%'
    or description like '%tencel%'
  );
""",
    swinno_db,
)

sni_codes = pd.read_sql_query(
    """
select
*
from 
sni_codes
""",
    swinno_db,
)

sni_codes = sni_codes.rename(columns={"code": "use_sector"})

swinno = pd.read_sql_query(
    """
select sinno_id, year_of_commercialization as year, innovation_name_in_swedish as name
from innovation;
""",
    swinno_db,
)

swinno["bioeconomy"] = swinno["sinno_id"].isin(bioeconomy["sinno_id"].unique())

def plot_innovation_counts_by_sector(df, title):
    df = df.copy()
    df["sector"] = df["use_sector"].astype(str).str[:2]
    df = (
        df.groupby("sector")
        .nunique()["sinno_id"]
        .sort_values(ascending=False)
        .reset_index()
    )
    df = df.rename(columns={"sinno_id": "count"})
    df = df.join(sni_codes.set_index("use_sector"), on="sector", how="left")
    df = df[["count", "sector", "label"]]

    chart = (
        alt.Chart(df)
        .mark_bar()
        .encode(
            x=alt.X("count:Q", title="Number of innovations"),
            y=alt.Y("label:N", title="Sector", sort="-x"),
            tooltip=["count"],
        )
        .properties(title=title, height=700)
    )

    return chart

plot_innovation_counts_by_sector(bioeconomy, "Innovation counts by sector in SWINNO")

```

# Trend

```{python}
# | label: plot bioeconomy share
def calculate_bioeconomy_share(df):
    df = df.copy()
    df = df[["sinno_id", "year", "bioeconomy"]]
    df = (
        df.groupby("year")
        .agg(
            total_count=("sinno_id", "nunique"), bioeconomy_count=("bioeconomy", "sum")
        )
        .reset_index()
    )
    df["bioeconomy_share"] = df["bioeconomy_count"] / df["total_count"]

    return df

def plot_innovation_counts_by_year(df, title):
    df = calculate_bioeconomy_share(df)

    base = alt.Chart(df).encode(alt.X("year:O", axis=alt.Axis(title="Year")))

    count = (
        base.mark_line(color="red", opacity=0.6)
        .encode(
            alt.Y("bioeconomy_count:Q", title="Count of Bioeconomy Innovations"),
            tooltip=["bioeconomy_count", "bioeconomy_share", "year"],
        )
        .interactive()
    )

    share = (
        base.mark_line()
        .encode(
            alt.Y("bioeconomy_share:Q", title="Share of Bioeconomy Innovations"),
            tooltip=["bioeconomy_count", "bioeconomy_share", "year"],
        )
        .interactive()
    )

    chart = (
        alt.layer(count, share)
        .resolve_scale(y="independent")
        .properties(title=title, width=900)
        .configure_title(font="Inconsolata")
        .configure_axis(labelFont="Inconsolata", titleFont="Inconsolata")
    )

    return chart

plot_innovation_counts_by_year(swinno, "Innovation counts by year in SWINNO")
```

# Three Visions for a Bioeconomy

:::: columns
::: {.column width="33%"}
![Bio-Technology Vision](../assets/Biotech.jpg){.fragment}
:::

::: {.column width="33%"}
![Bio-Resource Vision](../assets/Bioresource.jpg){.fragment}
:::

::: {.column width="33%"}
![Bio-Ecology Vision](../assets/Bioecology.jpg){.fragment}
:::
::::

[@bugge2016WhatBioeconomyReview]{.slide-cite style="font-size: x-small"}

## Bioeconomy Vision Classification{.smaller}

| @bugge2016WhatBioeconomyReview | Biotechnology | Bioresource | Bioecology |
|---|---|---|---|
| Aim &   Objectives | Economic growth & job   creation | Economic growth &   sustainability | Sustainability, biodiversity,   conservation of ecosystems, avoiding soil degradation |
| Value   Creation | Application of biotechnology,   commercialisation of research & technology | Conversion and upgrading of   bio-resources (process oriented) | Development of integrated   production systems and high-quality products with territorial identity |
| Drivers   & mediators of innovation | R & D, patents, TTOs,   Research councils and funders (Science push, linear model) | Interdisciplinary, optimisation   of land use, include degraded land in the production of biofuels, use and   availability of bio-resources, waste management, engineering, science &   market (Interactive & networked production mode) | Identification of favourable   organic agro-ecological practices, ethics, risk, transdisciplinary   sustainability, ecological interactions, re-use & recycling of waste,   land use, (Circular and self-sustained production mode) |

| @vivien2019HijackingBioeconomy | Type II | Type III | Type I |
|---|---|---|---|
|  | A science-based economy driven   by industrial biotechnology<br>     <br>     The cell is a factory | Biomass replaces fossil fuels   and mining to produce energy and materials<br>     <br>     Biorefining at the heart of ecological transition (multilevel perspective). | An ecological economy, that is   compatible with the biosphere<br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
| Codes | 1 | 2 | 3 |

## Preliminary Results

```{python}
bv_df = pd.read_sql(
    """SELECT bv.sinno_id, bv.bioeconomy_vision, i.year_of_commercialization AS "i.year"
FROM bioeconomy_visions AS bv
JOIN innovation AS i ON i.sinno_id = bv.sinno_id
WHERE bv.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%");
""",
    swinno_db,
)
```

```{python}

display(
    Markdown(
        f"<span style='font-size:10rem'>{bv_df.loc[:, 'sinno_id'].nunique()}</span> unique forest bioeconomy innovations classified"
    )
)

```

## More

```{python}
ei_count = pd.read_sql(
    """
SELECT
  codes.Category AS Type,
  COUNT(ei.innovation_type) AS Count
FROM
  eco_innovations as ei
  JOIN [classification_codes] AS codes ON codes.code = ei.innovation_type
WHERE ei.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
GROUP BY
  innovation_type
ORDER BY
  Count DESC;
""",
    swinno_db,
)

ei_count_0 = pd.read_sql(
    """
SELECT
  codes.Category AS Type,
  COUNT(ei.innovation_type) AS Count
FROM
  eco_innovations as ei
  JOIN [classification_codes] AS codes ON codes.code = ei.innovation_type
WHERE
  NOT EXISTS(
    SELECT * 
    FROM
      eco_innovations AS ei2
    WHERE
      ei2.sinno_id = ei.sinno_id
      AND ei2.innovation_type = "000"
  )
AND ei.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
GROUP BY
  Type
ORDER BY
  Count DESC;
""",
    swinno_db,
)

bv_count = pd.read_sql(
    """
SELECT
  codes.Category AS Vision,
  COUNT(bioeconomy_vision) AS Count
FROM
  bioeconomy_visions
  JOIN [classification_codes] AS codes ON codes.code = bioeconomy_visions.bioeconomy_vision
WHERE
  bioeconomy_visions.sinno_id NOT IN (
    SELECT
      sinno_id
    FROM
      categorization_notes
    WHERE
      notes NOT LIKE "%not forest%"
  )
GROUP BY
  Vision
ORDER BY
  Count DESC;
   """,
    swinno_db,
)

bv_count_zero = pd.read_sql(
    """
SELECT
  codes.Category AS Vision,
  COUNT(bioeconomy_vision) AS Count
FROM
  bioeconomy_visions
  JOIN [classification_codes] AS codes ON codes.code = bioeconomy_visions.bioeconomy_vision
WHERE
  NOT EXISTS (
    SELECT
      *
    FROM
      bioeconomy_visions AS bv2
    WHERE
      bv2.sinno_id = bioeconomy_visions.sinno_id
      AND bv2.bioeconomy_vision = 0
  )
  AND bioeconomy_visions.sinno_id NOT IN (
    SELECT
      sinno_id
    FROM
      categorization_notes
    WHERE
      notes NOT LIKE "%not forest%"
  )
GROUP BY
  Vision
ORDER BY
  Count DESC;
""",
    swinno_db,
)
```

## Count of Bioeconomy Visions
```{python}
bv_counts = pd.merge(bv_count, bv_count_zero, on="Vision", how="left")

bv_counts.columns = ["Vision", "Count", "Count Excl. Unsure"]
bv_counts.loc[:, "Count Excl. Unsure"] = (
    bv_counts["Count Excl. Unsure"].fillna(0).astype(int)
)

bv_counts.style.set_properties(**{"text-align": "right"}).hide_index()

```

## Count of Eco-Innovations

```{python}
# | label: tbl-ei-article-count
# | tbl-cap: Count of Eco-Innovations Using Swinno Data

ei_counts = pd.merge(ei_count, ei_count_0, on="Type", how="left")

ei_counts.columns = ["Type", "Count", "Count Excl. Unsure"]
ei_counts.loc[:, "Count Excl. Unsure"] = (
    ei_counts["Count Excl. Unsure"].fillna(0).astype(int)
)

ei_counts.style.set_properties(**{"text-align": "right"}).hide_index()
```

## Bioeconomy Vision Heatmap

```{python}
# | label: fig-bva-sectors
# | fig-cap: Heatmap of Bioeconomy Visions by Sector

bv_sectors = pd.read_sql(
    """
SELECT
  c.Category AS vision,
  sni_codes.label AS sector,
  count(bv.bioeconomy_vision) AS count
FROM
  bioeconomy_visions AS bv
  join use_sectors ON bv.sinno_id = use_sectors.[sinno_id]
  join sni_codes ON SUBSTR(use_sectors.use_sector, 1, 2) = sni_codes.code
  join [classification_codes] AS c ON bv.bioeconomy_vision = c.Code
WHERE
  NOT EXISTS(
    SELECT
      *
    FROM
      bioeconomy_visions AS bv2
    WHERE
      bv2.sinno_id = bv.sinno_id
      AND bv2.bioeconomy_vision = 0
  )
  AND bv.sinno_id NOT IN (
    SELECT
      sinno_id
    FROM
      categorization_notes
    WHERE
      notes NOT LIKE "%not forest%"
  )
GROUP BY
  sector,
  bioeconomy_vision;
""",
    swinno_db,
)

bv_matrix = prepare_heatmap_data(
    bv_sectors, index_col="sector", cols_col="vision", values_col="count"
)

plot_heatmap(bv_matrix, "Count Bioeconomy Vision by Sector Excluding Uncertains")
```

## Examples

<!---7495001 	good example of both bv2 and bv3 because of high integration of production system and conversion and upgrading of resource

7428001 	good example for innovation which could go either way

9048001 	good bv 2 example, good example of bv1

--->

## Uncertain Cases

```{python}
uncertains = pd.read_sql(
    """
SELECT
DISTINCT
  i.sinno_id,
  i.innovation_name_in_swedish AS name,
  i.description_in_swedish AS description,
  i.additional_information_if_origin_new_scientific_discovery || i.additional_information_if_origin_new_technologies_or_materials || i.additional_info_if_origin_official_regulation_legislation_and_standards || i.additional_information_if_origin_solution_for_a_problem || i.additional_information_if_origin_performance || i.additional_information_if_origin_other AS info,
  i.year_of_commercialization AS year
FROM
  innovation i
  JOIN use_sectors us ON i.sinno_id = us.sinno_id
  JOIN bioeconomy_visions bv on i.sinno_id = bv.sinno_id
WHERE i.sinno_id in (select distinct sinno_id from bioeconomy_visions where bioeconomy_vision = "0");
""",
    swinno_db,
)

display(Markdown(f"{uncertain["sinno_id"].nunique()} cases are marked as unsure bioeconomy visions."))
```

### 2023-09-25 docs/presentations/20230928-sehm_presentation.qmd
add presentations

---
title: Shades of Green
subtitle: Direction of the Swedish Forest Innovation System 1970-2020
author: Jonas Kreutzer
date: last-modified
abstract: |

csl: https://zotero.org/styles/apa?source=1
bibliography: [../../swinno-shades-green.bib]

format:
  revealjs:
    theme: [../assets/revealme.scss]
    transition: none
    title-slide-attributes:
      data-background-image: '../assets/ground-fog-valley-black-forest-sunset.jpg'
      data-background-size: fit
      data-background-opacity: '0.3'
    embed-resources: true

execute:
    freeze: auto
    echo: false
jupyter: python3
---

```{python}
# | label: Imports

import pandas as pd
import numpy as np
from IPython.display import display, Markdown
import matplotlib.pyplot as plt
import matplotlib.ticker as mtick
import seaborn as sns

from src.figures import *
from src.utils import get_project_root, format_table
from src.swinno_helpers import connect_swinno_db

ROOT = get_project_root()
conn = connect_swinno_db()

# set all text on plots to be monospace with rcParams
plt.rcParams["font.family"] = "Space Grotesk"

sns.set_theme(style="white")

# full code labels for merging

codes_df = pd.read_sql("SELECT * FROM classification_codes", conn)

codes_df["Category"] = codes_df["Category"].str.strip()
uncertain_eco = {"Code": "000", "Category": "Unsure"}
codes_df = codes_df.append(uncertain_eco, ignore_index=True)

codes_df = codes_df[~codes_df["Code"].duplicated(keep="last")]

codes_df = codes_df.set_index("Code")
```

```{python}
# | label: Query bioeconomy

bioeconomy = pd.read_sql(
    """
select
  i.sinno_id,
  i.innovation_name_in_swedish AS name,
  i.description_in_swedish AS description,
  i.additional_information_if_origin_new_scientific_discovery || i.additional_information_if_origin_new_technologies_or_materials || i.additional_info_if_origin_official_regulation_legislation_and_standards || i.additional_information_if_origin_solution_for_a_problem || i.additional_information_if_origin_performance || i.additional_information_if_origin_other AS info,
  i.year_of_commercialization AS year,
  us.use_sector
from
  innovation i
  join use_sectors us on i.sinno_id = us.sinno_id
where
  (
    us.use_sector like '02%'
    or us.use_sector like '20%'
    or us.use_sector like '21%'
    or us.use_sector like '36%'
    or product_code like '02%'
    or product_code like '20%'
    or product_code like '21%'
    or product_code like '36%'
  )
  or (
    description like '%virke%'
    or description like '%cellulos%'
    or description like '%lignin%'
    or description like '%spån%'
    or description like '%bark%'
    or description like '%levulinsyra%'
    or description like '%furfural%'
    or description like '%svarttjära%'
    or description like '%svartlut%'
    or description like '%växtbas%'
    or description like '%ved%'
    or description like '%trä%'
    or description like '%skog%'
    or description like '%biobränsle%'
    or description like '%biologisk%'
    or description like '%nedbrytbar%'
    or description like '%papper%'
    or description like '%pappret%'
    or description like '%karton%'
    or description like '%tencel%'
  );
""",
    conn,
)

sni_codes = pd.read_sql_query(
    """
select
*
from 
sni_codes
""",
    conn,
)

sni_codes = sni_codes.rename(columns={"code": "use_sector"})

swinno = pd.read_sql_query(
    """
select sinno_id, year_of_commercialization as year, innovation_name_in_swedish as name
from innovation;
""",
    conn,
)

swinno["bioeconomy"] = swinno["sinno_id"].isin(bioeconomy["sinno_id"].unique())

```

```{python}
# | label: Query eco-innovation

eco_innovations = pd.read_sql(
    """
SELECT
  sinno_id,
  innovation_type as innovation_type_code,
  c.Category as innovation_type
FROM
  eco_innovations
JOIN classification_codes AS c ON c.Code = eco_innovations.innovation_type
WHERE 
innovation_type NOT IN (701, 602, 601, 501, 111, 109, 108)
;
""",
    conn,
)

eco_innovations["innovation_type"] = eco_innovations["innovation_type"].str.strip()

```

```{python}
# | label: Calculate Bioeconomy Share

def calculate_bioeconomy_share(swinno):
    df = swinno.copy()
    df = df[["sinno_id", "year", "bioeconomy"]]
    df = (
        df.groupby("year")
        .agg(
            total_count=("sinno_id", "nunique"), bioeconomy_count=("bioeconomy", "sum")
        )
        .reset_index()
    )
    df["bioeconomy_share"] = df["bioeconomy_count"] / df["total_count"]

    return df

bioeconomy_share = calculate_bioeconomy_share(swinno)
```

## Bioeconomy Demarkations {.smaller}

::::{.columns}

:::{.column width="50%"}
**Using producer / user codes
**

|SNI Code | Description                            |
|--------:|:---------------------------------------|
|02       | Forestry and related services
|20       | Wood and wood product manufacturing except furniture
|21       | Pulp, paper and paper product manufacturing
|36       | Furniture manufacturing; other manufacturing
:::

:::{.column width="50%"}
**Adding keywords of value chain**

    - "virke",
    - "cellulos",
    - "lignin",
    - "spån", 
    - "bark",
    - "levulinsyra" (Levulinic acid),
    - "furfural" (Furfural),
    - "svarttjära",
    - "svartlut",
    - "växtbas",
    - "ved",
    - "trä",
    - "skog",
    - "papper",
    - "biobränsle",
    - "biologiskt",
    - "nedbrytbar",
    - "papper",
    - "pappret",
    - "karton",
    - "tencel",
:::

::::

****

## 817 Innovations in Forest Bioeconomy (Uncleaned)

## Count of Bioeconomy Innovation
### [Yearly]{.text-royal-blue-800} and [5 Year MA]{.text-gray-600}
```{python}
# label: fig-bioeconomy_count
#
#

plot_bioeconomy_trends_interactive(
    bioeconomy_share, y="bioeconomy_count"
).update_layout(yaxis_range=[0, 35]).show()

```

## Share of Bioeconomy Innovation{.smaller}
### [Yearly]{.text-royal-blue-800} and [5 Year MA]{.text-gray-600} 
```{python}

plot_bioeconomy_trends_interactive(bioeconomy_share, y="bioeconomy_share").update_yaxes(
    range=[0, 0.35], tickformat=".0%"
).show()

```

# Three Visions for a Bioeconomy

:::: columns
::: {.column width="33%"}
![Bio-Technology Vision](../assets/Biotech.jpg){.fragment}
:::

::: {.column width="33%"}
![Bio-Resource Vision](../assets/Bioresource.jpg){.fragment}
:::

::: {.column width="33%"}
![Bio-Ecology Vision](../assets/Bioecology.jpg){.fragment}
:::
::::

[@bugge2016WhatBioeconomyReview]{.slide-cite style="font-size: x-small"}

## Bioeconomy Vision Classification{.smaller .scrollable}

| @bugge2016WhatBioeconomyReview | Biotechnology | Bioresource | Bioecology |
|---|---|---|---|
| Aim &   Objectives | Economic growth & job   creation | Economic growth &   sustainability | Sustainability, biodiversity,   conservation of ecosystems, avoiding soil degradation |
| Value   Creation | Application of biotechnology,   commercialisation of research & technology | Conversion and upgrading of   bio-resources (process oriented) | Development of integrated   production systems and high-quality products with territorial identity |
| Drivers   & mediators of innovation | R & D, patents, TTOs,   Research councils and funders (Science push, linear model) | Interdisciplinary, optimisation   of land use, include degraded land in the production of biofuels, use and   availability of bio-resources, waste management, engineering, science &   market (Interactive & networked production mode) | Identification of favourable   organic agro-ecological practices, ethics, risk, transdisciplinary   sustainability, ecological interactions, re-use & recycling of waste,   land use, (Circular and self-sustained production mode) |

| @vivien2019HijackingBioeconomy | Type II | Type III | Type I |
|---|---|---|---|
|  | A science-based economy driven   by industrial biotechnology<br>     <br>     The cell is a factory | Biomass replaces fossil fuels   and mining to produce energy and materials<br>     <br>     Biorefining at the heart of ecological transition (multilevel perspective). | An ecological economy, that is   compatible with the biosphere <br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
| Codes | 1 | 2 | 3 |

## Preliminary Results

```{python}
bv_df = pd.read_sql(
    """SELECT bv.sinno_id, bv.bioeconomy_vision, i.year_of_commercialization AS "i.year"
FROM bioeconomy_visions AS bv
JOIN innovation AS i ON i.sinno_id = bv.sinno_id
WHERE bv.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
OR bv.sinno_id NOT IN (SELECT sinno_id FROM bioeconomy_visions WHERE bioeconomy_vision <> 0);
""",
    conn,
)
```

::::{.columns}
:::{.column width='80%'}
```{python}

display(
    Markdown(
        f"<span class='text-primary' style='font-size:15rem; font-weight:900'>{bv_df.loc[:, 'sinno_id'].nunique()}</span>"
    )
)

```

:::

:::{.column width='20%'}
unique forest bioeconomy innovations classified
:::

::::

## Examples

## Bioeconomy Vision by Year

```{python}

bv_year = pd.read_sql(
    """
SELECT
  codes.Category AS Vision,
  COUNT(bioeconomy_vision) AS Count,
  i.year_of_commercialization AS Year
FROM
  bioeconomy_visions
  JOIN classification_codes AS codes ON codes.code = bioeconomy_visions.bioeconomy_vision
  JOIN innovation AS i on i.sinno_id = bioeconomy_visions.sinno_id
WHERE
  NOT EXISTS (
    SELECT
      *
    FROM
      bioeconomy_visions AS bv2
    WHERE
      bv2.sinno_id = bioeconomy_visions.sinno_id
      AND bv2.bioeconomy_vision = 0
  )
  AND bioeconomy_visions.sinno_id NOT IN (
    SELECT
      sinno_id
    FROM
      categorization_notes
    WHERE
      notes NOT LIKE "%not forest%"
  )
GROUP BY
Year, Vision
ORDER BY
  Year ASC,
  Count DESC;
  """,
    conn,
)

def plot_bv_trends(df, title):
    sns.set_style("white")
    chart = sns.lineplot(
        data=df,
        x="Year",
        y="Count",
        hue="Vision",
        palette="colorblind",
        linewidth=2.5,
    )
    chart.set_title(title, fontdict={"fontsize": 16, "fontweight": "bold"})
    chart.set_xlabel("Year", fontdict={"fontsize": 12, "fontweight": "bold"})
    chart.set_ylabel("Count", fontdict={"fontsize": 12, "fontweight": "bold"})
    chart.tick_params(labelsize=10)
    plt.legend(title="Vision", fontsize=10, title_fontsize=12)
    plt.tight_layout()

    # Customize x-axis tick labels
    years = df["Year"].unique()
    chart.set_xticks(years[::5])
    chart.set_xticklabels(years[::5], rotation=45, ha="right")

    plt.show(chart)

plot_bv_trends(bv_year, title="")

```

## Bioeconomy Vision Count by Sector Excluding Uncertains {.smaller}
```{python}
bv_sectors = pd.read_sql(
    """
SELECT
  c.Category AS vision,
  sni_codes.label AS sector,
  count(bv.bioeconomy_vision) AS count
FROM
  bioeconomy_visions AS bv
  join use_sectors ON bv.sinno_id = use_sectors.[sinno_id]
  join sni_codes ON SUBSTR(use_sectors.use_sector, 1, 2) = sni_codes.code
  join [classification_codes] AS c ON bv.bioeconomy_vision = c.Code
WHERE
  NOT EXISTS(
    SELECT
      *
    FROM
      bioeconomy_visions AS bv2
    WHERE
      bv2.sinno_id = bv.sinno_id
      AND bv2.bioeconomy_vision = 0
  )
  AND bv.sinno_id NOT IN (
    SELECT
      sinno_id
    FROM
      categorization_notes
    WHERE
      notes NOT LIKE "%not forest%"
  )
GROUP BY
  sector,
  bioeconomy_vision;
""",
    conn,
)

bv_matrix = prepare_heatmap_data(
    bv_sectors, index_col="sector", cols_col="vision", values_col="count"
)

plot_heatmap(bv_matrix)
```

## Uncertain Cases

```{python}
uncertains = pd.read_sql(
    """
SELECT
DISTINCT
  i.sinno_id,
  i.innovation_name_in_swedish AS name,
  i.description_in_swedish AS description,
  i.additional_information_if_origin_new_scientific_discovery || i.additional_information_if_origin_new_technologies_or_materials || i.additional_info_if_origin_official_regulation_legislation_and_standards || i.additional_information_if_origin_solution_for_a_problem || i.additional_information_if_origin_performance || i.additional_information_if_origin_other AS info,
  i.year_of_commercialization AS year
FROM
  innovation i
  JOIN use_sectors us ON i.sinno_id = us.sinno_id
  JOIN bioeconomy_visions bv on i.sinno_id = bv.sinno_id
WHERE i.sinno_id in (select distinct sinno_id from bioeconomy_visions where bioeconomy_vision = "0");
""",
    conn,
)

display(
    Markdown(
        f"{uncertains['sinno_id'].nunique()} cases are marked as unsure bioeconomy visions."
    )
)
```

## References

### 2023-09-25 docs/presentations/presentation.qmd
add presentations

--- 2023-09-25 docs/presentations/20230829-Umea_workshop.qmd
@@
 
-format: revealjs
-# format:
-#   revealjs:
-#     theme: [../assets/revealme.scss]
-#     transition: none
-#     title-slide-attributes:
-#       data-background-image: '../assets/ground-fog-valley-black-forest-sunset.jpg'
-#       data-background-size: fit
-#       data-background-opacity: '0.3'
-#     embed-resources: true
+format:
+  revealjs:
+    theme: [../assets/revealme.scss]
+    transition: none
+    title-slide-attributes:
+      data-background-image: '../assets/ground-fog-valley-black-forest-sunset.jpg'
+      data-background-size: fit
+      data-background-opacity: '0.3'
+    embed-resources: true
 
@@
 import altair as alt
+import plotly.graph_objs as go
 
@@
 
+# static
+# def plot_heatmap(data, title):
+#     fig, ax = plt.subplots(1, 1, figsize=(10, 8))
+#     ax = sns.heatmap(
+#         data, annot=True, cmap="Blues", fmt="g", cbar=False, annot_kws={"fontsize": 14}
+#     )
+#     ax.set_xticklabels(ax.get_xticklabels(), rotation=45, ha="right")
+
+#     plt.show()
+
 def plot_heatmap(data, title):
-    fig, ax = plt.subplots(1, 1)
-    ax = sns.heatmap(
-        data, annot=True, cmap="Blues", fmt="g", cbar=False, annot_kws={"fontsize": 10}
+    fig = go.Figure(
+        go.Heatmap(
+            z=data.values,
+            x=data.columns,
+            y=data.index,
+            colorscale="Blues",
+            showscale=False,
+            hovertemplate="%{x}<br>%{y}<br>%{z}",  # Customize hover text
         )
-    ax.set(title=title, ylabel=None, xlabel=None)
-    ax.set_xticklabels(ax.get_xticklabels(), rotation=45, ha="right")
+    )
 
-    plt.show()
+    fig.update_layout(
+        template="plotly_white",
+        title=title,
+        font=dict(family="Inconsolata"),
+        width=900,
+        xaxis=dict(tickangle=-45),
+    )
+
+    fig.update_xaxes(side="top")
+    fig.show()
 
@@
 
-## First Estimation of Forest Bioeconomy Innovativeness
+****
 
@@
 
+# display(
+#     Markdown(
+#         f"## {swinno['bioeconomy'].sum()} Innovations in Forest Bioeconomy (Uncleaned)"
+#     )
+# )
+
+```
+
+## 817 Innovations in Forest Bioeconomy (Uncleaned)
+
+```{python}
 def plot_innovation_counts_by_sector(df, title):
@@
         .encode(
-            alt.Y("bioeconomy_count:Q", title="Count of Bioeconomy Innovations"),
+            alt.Y(
+                "bioeconomy_count:Q",
+                axis=alt.Axis(
+                    title="Count of Bioeconomy Innovations", titleColor="red"
+                ),
+            ),
             tooltip=["bioeconomy_count", "bioeconomy_share", "year"],
@@
         .encode(
-            alt.Y("bioeconomy_share:Q", title="Share of Bioeconomy Innovations"),
+            alt.Y(
+                "bioeconomy_share:Q",
+                axis=alt.Axis(
+                    title="Share of Bioeconomy Innovations", titleColor="blue"
+                ),
+            ),
             tooltip=["bioeconomy_count", "bioeconomy_share", "year"],
@@
 
-## Bioeconomy Vision Classification{.smaller}
+## Bioeconomy Vision Classification{.smaller .scrollable}
 
@@
 |  | A science-based economy driven   by industrial biotechnology<br>     <br>     The cell is a factory | Biomass replaces fossil fuels   and mining to produce energy and materials<br>     <br>     Biorefining at the heart of ecological transition (multilevel perspective).
-| An ecological economy, that is   compatible with the biosphere<br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
+| An ecological economy, that is   compatible with the biosphere <br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
 | Codes | 1 | 2 | 3 |
@@
 JOIN innovation AS i ON i.sinno_id = bv.sinno_id
-WHERE bv.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%");
+WHERE bv.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
+OR bv.sinno_id NOT IN (SELECT sinno_id FROM bioeconomy_visions WHERE bioeconomy_vision <> 0);
 """,
@@
 
+::::{.columns}
+:::{.column width='80%'}
 ```{python}
@@
     Markdown(
-        f"<span style='font-size:10rem'>{bv_df.loc[:, 'sinno_id'].nunique()}</span> unique forest bioeconomy innovations classified"
+        f"<span class='text-primary' style='font-size:15rem; font-weight:900'>{bv_df.loc[:, 'sinno_id'].nunique()}</span>"
     )
@@
 
-## More
+:::
+
+:::{.column width='20%'}
+unique forest bioeconomy innovations classified
+:::
+
+::::
+
+## Examples
+
+## Bioeconomy Vision by Year
 
 ```{python}
-ei_count = pd.read_sql(
-    """
-SELECT
-  codes.Category AS Type,
-  COUNT(ei.innovation_type) AS Count
-FROM
-  eco_innovations as ei
-  JOIN [classification_codes] AS codes ON codes.code = ei.innovation_type
-WHERE ei.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
-GROUP BY
-  innovation_type
-ORDER BY
-  Count DESC;
-""",
-    swinno_db,
-)
 
-ei_count_0 = pd.read_sql(
-    """
-SELECT
-  codes.Category AS Type,
-  COUNT(ei.innovation_type) AS Count
-FROM
-  eco_innovations as ei
-  JOIN [classification_codes] AS codes ON codes.code = ei.innovation_type
-WHERE
-  NOT EXISTS(
-    SELECT * 
-    FROM
-      eco_innovations AS ei2
-    WHERE
-      ei2.sinno_id = ei.sinno_id
-      AND ei2.innovation_type = "000"
-  )
-AND ei.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
-GROUP BY
-  Type
-ORDER BY
-  Count DESC;
-""",
-    swinno_db,
-)
-
-bv_count = pd.read_sql(
+bv_year = pd.read_sql(
     """
@@
   codes.Category AS Vision,
-  COUNT(bioeconomy_vision) AS Count
+  COUNT(bioeconomy_vision) AS Count,
+  i.year_of_commercialization AS Year
 FROM
   bioeconomy_visions
-  JOIN [classification_codes] AS codes ON codes.code = bioeconomy_visions.bioeconomy_vision
-WHERE
-  bioeconomy_visions.sinno_id NOT IN (
-    SELECT
-      sinno_id
-    FROM
-      categorization_notes
-    WHERE
-      notes NOT LIKE "%not forest%"
-  )
-GROUP BY
-  Vision
-ORDER BY
-  Count DESC;
-   """,
-    swinno_db,
-)
-
-bv_count_zero = pd.read_sql(
-    """
-SELECT
-  codes.Category AS Vision,
-  COUNT(bioeconomy_vision) AS Count
-FROM
-  bioeconomy_visions
-  JOIN [classification_codes] AS codes ON codes.code = bioeconomy_visions.bioeconomy_vision
+  JOIN classification_codes AS codes ON codes.code = bioeconomy_visions.bioeconomy_vision
+  JOIN innovation AS i on i.sinno_id = bioeconomy_visions.sinno_id
 WHERE
@@
 GROUP BY
-  Vision
+Year, Vision
 ORDER BY
+  Year ASC,
   Count DESC;
@@
 )
-```
 
-## Count of Bioeconomy Visions
-```{python}
-bv_counts = pd.merge(bv_count, bv_count_zero, on="Vision", how="left")
+def plot_bv_trends(df, title):
+    sns.set_style("white")
+    chart = sns.lineplot(
+        data=df,
+        x="Year",
+        y="Count",
+        hue="Vision",
+        palette="colorblind",
+        linewidth=2.5,
+    )
+    chart.set_title(title, fontdict={"fontsize": 16, "fontweight": "bold"})
+    chart.set_xlabel("Year", fontdict={"fontsize": 12, "fontweight": "bold"})
+    chart.set_ylabel("Count", fontdict={"fontsize": 12, "fontweight": "bold"})
+    chart.tick_params(labelsize=10)
+    plt.legend(title="Vision", fontsize=10, title_fontsize=12)
+    plt.tight_layout()
 
-bv_counts.columns = ["Vision", "Count", "Count Excl. Unsure"]
-bv_counts.loc[:, "Count Excl. Unsure"] = (
-    bv_counts["Count Excl. Unsure"].fillna(0).astype(int)
-)
+    # Customize x-axis tick labels
+    years = df["Year"].unique()
+    chart.set_xticks(years[::5])
+    chart.set_xticklabels(years[::5], rotation=45, ha="right")
 
-bv_counts.style.set_properties(**{"text-align": "right"}).hide_index()
+    plt.show(chart)
+
+plot_bv_trends(bv_year, title="")
 
@@
 
-## Count of Eco-Innovations
-
+## Bioeconomy Vision Count by Sector Excluding Uncertains {.smaller}
 ```{python}
-# | label: tbl-ei-article-count
-# | tbl-cap: Count of Eco-Innovations Using Swinno Data
-
-ei_counts = pd.merge(ei_count, ei_count_0, on="Type", how="left")
-
-ei_counts.columns = ["Type", "Count", "Count Excl. Unsure"]
-ei_counts.loc[:, "Count Excl. Unsure"] = (
-    ei_counts["Count Excl. Unsure"].fillna(0).astype(int)
-)
-
-ei_counts.style.set_properties(**{"text-align": "right"}).hide_index()
-```
-
-## Bioeconomy Vision Heatmap
-
-```{python}
-# | label: fig-bva-sectors
-# | fig-cap: Heatmap of Bioeconomy Visions by Sector
-
 bv_sectors = pd.read_sql(
@@
 
-plot_heatmap(bv_matrix, "Count Bioeconomy Vision by Sector Excluding Uncertains")
+plot_heatmap(bv_matrix, "")
 ```
-
-## Examples
-
-<!---7495001 	good example of both bv2 and bv3 because of high integration of production system and conversion and upgrading of resource
-
-7428001 	good example for innovation which could go either way
-
-9048001 	good bv 2 example, good example of bv1
-
---->
 
@@
 
-display(Markdown(f"{uncertain["sinno_id"].nunique()} cases are marked as unsure bioeconomy visions."))
+display(
+    Markdown(
+        f"{uncertains['sinno_id'].nunique()} cases are marked as unsure bioeconomy visions."
+    )
+)
 ```
+
+## References

### 2023-09-25 docs/presentations/20230928-sehm_presentation.qmd
update interactive plots of bioeconomy trends

@@
 ## Count of Bioeconomy Innovation
-### [Yearly]{.text-royal-blue-800} and [5 Year MA]{.text-gray-600}
+### [5 Year MA]{.text-royal-blue-800} and [Yearly]{.text-gray-600}
 ```{python}
@@
 ## Share of Bioeconomy Innovation{.smaller}
-### [Yearly]{.text-royal-blue-800} and [5 Year MA]{.text-gray-600} 
+### [5 Year MA]{.text-royal-blue-800} and [Yearly]{.text-gray-600}
 ```{python}

### 2023-09-25 src/figures.py

@@ -1,8 +1,103 @@
 import pandas as pd
 import matplotlib.pyplot as plt
 import seaborn as sns
-import plotly.graph_objects as po
+import plotly.graph_objects as go
 import plotly.express as px
+import plotly.io as pio
+
+# Set global options for all Plotly plots
+pio.templates.default = "presentation"
+
+royal_blue_800 = "#283aa4"
+gray_600 = "#878f97"
+
+def plot_bioeconomy_trends_interactive(data, y):
+    fig = go.Figure()
+
+    # Add yearly data to plot
+    fig.add_trace(
+        go.Scatter(
+            x=data["year"],
+            y=data[y],
+            mode="lines",
+            name="Yearly",
+            line=dict(dash="dash", color=gray_600),
+        )
+    )
+
+    # Add rolling mean data to plot
+    rolling_mean = data[y].rolling(window=5).mean()
+    fig.add_trace(
+        go.Scatter(
+            x=data["year"],
+            y=rolling_mean,
+            mode="lines",
+            name="5-year moving average",
+            line=dict(color=royal_blue_800),
+        )
+    )
+
+    fig.update_xaxes(type="date")
+    fig.update_layout(
+        showlegend=False,
+        xaxis_range=["1965", "2025"],
+        font=dict(family="Atkinson Hyperlegible"),
+        height=550,
+        hovermode="x unified",
+    )
+
+    # Show plot
+    return fig
+
+
+def plot_count(data):
+    fig, ax = plt.subplots()
+    ax.plot(
+        data["year"],
+        data["bioeconomy_count"],
+        label="Yearly",
+    )
+
+        # Calculate rolling mean and plot on same axis
+    rolling_mean = data["bioeconomy_count"].rolling(window=5).mean()
+    ax.plot(
+        data["year"],
+        rolling_mean,
+        label="5-year moving average",
+        linestyle="dashed"
+    )
+
+    ax.set_xticks(data["year"][::10])
+    ax.tick_params(labelsize=10)
+
+    ax.spines["top"].set_visible(False)
+    ax.spines["right"].set_visible(False)
+    ax.legend(frameon=False)
+
+    plt.show()
+
+
+def plot_share(data):
+    fig, ax = plt.subplots()
+    ax.plot(data["year"], data["bioeconomy_share"], label="Yearly")
+    ax.set_xticks(data["year"][::10])
[…]

### 2023-09-25 docs/reports/paper/paper1.qmd
fix remove repeated def of  bioeco share plot

@@
 
-def plot_count(data):
-    fig, ax = plt.subplots()
-    ax.plot(
-        data["year"],
-        data["bioeconomy_count"],
-        label="Yearly",
-    )
-
-        # Calculate rolling mean and plot on same axis
-    rolling_mean = data["bioeconomy_count"].rolling(window=5).mean()
-    ax.plot(
-        data["year"],
-        rolling_mean,
-        label="5-year moving average",
-        linestyle="dashed"
-    )
-
-    ax.set_xticks(data["year"][::10])
-    ax.tick_params(labelsize=10)
-
-    ax.spines["top"].set_visible(False)
-    ax.spines["right"].set_visible(False)
-    ax.legend(frameon=False)
-
-    plt.show()
-
-def plot_share(data):
-    fig, ax = plt.subplots()
-    ax.plot(data["year"], data["bioeconomy_share"], label="Yearly")
-    ax.set_xticks(data["year"][::10])
-    ax.yaxis.set_major_formatter(mtick.PercentFormatter(xmax=1.0))
-    ax.tick_params(which="major", labelsize=10)
-
-    # Calculate rolling mean and plot on same axis
-    rolling_mean = data["bioeconomy_share"].rolling(window=5).mean()
-    ax.plot(
-        data["year"],
-        rolling_mean,
-        label="5-year moving average",
-        linestyle="dashed"
-    )
-    ax.set_ylim((0, .5))
-    ax.legend(frameon=False)
-    ax.spines["top"].set_visible(False)
-    ax.spines["right"].set_visible(False)
-    plt.show()
-
 plot_count(bioeconomy_share)

### 2023-09-25 docs/reports/paper/paper1.qmd
update sum_bioeconomy name in paper1.qmd

@@
 
-codes_df = pd.read_sql(
-    "SELECT * FROM classification_codes", conn
-)
+codes_df = pd.read_sql("SELECT * FROM classification_codes", conn)
 
 codes_df["Category"] = codes_df["Category"].str.strip()
-uncertain_eco = {"Code":"000", "Category":"Unsure"}
+uncertain_eco = {"Code": "000", "Category": "Unsure"}
 codes_df = codes_df.append(uncertain_eco, ignore_index=True)
 
-codes_df = codes_df[~codes_df["Code"].duplicated(keep='last')
-]
+codes_df = codes_df[~codes_df["Code"].duplicated(keep="last")]
 
-codes_df =codes_df.set_index("Code")
+codes_df = codes_df.set_index("Code")
 ```
@@
 ```{python}
-#| label: Total SWINNO Innovation
-#| output: asis
+# | label: Total SWINNO Innovation
+# | output: asis
 
@@
 
-display(Markdown(
+display(
+    Markdown(
         f"The SWINNO database used here contains {n_swinno} innovations commercialized between {year_min} and {year_max}.
@@
 This method relies on the judgment of editiors and journalists at trade journals and hence captures what can be described as significant innovation."
-))
+    )
+)
 ```
@@
 ```{python}
-#| label: Number Bioeconomy Innovation
-#| output: asis
+# | label: Number Bioeconomy Innovation
+# | output: asis
 
-sum_bioeconomy_visions = swinno["bioeconomy"].sum()
+sum_bioeconomy_innovations = swinno["bioeconomy"].sum()
 
@@
     Markdown(
-        f"Querying the SWINNO database for forest based bioeconomy innovations initally yields {sum_bioeconomy_visions} innovations.
+        f"Querying the SWINNO database for forest based bioeconomy innovations initally yields {sum_bioeconomy_innovations} innovations.
 This number includes false positives from substrings matching keywords included in the search.
@@
 ```{python}
-#| label: Query eco-innovation
+# | label: Query eco-innovation
 
@@
 ```{python}
-#| label: Calculate Bioeconomy Share
+# | label: Calculate Bioeconomy Share
 
@@
 count_certain_bioeconomy_visions = (
-    certain_bioeconomy_visions
-    .groupby(["year", "bioeconomy_vision"])
+    certain_bioeconomy_visions.groupby(["year", "bioeconomy_vision"])
     .size()
@@
 
-total_certain_bioeconomy_visions = (count_certain_bioeconomy_visions
-        .groupby("bioeconomy_vision")
+total_certain_bioeconomy_visions = (
+    count_certain_bioeconomy_visions.groupby("bioeconomy_vision")
     .agg("sum", numeric_only=True)
@@
 ```{python}
-#| label: Number of Uncertains
-#| output: asis
+# | label: Number of Uncertains
+# | output: asis
 
@@
 
-n_certain_bioeconomy_visions = certain_bioeconomy_visions['sinno_id'].nunique()
+n_certain_bioeconomy_visions = certain_bioeconomy_visions["sinno_id"].nunique()
 
-certain_bioeconomy_visions_min_year = certain_bioeconomy_visions['year'].min()
+certain_bioeconomy_visions_min_year = certain_bioeconomy_visions["year"].min()
 
-certain_bioeconomy_visions_max_year = certain_bioeconomy_visions['year'].max()
+certain_bioeconomy_visions_max_year = certain_bioeconomy_visions["year"].max()
 
@@
 Note that the sum by category will not yield the total number of classified innovations, as one innovation can be associated with more than one bioeconomy vision."
-    ))
+    )
+)
 
@@
 
-n_bioecology_vision = count_certain_bioeconomy_visions.loc[count_certain_bioeconomy_visions["bioeconomy_vision"] == "Bioecology Vision", "count"].sum()
+n_bioecology_vision = count_certain_bioeconomy_visions.loc[
+    count_certain_bioeconomy_visions["bioeconomy_vision"] == "Bioecology Vision",
+    "count",
+].sum()
 
@@
 ```{python}
-#| label: fig-eco_innovation_by_bioeconomy_vision
-#| fig-cap: Count of Eco-Innovation Types by Bioeconomy Vision
-p = plot_eco_innovation_by_bioeconomy_vision(eco_innovation_by_bioeconomy_visions, codes_df)
+# | label: fig-eco_innovation_by_bioeconomy_vision
+# | fig-cap: Count of Eco-Innovation Types by Bioeconomy Vision
+p = plot_eco_innovation_by_bioeconomy_vision(
+    eco_innovation_by_bioeconomy_visions, codes_df
+)
 

### 2023-09-25 docs/presentations/20230928-sehm_presentation.qmd
update bioeconomy stats to dynamic in sehm_pres

@@
 
-## 817 Innovations in Forest Bioeconomy (Uncleaned)
+```{python}
+# | label: Sum bioeconomy visions
+# | output: asis
+
+sum_bioeconomy_innovations = swinno["bioeconomy"].sum()
+
+display(
+    Markdown(
+        f"# {sum_bioeconomy_innovations} Innovations in Forest Bioeconomy (Uncleaned){{.inverted}}"
+    )
+)
+
+```
 
@@
 | An ecological economy, that is   compatible with the biosphere <br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
-| Codes | 1 | 2 | 3 |
 

### 2023-09-25 docs/presentations/20230928-sehm_presentation.qmd
remove old bioeconomy certain query

@@
 
-bv_year = pd.read_sql(
-    """
-SELECT
-  codes.Category AS Vision,
-  COUNT(bioeconomy_vision) AS Count,
-  i.year_of_commercialization AS Year
-FROM
-  bioeconomy_visions
-  JOIN classification_codes AS codes ON codes.code = bioeconomy_visions.bioeconomy_vision
-  JOIN innovation AS i on i.sinno_id = bioeconomy_visions.sinno_id
-WHERE
-  NOT EXISTS (
-    SELECT
-      *
-    FROM
-      bioeconomy_visions AS bv2
-    WHERE
-      bv2.sinno_id = bioeconomy_visions.sinno_id
-      AND bv2.bioeconomy_vision = 0
-  )
-  AND bioeconomy_visions.sinno_id NOT IN (
-    SELECT
-      sinno_id
-    FROM
-      categorization_notes
-    WHERE
-      notes NOT LIKE "%not forest%"
-  )
-GROUP BY
-Year, Vision
-ORDER BY
-  Year ASC,
-  Count DESC;
-  """,
-    conn,
-)
-
-def plot_bv_trends(df, title):
-    sns.set_style("white")
-    chart = sns.lineplot(
-        data=df,
-        x="Year",
-        y="Count",
-        hue="Vision",
-        palette="colorblind",
-        linewidth=2.5,
-    )
-    chart.set_title(title, fontdict={"fontsize": 16, "fontweight": "bold"})
-    chart.set_xlabel("Year", fontdict={"fontsize": 12, "fontweight": "bold"})
-    chart.set_ylabel("Count", fontdict={"fontsize": 12, "fontweight": "bold"})
-    chart.tick_params(labelsize=10)
-    plt.legend(title="Vision", fontsize=10, title_fontsize=12)
-    plt.tight_layout()
-
-    # Customize x-axis tick labels
-    years = df["Year"].unique()
-    chart.set_xticks(years[::5])
-    chart.set_xticklabels(years[::5], rotation=45, ha="right")
-
-    plt.show(chart)
-
 plot_bv_trends(bv_year, title="")
@@
 
-```{python}
-uncertains = pd.read_sql(
-    """
-SELECT
-DISTINCT
-  i.sinno_id,
-  i.innovation_name_in_swedish AS name,
-  i.description_in_swedish AS description,
-  i.additional_information_if_origin_new_scientific_discovery || i.additional_information_if_origin_new_technologies_or_materials || i.additional_info_if_origin_official_regulation_legislation_and_standards || i.additional_information_if_origin_solution_for_a_problem || i.additional_information_if_origin_performance || i.additional_information_if_origin_other AS info,
-  i.year_of_commercialization AS year
-FROM
-  innovation i
-  JOIN use_sectors us ON i.sinno_id = us.sinno_id
-  JOIN bioeconomy_visions bv on i.sinno_id = bv.sinno_id
-WHERE i.sinno_id in (select distinct sinno_id from bioeconomy_visions where bioeconomy_vision = "0");
-""",
-    conn,
-)
-
-display(
-    Markdown(
-        f"{uncertains['sinno_id'].nunique()} cases are marked as unsure bioeconomy visions."
-    )
-)
-```
-
 ## References

### 2023-09-25 docs/presentations/20230928-sehm_presentation.qmd
add bv trends fig in sehm_pres

@@
 from IPython.display import display, Markdown
-import matplotlib.pyplot as plt
-import matplotlib.ticker as mtick
-import seaborn as sns
 
@@
 eco_innovations["innovation_type"] = eco_innovations["innovation_type"].str.strip()
+
+```
+
+```{python}
+# | label: Query bioeconomy visions
+
+bioeconomy_visions = pd.read_sql(
+    """
+SELECT bv.sinno_id
+, codes.Category AS bioeconomy_vision
+, i.year_of_commercialization AS "year"
+FROM bioeconomy_visions AS bv
+JOIN [classification_codes] AS codes ON codes.code = bv.bioeconomy_vision
+JOIN innovation AS i ON i.sinno_id = bv.sinno_id
+WHERE bv.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
+UNION
+SELECT DISTINCT(ei.sinno_id)
+, 'Bioecology Vision' AS bioeconomy_vision
+, i.year_of_commercialization AS "i.year"
+FROM eco_innovations AS ei
+JOIN innovation AS i ON i.sinno_id = ei.sinno_id
+WHERE ei.innovation_type IN (206, 107)
+AND ei.sinno_id NOT IN (
+  SELECT ei2.sinno_id
+  FROM eco_innovations as ei2
+  WHERE ei2.innovation_type IN (999, '000')
+  );
+-- adds recycling innovation to bioecology
+""",
+    conn,
+)
+
+bioeconomy_visions.loc[
+    bioeconomy_visions["bioeconomy_vision"] == "Not Bioeconomy",
+    ["bioeconomy_vision"],
+] = "No Bioeconomy Vision"
+
+uncertain_bioeconomy_visions = bioeconomy_visions.query("bioeconomy_vision == 'Unsure'")
+
+certain_bioeconomy_visions = bioeconomy_visions.loc[
+    ~bioeconomy_visions["sinno_id"].isin(uncertain_bioeconomy_visions["sinno_id"]), :
+]
+
+count_certain_bioeconomy_visions = (
+    certain_bioeconomy_visions.groupby(["year", "bioeconomy_vision"])
+    .size()
+    .reset_index(name="count")
+)
+
+total_certain_bioeconomy_visions = (
+    count_certain_bioeconomy_visions.groupby("bioeconomy_vision")
+    .agg("sum", numeric_only=True)
+    .sort_values("count", ascending=False)
+    .reset_index()
+)
 
@@
 
-```{python}
-bv_df = pd.read_sql(
-    """SELECT bv.sinno_id, bv.bioeconomy_vision, i.year_of_commercialization AS "i.year"
-FROM bioeconomy_visions AS bv
-JOIN innovation AS i ON i.sinno_id = bv.sinno_id
-WHERE bv.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
-OR bv.sinno_id NOT IN (SELECT sinno_id FROM bioeconomy_visions WHERE bioeconomy_vision <> 0);
-""",
-    conn,
-)
-```
-
 ::::{.columns}
@@
 
+n_certain_bioeconomy_visions = certain_bioeconomy_visions["sinno_id"].nunique()
+
 display(
     Markdown(
-        f"<span class='text-primary' style='font-size:15rem; font-weight:900'>{bv_df.loc[:, 'sinno_id'].nunique()}</span>"
+        f"<span class='text-primary' style='font-size:15rem; font-weight:900'>{n_certain_bioeconomy_visions}</span>"
     )
@@
 
-## Bioeconomy Vision by Year
+## Trends in Bioeconomy Visions{.smaller}
+### [No Bioeconomy]{.text-gray-600}, [Bioecology]{style="color:#cc78bc;"}, [Bioresource]{style="color:#de8f05;"} and [Biotechnoloy]{style="color:#029e73;"}{.smaller}
 
 ```{python}
+# | label: fig-bv_trends
+# | fig-cap: '[No Bioeconomy]{.text-gray-600}, [Bioecology]{style="color:#cc78bc;"}, [Bioresource]{style="color:#de8f05;"} and [Biotechnoloy]{style="color:#029e73"}'
 
-plot_bv_trends(bv_year, title="")
+color_map = {
+    "No Bioeconomy Vision": "rgb(0.8, 0.8, 0.8)",  # Gray
+    "Bioecology Vision": "#cc78bc",  # Purple
+    "Bioresource Vision": "#de8f05",  # Orange
+    "Biotechnology Vision": "#029e73",  # Green
+}
+
+def plot_bv_trends_interactive(df):
+    traces = []
+
+    for bv, bioeconomy_vision in df.groupby("bioeconomy_vision"):
+        traces.append(
+            go.Scatter(
+                x=bioeconomy_vision["year"],
+                y=bioeconomy_vision["count"],
+                name=bv,
+                mode="lines",
+                line=dict(color=color_map[bv]),  # Set line color based on the color_map
+            )
+        )
+
+    fig = go.Figure(data=traces)
+    fig.update_layout(showlegend=False, hovermode="x unified", height=550)
+    fig.show()
+
+plot_bv_trends_interactive(count_certain_bioeconomy_visions)
 

### 2023-09-25 docs/presentations/20230928-sehm_presentation.qmd
update bv trends fig in sehm_pres and figures.py

@@
 ## Trends in Bioeconomy Visions{.smaller}
-### [No Bioeconomy]{.text-gray-600}, [Bioecology]{style="color:#cc78bc;"}, [Bioresource]{style="color:#de8f05;"} and [Biotechnoloy]{style="color:#029e73;"}{.smaller}
+### [No Vision]{.text-gray-600}, [Bioecology]{style="color:#cc78bc;"}, [Bioresource]{style="color:#de8f05;"} and [Biotechnoloy]{style="color:#029e73;"}{.smaller}
 
@@
 
-color_map = {
-    "No Bioeconomy Vision": "rgb(0.8, 0.8, 0.8)",  # Gray
-    "Bioecology Vision": "#cc78bc",  # Purple
-    "Bioresource Vision": "#de8f05",  # Orange
-    "Biotechnology Vision": "#029e73",  # Green
-}
-
-def plot_bv_trends_interactive(df):
-    traces = []
-
-    for bv, bioeconomy_vision in df.groupby("bioeconomy_vision"):
-        traces.append(
-            go.Scatter(
-                x=bioeconomy_vision["year"],
-                y=bioeconomy_vision["count"],
-                name=bv,
-                mode="lines",
-                line=dict(color=color_map[bv]),  # Set line color based on the color_map
-            )
-        )
-
-    fig = go.Figure(data=traces)
-    fig.update_layout(showlegend=False, hovermode="x unified", height=550)
-    fig.show()
-
-plot_bv_trends_interactive(count_certain_bioeconomy_visions)
+plot_bv_trends_interactive(count_certain_bioeconomy_visions).show()
 

### 2023-09-25 src/figures.py

@@ -10,43 +10,36 @@ pio.templates.default = "presentation"
 
 royal_blue_800 = "#283aa4"
 gray_600 = "#878f97"
+def plot_bv_trends_interactive(df):
+  
+    color_map = {
+        "No Bioeconomy Vision": "rgb(0.8, 0.8, 0.8)",  # Gray
+        "Bioecology Vision": "#cc78bc",  # Purple
+        "Bioresource Vision": "#de8f05",  # Orange
+        "Biotechnology Vision": "#029e73",  # Green
+    }
 
-def plot_bioeconomy_trends_interactive(data, y):
-    fig = go.Figure()
+    traces = []
 
-    # Add yearly data to plot
-    fig.add_trace(
-        go.Scatter(
-            x=data["year"],
-            y=data[y],
-            mode="lines",
-            name="Yearly",
-            line=dict(dash="dash", color=gray_600),
+    for bv, bioeconomy_vision in df.groupby("bioeconomy_vision"):
+        traces.append(
+            go.Scatter(
+                x=bioeconomy_vision["year"],
+                y=bioeconomy_vision["count"],
+                name=bv,
+                mode="markers+lines",
+                line=dict(color=color_map[bv]),  # Set line color based on the color_map
+            )
         )
-    )
 
-    # Add rolling mean data to plot
-    rolling_mean = data[y].rolling(window=5).mean()
-    fig.add_trace(
-        go.Scatter(
-            x=data["year"],
-            y=rolling_mean,
-            mode="lines",
-            name="5-year moving average",
-            line=dict(color=royal_blue_800),
-        )
-    )
-
-    fig.update_xaxes(type="date")
+    fig = go.Figure(data=traces)
     fig.update_layout(
         showlegend=False,
-        xaxis_range=["1965", "2025"],
-        font=dict(family="Atkinson Hyperlegible"),
-        height=550,
         hovermode="x unified",
+        height=550,
+        yaxis_range=[0, 1],
+        xaxis_range=["1965", "2015"],
     )
-
-    # Show plot
     return fig
 
 
@@ -145,17 +138,6 @@ def plot_bv_trends(df):
     sns.despine()
 
     plt.show()
-
-def plot_bv_trends_interactive(df):
-    fig = px.line(df, x="year", y="count", color="bioeconomy_vision",
-                  color_discrete_sequence=px.colors.qualitative.Light24)
-    fig.update_layout(xaxis_tickangle=-45, xaxis_tickfont_size=10,
[…]

### 2023-09-26 src/figures.py
readded plot_bioeconomy_trends_interactive

@@ -6,42 +6,42 @@ import plotly.express as px
 import plotly.io as pio
 
 # Set global options for all Plotly plots
-pio.templates.default = "presentation"
 
 royal_blue_800 = "#283aa4"
 gray_600 = "#878f97"
-def plot_bv_trends_interactive(df):
-  
-    color_map = {
-        "No Bioeconomy Vision": "rgb(0.8, 0.8, 0.8)",  # Gray
-        "Bioecology Vision": "#cc78bc",  # Purple
-        "Bioresource Vision": "#de8f05",  # Orange
-        "Biotechnology Vision": "#029e73",  # Green
-    }
+def plot_bioeconomy_trends_interactive(data, y):
+    fig = go.Figure()
 
-    traces = []
-
-    for bv, bioeconomy_vision in df.groupby("bioeconomy_vision"):
-        traces.append(
-            go.Scatter(
-                x=bioeconomy_vision["year"],
-                y=bioeconomy_vision["count"],
-                name=bv,
-                mode="markers+lines",
-                line=dict(color=color_map[bv]),  # Set line color based on the color_map
-            )
+    fig.add_trace(
+        go.Scatter(
+            x=data["year"],
+            y=data[y],
+            mode="lines",
+            name="Yearly",
+            line=dict(dash="dash", color=gray_600),
         )
+    )
 
-    fig = go.Figure(data=traces)
+    # Rolling mean
+    rolling_mean = data[y].rolling(window=5).mean()
+    fig.add_trace(
+        go.Scatter(
+            x=data["year"],
+            y=rolling_mean,
+            mode="lines",
+            name="5-Year Moving Average",
+            line=dict(color=royal_blue_800)
+        )
+    )
+    fig.update_xaxes(type="date")
     fig.update_layout(
         showlegend=False,
-        hovermode="x unified",
+        xaxis_range=["1965", "2025"]
+        font=dict(family="Atkinson Hyperlegible"),
         height=550,
-        yaxis_range=[0, 1],
-        xaxis_range=["1965", "2015"],
+        hovermode="x unified"
     )
     return fig
-
 
 def plot_count(data):
     fig, ax = plt.subplots()
@@ -138,6 +138,40 @@ def plot_bv_trends(df):
     sns.despine()
 
     plt.show()
+
+
+def plot_bv_trends_interactive(df):
+  
+    color_map = {
+        "No Bioeconomy Vision": "rgb(0.8, 0.8, 0.8)",  # Gray
[…]

### 2023-09-26 src/figures.py
update interactive labels on bv trends in figures.py

@@ -36,11 +36,12 @@ def plot_bioeconomy_trends_interactive(data, y):
     fig.update_xaxes(type="date")
     fig.update_layout(
         showlegend=False,
-        xaxis_range=["1965", "2025"]
+        xaxis_range=["1965", "2025"],
         font=dict(family="Atkinson Hyperlegible"),
         height=550,
         hovermode="x unified"
     )
+
     return fig
 
 def plot_count(data):
@@ -139,9 +140,7 @@ def plot_bv_trends(df):
 
     plt.show()
 
-
 def plot_bv_trends_interactive(df):
-  
     color_map = {
         "No Bioeconomy Vision": "rgb(0.8, 0.8, 0.8)",  # Gray
         "Bioecology Vision": "#cc78bc",  # Purple
@@ -159,19 +158,25 @@ def plot_bv_trends_interactive(df):
                 name=bv,
                 mode="markers+lines",
                 line=dict(color=color_map[bv]),  # Set line color based on the color_map
+                visible="legendonly",
             )
         )
 
     fig = go.Figure(data=traces)
     fig.update_layout(
-        showlegend=False,
+        legend=dict(
+            orientation="h",
+            yanchor="bottom",
+            y=1.02,
+            xanchor="left",
+            font=dict(size=30, family="Atkinson Hyperlegible"),
+        ),
         hovermode="x unified",
         height=550,
-        yaxis_range=[0, 1],
         xaxis_range=["1965", "2015"],
+        font=dict(family="Atkinson Hyperlegible"),
     )
     return fig
-
 
 def plot_heatmap(data):
     fig, ax = plt.subplots(1, 1)

### 2023-09-26 docs/presentations/20230928-sehm_presentation.qmd
add interactive ecoinnovation plot

@@
 
+# Three Visions for a Bioeconomy
+
+:::: columns
+::: {.column width="33%"}
+![Bio-Technology Vision](../assets/Biotech.jpg){.fragment}
+:::
+
+::: {.column width="33%"}
+![Bio-Resource Vision](../assets/Bioresource.jpg){.fragment}
+:::
+
+::: {.column width="33%"}
+![Bio-Ecology Vision](../assets/Bioecology.jpg){.fragment}
+:::
+::::
+
+[@bugge2016WhatBioeconomyReview]{.slide-cite style="font-size: x-small"}
+
+## Research Questions
+
+1.
+2.
+3.
+
 ```{python}
@@
 
-# set all text on plots to be monospace with rcParams
-plt.rcParams["font.family"] = "Space Grotesk"
-
-sns.set_theme(style="white")
+pio.templates.default = "presentation"
 
@@
 
-## Bioeconomy Demarkations {.smaller}
+## Bioeconomy Innovation Definition {.smaller}
 
@@
 
-****
+[based on @wolfslehner2016ForestBioeconomyNew]{.slide-cite}
 
@@
 # label: fig-bioeconomy_count
-#
-#
 
@@
 
-# Three Visions for a Bioeconomy
+## Bioeconomy Vision Classification{.smaller .scrollable}
+@bugge2016WhatBioeconomyReview
 
-:::: columns
-::: {.column width="33%"}
-![Bio-Technology Vision](../assets/Biotech.jpg){.fragment}
-:::
-
-::: {.column width="33%"}
-![Bio-Resource Vision](../assets/Bioresource.jpg){.fragment}
-:::
-
-::: {.column width="33%"}
-![Bio-Ecology Vision](../assets/Bioecology.jpg){.fragment}
-:::
-::::
-
-[@bugge2016WhatBioeconomyReview]{.slide-cite style="font-size: x-small"}
-
-## Bioeconomy Vision Classification{.smaller .scrollable}
-
-| @bugge2016WhatBioeconomyReview | Biotechnology | Bioresource | Bioecology |
+|  | Biotechnology | Bioresource | Bioecology |
 |---|---|---|---|
@@
 
-| @vivien2019HijackingBioeconomy | Type II | Type III | Type I |
+<style>
+  tbl {
+    fontsize=9;
+  }
+</style>
+
+## Bioeconomy Vision Classification {.smaller}
+@vivien2019HijackingBioeconomy
+
+|  | Type II | Type III | Type I |
 |---|---|---|---|
@@
 
-::::{.columns}
-:::{.column width='80%'}
 ```{python}
@@
     Markdown(
-        f"<span class='text-primary' style='font-size:15rem; font-weight:900'>{n_certain_bioeconomy_visions}</span>"
+        f"<span class='r-fit-text text-royal-blue-600'>{n_certain_bioeconomy_visions}</span> "
     )
@@
 
-:::
-
-:::{.column width='20%'}
 unique forest bioeconomy innovations classified
-:::
-
-::::
-
-## Examples
 
 ## Trends in Bioeconomy Visions{.smaller}
-### [No Vision]{.text-gray-600}, [Bioecology]{style="color:#cc78bc;"}, [Bioresource]{style="color:#de8f05;"} and [Biotechnoloy]{style="color:#029e73;"}{.smaller}
+<!-- ### [No Vision]{.text-gray-600}, [Bioecology]{style="color:#cc78bc;"}, [Bioresource]{style="color:#de8f05;"} and [Biotechnoloy]{style="color:#029e73;"}{.smaller} -->
 
 ```{python}
-# | label: fig-bv_trends
-# | fig-cap: '[No Bioeconomy]{.text-gray-600}, [Bioecology]{style="color:#cc78bc;"}, [Bioresource]{style="color:#de8f05;"} and [Biotechnoloy]{style="color:#029e73"}'
+# | label: Trends in Bioeconomy Visions Figure
 
 plot_bv_trends_interactive(count_certain_bioeconomy_visions).show()
+
+```
+
+```{python}
+# | output: asis
+
+recycle_innovation = pd.read_sql(
+    """
+SELECT *
+FROM classification_codes
+LEFT JOIN eco_innovations ON classification_codes.Code = eco_innovations.innovation_type
+WHERE eco_innovations.innovation_type = 107 OR eco_innovations.innovation_type = 206 AND sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes LIKE '%not forest%')
+""",
+    conn,
+)
+
+recycle_innovation_count = recycle_innovation["sinno_id"].nunique()
+
+n_bioecology_vision = count_certain_bioeconomy_visions.loc[
+    count_certain_bioeconomy_visions["bioeconomy_vision"] == "Bioecology Vision",
+    "count",
+].sum()
+
+display(
+    Markdown(
+        f"""
+        :::{{.notes}} 
+          {recycle_innovation_count} innovations involving either a product or process innovation related to recyling.
+Hence, {n_bioecology_vision - recycle_innovation_count} innovation are associated with the bioecology vision, without using or facilitating recycling.
+        :::
+        """
+    )
+)
+
+```
+
+## Eco Innovations
+
+```{python}
+
+eco_innovation_merged_bv = certain_bioeconomy_visions.merge(
+    eco_innovations, on="sinno_id", validate="many_to_many"
+)
+
+eco_innovation_by_bioeconomy_visions = (
+    eco_innovation_merged_bv.groupby(["bioeconomy_vision", "innovation_type_code"])
+    .size()
+    .to_frame("count")
+    .reset_index()
+)
+
+eco_innovation_by_bioeconomy_visions["innovation_type_group"] = np.where(
+    eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0] == "1",
+    "Product",
+    np.where(
+        eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0] == "2",
+        "Process",
+        np.where(
+            eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0] == "3",
+            "Organizational",
+            np.where(
+                eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0]
+                == "4",
+                "Marketing",
+                "No Eco-Innovation",
+            ),
+        ),
+    ),
+)
+```
+
+```{python}
+plot_eco_innovation_by_bioeconomy_vision_interactive(
+    eco_innovation_by_bioeconomy_visions
+).show()
 
@@
 ## References
+
+:::{#refs}
+:::
+
+# Appendix
+
+## Count of all SWINNO
+### [5 Year MA]{.text-royal-blue-800} and [Yearly]{.text-gray-600}
+
+```{python}
+
+plot_bioeconomy_trends_interactive(bioeconomy_share, "total_count")
+```
+
+## Count of Bioeconomy Visions
+
+```{python}
+
+display(
+    Markdown(
+        count_certain_bioeconomy_visions.groupby("bioeconomy_vision")
+        .agg("sum", numeric_only=True)
+        .sort_values("count", ascending=False)
+        .reset_index()
+        .rename(columns={"bioeconomy_vision": "Vision", "count": "Count"})
+        .to_markdown(index=False)
+    )
+)
+
+```

### 2023-09-26 src/figures.py

@@ -287,3 +287,38 @@ def plot_eco_innovation_by_bioeconomy_vision(data, label_lookup):
     # plt.show()
 
     return fig
+
+
+def plot_eco_innovation_by_bioeconomy_vision_interactive(data):
+    label_map = create_label_mapping(codes_df)
+
+    data["innovation_type_label"] = data["innovation_type_code"].map(label_map)
+
+    fig = px.bar(
+        x=data["count"],
+        y=data["innovation_type_label"],
+        color=data["innovation_type_group"],
+        facet_col=data["bioeconomy_vision"],
+        facet_col_wrap=4,
+    )
+
+    fig.for_each_annotation(
+        lambda a: a.update(text=a.text.split("=")[-1].strip(" Vision"))
+    )
+    fig.update_layout(
+        showlegend=False,
+        font=dict(family="Atkinson Hyperlegible"),
+        margin=dict(l=150),
+        yaxis=dict(autorange="reversed"),
+    )
+    fig.update_traces(hovertemplate="<b>%{x}</b> %{y} Innovations")
+
+    # hide subplot y-axis titles and x-axis titles
+    for axis in fig.layout:
+        if type(fig.layout[axis]) == go.layout.YAxis:
+            fig.layout[axis].title.text = ""
+            fig.layout[axis].tickfont = dict(size=10)
+        if type(fig.layout[axis]) == go.layout.XAxis:
+            fig.layout[axis].title.text = ""
+            fig.layout[axis].tickfont = dict(size=12)
+    return fig

### 2023-09-26 docs/presentations/20230928-sehm_presentation.qmd
update data slide and add result section in sehm_pres

@@
 
+# Data{.inverted}
+```{python}
+# | label: Data
+n_swinno = swinno["sinno_id"].nunique()
+
+display(
+    Markdown(
+        f"""
+  ## SWINNO -- A Swedish Innovation Output Database with {n_swinno} Innovations{{.smaller}}
+
+  ### [5 Year MA]{{.text-royal-blue-800}} and [Yearly]{{.text-gray-600}}
+  """
+    )
+)
+
+plot_bioeconomy_trends_interactive(bioeconomy_share, "total_count")
+
+```
+
 ## Bioeconomy Innovation Definition {.smaller}
@@
 ```
+
+# Results{.inverted}
 
@@
 
-## Count of all SWINNO
-### [5 Year MA]{.text-royal-blue-800} and [Yearly]{.text-gray-600}
-
-```{python}
-
-plot_bioeconomy_trends_interactive(bioeconomy_share, "total_count")
-```
-
 ## Count of Bioeconomy Visions

### 2023-09-26 docs/presentations/20230928-sehm_presentation.qmd
fix new labels in interactive eco-innovation plot

@@
 
-## Research Questions
+## Research Questions {.fragment}
 
-1.
-2.
-3.
+1. How innovative is the Swedish forest bioeconomy?
+2. What visions of a bioeconomy are represented?
+3. What eco-innovations are developed?
 
@@
 plot_eco_innovation_by_bioeconomy_vision_interactive(
-    eco_innovation_by_bioeconomy_visions
+    eco_innovation_by_bioeconomy_visions, codes_df
 ).show()

### 2023-09-26 src/figures.py

@@ -289,7 +289,7 @@ def plot_eco_innovation_by_bioeconomy_vision(data, label_lookup):
     return fig
 
 
-def plot_eco_innovation_by_bioeconomy_vision_interactive(data):
+def plot_eco_innovation_by_bioeconomy_vision_interactive(data, codes_df):
     label_map = create_label_mapping(codes_df)
 
     data["innovation_type_label"] = data["innovation_type_code"].map(label_map)

### 2023-09-26 docs/presentations/20230928-sehm_presentation.qmd
update presentation formatting

@@
 
-# Three Visions for a Bioeconomy
+## Three Visions for a Bioeconomy
 
@@
 
-## Research Questions {.fragment}
+## Research Questions
 
+:::{.incremental}
 1. How innovative is the Swedish forest bioeconomy?
@@
 3. What eco-innovations are developed?
+:::
 
@@
 
-# Data{.inverted}
+# Data {.inverted}
+
 ```{python}
 # | label: Data
+# | output: asis
+
 n_swinno = swinno["sinno_id"].nunique()
@@
     Markdown(
-        f"""
-  ## SWINNO -- A Swedish Innovation Output Database with {n_swinno} Innovations{{.smaller}}
-
-  ### [5 Year MA]{{.text-royal-blue-800}} and [Yearly]{{.text-gray-600}}
-  """
+        f"## SWINNO -- A Swedish Innovation Output Database with {n_swinno} Innovations"
+        + "{.smaller}"
     )
 )
+
+```
+
+### [5 Year MA]{.text-royal-blue-800} and [Yearly]{.text-gray-600}
+
+```{python}
 
@@
 
-```{python}
-# | label: Sum bioeconomy visions
-# | output: asis
+## Count of Bioeconomy Innovation
+###  Uncleaned [5 Year MA]{.text-royal-blue-800} and [Yearly]{.text-gray-600}
 
-sum_bioeconomy_innovations = swinno["bioeconomy"].sum()
-
-display(
-    Markdown(
-        f"# {sum_bioeconomy_innovations} Innovations in Forest Bioeconomy (Uncleaned){{.inverted}}"
-    )
-)
-
-```
-
-# Results{.inverted}
-
-## Count of Bioeconomy Innovation
-### [5 Year MA]{.text-royal-blue-800} and [Yearly]{.text-gray-600}
 ```{python}
@@
 
+```{python}
+# | label: Sum bioeconomy visions
+# | output: asis
+
+sum_bioeconomy_innovations = swinno["bioeconomy"].sum()
+
+display(
+    Markdown(
+        f"{sum_bioeconomy_innovations}{{.text-royal-blue-400}} Total Innovations in Forest Bioeconomy (Uncleaned)"
+    )
+)
+
+```
+
 ## Share of Bioeconomy Innovation{.smaller}
-### [5 Year MA]{.text-royal-blue-800} and [Yearly]{.text-gray-600}
+###  Uncleaned  [5 Year MA]{.text-royal-blue-800} and [Yearly]{.text-gray-600}
 ```{python}
@@
 
-## Bioeconomy Vision Classification{.smaller .scrollable}
+## Bioeconomy Vision Classification {.smaller}
 @bugge2016WhatBioeconomyReview
 
-|  | Biotechnology | Bioresource | Bioecology |
+|  | [Biotechnology]{.text-colorblind-green} | [Bioresource]{.text-colorblind-orange} | [Bioecology]{.text-colorblind-purple} |
 |---|---|---|---|
@@
 | Drivers   & mediators of innovation | R & D, patents, TTOs,   Research councils and funders (Science push, linear model) | Interdisciplinary, optimisation   of land use, include degraded land in the production of biofuels, use and   availability of bio-resources, waste management, engineering, science &   market (Interactive & networked production mode) | Identification of favourable   organic agro-ecological practices, ethics, risk, transdisciplinary   sustainability, ecological interactions, re-use & recycling of waste,   land use, (Circular and self-sustained production mode) |
-
-<style>
-  tbl {
-    fontsize=9;
-  }
-</style>
+: {.tiny}
 
@@
 
-|  | Type II | Type III | Type I |
+|  | [Type II]{.text-colorblind-green} | [Type III]{.text-colorblind-orange} | [Type I]{.text-colorblind-purple} |
 |---|---|---|---|
-|  | A science-based economy driven   by industrial biotechnology<br>     <br>     The cell is a factory | Biomass replaces fossil fuels   and mining to produce energy and materials<br>     <br>     Biorefining at the heart of ecological transition (multilevel perspective).
-| An ecological economy, that is   compatible with the biosphere <br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
+|  | A science-based economy driven  by industrial biotechnology| Biomass replaces fossil fuels   and mining to produce energy and materials| An ecological economy, that is compatible with the biosphere |
+| | The cell is a factory | Biorefining at the heart of ecological transition (multilevel perspective).
+| Counter-expertise rather than concrete technical solutions <br> Criticism from  social groups who remain at the margins of decisionmaking centers |
+:  {.small}
 
-## Preliminary Results
+# Preliminary Results {.inverted}
 
@@
     Markdown(
-        f"<span class='r-fit-text text-royal-blue-600'>{n_certain_bioeconomy_visions}</span> "
+        f"## <span class='r-fit-text text-royal-blue-600'>{n_certain_bioeconomy_visions}</span> unique forest bioeconomy innovations classified"
     )
@@
 # | output: asis
+# | include: false
 
@@
 ```
+
+::: notes
+39 innovations involving either a product or process innovation related to recyling. Hence, 22 innovation are associated with the bioecology vision, without using or facilitating recycling.
+:::
 

### 2023-09-26 docs/presentations/20230928-sehm_presentation.qmd
fix headers in sehm_pres

@@
       data-background-size: fit
-      data-background-opacity: '0.3'
+      data-background-opacity: '0.45'
     embed-resources: true
@@
 
+****
+
 ```{python}
@@
     Markdown(
-        f"## SWINNO -- A Swedish Innovation Output Database with {n_swinno} Innovations"
-        + "{.smaller}"
+        f"### SWINNO -- A Swedish Innovation Output Database with {n_swinno} Innovations"
     )
@@
 ```
+
+[@sjoo2014DatabaseSwedishInnovations]{.slide-cite}
+
+<style>
+  h2 {font-size=3rem;}
+</style>
 
@@
 ```{python}
+# | output: asis
 
@@
 
-display(
-    Markdown(
-        f"## <span class='r-fit-text text-royal-blue-600'>{n_certain_bioeconomy_visions}</span> unique forest bioeconomy innovations classified"
-    )
-)
+text = f"""<span class="r-fit-text">after<b> {n_certain_bioeconomy_visions}</b> unique forest innovations classified</span>"""
+
+display(Markdown(text))
 
 ```
-
-unique forest bioeconomy innovations classified
 

### 2024-01-29 docs/reports/paper/paper1.qmd
Refactor SQL query to exclude non-forest innovation

@@
     """ 
-select
+    SELECT distinct sinno_id from (SELECT
     i.sinno_id,
     i.innovation_name_in_swedish AS name,
+    i.literal_name_of_innovating_firm_sinnos_common_name_sinno_id AS firm,
+    i.literal_name_of_collaborator_sinnos_common_name_sinno_id AS collaborator_0,
+    i.literal_name_of_collaborator_sinnos_common_name_sinno_id1 AS collaborator_1,
+    i.literal_name_of_collaborator_sinnos_common_name_sinno_id2 AS collaborator_2,
+    i.literal_name_of_collaborator_sinnos_common_name_sinno_id3 AS collaborator_3,
+    i.literal_name_of_collaborator_sinnos_common_name_sinno_id4 AS collaborator_4,
+    i.literal_name_of_collaborator_sinnos_common_name_sinno_id5 AS collaborator_5,
+    i.literal_name_of_collaborator_sinnos_common_name_sinno_id6 AS collaborator_6,
+    i.literal_name_of_collaborator_sinnos_common_name_sinno_id7 AS collaborator_7,
+    i.literal_name_of_collaborator_sinnos_common_name_sinno_id8 AS collaborator_8,
     i.description_in_swedish AS description,
-  i.additional_information_if_origin_new_scientific_discovery || i.additional_information_if_origin_new_technologies_or_materials || i.additional_info_if_origin_official_regulation_legislation_and_standards || i.additional_information_if_origin_solution_for_a_problem || i.additional_information_if_origin_performance || i.additional_information_if_origin_other AS info,
     i.year_of_commercialization AS year,
     us.use_sector
-from
+    FROM
     innovation i
-  join use_sectors us on i.sinno_id = us.sinno_id
-where
+    JOIN use_sectors us ON i.sinno_id = us.sinno_id
+    WHERE
     (
-    us.use_sector like '02%'
-    or us.use_sector like '20%'
-    or us.use_sector like '21%'
-    or us.use_sector like '36%'
-    or product_code like '02%'
-    or product_code like '20%'
-    or product_code like '21%'
-    or product_code like '36%'
+        us.use_sector LIKE '02%'
+        OR us.use_sector LIKE '20%'
+        OR us.use_sector LIKE '21%'
+        OR us.use_sector LIKE '36%'
+        OR product_code LIKE '02%'
+        OR product_code LIKE '20%'
+        OR product_code LIKE '21%'
+        OR product_code LIKE '36%'
     )
-  or (
-    description like '%virke%'
-    or description like '%cellulos%'
-    or description like '%lignin%'
-    or description like '%spån%'
-    or description like '%bark%'
-    or description like '%levulinsyra%'
-    or description like '%furfural%'
-    or description like '%svarttjära%'
-    or description like '%svartlut%'
-    or description like '%växtbas%'
-    or description like '%ved%'
-    or description like '%trä%'
-    or description like '%skog%'
-    or description like '%biobränsle%'
-    or description like '%biologisk%'
-    or description like '%nedbrytbar%'
-    or description like '%papper%'
-    or description like '%pappret%'
-    or description like '%karton%'
-    or description like '%tencel%'
-  );
+    OR (
+        description LIKE '%virke%'
+        OR description LIKE '%cellulos%'
+        OR description LIKE '%lignin%'
+        OR description LIKE '%spån%'
+        OR description LIKE '%bark%'
+        OR description LIKE '%levulinsyra%'
+        OR description LIKE '%furfural%'
+        OR description LIKE '%svarttjära%'
+        OR description LIKE '%svartlut%'
+        OR description LIKE '%växtbas%'
+        OR description LIKE '%ved%'
+        OR description LIKE '%trä%'
+        OR description LIKE '%skog%'
+        OR description LIKE '%biobränsle%'
+        OR description LIKE '%biologisk%'
+        OR description LIKE '%nedbrytbar%'
+        OR description LIKE '%papper%'
+        OR description LIKE '%pappret%'
+        OR description LIKE '%karton%'
+        OR description LIKE '%tencel%'
+    )
+    AND i.sinno_id not in (select sinno_id from categorization_notes
+    where notes not like "%not forest%"));
 """,

### 2024-03-13 docs/reports/paper/paper1.qmd
migration commit

@@
 format:
-    wordcount-pdf: 
+  hikmah-manuscript-pdf: 
     papersize: A4
-        linestretch: 1.5
 execute:
@@
 year_max = swinno["year"].max()
+```
 
-display(
-    Markdown(
-        f"The SWINNO database used here contains {n_swinno} innovations commercialized between {year_min} and {year_max}.
+The SWINNO database used here contains `{python} n_swinno` innovations commercialized between `{python} year_min` and `{python} year_max`.
 It is constructed according to  literature-based innovation output (LBIO) method using Swedish source material [@sjoo2014DatabaseSwedishInnovations].
-This method relies on the judgment of editiors and journalists at trade journals and hence captures what can be described as significant innovation."
-    )
-)
-```
+This method relies on the judgment of editiors and journalists at trade journals and hence captures what can be described as significant innovation.
 
@@
 
-classification_examples = pd.read_sql(
-    """SELECT
-  n.sinno_id,
-  i.description_in_swedish,
-  GROUP_CONCAT(c.Category, ' & ') AS Categories,
-  n.notes
-FROM
-  categorization_notes n
-  JOIN innovation i ON i.sinno_id = n.sinno_id
-  JOIN bioeconomy_visions bv
-  ON n.sinno_id = bv.sinno_id
-  JOIN classification_codes c ON bv.bioeconomy_vision = c.Code
-WHERE
-  n.notes LIKE "%good example%"
-GROUP BY
-  n.sinno_id,
-  i.description_in_swedish,
-  n.notes""",
-    conn,
-)
+# classification_examples = pd.read_sql(
+#     """SELECT
+#   n.sinno_id,
+#   i.description_in_swedish,
+#   GROUP_CONCAT(c.Category, ' & ') AS Categories,
+#   n.notes
+# FROM
+#   categorization_notes n
+#   JOIN innovation i ON i.sinno_id = n.sinno_id
+#   JOIN bioeconomy_visions bv
+#   ON n.sinno_id = bv.sinno_id
+#   JOIN classification_codes c ON bv.bioeconomy_vision = c.Code
+# WHERE
+#   n.notes LIKE "%good example%"
+# GROUP BY
+#   n.sinno_id,
+#   i.description_in_swedish,
+#   n.notes""",
+#     conn,
+# )
 
-classification_examples_out = (
-    classification_examples.copy()
-    .loc[:, ["Categories", "description_in_swedish"]]
-    .iloc[2:]
-)
+# classification_examples_out = (
+#     classification_examples.copy()
+#     .loc[:, ["Categories", "description_in_swedish"]]
+#     .iloc[2:]
+# )
 
-rowIDs = range(1, len(classification_examples_out) + 1)
-display(
-    Markdown(
-        classification_examples_out.rename(
-            columns={"description_in_swedish": "Short Description"}
-        ).to_markdown(index=rowIDs)
-    )
-)
+# rowIDs = range(1, len(classification_examples_out) + 1)
+# display(
+#     Markdown(
+#         classification_examples_out.rename(
+#             columns={"description_in_swedish": "Short Description"}
+#         ).to_markdown(index=rowIDs)
+#     )
+# )
 
@@
 
-display(
-    Markdown(
-        f"Querying the SWINNO database for forest based bioeconomy innovations initally yields {sum_bioeconomy_innovations} innovations.
+```
+
+Querying the SWINNO database for forest based bioeconomy innovations initally yields `{python} sum_bioeconomy_innovations` innovations.
 This number includes false positives from substrings matching keywords included in the search.
 While there is no indication that these false positives are non-randomly distributed across time and sectors, the following analysis for bioeconomy innovation trends at large needs to be viewed in the light of this being unlceaned data.
-Where the analysis turns to specific bioeconomy visions, only clean and certain cases are discussed."
-    )
-)
-```
+Where the analysis turns to specific bioeconomy visions, only clean and certain cases are discussed.
 
@@
 ].values[0]
-
-display(
-    Markdown(
-        f"@fig-bioeconomy_share shows the count (a) and share relative to all innovation contained in SWINNO (b) of bioeconomy innovation by year and as five year moving averages.
-From visual inspection alone, two features of the forest based bioeconomy innovation system emerge.
-First, two periods of high output can be identified from 1970-1990 and from 2005 to approximately 2015.
-The period in between 1990 and 2005 is characterized by low average output with high volatility.
-Second, the forest based bioeconomy has failed to gain in importance relative to other innovation captures by SWINNO.
-From a high of {share_max:.2%} in {share_max_year} to {share_latest:.2%} in {share_last_year}."
-    )
-)
 
@@
 
-sectors = pd.read_sql(
-    """
-SELECT
-  bv.sinno_id as sinno_id,
-  c.Category AS vision,
-  sni_codes.label AS sector
-FROM
-  bioeconomy_visions AS bv
-  JOIN use_sectors ON bv.sinno_id = use_sectors.[sinno_id]
-  JOIN sni_codes ON SUBSTR(use_sectors.use_sector, 1, 2) = sni_codes.code
-  JOIN classification_codes AS c ON bv.bioeconomy_vision = c.Code;
-""",
-    conn,
-)
+# sectors = pd.read_sql(
+#     """
+# SELECT
+#   bv.sinno_id as sinno_id,
+#   c.Category AS vision,
+#   sni_codes.label AS sector
+# FROM
+#   bioeconomy_visions AS bv
+#   JOIN use_sectors ON bv.sinno_id = use_sectors.[sinno_id]
+#   JOIN sni_codes ON SUBSTR(use_sectors.use_sector, 1, 2) = sni_codes.code
+#   JOIN classification_codes AS c ON bv.bioeconomy_vision = c.Code;
+# """,
+#     conn,
+# )
 
-sectors = sectors.loc[
-    sectors["sinno_id"].isin(set(certain_bioeconomy_visions["sinno_id"])), :
-]
+# sectors = sectors.loc[
+#     sectors["sinno_id"].isin(set(certain_bioeconomy_visions["sinno_id"])), :
+# ]
 
-sectors.loc[sectors["vision"] == "Not Bioeconomy", "vision"] = "No Bioeconomy Vision"
+# sectors.loc[sectors["vision"] == "Not Bioeconomy", "vision"] = "No Bioeconomy Vision"
 
-sector_counts = (
-    sectors.groupby(["sector", "vision"]).size().to_frame("count").reset_index()
-)
+# sector_counts = (
+#     sectors.groupby(["sector", "vision"]).size().to_frame("count").reset_index()
+# )
 
-short_sector_labels = {
-    "Manufacture of other transport equipment": "Transport Equipment",
-    "Manufacture of motor vehicles, trailers and semi-trailers": "Motor Vehicle Manufacturing",
-    "Manufacture of furniture; manufacturing n.e.c.": "Furniture Manufacturing",
-    "Manufacture of fabricated metal products, except machinery and equipment": "Metal Products Manufacturing",
-    "Manufacture of machinery and equipment n.e.c.": "Machinery and Equipment",
-    "Construction": "Construction",
-    "Manufacture of electrical machinery and apparatus n.e.c.": "Electrical Machinery Manufacturing",
-    "Manufacture of radio, television and communication equipment and apparatus": "Communication Equipment",
-    "Manufacture of rubber and plastic products": "Rubber and Plastic Products",
-    "Forestry, logging and related service activities": "Forestry and Logging",
-    "Manufacture of wood and of products of wood and cork, except furniture; manufacture of articles of straw and plaiting materials": "Wood and Cork Products",
-    "Agriculture, hunting and related service activities": "Agriculture and Hunting",
-    "Land transport; transport via pipelines": "Land Transport",
-    "Supporting and auxiliary transport activities; activities of travel agencies": "Transport Services",
-    "Water transport": "Water Transport",
-    "Air transport": "Air Transport",
-    "Manufacture of medical, precision and optical instruments, watches and clocks": "Precision Instruments",
-    "Recycling": "Recycling",
-    "Manufacture of basic metals": "Basic Metals Manufacturing",
-    "Public administration and defence; compulsory social security": "Public Administration",
-    "Electricity, gas, steam and hot water supply": "Utilities",
-    "Sewage and refuse disposal, sanitation and similar activities": "Waste Management",
-    "Manufacture of chemicals and chemical products": "Chemical Products",
-    "Manufacture of pulp, paper and paper products": "Paper Products",
-    "Manufacture of coke, refined petroleum products and nuclear fuel": "Petroleum Products",
-    "Collection, purification and distribution of water": "Water Services",
-    "Publishing, printing and reproduction of recorded media": "Publishing and Media",
-    "Other mining and quarrying": "Mining and Quarrying",
-    "Mining of metal ores": "Metal Ore Mining",
-    "Tanning and dressing of leather; manufacture of luggage, handbags, saddlery, harness and footwear": "Leather and Footwear",
-    "Manufacture of food products and beverages": "Food and Beverages",
-    "Manufacture of other non-metallic mineral products": "Non-metallic Products",
-    "Wholesale trade and commission trade, except of motor vehicles and motorcycles": "Wholesale Trade",
-    "Other business activities": "Business Services",
-    "Research and development": "R&D",
-    "Manufacture of textiles": "Textile Manufacturing",
-    "Financial intermediation, except insurance and pension funding": "Financial Services",
-    "Education": "Education",
-    "Health and social work": "Healthcare",
-    "Fishing, operation of fish hatcheries and fish farms; service activities incidental to fishing": "Fishing and Aquaculture",
-    "Recreational, cultural and sporting activities": "Recreation and Culture",
-    "Extraction of crude petroleum and natural gas; service activities incidental to oil and gas extraction excluding surveying": "Oil and Gas Extraction",
-    "Post and telecommunications": "Post and Telecom",
-    "Sale, maintenance and repair of motor vehicles and motorcycles; retail sale of automotive fuel": "Automotive Services",
-    "Manufacture of wearing apparel; dressing and dyeing of fur": "Apparel Manufacturing",
-}
+# short_sector_labels = {
+#     "Manufacture of other transport equipment": "Transport Equipment",
+#     "Manufacture of motor vehicles, trailers and semi-trailers": "Motor Vehicle Manufacturing",
+#     "Manufacture of furniture; manufacturing n.e.c.": "Furniture Manufacturing",
+#     "Manufacture of fabricated metal products, except machinery and equipment": "Metal Products Manufacturing",
+#     "Manufacture of machinery and equipment n.e.c.": "Machinery and Equipment",
+#     "Construction": "Construction",
+#     "Manufacture of electrical machinery and apparatus n.e.c.": "Electrical Machinery Manufacturing",
+#     "Manufacture of radio, television and communication equipment and apparatus": "Communication Equipment",
+#     "Manufacture of rubber and plastic products": "Rubber and Plastic Products",
+#     "Forestry, logging and related service activities": "Forestry and Logging",
+#     "Manufacture of wood and of products of wood and cork, except furniture; manufacture of articles of straw and plaiting materials": "Wood and Cork Products",
+#     "Agriculture, hunting and related service activities": "Agriculture and Hunting",
+#     "Land transport; transport via pipelines": "Land Transport",
+#     "Supporting and auxiliary transport activities; activities of travel agencies": "Transport Services",
+#     "Water transport": "Water Transport",
+#     "Air transport": "Air Transport",
+#     "Manufacture of medical, precision and optical instruments, watches and clocks": "Precision Instruments",
+#     "Recycling": "Recycling",
+#     "Manufacture of basic metals": "Basic Metals Manufacturing",
+#     "Public administration and defence; compulsory social security": "Public Administration",
+#     "Electricity, gas, steam and hot water supply": "Utilities",
+#     "Sewage and refuse disposal, sanitation and similar activities": "Waste Management",
+#     "Manufacture of chemicals and chemical products": "Chemical Products",
+#     "Manufacture of pulp, paper and paper products": "Paper Products",
+#     "Manufacture of coke, refined petroleum products and nuclear fuel": "Petroleum Products",
+#     "Collection, purification and distribution of water": "Water Services",
+#     "Publishing, printing and reproduction of recorded media": "Publishing and Media",
+#     "Other mining and quarrying": "Mining and Quarrying",
+#     "Mining of metal ores": "Metal Ore Mining",
+#     "Tanning and dressing of leather; manufacture of luggage, handbags, saddlery, harness and footwear": "Leather and Footwear",
+#     "Manufacture of food products and beverages": "Food and Beverages",
+#     "Manufacture of other non-metallic mineral products": "Non-metallic Products",
+#     "Wholesale trade and commission trade, except of motor vehicles and motorcycles": "Wholesale Trade",
+#     "Other business activities": "Business Services",
+#     "Research and development": "R&D",
+#     "Manufacture of textiles": "Textile Manufacturing",
+#     "Financial intermediation, except insurance and pension funding": "Financial Services",
+#     "Education": "Education",
+#     "Health and social work": "Healthcare",
+#     "Fishing, operation of fish hatcheries and fish farms; service activities incidental to fishing": "Fishing and Aquaculture",
+#     "Recreational, cultural and sporting activities": "Recreation and Culture",
+#     "Extraction of crude petroleum and natural gas; service activities incidental to oil and gas extraction excluding surveying": "Oil and Gas Extraction",
+#     "Post and telecommunications": "Post and Telecom",
+#     "Sale, maintenance and repair of motor vehicles and motorcycles; retail sale of automotive fuel": "Automotive Services",
+#     "Manufacture of wearing apparel; dressing and dyeing of fur": "Apparel Manufacturing",
+# }
 
-sector_counts = sector_counts.sort_values("count", ascending=False)
+# sector_counts = sector_counts.sort_values("count", ascending=False)
 
-sector_counts["sector"] = sector_counts["sector"].map(short_sector_labels)
+# sector_counts["sector"] = sector_counts["sector"].map(short_sector_labels)
 
-heatmap_data = prepare_heatmap_data(sector_counts, "sector", "vision", "count")
+# heatmap_data = prepare_heatmap_data(sector_counts, "sector", "vision", "count")
 
-plot_heatmap(heatmap_data)
+# plot_heatmap(heatmap_data)
 

### 2024-03-13 paper.qmd
migration commit

--- 2024-03-13 docs/reports/paper/paper1.qmd
@@
 ---
-title: Shades of Green in the Swedish Forest-Based Bioeconomy Transition 
+title: Directionality and Direction of Innovation in the Swedish Forest-Based Bioeconomy Transition 
 subtitle: Draft
@@
 csl: https://zotero.org/styles/apa?source=1
-bibliography: [../../swinno-shades-green.bib]
+bibliography: [swinno-shades-green.bib]
 link-citations: true
@@
 number-sections: true
+
+execute: 
+  warning: false
+  error: false
+  echo: false
+
 format:
-  hikmah-manuscript-pdf: 
+    pdf:
         papersize: A4
-execute:
-    include: true
-    echo: false
-    warning: false
+        linestretch: 1.25
 ---
@@
 
-The Baltic forests are facing a trilemma of expectations to mitigate climate change, preserve biodiversity, render other ecosystem services and provide the resources for a future bioeconomy [@hogbom2021TrilemmaNordicBaltic].
-Balancing these expectations includes managing conflicting goals and hence difficult trade-offs between ecosystem services. 
-Transitioning to a bioeconomy, meaning an economy based on bio-based rather than fossil-based inputs is set to increase the pressure on Baltic forests in the coming years [@hogbom2021TrilemmaNordicBaltic].
-In recent years transitioning to a bioeconomy has gained relevance among policymakers, practitioners and researchers. 
-Connected to this are hopes of increased resource efficiency and renewability.
-In a report commissioned by the Swedish Government, the @2012StrategyBiobasedEconomy defines a vision of a Swedish bioeconomy based on two central pillars.
+<!-- Paragraphs 1-2: Motivation.
+After reading these paragraphs a reader in any field of economics should believe that if you answer your research question your paper will make an important contribution.
+-->
+Discussions about transitions to a bioeconomy are a hot topic of research and policy at the moment.
+Different visions of the endgoal exists, stressing needs for ecological, resource or technological aims.
+Actors in the transition stress the need for innovation to transition to a bioeconomy.
+Yet, the actual efforts undertaken to realize these visions remain underexplored.
+
+<!-- • Paragraphs 3-4: Challenges.
+These paragraphs explain why your research question has not already been answered, i.e., what are the central challenges a researcher must tackle to answer this question.
+-->
+Rigorously operationalizing directionality is challening due to a plurality of possible development paths [@stirling2011PluralisingProgressIntegrative].
+Stirling argues that innovation is better understood as a vector of possibilities rather than a scalar decision between clearly defined alternatives.
+<!-- Lastly, Stirling (2011, 2009, 2008) is a clear voice pointing out a slightly different implication of directionality.
+Arguably, he introduced the concept to the field of sustainability transitions research, as part of broader critique of one-dimensional growth ideology.
+In line with a multidimensional view on sociotechnical change, he stresses that innovation is better viewed as “a vector, rather than just a scalarquantity” (Stirling, 2011, p.
+84, 2008, p.
+263).
+But he goes further and makes a passionate and eloquent call for a more pluralistic perspective on technological innovation.
+
+@andersson2021OutcomesDirectionalityMorphology 111 -->
+Where innovation indicators are concerned, directionality is frequently constrained to a dichotomous choice by the innovator, for example, to produce an environmental innovation or to produce a conventional innovation.
+In reality, however, innovators face a much more uncertain decision with many possible paths, of which some might be exclusionary.
+Additionally, these studies frequently focus on patents as proxies for innovation acitvity.
+Patents, however, suffer from the flaw that they may be filed for strategic reasons for example, to prevent a competitor from accessing a technolgy.
+Care should therefore be taken to infer the direction of a transition from patent fillings alone.
+
+Previous research from transition studies and political sciences has been more succesful in illuminating the pluarlity of possible tranisiton directions.
+However, the basis for this is usually confined to the analysis of discourse, which can illuminate the intent, but as the saying goes "Talk is cheap".
+Without investigating what actors are willing to implement discourse analysis remains a poor measure of directionality in innovation directionality.
+
+A central challenge in assessing the directionality of transitions lies in lack of high quality data that allows to unpack which of the intended directions are seriously pursued by innovators.
+
+<!-- • Paragraph 5: This Paper. This paragraph states in a nutshell what the paper accomplishes and how. -->
+This paper adresses this issue by using a unique database of significant Swedish innnovations to examine trends in the transition to a forest based bioeconomy.
+Technical source documents are qualitatively coded regarding their alignment with different visions of a bioeconomy.
+The result paints a detailed picture of activities to achieve competing visions and through narrative analysis highlights how development blocks enable and prevent certain transition directions.
+
+<!-- • Paragraphs 6-7: Model. Summarize the key formal assumptions you will maintain in your analysis. -->
+The core assumption needed for this paper is that innovations can be aligned with visions of transitions without requiring that the vision is already explicitly formalized.
+
+<!-- • Paragraphs 8-9: Data. Explain where you obtain your data and how you measure the concepts that are central to your study. -->
+Bioeconomy innovations are identified through SNI codes of user and producer sectors as well as keywords in innovation descriptions from the SWINNO database.
+SWINNO consists of technical newspaper articles which report on innovations for a professional audience and hence offer high quality data on the type, function, construction and aims of or motivation behind innovations.
+Because these articles are published in edited professional journals they represent innovations of significance for the market and hence accurately capture the direction of search deemed most promising by the innovation system.
+
+<!-- • Paragraphs 10-11: Methods. Explain how you take your model to the data and how you overcome the challenges you raised in paragraphs 3-4. -->
+The inclusion of an innovation is based on the value chain of the forest based bioeconomy [@wolfslehner2016ForestBioeconomyNew].
+I develop a classification scheme build on visions of bioeconomy transition outcomes and eco-innovation types and apply it to the database sources.
+This allows me to analyze innovation as a vector rather than a scalar as called for by @stirling2010MulticriteriaDiversityAnalysis.
+
+<!-- • Paragraphs 12-13: Findings. Describe the key findings. Make sure they connect clearly to the motiva-
+tion in paragraphs 1-2. -->
+I find that the Swedish bioeconomy innovation output has declined over the last 50 years.
+Of the three different visions the bioresource vision is the most pronounced, in line with expecatations from discourse analysis and the Swedish bioeconomy strategy.
+I show, however, also that the strategy should not be seen as visionary, as the innovation system had been producing bioeconomy innovation and even bioresource aligned innovation long before the strategy was formulated.
+This points to a strong interconnection between path dependence and the creation of directionality in the policy space.
+Suprisingly, I find that the majority of innovations related to the bioeconomy cannot be classified as aligning with one of the three visions.
+Instead they represent general improvements to machinery and tools used in various segments of the forest based bioeconomy.
+
+The bioresource vision is rooted in a development block to respond to landscape pressure of energy scarcity during the 1970s oil crises.
+Examples of a strong bioecological vision of high value added products with regional identity exists, but lack coordination and remain in the niche.
+
+<!-- • Paragraphs 14-15: Literature.
+Lay out the two main ways your paper contributes to the literature.
+Each paragraph should center around one contribution and should explain precisely how your paper differs from the most closely related recent work.
+-->
+This paper adds to the literature on directionality by providing a deep insight into the direction of search for an ongoing sustainability transition.
+<!-- From an empirical perspective, this is essentially about furthering the current understanding of how sociotechnical systems grow as a result of historical path dependencies, system-internal dynamics and external influence, towards a more elaborate understanding of how these mechanisms influence how system configurations change over time.
+From a conceptual perspective, it is rather about enabling such investigations, by developing analytical frameworks that capture not only the pace, but also the directionality of innovation processes.
+For example, there may be a need to advancethe commonly used technological innovation systems framework, since the typology of functional processes is geared towards investigating system growth in a way that hides the potential for vastly different technical, social and spatial development trajectories (Andersson, 2020).
+@andersson2021OutcomesDirectionalityMorphology [127-128] -->
+
+It also adds to the literature about bioeconomy transitions specifically by measuring progress towards different visions of a bioeconomy for the first time using innovations produced by the Swedish forest based bioeconomy.
+
+# Directionality, Direction and the Transition to a Bioeconomy {#sec-literature}
+
+This section reviews the literature on transitions to a bioeconomy, introduces competing visions of desired outcomes of this transition, central actors in the Swedish discourse and key concepts relating to the governance of sustainability transitions such as the transition to a bioeconomy, namely directionality and the direction of search within innovation systems.
+
+## Clarification of Bioeconomy Concept
+
+The bioeconomy has gained relevance among policymakers, practitioners and researchers alike.
+Broadly speaking, the bioeconomy refers to an economy in which raw materials for energy and production processes are provided from biological instead of fossil sources.
+<!-- TODO: citation -->
+Hopes of increased resource efficiency and renewability as well as promised economic growth especially in regions rich in biological resources abound [@el-chichakli2016PolicyFiveCornerstones].
+However, a such a profound change to economic systems will not be possible without balancing significant tradeoffs between the needs and desires of different human and non-human stakeholders.
+@hogbom2021TrilemmaNordicBaltic summarize these tradeoffs as a trilemma between forests ability to provide sufficent wood to replace fossil based materials, carbon sequestration and storage, and the delivery of other eco-system services such as forest biodiversity.
+Resolving these tradeoffs is challening, as goal conflicts between different development pathways make simulatneous progress on social, ecological and economic goals difficult [@bennich2021BiobasedEconomy2030].
+
+As one of the leading forestry nations in the world, Sweden sees substantial economic and political potential in advancing the bioeconomy .
+In a report commissioned by the Swedish Government, the @2012StrategyBiobasedEconomy lays out the plan for a Swedish bioeconomy based on two central pillars.
 It is to be an economy based on,
@@
 Notions of a biotechnology and bio-resource oriented bioeconomy are widespread and dominate discourse at various levels [@dieken2021MultitudesBioeconomiesSystematic].
-Despite -- or perhaps because of -- bioeconomy lacking a strong definition, the term is used to refer to different concepts, with different aims and objectives, foci, normative assumptions and hence implications for society and environment [@bugge2016WhatBioeconomyReview].
+
+### Bioeconomic Visions {#sec-bioeconomy-visions}
 
@@
 
+An overview or more likely more elaborate description of @vivien2019HijackingBioeconomy and @bugge2016WhatBioeconomyReview here.
+Tables as placeholders for now.
+
+| | Biotechnology | Bioresource | Bioecology |
+|---|---|---|---|
+| Aim &   Objectives | Economic growth & job   creation | Economic growth &   sustainability | Sustainability, biodiversity,   conservation of ecosystems, avoiding soil degradation |
+| Value   Creation | Application of biotechnology,   commercialisation of research & technology | Conversion and upgrading of   bio-resources (process oriented) | Development of integrated   production systems and high-quality products with territorial identity |
+| Drivers   & mediators of innovation | R & D, patents, TTOs,   Research councils and funders (Science push, linear model) | Interdisciplinary, optimisation   of land use, include degraded land in the production of biofuels, use and   availability of bio-resources, waste management, engineering, science &   market (Interactive & networked production mode) | Identification of favourable   organic agro-ecological practices, ethics, risk, transdisciplinary   sustainability, ecological interactions, re-use & recycling of waste,   land use, (Circular and self-sustained production mode) |
+: Table adapted from @bugge2016WhatBioeconomyReview [10] {#tbl-bugge}
+
+|  | Type II | Type III | Type I |
+|---|---|---|---|
+|  | A science-based economy driven   by industrial biotechnology<br>     <br>     The cell is a factory | Biomass replaces fossil fuels   and mining to produce energy and materials<br>     <br>     Biorefining at the heart of ecological transition (multilevel perspective).
+| An ecological economy, that is   compatible with the biosphere<br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
+: Table adapted from @vivien2019HijackingBioeconomy [194] {#tbl-vivien}
+
 The exact role played by forests in a bioeconomy transition depends on the desired vision [@kleinschmit2014ShadesGreenSocial].
 There is no doubt that the Swedish vision places high demands on forests as primary providers of the required biomass.
+Or, as @belandlindahl2015SwedishForestryModel call it a forest sector that operates under a "more of everything" paradigm.
 Both in the strategy formulated in @research2012StrategyBiobasedEconomy and in public discourse, a small, well-connected group of actors center the bio-resource vision as legitimate and desirable [@holmgren2022WhoseTransformationThis].
-The result is a forest sector which operates under a "more of everything" paradigm [@belandlindahl2015SwedishForestryModel].
 
-## Innovation for a Bioeconomy {#sec-innovation_bioeconomy}
+A paragraph here explaining the detailed findings of @holmgren2022WhoseTransformationThis discourse network here.
 
-Apart from the substitution of fossil resources with bio-based resources, a central narrative used in the Swedish transition is that, "\[t\]echnological innovation is key to a greener future" [@holmgren2022WhoseTransformationThis 42].
-According to @jankovsky2021InnovationsForestBioeconomy there are four types of innovation crucial for a forest-based bioeconomy:
+Despite -- or perhaps because of -- bioeconomy lacking a strong definition, the term is used to refer to different concepts, with different aims and objectives, foci, normative assumptions and hence implications for society and environment [@bugge2016WhatBioeconomyReview].
+Common ground in ambitions to realize a bioeconomy is the need for diverse types of social, technological and organizational innovation [@el-chichakli2016PolicyFiveCornerstones].
+According to @jankovsky2021InnovationsForestBioeconomy there are four types of innovation crucial for a forest-based bioeconomy: (i) substitute products, (ii) new (bio-based) processes, (iii) new (bio-based) products, (iv) new behavior.
+The need for innovation also features large in Swedish bioeconomy discourse and narratives therein [@bauer2017TechnologicalInnovationSystems, @bennich2021BiobasedEconomy2030], as, "\[t\]echnological innovation is key to a greener future" [@holmgren2022WhoseTransformationThis 42].
 
-1.  Substitute Products,
-2.  New (bio-based) Processes,
-3.  New (bio-based) Products,
-4.  New Behavior.
+## Visions Provide Directionality to Innovation for Transitions
 
-Currently, most research into bioeconomy innovation relies on patent data or qualitative case studies [e.g., @bennich2021BiobasedEconomy2030; @jander2019MonitoringTransitionBioeconomy].
-Unfortunately, these approaches are not without their flaws.
-Concentrating on patents might overemphasize the significance of the biotechnology vision, which predominantly relies on a science-oriented innovation approach driven by patents [@bugge2016WhatBioeconomyReview].
-On the other hand, qualitative case studies, may miss broader trends and features of a bioeconomy transition.
+Visions, narratives, mental models and expectations about technological innovation play a central role in social science sustainability research such as the literature on innovation systems, transition studies or economic policy models such as Mission Oriented Innovation Policy.
 
-Public, scientific and political debate of bioeconomies has brought forth a number of discourse analyses.
-Both @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy conclude that the concept of a bioeconomy conceals different, possibly conflicting visions of transformation outcomes.
+<!-- ### Plurality of Possible Development Pathways
+Not infinite, as transition already embedded in time and space [@andersson2021OutcomesDirectionalityMorphology], but more than scalar @stirling2010MulticriteriaDiversityAnalysis, @stirling2011PluralisingProgressIntegrative, @bergek2015TechnologicalInnovationSystems -->
+
+### Directionality
+
+As innovation policy took a normative turn under which innovation is needed to solve grand societal challenges [@schot2018ThreeFramesInnovation], the question of how to steer innovation into desired directions became a central issue in innovation research [@kohler2019AgendaSustainabilityTransitions].
+Innovation turned from a scalar concept, where the central questions where how to promote the right amount of innovation, into a question of influencing the vector in which society develops through innovation [@stirling2011PluralisingProgressIntegrative].
+But, as innovation systems are embedded in a spatial and temporal context they tend to develop along trajectories characterized by path dependency [@kemp2022DiversifyingDeepTransitions, @bergek2015TechnologicalInnovationSystems].
+
+Two concepts are central to understanding how innovation is supposed to be guided based on normative decisions between different feasible paths, both serve as the background concepts in this and many related studies on sustainability transitions.
+The multilevel perspective (MLP) introduced by [@geels2007TypologySociotechnicalTransition, @geels2010OntologiesSociotechnicalTransitions] and the literature on innovation systems at their diverse levels of analysis such as technological, regional, or national levels.
+<!-- TODO add citations. -->
+@hekkert2007FunctionsInnovationSystems provide a framework of intervention in innovation on a systemic level, which highlights the importance of guidance of search as one of seven functions innovation system must provide.
+They explain guidance of search as, "those activities within the innovation system that can positively affect the visibility and clarity of specific wants among technology users" (423).
+It is changing preferences in society, which through market and government influence and, "..
+an interactive and cumulative process of exchanging ideas between technology producers, technology users, and many other actors" (423) decides which direction of technological development is most promising.
+@bauer2017TechnologicalInnovationSystems adapts the innovation system functions to the specific case of biorefineries in Sweden and define the influence on the direction of search as "the incentives and/or pressures for organizations to enter the technological field.
+These may come in the form of visions, expectations of growth potential, regulation, policy targets, standards, articulation of demand from leading customers, crises in current business, etc." (4).
+
+<!-- 
+The most accepted view in sustainability transitions studies is that system change is driven by a combination of exogenous and endogenous driving forces.
+Moreover, the external shocks ‘do not mechanically impact niches and regimes, but need to be perceived and translated by actors to exert influence’ (Geels & Schot, 2007, p.
+404).
+This implies that the actual directions of change are shaped by strategic action (Grillitsch et al., 2018; Markard & Truffer, 2008; Pacheco et al., 2010; Smith et al., 2010; Yap & Truffer, 2019).
+Although there are many studies on the role of agency in shaping sustainability transitions, there is still limited understanding about how this agency shapes the directionality of the change process.
+@yang2022ShapingDirectionalitySustainability [752]
+-->
+
+<!-- 
+The directionality battle is not about whether the new (niche actors) will win over the old (regime actors).
+In our research we do not want to tie regime actors upfront to a strategy of maintaining institutions (defending the regime) while niche actors do the creating (building niches) and disrupting work (destabilizing regimes).
+Rather, battles about the actual course of action may happen equally among regime actors within a prevailing regime as among actors supporting (potentially manifold) niches.
+Such a view accounts for a situation in which regime actors may operate in the niche and have an interest in promoting niches, while niche actors may not want to destroy the regime and prefer to operate at the niche level only.
+In the end, the unfolding directionality will be the result of interaction among actors with different degrees of incumbency (Jørgensen, 2012; Yap & Truffer, 2019).
+The fact that regime actors are not just defending the status quo has also been recognized in neo-institutional literature.
+The seminal work of Leblebici et al.
+(1991) emphasized that internal institutional contradictions may emerge as a starting point for dominant actors to engage with institutional change.
+In transition studies, Fünfschilling and Truffer (2014) elaborated how different institutional logics in a regime may create tensions within and among actors who are incumbents in the prevailing regime.
+We have therefore to account for a multiplicity of institutional work strategies of a multitude of actors, which are more or less tied to the prevailing regime structures.
+
+@yang2022ShapingDirectionalitySustainability [754] -->
+
+<!-- Our analysis supports that in the context of transformative industrial change, companies’ innovation activities change directionality.
+Companies’ central roles in the RIS still lie in exploiting, combining, and recombining knowledge and turning it into innovations (Autio, 1998; Isaksen & Trippl, 2016).
+The influence of the direction of search and legitimacy of change provided by pioneering organizations (Grillitsch & Sotarauta, 2020), such as Paper Province, additionally supports their identification with the environmental debate.
+
+@martin2023ReconsideringActorRoles 1643 -->
+
+@yang2022ShapingDirectionalitySustainability too highlight the importance of visions in directionality over transformations, stressing the importance of shared visions between niche and regime actors.
+They argue that a stretch-and-transform pattern is more likely if niche actors are given a leading role and core regime actors obtain new roles and identites.
+"The objective in the stretching and transforming form of niche empowerment is to convince the wider world that the rules of the game need to be changed.", @smith2012WhatProtectiveSpace [17].
+
+Visions as shared imaginaries, mental models and aspirations,
+narratives as their embodiment in discourse [@bauer2018NarrativesBiorefineryInnovation, @holmgren2022WhoseTransformationThis], technology as their material realization.
+
+@konnola2021TransformativeGovernanceInnovation highlight the importance of shared visions to provide a normative direction and purpose to a system.
+
+@grillitsch2019InnovationPolicySystemwidea illustrate that visions alone are not sufficent to provide clear directionality.
+They must be specific enough and with significant buyin from actors, which may be difficult to obtain with vision statements created elsewhere than within the system.
+
+<!-- Especially, this study indicates that the build-up of shared visions across niche and regime actors is key, and when these shared visions allow for a leading role of niche actors combined with openings for new roles and identities of core regime actors, the emergence of a stretch-and-transform pattern is more likely.
+
+@yang2022ShapingDirectionalitySustainability [766] -->
+
+<!-- Key paper cited in every paper on directionoality
+@weber2012LegitimizingResearchTechnology -->
+
+Instead of linear view, all levels interact to provide directionality, i.e. influencing the direction of search.
+Other influences on the direction of search exists as well, such as landscape pressures, positive and negative pressures, windows of opportunties etc.
+Top down visions and bottom up pragmatism interact to shape transition dynamics [@bolton2019GrandVisionsPragmatic].
+Operationalizing directionality hence difficult.
+Few notable attempts such as
+@cappellano2022WhereInnovationMeets operationalize directionality as a measure of R&I expenditure earmarked for specific transition objectives.
+Other attempts to measure the directionality and progress towards a bioeconomy @jander2019MonitoringTransitionBioeconomy and @jander2020MonitoringBioeconomyTransitions, but lack the plurality of innovation.
+
+<!-- As Stirling (2011) emphasizes, transition pathways are contingent.
+They involve contestations between different worldviews and interests.
+Instead of focusing entirely on the cascades of innovation that converge into clear-cut and singular regime shifts, transition analysis should remain attentive to the multiplicity of directions that a transition process may take.
+To be sure, the DT frameworkworks with an evolutionary logic that is far from deterministic.
+It addresses directionality explicitly.
+Schot & Kanger (2018, p.
+8) consider how the diversity of current socio-technical niches make for undercurrents that contain the seeds for alternative, sustainable and just directions of development.
+Kanger & Schot (2018, pp.
+15-16) describe a ‘portfolio of directionality’ that, throughout subsequent surges and notwithstanding the alignments across sectors, becomes broader over time.
+Along the formation of ‘regimes’ and ‘meta-regimes’, niche alternatives are not considered to just perish or become absorbed.
+Instead, these alternative directions may manifest after periods of hibernation, or they break through locally.
+@kemp2022DiversifyingDeepTransitions [112] -->
+
+<!-- However, the DT framework displays a rather reductionist understanding of directionality.
+It reduces the matter to analytical openness, minding the evolutionary multiplicity of a transformation trajectory (e.g.
+towards, or away from, mass production).
+The normative reflection remains tied to a particular direction that a transition process should be taking (K ̈ ohler et al., 2019).
+Yet what is needed is a more fundamentalattending to the multiplicityof possible socio-technical configurations that may emerge,to the different process dynamics that may unfold, and to the different (competing) normative appraisals – by stakeholders and by analysts - of outcomes (Pel et al., 2020).
+Stirling (2009, 2011) underlined the need to unfold evolutionary diversity in a way that clarifies the normative implications of this diversity.To articulate directionality is to increaseevaluative diversity (Stirling, 2008: 268), opening up the wickedness and the ambiguities of transition issues.
+@kemp2022DiversifyingDeepTransitions [113] -->
+
+<!-- Do I need patents to actually show the entire funnel? -->
+
 Recent scholarship has called for advancing research fronts instead of remaining entrenched in contested discourse positions [@allain2022BioeconomicsVsBioeconomy].
-
-This paper takes up this call and asks what bioeconomy the Swedish innovation system pursues.
-A unique database of actual innovation output is first used to answer how innovative the system has been between 1970 and 2020.
-The database sources are then used to provide a qualitative insight into the extent to which specific bioeconomy visions appear in commercialized innovation and the extent to which they include eco-innovations which could lessen environmental impact.
+Here is how I propose to measure how the visions have provided directionality to the Swedish Transition to a bioeconomy.
+Directionality as influencing the direction of search.
+I then measure the effectivness of the directionality provided by different visions in actually producing innovation in the direction of search.
+The emergence of visions is at least partially grounded in development paths taken by system actors.
+These then serve as important vehicles to create legitimacy of the chosen paths and further provide directionality for the search.
 
@@
 
+Based on the analyses of competing bioeconomy visions in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy, I develop and apply a matrix of characterstics for each vision.
+<!-- TODO: Matrix That matrix is presented in @tbl- -->
+Reading the source texts of innovations identified from SWINNO, I classify an innovation as being *aligned* with a vision if it fulfills at least one criteria in the matrix.
+This means that one innovation can be aligned with more than one vision at a time, for example, if it uses biological processes such as bacteria (biotechnology) to refine a wood based material to replace a fossil input elsewhere in the economy (bioresource vision).
+Not all innovations fullfill at least one criteria, yet still originate or are used within the analytical boundaries of the forest based bioeconomy.
+I classify these innovations as Vision Neutral.
+Examples include innovations aimed at the health and saftey of employees 
+<!-- TODO enter source -->
+, or general improvements to the longevity of materials and tools used somewhere in the value chain of the forest bioeconomy.
+The majority of innovations in this category are of the latter nature.
+
 Innovations were classified as belonging to a bioeconomy vision if they fulfilled one, or more criteria outlined in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy.
 @tbl-bugge and @tbl-vivien show the summary of bioeconomy visions presented in the cited papers.
+
+**DECISION MATRIX HERE**
+
 For the remainder of the paper the nomenclature of @bugge2016WhatBioeconomyReview is adopted.
@@
 
-| | Biotechnology | Bioresource | Bioecology |
-|---|---|---|---|
-| Aim &   Objectives | Economic growth & job   creation | Economic growth &   sustainability | Sustainability, biodiversity,   conservation of ecosystems, avoiding soil degradation |
-| Value   Creation | Application of biotechnology,   commercialisation of research & technology | Conversion and upgrading of   bio-resources (process oriented) | Development of integrated   production systems and high-quality products with territorial identity |
-| Drivers   & mediators of innovation | R & D, patents, TTOs,   Research councils and funders (Science push, linear model) | Interdisciplinary, optimisation   of land use, include degraded land in the production of biofuels, use and   availability of bio-resources, waste management, engineering, science &   market (Interactive & networked production mode) | Identification of favourable   organic agro-ecological practices, ethics, risk, transdisciplinary   sustainability, ecological interactions, re-use & recycling of waste,   land use, (Circular and self-sustained production mode) |
-: Table adapted from @bugge2016WhatBioeconomyReview [10] {#tbl-bugge}
-
-|  | Type II | Type III | Type I |
-|---|---|---|---|
-|  | A science-based economy driven   by industrial biotechnology<br>     <br>     The cell is a factory | Biomass replaces fossil fuels   and mining to produce energy and materials<br>     <br>     Biorefining at the heart of ecological transition (multilevel perspective).
-| An ecological economy, that is   compatible with the biosphere<br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
-: Table adapted from @vivien2019HijackingBioeconomy [194] {#tbl-vivien}
-
-@tbl-vision_examples shows five examples of classified innovation.
+@tbl-vision-examples shows five examples of classified innovation.
 This selection also highlights that multiple labels can be attached to one innovation.
@@
 This is the case for innovation which brings general technological improvements, purely advances economic performance, or relates to a general purpose technology such as appliance of computers.
-These innovations are labeled as No Bioeconomy Vision, which does not imply that the innovation is not used, or originates in the forest based bioeconomy as defined above.
+These innovations are labeled as Vision Neutral, which does not imply that the innovation is not used, or originates in the forest based bioeconomy as defined above.
 An illustrative example of such an innovation are improved work boots, which promote safety of forest workers but do not correspond with any of the criteria in @tbl-bugge or @tbl-vivien.
@@
 ```{python}
-# | label: tbl-vision_examples
+# | label: tbl-vision-examples
 # | tbl-cap: "List of Eco-Innovation Labels"
@@
 For example, some innovations aim to protect harvested timber, either at harvest, during transport or storage.
-While @tbl-full_codes shows all new added labels, in the following text those that do not clearly represent an eco-innovation -- following @garcia-granero2018EcoinnovationMeasurementReview "…fewer adverse effects on the environment and more efficient use of resources" (p.
+While @tbl-full-codes shows all new added labels, in the following text those that do not clearly represent an eco-innovation -- following @garcia-granero2018EcoinnovationMeasurementReview "…fewer adverse effects on the environment and more efficient use of resources" (p.
 305) -- are excluded.
@@
 ```{python}
-# | label: fig-bioeconomy_share
+# | label: fig-bioeconomy-share
 # | fig-cap: "Bioeconomy Innovation Trends in SWINNO (Uncleaned)"
@@
     ["bioeconomy_vision"],
-] = "No Bioeconomy Vision"
+] = "Vision Neutral"
 
@@
 ```{python}
-# | label: fig-bioeconomy_trends
+# | label: fig-bioeconomy-trends
 # | fig-cap: "Count of Bioeconomy Visions by Year"
@@
 
-@fig-bioeconomy_trends illustrates that the peaks in innovation output, identified in @fig-bioeconomy_share and discussed above, were driven by innovation related to the Bioresource Vision as well as No Bioeconomy Vision.
+@fig-bioeconomy-trends illustrates that the peaks in innovation output, identified in @fig-bioeconomy-share and discussed above, were driven by innovation related to the Bioresource Vision as well as Vision Neutral.
 While the Bioresource Visions did strongly contribute to the first peak, the second appears to have been primarily driven by innovation which cannot be associated as belonging to any vision of a bioeconomy.
@@
 ```{python}
-# | label: fig-vision_heatmap
+# | label: fig-vision-heatmap
 # | fig-caption: "Sector's Contribution to Bioeconomy Visions"
@@
 
-# sectors.loc[sectors["vision"] == "Not Bioeconomy", "vision"] = "No Bioeconomy Vision"
+# sectors.loc[sectors["vision"] == "Not Bioeconomy", "vision"] = "Vision Neutral"
 
@@
 While a quantitative assessment of technological impact on the environment comes with a host of challenges, the labelling of eco-innovations at least allows a glimpse at what could be expected from each vision.
-@fig-eco_innovation_by_bioeconomy_vision shows the count of eco-innovation associated with each bioeconomy vision by type.
+@fig-eco-innovation-bioeconomy-vision shows the count of eco-innovation associated with each bioeconomy vision by type.
 As before, there is no weighting, which is why the total sum of all eco-innovation across all bioeconomy visions outnumbers the sum of innovations in the bioeconomy.
 A single innovation can be more than one eco-innovation type, for example, if it reduces the amount of raw material needed by recycling waste.
-The picture of innovation in the forest bioeconomy which @fig-eco_innovation_by_bioeconomy_vision suggests fits previous literature and what can be expected from LBIO data.
+The picture of innovation in the forest bioeconomy which @fig-eco-innovation-bioeconomy-vision suggests fits previous literature and what can be expected from LBIO data.
 Product and Process innovation are by far the most frequent across all visions.
@@
 ```{python}
-# | label: fig-eco_innovation_by_bioeconomy_vision
+# | label: fig-eco-innovation-bioeconomy-vision
 # | fig-cap: Count of Eco-Innovation Types by Bioeconomy Vision
@@
 
+# Discussion {#sec-discussion}
+
+While the bioeconomy innovation system can be viewed as a niche in the transition to a economic system which does not rely on fossil based technologies, this analysis highlights the layered nature of contested transitions.
+Within the bioeconomic innovation system, the different visions could reasonably be conceived of as niches themselves, vying not only for importance over each other, but even for providing sufficient direction to overcome general purpose improvements to and from a regime system.
+
+## Limitations {#sec-limitations}
+
+# Conclusion {#sec-conclusion}
+
 {{< pagebreak >}}
-
 # References {.unnumbered}
@@
 ```{python}
-# | label: tbl-full_codes
+# | label: tbl-full-codes
 # | tbl-cap: "List of Eco-Innovation Labels"

### 2024-03-13 scripts/calculate_bioeconomy_share_data.py

@@ -1,0 +1,146 @@
+# | label: Imports
+
+import pandas as pd
+import numpy as np
+from pathlib import Path
+from src.figures import *
+from src.utils import get_project_root, format_table
+from src.swinno_helpers import connect_swinno_db
+
+ROOT = get_project_root()
+conn = connect_swinno_db()
+
+# set all text on plots to be monospace with rcParams
+plt.rcParams["font.family"] = "Space Grotesk"
+
+sns.set_theme(style="white")
+
+# full code labels for merging
+
+
+codes_df = pd.read_sql("SELECT * FROM classification_codes", conn)
+
+codes_df["Category"] = codes_df["Category"].str.strip()
+uncertain_eco = {"Code": "000", "Category": "Unsure"}
+codes_df = codes_df.append(uncertain_eco, ignore_index=True)
+
+codes_df = codes_df[~codes_df["Code"].duplicated(keep="last")]
+
+codes_df = codes_df.set_index("Code")
+
+# | label: Query bioeconomy
+
+bioeconomy = pd.read_sql(
+    """
+select
+  i.sinno_id,
+  i.innovation_name_in_swedish AS name,
+  i.description_in_swedish AS description,
+  i.additional_information_if_origin_new_scientific_discovery || i.additional_information_if_origin_new_technologies_or_materials || i.additional_info_if_origin_official_regulation_legislation_and_standards || i.additional_information_if_origin_solution_for_a_problem || i.additional_information_if_origin_performance || i.additional_information_if_origin_other AS info,
+  i.year_of_commercialization AS year,
+  us.use_sector
+from
+  innovation i
+  join use_sectors us on i.sinno_id = us.sinno_id
+where
+  (
+    us.use_sector like '02%'
+    or us.use_sector like '20%'
+    or us.use_sector like '21%'
+    or us.use_sector like '36%'
+    or product_code like '02%'
+    or product_code like '20%'
+    or product_code like '21%'
+    or product_code like '36%'
+  )
+  or (
+    description like '%virke%'
+    or description like '%cellulos%'
+    or description like '%lignin%'
+    or description like '%spån%'
+    or description like '%bark%'
+    or description like '%levulinsyra%'
+    or description like '%furfural%'
+    or description like '%svarttjära%'
+    or description like '%svartlut%'
+    or description like '%växtbas%'
[…]

### 2024-03-13 src/figures.py

@@ -1,5 +1,6 @@
 import pandas as pd
 import matplotlib.pyplot as plt
+import matplotlib.ticker as mtick
 import seaborn as sns
 import plotly.graph_objects as go
 import plotly.express as px
@@ -9,6 +10,8 @@ import plotly.io as pio
 
 royal_blue_800 = "#283aa4"
 gray_600 = "#878f97"
+
+
 def plot_bioeconomy_trends_interactive(data, y):
     fig = go.Figure()
 
@@ -30,7 +33,7 @@ def plot_bioeconomy_trends_interactive(data, y):
             y=rolling_mean,
             mode="lines",
             name="5-Year Moving Average",
-            line=dict(color=royal_blue_800)
+            line=dict(color=royal_blue_800),
         )
     )
     fig.update_xaxes(type="date")
@@ -39,10 +42,11 @@ def plot_bioeconomy_trends_interactive(data, y):
         xaxis_range=["1965", "2025"],
         font=dict(family="Atkinson Hyperlegible"),
         height=550,
-        hovermode="x unified"
+        hovermode="x unified",
     )
 
     return fig
+
 
 def plot_count(data):
     fig, ax = plt.subplots()
@@ -52,13 +56,10 @@ def plot_count(data):
         label="Yearly",
     )
 
-        # Calculate rolling mean and plot on same axis
+    # Calculate rolling mean and plot on same axis
     rolling_mean = data["bioeconomy_count"].rolling(window=5).mean()
     ax.plot(
-        data["year"],
-        rolling_mean,
-        label="5-year moving average",
-        linestyle="dashed"
+        data["year"], rolling_mean, label="5-year moving average", linestyle="dashed"
     )
 
     ax.set_xticks(data["year"][::10])
@@ -81,12 +82,9 @@ def plot_share(data):
     # Calculate rolling mean and plot on same axis
     rolling_mean = data["bioeconomy_share"].rolling(window=5).mean()
     ax.plot(
-        data["year"],
-        rolling_mean,
-        label="5-year moving average",
-        linestyle="dashed"
+        data["year"], rolling_mean, label="5-year moving average", linestyle="dashed"
     )
-    ax.set_ylim((0, .5))
+    ax.set_ylim((0, 0.5))
     ax.legend(frameon=False)
     ax.spines["top"].set_visible(False)
     ax.spines["right"].set_visible(False)
@@ -94,12 +92,13 @@ def plot_share(data):
 
 
[…]

### 2024-03-14 paper.qmd
fix migration issues to mac

@@
 # full code labels for merging
-
 codes_df = pd.read_sql("SELECT * FROM classification_codes", conn)
@@
 codes_df["Category"] = codes_df["Category"].str.strip()
-uncertain_eco = {"Code": "000", "Category": "Unsure"}
-codes_df = codes_df.append(uncertain_eco, ignore_index=True)
+uncertain_eco = {"Code": ["000"], "Category": ["Unsure"]}
+codes_df = pd.concat([codes_df, pd.DataFrame(uncertain_eco)], ignore_index=True)
 

### 2024-03-14 src/swinno_helpers.py

@@ -1,6 +1,7 @@
 import pandas as pd
 import numpy as np
 from sqlalchemy import create_engine
+from pathlib import Path
 from src.utils import get_project_root
 
 
@@ -78,7 +79,7 @@ def clean_import(file_type, file_path, **kwargs):
 
 
 def connect_swinno_db():
-    database_dir = get_project_root().parent.absolute()
-    database_uri = f"sqlite:///{database_dir}/swinno-db/data/swinno.db"
+    database_dir = "/Users/research/Library/CloudStorage/OneDrive-LundUniversity/research/swinno-db"
+    database_uri = f"sqlite:///{database_dir}/data/swinno.db"
     engine = create_engine(database_uri)
     return engine

### 2024-03-14 tests/test_connect_db.py

@@ -1,0 +1,29 @@
+import pytest
+from pathlib import Path
+from sqlalchemy import create_engine
+
+from src.swinno_helpers import connect_swinno_db
+
+
+def test_connect_swinno_db():
+    # Mock the get_project_root function
+    def mock_get_project_root():
+        return Path("/path/to/project/root")
+
+    # Mock the create_engine function
+    def mock_create_engine(database_uri):
+        return "Mocked engine"
+
+    # Monkey patch the get_project_root and create_engine functions
+    connect_swinno_db.get_project_root = mock_get_project_root
+    connect_swinno_db.create_engine = mock_create_engine
+
+    # Call the function and assert the result
+    engine = connect_swinno_db()
+    assert engine == "Mocked engine"
+
+    # Restore the original functions
+    connect_swinno_db.get_project_root = Path(get_project_root().parent.absolute())
+    connect_swinno_db.create_engine = create_engine
+
+    # Add more test cases if needed

### 2024-04-03 paper.qmd
update text

@@
     pdf:
+        documentclass: scrartcl
         papersize: A4
@@
 
-The bioeconomy has gained relevance among policymakers, practitioners and researchers alike.
+Although the bioecononmy has gained importance as a research, policy and industry focus, it remains unclear what exactly it entails.
 Broadly speaking, the bioeconomy refers to an economy in which raw materials for energy and production processes are provided from biological instead of fossil sources.
@@
 Hopes of increased resource efficiency and renewability as well as promised economic growth especially in regions rich in biological resources abound [@el-chichakli2016PolicyFiveCornerstones].
-However, a such a profound change to economic systems will not be possible without balancing significant tradeoffs between the needs and desires of different human and non-human stakeholders.
+However, a such a profound change to economic systems will not be possible without balancing significant tradeoffs between the needs and desires of different human and non-human stakeholders [@dieken2021MultitudesBioeconomiesSystematic].
 @hogbom2021TrilemmaNordicBaltic summarize these tradeoffs as a trilemma between forests ability to provide sufficent wood to replace fossil based materials, carbon sequestration and storage, and the delivery of other eco-system services such as forest biodiversity.
-Resolving these tradeoffs is challening, as goal conflicts between different development pathways make simulatneous progress on social, ecological and economic goals difficult [@bennich2021BiobasedEconomy2030].
+Resolving these tradeoffs is challenging, as goal conflicts between different development pathways make simulatneous progress on social, ecological and economic goals difficult [@bennich2021BiobasedEconomy2030].
 
@@
 These pillars reflect the two most influential visions of the bioeconomy concept.
-Notions of a biotechnology and bio-resource oriented bioeconomy are widespread and dominate discourse at various levels [@dieken2021MultitudesBioeconomiesSystematic].
+Notions of a biotechnology and bio-resource oriented bioeconomy are widespread, dominate discourse at various levels [@dieken2021MultitudesBioeconomiesSystematic] and form the main industrial visions [@befort2020GoingDefinitionsUnderstand].
+A third vision, centered around ecology and an economy compatible with the biosphere acts as a counter point.
 
@@
 
-In more recent years, visions of a bio-resource bioeconomy have taken over as the leading narrative.
+In more recent years, visions of a bio-resource bioeconomy have taken over as the leading narrative [@allain2022BioeconomicsVsBioeconomy].
 Unlike the biotechnology vision, this conceptualization does not only focus on economic growth and job creation, but attempts to include ecological sustainability challenges [@bugge2016WhatBioeconomyReview].
@@
 
-An overview or more likely more elaborate description of @vivien2019HijackingBioeconomy and @bugge2016WhatBioeconomyReview here.
-Tables as placeholders for now.
-
-| | Biotechnology | Bioresource | Bioecology |
-|---|---|---|---|
-| Aim &   Objectives | Economic growth & job   creation | Economic growth &   sustainability | Sustainability, biodiversity,   conservation of ecosystems, avoiding soil degradation |
-| Value   Creation | Application of biotechnology,   commercialisation of research & technology | Conversion and upgrading of   bio-resources (process oriented) | Development of integrated   production systems and high-quality products with territorial identity |
-| Drivers   & mediators of innovation | R & D, patents, TTOs,   Research councils and funders (Science push, linear model) | Interdisciplinary, optimisation   of land use, include degraded land in the production of biofuels, use and   availability of bio-resources, waste management, engineering, science &   market (Interactive & networked production mode) | Identification of favourable   organic agro-ecological practices, ethics, risk, transdisciplinary   sustainability, ecological interactions, re-use & recycling of waste,   land use, (Circular and self-sustained production mode) |
-: Table adapted from @bugge2016WhatBioeconomyReview [10] {#tbl-bugge}
-
-|  | Type II | Type III | Type I |
-|---|---|---|---|
-|  | A science-based economy driven   by industrial biotechnology<br>     <br>     The cell is a factory | Biomass replaces fossil fuels   and mining to produce energy and materials<br>     <br>     Biorefining at the heart of ecological transition (multilevel perspective).
-| An ecological economy, that is   compatible with the biosphere<br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
-: Table adapted from @vivien2019HijackingBioeconomy [194] {#tbl-vivien}
-
 The exact role played by forests in a bioeconomy transition depends on the desired vision [@kleinschmit2014ShadesGreenSocial].
@@
 
+For the remainder of the paper the nomenclature of @bugge2016WhatBioeconomyReview is adopted.
+Vivien et al.'s [-@vivien2019HijackingBioeconomy] Type I bioeconomy is taken to correspond with the bioecology vision, Type II and Type III to correspond with the bioresource and biotechnology vision, respectively.
+The analysis of bioeconomy discourse which revealed different visions, highlights that the visions have emerged at different times and with different actors at their centers [@vivien2019HijackingBioeconomy, @bugge2016WhatBioeconomyReview].
+@vivien2019HijackingBioeconomy trace the introduction of the term bioeconomics to the 1920s and a work on fishery by the Russian biologist Baranoff.
+The term bioeoconomics which they label as Type I bioeconomy, was introduced by Georgescu-Roegen, in -@georgescu-roegen1975BioeconomicAspectsEntropy.
+@vivien2019HijackingBioeconomy argue that the term biotechnology gradually emerged between 1990 and 2000, in the wake of enthusiasm about biotechnological advances, such as the discovery of DNA.
+It gained popularity and international importance when the OECD began increasingly supporting and reshaping its meaning at the turn of the millennium [@vivien2019HijackingBioeconomy].
+According to them, the last concept to enter the discourse was the bioresource oriented bioeconomy.
+While biomass has been used throughout history in various ways, the renewed interest with the intention to replace fossil inputs, for example in European Commission communication since 2010.
+However, @vivien2019HijackingBioeconomy point to the long history of biorefinery as a core aspect of this vision, which was discussed as early as the 1930s.
+Hence, while the organization of the discourse into the labels biotechnology, bioecology and bioresource vision is a very recent research outcome, the underlying concepts have a richer history.
+
 ## Visions Provide Directionality to Innovation for Transitions
@@
 
-### Directionality
+### Directionality {#sec-directionality}
 
@@
 
-<!-- Key paper cited in every paper on directionoality
-@weber2012LegitimizingResearchTechnology -->
+@weber2012LegitimizingResearchTechnology synthesis policy insights from innovation systems and multilevel perspective research into a failures framework.
+Therein, directionality failure is understood as the failure to provide direction to the process of search and unclear or missing setting of collective priorities.
+They caution that it should not be confused with shortsighted or flawed anticipation of market developments, as this myopia lacks a view on the direction of change.
+To overcome this failure, that is, to provide directionality, they argue, guiding orientation needs to be translated and intermediated on two levels.
+External requirements placed on the innovation system need to be absorbed after which they need to be interpreted in a process of negotiation between the different actors.
+Reaching this collective understanding of direction, @weber2012LegitimizingResearchTechnology [p. 1042] write, is in practice, "achieved by establishing shared future visions".
 
@@
 
-An important limitation of LBIO measures, especially for research into forest sector innovation, is a bias against non-commercialized innovation.
-New processes, or new organizational approaches for example, will only be captured if they are sold to another company.
-This is especially sensitive for bioeconomy research as, for example, new methods or plans for managing forest stands are less likely to be included.
-Relevant product innovation which facilitates these plans, for example products related to the planting of trees, on the other hand, are captured.
-A likely consequence of this limitation is that the number of innovation related to the bioecology vision are under counted, as this vision highlights "Counter-expertise rather than concrete technical solutions" [@vivien2019HijackingBioeconomy 194].
-Nonetheless, a few innovations associated with this vision were already identified in the database which will be discussed later.
-
 But, measuring innovation with LBIO has also substantial advantages.
@@
 Based on the analyses of competing bioeconomy visions in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy, I develop and apply a matrix of characterstics for each vision.
+ Visions based on @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy, biotechnology definition based on @yeung2019CurrentResearchBiotechnology. 
 <!-- TODO: Matrix That matrix is presented in @tbl- -->
@@
 
-Innovations were classified as belonging to a bioeconomy vision if they fulfilled one, or more criteria outlined in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy.
-@tbl-bugge and @tbl-vivien show the summary of bioeconomy visions presented in the cited papers.
-
-**DECISION MATRIX HERE**
-
-For the remainder of the paper the nomenclature of @bugge2016WhatBioeconomyReview is adopted.
-Vivien et al.'s [-@vivien2019HijackingBioeconomy] Type I bioeconomy is taken to correspond with the bioecology vision, Type II and Type III to correspond with the bioresource and biotechnology vision, respectively.
-The analysis of bioeconomy discourse which revealed different visions, highlights that the visions have emerged at different times and with different actors at their centers [@vivien2019HijackingBioeconomy, @bugge2016WhatBioeconomyReview].
-@vivien2019HijackingBioeconomy trace the introduction of the term bioeconomics to the 1920s and a work on fishery by the Russian biologist Baranoff.
-The term bioeoconomics which they label as Type I bioeconomy, was introduced by Georgescu-Roegen, in [-@georgescu-roegen1975BioeconomicAspectsEntropy].
-@vivien2019HijackingBioeconomy argue that the term biotechnology gradually emerged between 1990 and 2000, in the wake of enthusiasm about biotechnological advances, such as the discovery of DNA.
-It gained popularity and international importance when the OECD began increasingly supporting and reshaping its meaning at the turn of the millennium [@vivien2019HijackingBioeconomy].
-According to them, the last concept to enter the discourse was the bioresource oriented bioeconomy.
-While biomass has been used throughout history in various ways, the renewed interest with the intention to replace fossil inputs, for example in European Commission communication since 2010.
-However, @vivien2019HijackingBioeconomy point to the long history of biorefinery as a core aspect of this vision, which was discussed as early as the 1930s.
-Hence, while the organization of the discourse into the labels biotechnology, bioecology and bioresource vision is a very recent research outcome, the underlying concepts have a richer history.
-This alleviates concerns that using these labels retroactively confines the innovations to current policy terms.
-
-While following the criteria of @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy, a number of decisions were made concerning content of the summary presented in @tbl-bugge and @tbl-vivien, and the nature of the corresponding vision.
-For one, economic growth is cited as the aim and objective of biotechnology and bioresource visions by @bugge2016WhatBioeconomyReview.
-It can be argued that all innovation can contribute to economic growth, which is why these aspects were not given heightened importance in the classification decision.
-A second topic of judgment has to be the inclusion of innovation which increases the recyclability of a product, or uses recycled waste, water and material in the process.
-Based on the definition of a bioecology vision presented in @bugge2016WhatBioeconomyReview, these innovations should be included as contributing to this vision.
-The extent to which such innovation contributes in reality to a "[c]ircular and self-sustained production mode" [@bugge2016WhatBioeconomyReview 10], however, can be debated.
+{{< include _decision-matrix.md >}}
+: Decision matrix used to classify innovation. {#tbl-decision-matrix tbl-colwidths="[20,80]"}
 
@@
 
-Querying the SWINNO database for forest based bioeconomy innovations initally yields `{python} sum_bioeconomy_innovations` innovations.
+Querying the SWINNO database for forest based bioeconomy innovations initially yields `{python} sum_bioeconomy_innovations` innovations.
 This number includes false positives from substrings matching keywords included in the search.
-While there is no indication that these false positives are non-randomly distributed across time and sectors, the following analysis for bioeconomy innovation trends at large needs to be viewed in the light of this being unlceaned data.
+While there is no indication that these false positives are non-randomly distributed across time and sectors, the following analysis for bioeconomy innovation trends at large needs to be viewed in the light of this being uncleaned data.
 Where the analysis turns to specific bioeconomy visions, only clean and certain cases are discussed.
@@
 
+Within the bioeconomic innovation system, the different visions could reasonably be conceived of as niches themselves, vying not only for importance over each other, but even for providing sufficient direction to overcome general purpose improvements to and from a regime system.
 While the bioeconomy innovation system can be viewed as a niche in the transition to a economic system which does not rely on fossil based technologies, this analysis highlights the layered nature of contested transitions.
-Within the bioeconomic innovation system, the different visions could reasonably be conceived of as niches themselves, vying not only for importance over each other, but even for providing sufficient direction to overcome general purpose improvements to and from a regime system.
 
 ## Limitations {#sec-limitations}
+
+An important limitation of LBIO measures, especially for research into forest sector innovation, is a bias against non-commercialized innovation.
+New processes, or new organizational approaches for example, will only be captured if they are sold to another company.
+This is especially sensitive for bioeconomy research as, for example, new methods or plans for managing forest stands are less likely to be included.
+Relevant product innovation which facilitates these plans, for example products related to the planting of trees, on the other hand, are captured.
+A likely consequence of this limitation is that the number of innovation related to the bioecology vision are under counted, as this vision highlights "Counter-expertise rather than concrete technical solutions" [@vivien2019HijackingBioeconomy 194].
+Nonetheless, a few innovations associated with this vision were already identified in the database which will be discussed later.
 

### 2024-05-01 paper.qmd
update paper

@@
 ---
-title: Directionality and Direction of Innovation in the Swedish Forest-Based Bioeconomy Transition 
+title: Quantifying Directionality and Innovation Output in Sweden's Transition to a Forest-Based Bioeconomy
 subtitle: Draft
@@
   name: Jonas Kreutzer
-    affiliation: Department of Economic History, Lund University
+  affiliation: 'Department of Economic History, Lund University'
 date: last-modified
@@
   The transition to a bioeconomy is high on the agenda for research, industry and policy.
-Beyond conceptual uncertainties of what is meant by a bioeconomy, three different visions have emerged, differing in their aims and objectives, drivers and mediators of innovation and relation to the biosphere.
-The use of detailed innovation data from a Swedish innovation database, SWINNO, allows to capture the actual innovation output of the Swedish forest based bioeconomy innovation system.
-Qualitative coding of the underlying source material associates innovation from 1970 to 2010 with visions of a bioresource, biotechnology and / or bioecology bioeconomy.
-Preliminary results suggest that the innovation output of the forest based bioeconomy has decreased relative to the total output.
-Of the three visions, the bioresource vision has been the most dominant, especially in the 1970s and after 2005.
-At the same time, innovation which fit into multiple visions exist, suggesting that the opportunity space for future development is not yet locked in.
+Like other sustainability transition, the bioeconomy transition is characterized by contestation over desired outcomes and transition pathways.
+While shared visions are seen as pivotal in guiding transitions, the bioeconomy discourse is marked by the contention of three competing visions.
+The Swedish forest-based bioeconomy is well established as a case study in this arena.
+Yet, knowledge about innovation produced across the entire system is scarce.
+How innovative is this new economy, especially considering the central role of innovation in transition discourse?
+And what does commercialized innovation suggest for the directionality of the bioeconomy transition?
+The use of detailed innovation data from a Swedish literature-based innovation output database, SWINNO, allows to capture the actual innovation output of the Swedish forest-based bioeconomy innovation system.
+Qualitative coding of the source material links innovation from 1970 to 2021 with visions of a bioresource, biotechnology and / or bioecology bioeconomy.
+Findings reveal a notable decrease in the innovation output of the forest-based bioeconomy relative to Sweden's overall commercialized innovation.
+Notably, innovations associated with a bioresource vision surged during the energy crises of the 1970s, pre-dating their formalization in strategy documents, suggesting a significant role for path dependency in shaping shared visions.
+Concurrently, the presence of innovations aligning with multiple visions indicates an open opportunity space for future development across various directions.
+This research underscores the dynamic interplay between historical contexts, shared visions, and innovation trajectories within the bioeconomy transition, and questions the extent to which visions can provide novel directionality to transitions.
 
-csl: https://zotero.org/styles/apa?source=1
-bibliography: [swinno-shades-green.bib]
+csl: 'https://zotero.org/styles/apa?source=1'
+bibliography:
+  - swinno-shades-green.bib
 link-citations: true
@@
 number-sections: true
-
 execute:
@@
   echo: false
-
 format:
   pdf:
-        documentclass: scrartcl
-        papersize: A4
+    papersize: a4
     linestretch: 1.25
+    mainfont: "Adobe Garamond Pro"
+    sansfont: "Adobe Garamond Pro Semibold"
+    include-in-header: 
+        - text: \usepackage[a4paper, total={16.5cm,24.7cm}, top=2.5cm, ignorehead, centering, footskip=\footskip+4mm]{geometry}
+        - text: \raggedbottom 
+jupyter: python3
 ---
@@
 import numpy as np
+import scipy.stats as stats
+from pyhere import here
 from IPython.display import display, Markdown
@@
 import seaborn as sns
+from tabulate import tabulate
 
@@
 from src.utils import get_project_root, format_table
-from src.swinno_helpers import connect_swinno_db
+from src.data_handling import database
 
 ROOT = get_project_root()
-conn = connect_swinno_db()
+raw_data = here("data/raw")
+modified_data = here("data/modified")
+conn = database.connect_swinno_db()
 
@@
 sns.set_theme(style="white")
-
-# full code labels for merging
-codes_df = pd.read_sql("SELECT * FROM classification_codes", conn)
-
-codes_df["Category"] = codes_df["Category"].str.strip()
-uncertain_eco = {"Code": ["000"], "Category": ["Unsure"]}
-codes_df = pd.concat([codes_df, pd.DataFrame(uncertain_eco)], ignore_index=True)
-
-codes_df = codes_df[~codes_df["Code"].duplicated(keep="last")]
-
-codes_df = codes_df.set_index("Code")
 ```
@@
 ```{python}
-# | label: Query bioeconomy
-
-bioeconomy = pd.read_sql(
-    """ 
-    SELECT distinct sinno_id from (SELECT
-    i.sinno_id,
-    i.innovation_name_in_swedish AS name,
-    i.literal_name_of_innovating_firm_sinnos_common_name_sinno_id AS firm,
-    i.literal_name_of_collaborator_sinnos_common_name_sinno_id AS collaborator_0,
-    i.literal_name_of_collaborator_sinnos_common_name_sinno_id1 AS collaborator_1,
-    i.literal_name_of_collaborator_sinnos_common_name_sinno_id2 AS collaborator_2,
-    i.literal_name_of_collaborator_sinnos_common_name_sinno_id3 AS collaborator_3,
-    i.literal_name_of_collaborator_sinnos_common_name_sinno_id4 AS collaborator_4,
-    i.literal_name_of_collaborator_sinnos_common_name_sinno_id5 AS collaborator_5,
-    i.literal_name_of_collaborator_sinnos_common_name_sinno_id6 AS collaborator_6,
-    i.literal_name_of_collaborator_sinnos_common_name_sinno_id7 AS collaborator_7,
-    i.literal_name_of_collaborator_sinnos_common_name_sinno_id8 AS collaborator_8,
-    i.description_in_swedish AS description,
-    i.year_of_commercialization AS year,
-    us.use_sector
-    FROM
-    innovation i
-    JOIN use_sectors us ON i.sinno_id = us.sinno_id
-    WHERE
-    (
-        us.use_sector LIKE '02%'
-        OR us.use_sector LIKE '20%'
-        OR us.use_sector LIKE '21%'
-        OR us.use_sector LIKE '36%'
-        OR product_code LIKE '02%'
-        OR product_code LIKE '20%'
-        OR product_code LIKE '21%'
-        OR product_code LIKE '36%'
-    )
-    OR (
-        description LIKE '%virke%'
-        OR description LIKE '%cellulos%'
-        OR description LIKE '%lignin%'
-        OR description LIKE '%spån%'
-        OR description LIKE '%bark%'
-        OR description LIKE '%levulinsyra%'
-        OR description LIKE '%furfural%'
-        OR description LIKE '%svarttjära%'
-        OR description LIKE '%svartlut%'
-        OR description LIKE '%växtbas%'
-        OR description LIKE '%ved%'
-        OR description LIKE '%trä%'
-        OR description LIKE '%skog%'
-        OR description LIKE '%biobränsle%'
-        OR description LIKE '%biologisk%'
-        OR description LIKE '%nedbrytbar%'
-        OR description LIKE '%papper%'
-        OR description LIKE '%pappret%'
-        OR description LIKE '%karton%'
-        OR description LIKE '%tencel%'
-    )
-    AND i.sinno_id not in (select sinno_id from categorization_notes
-    where notes not like "%not forest%"));
-""",
-    conn,
-)
+#| label: Load Data
+raw_data = here("data/raw")
+bioeconomy = pd.read_csv(raw_data / "id_year_bioeconomy.csv")
+bioeconomy = bioeconomy.loc[bioeconomy["bioeconomy"] == 1]
 
@@
 
-swinno["bioeconomy"] = swinno["sinno_id"].isin(bioeconomy["sinno_id"].unique())
+swinno["bioeconomy"] = (
+    swinno["sinno_id"].astype(str).isin(bioeconomy["sinno_id"].astype(str).unique())
+)
 
+# full code labels for merging
+codes_df = pd.read_sql("SELECT * FROM classification_codes", conn)
+
+codes_df["Category"] = codes_df["Category"].str.strip()
+uncertain_eco = {"Code": ["000"], "Category": ["Unsure"]}
+codes_df = pd.concat([codes_df, pd.DataFrame(uncertain_eco)], ignore_index=True)
+
+codes_df = codes_df[~codes_df["Code"].duplicated(keep="last")]
+
+codes_df = codes_df.set_index("Code")
+
+eco_innovations = database.query_eco_innovations(conn)
+eco_innovations["innovation_type"] = eco_innovations["innovation_type"].str.strip()
+
+bioeconomy_visions = pd.read_csv(raw_data / "bioeconomy_visions.csv")
 ```
@@
 -->
-Discussions about transitions to a bioeconomy are a hot topic of research and policy at the moment.
-Different visions of the endgoal exists, stressing needs for ecological, resource or technological aims.
-Actors in the transition stress the need for innovation to transition to a bioeconomy.
-Yet, the actual efforts undertaken to realize these visions remain underexplored.
+At the heart of sustainability transitions lies the question how societal functions can be fulfilled within ecological boundaries [@kohler2019AgendaSustainabilityTransitions].
+Societal functions are provided through socio-technical systems, which consist of the interplay between the production, distribution, and use of technology [@geels2004SectoralSystemsInnovation].
+Transitions are complex, long-term processes of shifts between different systems, involving diverse actors who have to chose between multiple, open-ended pathways, uncertain of their technological, social and political success [@kohler2019AgendaSustainabilityTransitions].
+Further, success in the outcome and chosen pathway of sustainability transitions is contested based on the values, norms, and goals of  different actors and social groups.
+Hence, innovation cannot be seen as a scalar, but must rather be understood as a vector, as possessing and providing direction to change rather than being simply progress forward [@stirling2011PluralisingProgressIntegrative].
+@andersson2021OutcomesDirectionalityMorphology term these two properties _positive_ and _normative_ directionality, respectively.
+Positive directionality of innovation arises from the fact that technological change is path-dependent [@arthur1989CompetingTechnologiesIncreasing, @david1985ClioEconomicsQWERTY], while normative directionality is rooted in the assumption that without public policy, private actors lack intrinsic incentives to pursue sustainability [@kohler2019AgendaSustainabilityTransitions].
+
+A sustainability transition which has been gaining relevance for researchers, policy makers, and producers alike is the bioeconomy.
+Broadly speaking, it aims to radically shift economic systems from relying on fossil fuels for energy and material resource, to an economic system built on biobased materials, circular and renewable value chains and higher environmental integrity through innovation[@bugge2016WhatBioeconomyReview].
+As one of the main sources for biomass in a bioeconomy [@priefer2017PathwaysShapeBioeconomy], forests have become a key arena for the bioeconomy transition, highlighting tradeoffs between material provision, carbon storage and biodiversity [@hogbom2021TrilemmaNordicBaltic].
+Hence, underneath the bioeconomy umbrella, the concept and viable transition pathways are heavily contested [@allain2022BioeconomicsVsBioeconomy].
+Three different visions of a bioeconomy compete in discourse to provide normative directionality of the transition; the first vision centers on the provision of biomass and renewable resources, the second on the scientific development and application of biotechnology, while the third emphasizes ecological considerations[@bugge2016WhatBioeconomyReview, @vivien2019HijackingBioeconomy].
+The material effects of this transition, however, remain important research topics [@fischer2020SwedishForestsBioeconomy].
+How innovative is this new economy, especially considering the central role of innovation in transition discourse?
+And what does commercialized innovation suggest for the positive directionality of the bioeconomy transition?
+
+<!-- @fischer2020SwedishForestsBioeconomy also note the consensus builidng of Swedish bioeconomy discourse marginalizes non-mainstream opinions.
+I can show that this is rooted, at least in part, in the path-dependence on technology output.
+-->
 
@@
 Arguably, he introduced the concept to the field of sustainability transitions research, as part of broader critique of one-dimensional growth ideology.
-In line with a multidimensional view on sociotechnical change, he stresses that innovation is better viewed as “a vector, rather than just a scalarquantity” (Stirling, 2011, p.
+In line with a multidimensional view on sociotechnical change, he stresses that innovation is better viewed as “a vector, rather than just a scalar quantity” (Stirling, 2011, p.
 84, 2008, p.
@@
 @andersson2021OutcomesDirectionalityMorphology 111 -->
-Where innovation indicators are concerned, directionality is frequently constrained to a dichotomous choice by the innovator, for example, to produce an environmental innovation or to produce a conventional innovation.
-In reality, however, innovators face a much more uncertain decision with many possible paths, of which some might be exclusionary.
-Additionally, these studies frequently focus on patents as proxies for innovation acitvity.
-Patents, however, suffer from the flaw that they may be filed for strategic reasons for example, to prevent a competitor from accessing a technolgy.
-Care should therefore be taken to infer the direction of a transition from patent fillings alone.
 
-Previous research from transition studies and political sciences has been more succesful in illuminating the pluarlity of possible tranisiton directions.
-However, the basis for this is usually confined to the analysis of discourse, which can illuminate the intent, but as the saying goes "Talk is cheap".
-Without investigating what actors are willing to implement discourse analysis remains a poor measure of directionality in innovation directionality.
-
-A central challenge in assessing the directionality of transitions lies in lack of high quality data that allows to unpack which of the intended directions are seriously pursued by innovators.
+<!-- 1 The notion of directionality was introduced by Stirling (2008, 2009) as part of his call for opening up the process of technical change for alternative options.
+He builds on the broader idea that socio-technical change has a direction, choices are made between directions and actors gradually become blind to alternatives, which is a central tenet of much of the innovation studies literature.
+Our notion of directionality also draws on this idea (see also Weber and Rohracher, 2012: 1042–1043).
+@schot2018DeepTransitionsEmergence -->
 
@@
 ```{python}
-# | label: Total SWINNO Innovation
-# | output: asis
+#| label: Total SWINNO Innovation
+#| output: asis
 
@@
 
-{{< include _decision-matrix.md >}}
+| Corresponding Vision | Source article explicitly mentions that innovation:                                                   |
+|----------------------|-------------------------------------------------------------------------------------------------------|
+| Biotechnology        | Develops new, or modifies existing, organism, biological system, or process                            |
+|                      | Uses cellular or molecular process to solve problem or in production process                          |
+|                      | Applies biological agent in scientific or engineering based principle to produce product or service   |
+|                      | Based on molecular biology, biochemistry, cell biology, embryology, genetics or microbiology research |
+|                      | Provides method for research in the above areas                                                       |
+| Bioresource          | Biomass used to substitute other input                                                                |
+|                      | Refines biomass through mechanical, chemical or other process                                         |
+|                      | Converts and upgrades bioresource                                                                     |
+|                      | Optimises land use                                                                                    |
+|                      | Includes degraded land in production of biofuels                                                      |
+|                      | Intensifies use of biomass, including from waste                                                      |
+| Bioecology           | Prioritizes biodiversity, conservation, ecosystem or soil health                                      |
+|                      | Integrates production systems                                                                         |
+|                      | Develops high quality product with territorial identity                                               |
+|                      | Reuses or recycles waste                                                                              |
+|                      | States compatibility with biosphere as motivation or aim                                              |
+|                      | States need for social or ecological counter-expertise relative to other socio-technical innovations  |
+|                      | Promotes or uses a circular production mode                                                           |
+|                      | Promotes or uses a self-sustained production mode                                                     |
+
 : Decision matrix used to classify innovation. {#tbl-decision-matrix tbl-colwidths="[20,80]"}
@@
 
-```{python}
-# | label: tbl-vision-examples
-# | tbl-cap: "List of Eco-Innovation Labels"
-
-# classification_examples = pd.read_sql(
-#     """SELECT
-#   n.sinno_id,
-#   i.description_in_swedish,
-#   GROUP_CONCAT(c.Category, ' & ') AS Categories,
-#   n.notes
-# FROM
-#   categorization_notes n
-#   JOIN innovation i ON i.sinno_id = n.sinno_id
-#   JOIN bioeconomy_visions bv
-#   ON n.sinno_id = bv.sinno_id
-#   JOIN classification_codes c ON bv.bioeconomy_vision = c.Code
-# WHERE
-#   n.notes LIKE "%good example%"
-# GROUP BY
-#   n.sinno_id,
-#   i.description_in_swedish,
-#   n.notes""",
-#     conn,
-# )
-
-# classification_examples_out = (
-#     classification_examples.copy()
-#     .loc[:, ["Categories", "description_in_swedish"]]
-#     .iloc[2:]
-# )
-
-# rowIDs = range(1, len(classification_examples_out) + 1)
-# display(
-#     Markdown(
-#         classification_examples_out.rename(
-#             columns={"description_in_swedish": "Short Description"}
-#         ).to_markdown(index=rowIDs)
-#     )
-# )
-
-```
-
-### Eco-Innovation Types {#sec-eco_innovatoin_types}
+### Eco-Innovation Types {#sec-eco_innovation_types}
 
@@
 ```{python}
-# | label: Number Bioeconomy Innovation
-# | output: asis
+#| label: Number Bioeconomy Innovation
+#| output: asis
 
 sum_bioeconomy_innovations = swinno["bioeconomy"].sum()
-
 ```
@@
 ```{python}
-# | label: Query eco-innovation
-
-eco_innovations = pd.read_sql(
-    """
-SELECT
-  sinno_id,
-  innovation_type as innovation_type_code,
-  c.Category as innovation_type
-FROM
-  eco_innovations
-JOIN classification_codes AS c ON c.Code = eco_innovations.innovation_type
-WHERE 
-innovation_type NOT IN (701, 602, 601, 501, 111, 109, 108)
-;
-""",
-    conn,
-)
-
-eco_innovations["innovation_type"] = eco_innovations["innovation_type"].str.strip()
-
-```
-
-```{python}
-# | label: Calculate Bioeconomy Share
+#| label: Calculate Bioeconomy Share
 
@@
 bioeconomy_share = calculate_bioeconomy_share(swinno)
-
 ```
@@
 ```{python}
-# | label: Bioeconomy Trend Description
-# | output: asis
-
+#| label: Bioeconomy Trend Description
+#| output: asis
 share_max = bioeconomy_share["bioeconomy_share"].max()
@@
 ].values[0]
-
 ```
 
-```{python}
-# | label: fig-bioeconomy-share
-# | fig-cap: "Bioeconomy Innovation Trends in SWINNO (Uncleaned)"
-# | fig-subcap:
-# |   - "Bioeconomy Innovation Count"
-# |   - "Bioeconomy Innovation Share in %"
-# | layout-ncol: 1
-
-plot_count(bioeconomy_share)
-plot_share(bioeconomy_share)
-```
+![a.
+A total number of innovation registered in SWINNO database.
+b.
+Forest based bioeconomy innovations registered in SWINNO database.
+c.
+Percentage of forest based bioeconomy innovations to total innovation registered in SWINNO database.
+Shown are the 5 year moving averages for each time series.](results/figures/innovation_trends.pdf){#fig-innovation-trends}
 
@@
 ```{python}
-# | label: Query bioeconomy visions
-
-bioeconomy_visions = pd.read_sql(
-    """
-SELECT bv.sinno_id
-, codes.Category AS bioeconomy_vision
-, i.year_of_commercialization AS "year"
-FROM bioeconomy_visions AS bv
-JOIN [classification_codes] AS codes ON codes.code = bv.bioeconomy_vision
-JOIN innovation AS i ON i.sinno_id = bv.sinno_id
-WHERE bv.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
-UNION
-SELECT DISTINCT(ei.sinno_id)
-, 'Bioecology Vision' AS bioeconomy_vision
-, i.year_of_commercialization AS "i.year"
-FROM eco_innovations AS ei
-JOIN innovation AS i ON i.sinno_id = ei.sinno_id
-WHERE ei.innovation_type IN (206, 107)
-AND ei.sinno_id NOT IN (
-  SELECT ei2.sinno_id
-  FROM eco_innovations as ei2
-  WHERE ei2.innovation_type IN (999, '000')
-  );
--- adds recycling innovation to bioecology
-""",
-    conn,
-)
+#| label: Uncertain Visions
 
@@
 )
-
 ```
@@
 ```{python}
-# | label: Number of Uncertains
-# | output: asis
+#| label: Number of Uncertains
+#| output: asis
 
 n_uncertain_bioeconomy_visions = uncertain_bioeconomy_visions["sinno_id"].nunique()
-
-display(
-    Markdown(
-        f"For the following part, innovations which do not relate to the forest are excluded.
-Excluded are also the {n_uncertain_bioeconomy_visions} innovations whose source text did not allow labeling without significant uncertaintiy.
-These innovations are planned to undergo a revision by subject matter experts at a later point."
-    )
-)
 ```
 
+For the following part, innovations which do not relate to the forest are excluded.
+Excluded are also the `{python} n_uncertain_bioeconomy_visions` innovations whose source text did not allow labeling without significant uncertaintiy.
+These innovations are planned to undergo a revision by subject matter experts at a later point.
+
 ```{python}
-# | label: Number of Results
-# | output: asis
+#| label: Number of Results
+#| output: asis
 
@@
 ```{python}
-# | label: Count Recycling Innovation
+#| label: Count Recycling Innovation
 
@@
 ```{python}
-# | label: tbl-count_bioeconomy_visions
-# | tbl-cap: "Count of Bioeconomy Visions"
-# | output: asis
-# | include: false
+#| label: tbl-count_bioeconomy_visions
+#| tbl-cap: Count of Bioeconomy Visions
+#| output: asis
+#| include: false
 
@@
 
-```{python}
-# | label: fig-bioeconomy-trends
-# | fig-cap: "Count of Bioeconomy Visions by Year"
-
-plot_bv_trends(count_certain_bioeconomy_visions)
-
-```
+![Counts of innovation classification by Bioeconomy Vision Category.
+Innovations can be classified into multiple visions simulatenously.
+The sum of all vision counts does not match the sum of all bioeconomy innovations.](results/figures/vision_count.pdf){#fig-bioeconomy-trends}
 
@@
 
+Within the bioresource vision, two technological development blocks provide the basis for a substantial share of innovations: more efficient harvesting methods through technology and more efficient utilization of timber downstream.
+Write here about the pendelsag, about the forwarders, about the extent to which sustainability was incorporated but only to the extent that it allowed more profitable harvesting of wood.
+
+The role that the 1970s played in the development of core technologies for the bioresource vision can be illustrated in the fact that 7 of the 9 forwarder innovations recorded in SWINNO were commercialized between 1974 and 1982.
+Forwarders are machines used to retrieve felled timber from forest plots before loading them on heavier timber transport options.
+Forwarders replaced forest tractors or unmechanized timber retrival methods such as horse pulled sledges.
+@nordfjell2019TechnicalDevelopmentForwarders show that although they have become heavier, their core design has not changed dramatically.
+
+Perhaps a little bit about the uncertainty and promise that it would be better / not harmful to forest soil.
+
 ```{python}
-# | label: fig-vision-heatmap
-# | fig-caption: "Sector's Contribution to Bioeconomy Visions"
-# | include: false
+#| label: forwarder
+#| include: false
+forwarder_innovation = pd.read_sql(
+    "SELECT sinno_id, year_of_commercialization from innovation where description_in_swedish LIKE '%skotare%'",
+    conn,
+)
 
-# sectors = pd.read_sql(
-#     """
-# SELECT
-#   bv.sinno_id as sinno_id,
-#   c.Category AS vision,
-#   sni_codes.label AS sector
-# FROM
-#   bioeconomy_visions AS bv
-#   JOIN use_sectors ON bv.sinno_id = use_sectors.[sinno_id]
-#   JOIN sni_codes ON SUBSTR(use_sectors.use_sector, 1, 2) = sni_codes.code
-#   JOIN classification_codes AS c ON bv.bioeconomy_vision = c.Code;
-# """,
-#     conn,
-# )
-
-# sectors = sectors.loc[
-#     sectors["sinno_id"].isin(set(certain_bioeconomy_visions["sinno_id"])), :
-# ]
-
-# sectors.loc[sectors["vision"] == "Not Bioeconomy", "vision"] = "Vision Neutral"
-
-# sector_counts = (
-#     sectors.groupby(["sector", "vision"]).size().to_frame("count").reset_index()
-# )
-
-# short_sector_labels = {
-#     "Manufacture of other transport equipment": "Transport Equipment",
-#     "Manufacture of motor vehicles, trailers and semi-trailers": "Motor Vehicle Manufacturing",
-#     "Manufacture of furniture; manufacturing n.e.c.": "Furniture Manufacturing",
-#     "Manufacture of fabricated metal products, except machinery and equipment": "Metal Products Manufacturing",
-#     "Manufacture of machinery and equipment n.e.c.": "Machinery and Equipment",
-#     "Construction": "Construction",
-#     "Manufacture of electrical machinery and apparatus n.e.c.": "Electrical Machinery Manufacturing",
-#     "Manufacture of radio, television and communication equipment and apparatus": "Communication Equipment",
-#     "Manufacture of rubber and plastic products": "Rubber and Plastic Products",
-#     "Forestry, logging and related service activities": "Forestry and Logging",
-#     "Manufacture of wood and of products of wood and cork, except furniture; manufacture of articles of straw and plaiting materials": "Wood and Cork Products",
-#     "Agriculture, hunting and related service activities": "Agriculture and Hunting",
-#     "Land transport; transport via pipelines": "Land Transport",
-#     "Supporting and auxiliary transport activities; activities of travel agencies": "Transport Services",
-#     "Water transport": "Water Transport",
-#     "Air transport": "Air Transport",
-#     "Manufacture of medical, precision and optical instruments, watches and clocks": "Precision Instruments",
-#     "Recycling": "Recycling",
-#     "Manufacture of basic metals": "Basic Metals Manufacturing",
-#     "Public administration and defence; compulsory social security": "Public Administration",
-#     "Electricity, gas, steam and hot water supply": "Utilities",
-#     "Sewage and refuse disposal, sanitation and similar activities": "Waste Management",
-#     "Manufacture of chemicals and chemical products": "Chemical Products",
-#     "Manufacture of pulp, paper and paper products": "Paper Products",
-#     "Manufacture of coke, refined petroleum products and nuclear fuel": "Petroleum Products",
-#     "Collection, purification and distribution of water": "Water Services",
-#     "Publishing, printing and reproduction of recorded media": "Publishing and Media",
-#     "Other mining and quarrying": "Mining and Quarrying",
-#     "Mining of metal ores": "Metal Ore Mining",
-#     "Tanning and dressing of leather; manufacture of luggage, handbags, saddlery, harness and footwear": "Leather and Footwear",
-#     "Manufacture of food products and beverages": "Food and Beverages",
-#     "Manufacture of other non-metallic mineral products": "Non-metallic Products",
-#     "Wholesale trade and commission trade, except of motor vehicles and motorcycles": "Wholesale Trade",
-#     "Other business activities": "Business Services",
-#     "Research and development": "R&D",
-#     "Manufacture of textiles": "Textile Manufacturing",
-#     "Financial intermediation, except insurance and pension funding": "Financial Services",
-#     "Education": "Education",
-#     "Health and social work": "Healthcare",
-#     "Fishing, operation of fish hatcheries and fish farms; service activities incidental to fishing": "Fishing and Aquaculture",
-#     "Recreational, cultural and sporting activities": "Recreation and Culture",
-#     "Extraction of crude petroleum and natural gas; service activities incidental to oil and gas extraction excluding surveying": "Oil and Gas Extraction",
-#     "Post and telecommunications": "Post and Telecom",
-#     "Sale, maintenance and repair of motor vehicles and motorcycles; retail sale of automotive fuel": "Automotive Services",
-#     "Manufacture of wearing apparel; dressing and dyeing of fur": "Apparel Manufacturing",
-# }
-
-# sector_counts = sector_counts.sort_values("count", ascending=False)
-
-# sector_counts["sector"] = sector_counts["sector"].map(short_sector_labels)
-
-# heatmap_data = prepare_heatmap_data(sector_counts, "sector", "vision", "count")
-
-# plot_heatmap(heatmap_data)
-
+forwarder_innovation
 ```
 
+Within bioecology, few innovations aim to decommodify value chains.
+Dominated by technological development block to green the pollution of pulp and paper industry.
+
+In biotechnology, attempts to form a development block around new material such as tencel flaunder.
+
 ### How many eco-innovations are there?
-
-How does each vision contribute to lessening the environmental impact of the forest bioeconomy?
-While a quantitative assessment of technological impact on the environment comes with a host of challenges, the labelling of eco-innovations at least allows a glimpse at what could be expected from each vision.
-@fig-eco-innovation-bioeconomy-vision shows the count of eco-innovation associated with each bioeconomy vision by type.
-As before, there is no weighting, which is why the total sum of all eco-innovation across all bioeconomy visions outnumbers the sum of innovations in the bioeconomy.
-A single innovation can be more than one eco-innovation type, for example, if it reduces the amount of raw material needed by recycling waste.
-The picture of innovation in the forest bioeconomy which @fig-eco-innovation-bioeconomy-vision suggests fits previous literature and what can be expected from LBIO data.
-Product and Process innovation are by far the most frequent across all visions.
-The Bioresource Vision is heavily involved with improving raw-material utilization, while the Bioecology Vision is most frequently associated with recycling.
-At the very least, this suggests that the definitions for each Vision and the classification scheme are in agreement.
-
-A noteworthy feature of innovation belonging to No Bioresource Vision is that this category has by far the most Non-Eco innovations.
-For the sustainability transition of the forest bioeconomy this could imply that having some Vision of a bioeconomy creates more eco-innovation than pursuing other innovation paths.
 
@@
     eco_innovations, on="sinno_id", validate="many_to_many"
+)
+
+eco_innovation_merged_bv["innovation_type_group"] = np.where(
+    eco_innovation_merged_bv["innovation_type_code"].str[0] == "1",
+    "Product",
+    np.where(
+        eco_innovation_merged_bv["innovation_type_code"].str[0] == "2",
+        "Process",
+        np.where(
+            eco_innovation_merged_bv["innovation_type_code"].str[0] == "3",
+            "Organizational",
+            np.where(
+                eco_innovation_merged_bv["innovation_type_code"].str[0] == "4",
+                "Marketing",
+                "No Eco-Innovation",
+            ),
+        ),
+    ),
 )
@@
 ```{python}
-# | label: fig-eco-innovation-bioeconomy-vision
-# | fig-cap: Count of Eco-Innovation Types by Bioeconomy Vision
+# | label: chi-square-vision-eco
+other_sustainability = {501, 701, 108, 109, 111}
+eco_innovation_filtered = eco_innovation_merged_bv[
+    ~(
+        eco_innovation_merged_bv["innovation_type_code"]
+        .astype(int)
+        .isin(other_sustainability)
+    )
+]
+# Create the groups for the cross tab
+eco_innovation_filtered["eco_innovation"] = (
+    eco_innovation_filtered["innovation_type_group"] != "No Eco-Innovation"
+)
+eco_innovation_filtered["vision_neutral"] = (
+    eco_innovation_filtered["bioeconomy_vision"] == "Vision Neutral"
+)
+
+# Calculate the chi squared test
+vision_eco_cross_tab = (
+    eco_innovation_filtered.drop_duplicates(subset="sinno_id")
+    .groupby(["vision_neutral", "eco_innovation"])
+    .agg(count=("sinno_id", "nunique"))
+    .unstack()
+)
+chi2_stat, p_val, dof, expected = stats.chi2_contingency(vision_eco_cross_tab)
+```
+
+How does each vision contribute to lessening the environmental impact of the forest bioeconomy?
+While a quantitative assessment of technological impact on the environment comes with a host of challenges, the labelling of eco-innovations at least allows a glimpse at what could be expected from each vision.
+@fig-eco-innovation-bioeconomy-vision shows the count of eco-innovation associated with each bioeconomy vision by type.
+As before, there is no weighting, which is why the total sum of all eco-innovation across all bioeconomy visions outnumbers the sum of innovations in the bioeconomy.
+A single innovation can be more than one eco-innovation type, for example, if it reduces the amount of raw material needed by recycling waste.
+The picture of innovation in the forest bioeconomy which @fig-eco-innovation-bioeconomy-vision suggests fits previous literature and what can be expected from LBIO data.
+Product and Process innovation are by far the most frequent across all visions.
+The Bioresource Vision is heavily involved with improving raw-material utilization, while the Bioecology Vision is most frequently associated with recycling.
+At the very least, this suggests that the definitions for each Vision and the classification scheme are in agreement.
+
+A noteworthy feature of innovation belonging to No Bioresource Vision is that this category has by far the most Non-Eco innovations.
+The chi-squared test of independence reveals a statistically significant association between an innovation being aligned with a vision and it being an eco-innovation ($\chi^2 =$ `{python} round(chi2_stat, 3)` df = `{python} dof`, p < 0.05).
+For the sustainability transition of the forest bioeconomy this could imply that having some Vision of a bioeconomy creates more eco-innovation than pursuing other innovation paths.
+
+```{python}
+# | label: tbl-cross-table-eco-visions
+# | tbl-cap: Cross Table of Unique Innovations Classified as Vision Neutral And Eco-Innovation
+# Headers
+headers = ["Non-Eco-Innovation", "Eco-Innovation"]
+index = ["Bioeconomy Vision", "Vision Neutral"]
+
+# Generate the table
+cross_table = tabulate(
+    vision_eco_cross_tab.values, headers=headers, showindex=index, tablefmt="pipe"
+)
+
+display(Markdown(cross_table))
+```
+
+```{python}
+#| label: fig-eco-innovation-bioeconomy-vision
+#| fig-cap: Count of Eco-Innovation Types by Bioeconomy Vision
 p = plot_eco_innovation_by_bioeconomy_vision(
@@
 # TODO: perhaps a per cent display by group size would be interesting too?
-# TODO: order y by label groups, color bars
-
-# def barplot_eco_innovation_by_bioeconomy_vision(data):
-
 ```
@@
 <!-- TODO: find company -->
-<!-- , also included a hand drawn charicature of a hiker, who, taking a break on the last stump on a hill and watching the harvester at work, exclaims "Thankfully they left at least one standing". -->
+<!-- , also included a hand drawn charicature of a hiker, who, taking a break on the  last stump on a hill and watching the harvester at work, exclaims "Thankfully they left at least one standing".
+-->
 
@@
 {{< pagebreak >}}
+
 # References {.unnumbered}
@@
 ```{python}
-# | label: tbl-full-codes
-# | tbl-cap: "List of Eco-Innovation Labels"
+#| label: tbl-full-codes
+#| tbl-cap: List of Eco-Innovation Labels
 format_table(codes_df.iloc[:5, :])
-
 ```

### 2024-05-01 paper.qmd
update paper

@@
 Yet, knowledge about innovation produced across the entire system is scarce.
-How innovative is this new economy, especially considering the central role of innovation in transition discourse?
-And what does commercialized innovation suggest for the directionality of the bioeconomy transition?
 The use of detailed innovation data from a Swedish literature-based innovation output database, SWINNO, allows to capture the actual innovation output of the Swedish forest-based bioeconomy innovation system.
 Qualitative coding of the source material links innovation from 1970 to 2021 with visions of a bioresource, biotechnology and / or bioecology bioeconomy.
-Findings reveal a notable decrease in the innovation output of the forest-based bioeconomy relative to Sweden's overall commercialized innovation.
+Findings reveal a decrease in the innovation output of the forest-based bioeconomy relative to Sweden's overall commercialized innovation.
 Notably, innovations associated with a bioresource vision surged during the energy crises of the 1970s, pre-dating their formalization in strategy documents, suggesting a significant role for path dependency in shaping shared visions.
@@
 number-sections: true
+format:
+  pdf:
+    documentclass: scrartcl
+    papersize: a4
+    linestretch: 1.25
+    mainfont: "Adobe Garamond Pro"
+    sansfont: "Adobe Garamond Pro Semibold"
 execute:
@@
   echo: false
-format:
-  pdf:
-    papersize: a4
-    linestretch: 1.25
-    mainfont: "Adobe Garamond Pro"
-    sansfont: "Adobe Garamond Pro Semibold"
-    include-in-header: 
-        - text: \usepackage[a4paper, total={16.5cm,24.7cm}, top=2.5cm, ignorehead, centering, footskip=\footskip+4mm]{geometry}
-        - text: \raggedbottom 
 jupyter: python3
@@
 
+```{python}
+# | label: Total SWINNO Innovation
+
+n_swinno = swinno["sinno_id"].nunique()
+year_min = swinno["year"].min()
+year_max = swinno["year"].max()
+```
+
 # Introduction {#sec-introduction}
@@
 A sustainability transition which has been gaining relevance for researchers, policy makers, and producers alike is the bioeconomy.
-Broadly speaking, it aims to radically shift economic systems from relying on fossil fuels for energy and material resource, to an economic system built on biobased materials, circular and renewable value chains and higher environmental integrity through innovation[@bugge2016WhatBioeconomyReview].
+Broadly speaking, it aims to radically shift economic systems from relying on fossil fuels for energy and material resource, to an economic system built on biobased materials [@bennich2018BioBasedEconomyDynamics], circular and renewable value chains [@damato2017GreenCircularBio] and higher environmental integrity through innovation [@bugge2016WhatBioeconomyReview].
 As one of the main sources for biomass in a bioeconomy [@priefer2017PathwaysShapeBioeconomy], forests have become a key arena for the bioeconomy transition, highlighting tradeoffs between material provision, carbon storage and biodiversity [@hogbom2021TrilemmaNordicBaltic].
 Hence, underneath the bioeconomy umbrella, the concept and viable transition pathways are heavily contested [@allain2022BioeconomicsVsBioeconomy].
-Three different visions of a bioeconomy compete in discourse to provide normative directionality of the transition; the first vision centers on the provision of biomass and renewable resources, the second on the scientific development and application of biotechnology, while the third emphasizes ecological considerations[@bugge2016WhatBioeconomyReview, @vivien2019HijackingBioeconomy].
-The material effects of this transition, however, remain important research topics [@fischer2020SwedishForestsBioeconomy].
-How innovative is this new economy, especially considering the central role of innovation in transition discourse?
+Three different visions of a bioeconomy compete in discourse to provide normative directionality of the transition; the first vision centers on the provision of biomass and renewable resources, the second on the scientific development and application of biotechnology, while the third emphasizes ecological considerations[@bugge2016WhatBioeconomyReview; @vivien2019HijackingBioeconomy].
+While the discourse is well understood in terms of content [e.g., @bugge2016WhatBioeconomyReview, @vivien2019HijackingBioeconomy] its narratives [@bauer2018NarrativesBiorefineryInnovation], and actors [@holmgren2022WhoseTransformationThis], the material effects of this transition, however, remain important research topics [@fischer2020SwedishForestsBioeconomy].
+How many innovations are commercialized in this new economy, especially considering the central role of innovation in transition discourse?
 And what does commercialized innovation suggest for the positive directionality of the bioeconomy transition?
@@
 -->
-Rigorously operationalizing directionality is challening due to a plurality of possible development paths [@stirling2011PluralisingProgressIntegrative].
-Stirling argues that innovation is better understood as a vector of possibilities rather than a scalar decision between clearly defined alternatives.
-<!-- Lastly, Stirling (2011, 2009, 2008) is a clear voice pointing out a slightly different implication of directionality.
-Arguably, he introduced the concept to the field of sustainability transitions research, as part of broader critique of one-dimensional growth ideology.
-In line with a multidimensional view on sociotechnical change, he stresses that innovation is better viewed as “a vector, rather than just a scalar quantity” (Stirling, 2011, p.
-84, 2008, p.
-263).
-But he goes further and makes a passionate and eloquent call for a more pluralistic perspective on technological innovation.
+<!-- • Paragraph 5: This Paper. This paragraph states in a nutshell what the paper accomplishes and how. -->
+<!-- • Paragraphs 6-7: Model. Summarize the key formal assumptions you will maintain in your analysis. -->
+<!-- • Paragraphs 8-9: Data. Explain where you obtain your data and how you measure the concepts that are central to your study. -->
+<!-- • Paragraphs 10-11: Methods. Explain how you take your model to the data and how you overcome the challenges you raised in paragraphs 3-4. -->
 
-@andersson2021OutcomesDirectionalityMorphology 111 -->
-
-<!-- 1 The notion of directionality was introduced by Stirling (2008, 2009) as part of his call for opening up the process of technical change for alternative options.
-He builds on the broader idea that socio-technical change has a direction, choices are made between directions and actors gradually become blind to alternatives, which is a central tenet of much of the innovation studies literature.
-Our notion of directionality also draws on this idea (see also Weber and Rohracher, 2012: 1042–1043).
-@schot2018DeepTransitionsEmergence -->
-
-<!-- • Paragraph 5: This Paper. This paragraph states in a nutshell what the paper accomplishes and how. -->
-This paper adresses this issue by using a unique database of significant Swedish innnovations to examine trends in the transition to a forest based bioeconomy.
-Technical source documents are qualitatively coded regarding their alignment with different visions of a bioeconomy.
-The result paints a detailed picture of activities to achieve competing visions and through narrative analysis highlights how development blocks enable and prevent certain transition directions.
-
-<!-- • Paragraphs 6-7: Model. Summarize the key formal assumptions you will maintain in your analysis. -->
-The core assumption needed for this paper is that innovations can be aligned with visions of transitions without requiring that the vision is already explicitly formalized.
-
-<!-- • Paragraphs 8-9: Data. Explain where you obtain your data and how you measure the concepts that are central to your study. -->
-Bioeconomy innovations are identified through SNI codes of user and producer sectors as well as keywords in innovation descriptions from the SWINNO database.
-SWINNO consists of technical newspaper articles which report on innovations for a professional audience and hence offer high quality data on the type, function, construction and aims of or motivation behind innovations.
-Because these articles are published in edited professional journals they represent innovations of significance for the market and hence accurately capture the direction of search deemed most promising by the innovation system.
-
-<!-- • Paragraphs 10-11: Methods. Explain how you take your model to the data and how you overcome the challenges you raised in paragraphs 3-4. -->
-The inclusion of an innovation is based on the value chain of the forest based bioeconomy [@wolfslehner2016ForestBioeconomyNew].
-I develop a classification scheme build on visions of bioeconomy transition outcomes and eco-innovation types and apply it to the database sources.
-This allows me to analyze innovation as a vector rather than a scalar as called for by @stirling2010MulticriteriaDiversityAnalysis.
+To unpack the plurality of possible socio-technical development paths, places high demands on the data quality of innovation measurement.
+Although progress towards monitoring the bioeconomy transition has been made [@jander2019MonitoringTransitionBioeconomy; @jander2020MonitoringBioeconomyTransitions], assessing directionality is limited by a lack of innovation output measures [@wydra2020MeasuringInnovationBioeconomy].
+In this paper, I use SWINNO, a literature-based innovation output database capturing all significant commercialized innovation in Sweden [@sjoo2014DatabaseSwedishInnovations], to analyze the innovation output of the Swedish forest-based bioeconomy.
+I develop and apply a classification scheme to identify which bioeconomy vision an innovation produced or used in forestry or the downstream value chain aligns with.
 
@@
 tion in paragraphs 1-2. -->
-I find that the Swedish bioeconomy innovation output has declined over the last 50 years.
-Of the three different visions the bioresource vision is the most pronounced, in line with expecatations from discourse analysis and the Swedish bioeconomy strategy.
-I show, however, also that the strategy should not be seen as visionary, as the innovation system had been producing bioeconomy innovation and even bioresource aligned innovation long before the strategy was formulated.
+Results show that Swedish bioeconomy innovation output has declined over the last 50 years.
+Of the three different visions the bioresource vision is the most pronounced, in line with expectations from discourse analysis and the Swedish bioeconomy strategy.
+However, the strategy should not be seen as visionary, as the innovation system had been producing bioeconomy innovation and even bioresource aligned innovation from the 1970s onwards.
 This points to a strong interconnection between path dependence and the creation of directionality in the policy space.
-Suprisingly, I find that the majority of innovations related to the bioeconomy cannot be classified as aligning with one of the three visions.
-Instead they represent general improvements to machinery and tools used in various segments of the forest based bioeconomy.
-
-The bioresource vision is rooted in a development block to respond to landscape pressure of energy scarcity during the 1970s oil crises.
-Examples of a strong bioecological vision of high value added products with regional identity exists, but lack coordination and remain in the niche.
 
@@
 -->
-This paper adds to the literature on directionality by providing a deep insight into the direction of search for an ongoing sustainability transition.
+
+<!-- By analyzing the innovations produced by the forest-bioeconomy system in Sweden, this paper adds to existing literature on directionality in sustainability transitions and bioeconomy literature in two main ways.
+On one hand, using literature covering innovation output can be used to assess the directionality a sociotechnical system  -->
+
 <!-- From an empirical perspective, this is essentially about furthering the current understanding of how sociotechnical systems grow as a result of historical path dependencies, system-internal dynamics and external influence, towards a more elaborate understanding of how these mechanisms influence how system configurations change over time.
@@
 
-It also adds to the literature about bioeconomy transitions specifically by measuring progress towards different visions of a bioeconomy for the first time using innovations produced by the Swedish forest based bioeconomy.
+# Directionality and Visions of Bioeconomy {#sec-literature}
 
-# Directionality, Direction and the Transition to a Bioeconomy {#sec-literature}
+Previous literature has investigated directionality of socio-technical systems and emphasized the importance of generating shared visions as tools to provide normative directionality.
+Three distinct visions of desirable transition pathways and outcomes compete for normative directionality of the bioeconomy transition.
+This section introduces key concepts from this literature.
 
-This section reviews the literature on transitions to a bioeconomy, introduces competing visions of desired outcomes of this transition, central actors in the Swedish discourse and key concepts relating to the governance of sustainability transitions such as the transition to a bioeconomy, namely directionality and the direction of search within innovation systems.
-
-## Clarification of Bioeconomy Concept
-
-Although the bioecononmy has gained importance as a research, policy and industry focus, it remains unclear what exactly it entails.
-Broadly speaking, the bioeconomy refers to an economy in which raw materials for energy and production processes are provided from biological instead of fossil sources.
-<!-- TODO: citation -->
-Hopes of increased resource efficiency and renewability as well as promised economic growth especially in regions rich in biological resources abound [@el-chichakli2016PolicyFiveCornerstones].
-However, a such a profound change to economic systems will not be possible without balancing significant tradeoffs between the needs and desires of different human and non-human stakeholders [@dieken2021MultitudesBioeconomiesSystematic].
-@hogbom2021TrilemmaNordicBaltic summarize these tradeoffs as a trilemma between forests ability to provide sufficent wood to replace fossil based materials, carbon sequestration and storage, and the delivery of other eco-system services such as forest biodiversity.
-Resolving these tradeoffs is challenging, as goal conflicts between different development pathways make simulatneous progress on social, ecological and economic goals difficult [@bennich2021BiobasedEconomy2030].
-
-As one of the leading forestry nations in the world, Sweden sees substantial economic and political potential in advancing the bioeconomy .
-In a report commissioned by the Swedish Government, the @2012StrategyBiobasedEconomy lays out the plan for a Swedish bioeconomy based on two central pillars.
-It is to be an economy based on,
-
-> \[a\] sustainable production of biomass to enable increased use within a number of different sectors of society. The objective is to reduce climate effects and the use of fossil-based raw materials.
-
-With,
-
-> \[a\]n increased added value for biomass materials, concomitant with a reduction in energy consumption and recovery of nutrients and energy as additional end products.
-The objective is to optimize the value and contribution of ecosystem services to the economy.
-(p.
-9).
-
-These pillars reflect the two most influential visions of the bioeconomy concept.
-Notions of a biotechnology and bio-resource oriented bioeconomy are widespread, dominate discourse at various levels [@dieken2021MultitudesBioeconomiesSystematic] and form the main industrial visions [@befort2020GoingDefinitionsUnderstand].
-A third vision, centered around ecology and an economy compatible with the biosphere acts as a counter point.
-
-### Bioeconomic Visions {#sec-bioeconomy-visions}
-
-The application of biotechnology to promote economic growth and provide jobs lends the biotechnology vision its name [@bugge2016WhatBioeconomyReview].
-Science driven innovation features prominently in this vision, with collaborations between research institutions in academia and industry taking on an important role [@bugge2016WhatBioeconomyReview].
-@vivien2019HijackingBioeconomy present a review of the concept showing that the focus on science as an absolute motor of change also shaped policy recommendations, driven especially by the OECD, towards increasing the speed and diffusion of innovation through partnerships between new start-ups and establish pharmaceutical companies.
-As a policy narrative this vision was particularly potent and widespread at the turn of the 21st century.
-According to @vivien2019HijackingBioeconomy, the intellectual heritage of Schumpeter and Kontradieff play an important role in this vision.
-Technological breakthroughs are poised to solve economic and ecological challenges, for example through genetic engineering.
-Consequently, this vision has a weak sustainability notion, viewing sustainability challenges as temporary, and nature as substitutable given sufficient research and knowledge creation [@vivien2019HijackingBioeconomy].
-
-In more recent years, visions of a bio-resource bioeconomy have taken over as the leading narrative [@allain2022BioeconomicsVsBioeconomy].
-Unlike the biotechnology vision, this conceptualization does not only focus on economic growth and job creation, but attempts to include ecological sustainability challenges [@bugge2016WhatBioeconomyReview].
-Biological resources, not technologies, form the basis of a new economy.
-In essence, there are two ways in which bio-resources are poised to replace fossil resources: as substitution, e.g., biofuels replacing fossil fuels, and as high value added products, e.g., cross-laminated timber used as an alternative to concrete in construction [@vivien2019HijackingBioeconomy].
-To substitute fossil resources requires significant efforts in innovation.
-In contrast to the biotechnology vision, however, innovation is not confined to scientific advances from a narrow biochemical and pharmaceutical field.
-Instead, new and old actors are required to collaborate between and beyond established sectors [@bugge2016WhatBioeconomyReview].
-Although sustainability plays a more important role in this vision of the bioeconomy, this vision is considered an example of weak sustainability [@damato2017GreenCircularBio; @vivien2019HijackingBioeconomy].
-The reason is the focus of substitution and high biomass demands.
-The consumption of raw materials may even increase, not only for bio-based, but also for fossil-based resources [@asada2020BioeconomicTransitionProjecting].
-
-The exact role played by forests in a bioeconomy transition depends on the desired vision [@kleinschmit2014ShadesGreenSocial].
-There is no doubt that the Swedish vision places high demands on forests as primary providers of the required biomass.
-Or, as @belandlindahl2015SwedishForestryModel call it a forest sector that operates under a "more of everything" paradigm.
-Both in the strategy formulated in @research2012StrategyBiobasedEconomy and in public discourse, a small, well-connected group of actors center the bio-resource vision as legitimate and desirable [@holmgren2022WhoseTransformationThis].
-
-A paragraph here explaining the detailed findings of @holmgren2022WhoseTransformationThis discourse network here.
-
-Despite -- or perhaps because of -- bioeconomy lacking a strong definition, the term is used to refer to different concepts, with different aims and objectives, foci, normative assumptions and hence implications for society and environment [@bugge2016WhatBioeconomyReview].
-Common ground in ambitions to realize a bioeconomy is the need for diverse types of social, technological and organizational innovation [@el-chichakli2016PolicyFiveCornerstones].
-According to @jankovsky2021InnovationsForestBioeconomy there are four types of innovation crucial for a forest-based bioeconomy: (i) substitute products, (ii) new (bio-based) processes, (iii) new (bio-based) products, (iv) new behavior.
-The need for innovation also features large in Swedish bioeconomy discourse and narratives therein [@bauer2017TechnologicalInnovationSystems, @bennich2021BiobasedEconomy2030], as, "\[t\]echnological innovation is key to a greener future" [@holmgren2022WhoseTransformationThis 42].
-
-For the remainder of the paper the nomenclature of @bugge2016WhatBioeconomyReview is adopted.
-Vivien et al.'s [-@vivien2019HijackingBioeconomy] Type I bioeconomy is taken to correspond with the bioecology vision, Type II and Type III to correspond with the bioresource and biotechnology vision, respectively.
-The analysis of bioeconomy discourse which revealed different visions, highlights that the visions have emerged at different times and with different actors at their centers [@vivien2019HijackingBioeconomy, @bugge2016WhatBioeconomyReview].
-@vivien2019HijackingBioeconomy trace the introduction of the term bioeconomics to the 1920s and a work on fishery by the Russian biologist Baranoff.
-The term bioeoconomics which they label as Type I bioeconomy, was introduced by Georgescu-Roegen, in -@georgescu-roegen1975BioeconomicAspectsEntropy.
-@vivien2019HijackingBioeconomy argue that the term biotechnology gradually emerged between 1990 and 2000, in the wake of enthusiasm about biotechnological advances, such as the discovery of DNA.
-It gained popularity and international importance when the OECD began increasingly supporting and reshaping its meaning at the turn of the millennium [@vivien2019HijackingBioeconomy].
-According to them, the last concept to enter the discourse was the bioresource oriented bioeconomy.
-While biomass has been used throughout history in various ways, the renewed interest with the intention to replace fossil inputs, for example in European Commission communication since 2010.
-However, @vivien2019HijackingBioeconomy point to the long history of biorefinery as a core aspect of this vision, which was discussed as early as the 1930s.
-Hence, while the organization of the discourse into the labels biotechnology, bioecology and bioresource vision is a very recent research outcome, the underlying concepts have a richer history.
-
-## Visions Provide Directionality to Innovation for Transitions
-
-Visions, narratives, mental models and expectations about technological innovation play a central role in social science sustainability research such as the literature on innovation systems, transition studies or economic policy models such as Mission Oriented Innovation Policy.
-
-<!-- ### Plurality of Possible Development Pathways
-Not infinite, as transition already embedded in time and space [@andersson2021OutcomesDirectionalityMorphology], but more than scalar @stirling2010MulticriteriaDiversityAnalysis, @stirling2011PluralisingProgressIntegrative, @bergek2015TechnologicalInnovationSystems -->
-
-### Directionality {#sec-directionality}
+## Directionality {#sec-directionality}
 
@@
 
-Two concepts are central to understanding how innovation is supposed to be guided based on normative decisions between different feasible paths, both serve as the background concepts in this and many related studies on sustainability transitions.
-The multilevel perspective (MLP) introduced by [@geels2007TypologySociotechnicalTransition, @geels2010OntologiesSociotechnicalTransitions] and the literature on innovation systems at their diverse levels of analysis such as technological, regional, or national levels.
-<!-- TODO add citations. -->
 @hekkert2007FunctionsInnovationSystems provide a framework of intervention in innovation on a systemic level, which highlights the importance of guidance of search as one of seven functions innovation system must provide.
@@
 
+## Clarification of Bioeconomy Concept
+
+Although the bioecononmy has gained importance as a research, policy and industry focus, it remains unclear what exactly it entails.
+Broadly speaking, the bioeconomy refers to an economy in which raw materials for energy and production processes are provided from biological instead of fossil sources.
+<!-- TODO: citation -->
+Hopes of increased resource efficiency and renewability as well as promised economic growth especially in regions rich in biological resources abound [@el-chichakli2016PolicyFiveCornerstones].
+However, a such a profound change to economic systems will not be possible without balancing significant tradeoffs between the needs and desires of different human and non-human stakeholders [@dieken2021MultitudesBioeconomiesSystematic].
+@hogbom2021TrilemmaNordicBaltic summarize these tradeoffs as a trilemma between forests ability to provide sufficent wood to replace fossil based materials, carbon sequestration and storage, and the delivery of other eco-system services such as forest biodiversity.
+Resolving these tradeoffs is challenging, as goal conflicts between different development pathways make simulatneous progress on social, ecological and economic goals difficult [@bennich2021BiobasedEconomy2030].
+
+As one of the leading forestry nations in the world, Sweden sees substantial economic and political potential in advancing the bioeconomy .
+In a report commissioned by the Swedish Government, the @2012StrategyBiobasedEconomy lays out the plan for a Swedish bioeconomy based on two central pillars.
+It is to be an economy based on,
+
+> \[a\] sustainable production of biomass to enable increased use within a number of different sectors of society. The objective is to reduce climate effects and the use of fossil-based raw materials.
+
+With,
+
+> \[a\]n increased added value for biomass materials, concomitant with a reduction in energy consumption and recovery of nutrients and energy as additional end products.
+The objective is to optimize the value and contribution of ecosystem services to the economy.
+(p.
+9).
+
+These pillars reflect the two most influential visions of the bioeconomy concept.
+Notions of a biotechnology and bio-resource oriented bioeconomy are widespread, dominate discourse at various levels [@dieken2021MultitudesBioeconomiesSystematic] and form the main industrial visions [@befort2020GoingDefinitionsUnderstand].
+A third vision, centered around ecology and an economy compatible with the biosphere acts as a counter point.
+
+### Bioeconomic Visions {#sec-bioeconomy-visions}
+
+The application of biotechnology to promote economic growth and provide jobs lends the biotechnology vision its name [@bugge2016WhatBioeconomyReview].
+Science driven innovation features prominently in this vision, with collaborations between research institutions in academia and industry taking on an important role [@bugge2016WhatBioeconomyReview].
+@vivien2019HijackingBioeconomy present a review of the concept showing that the focus on science as an absolute motor of change also shaped policy recommendations, driven especially by the OECD, towards increasing the speed and diffusion of innovation through partnerships between new start-ups and establish pharmaceutical companies.
+As a policy narrative this vision was particularly potent and widespread at the turn of the 21st century.
+According to @vivien2019HijackingBioeconomy, the intellectual heritage of Schumpeter and Kontradieff play an important role in this vision.
+Technological breakthroughs are poised to solve economic and ecological challenges, for example through genetic engineering.
+Consequently, this vision has a weak sustainability notion, viewing sustainability challenges as temporary, and nature as substitutable given sufficient research and knowledge creation [@vivien2019HijackingBioeconomy].
+
+In more recent years, visions of a bio-resource bioeconomy have taken over as the leading narrative [@allain2022BioeconomicsVsBioeconomy].
+Unlike the biotechnology vision, this conceptualization does not only focus on economic growth and job creation, but attempts to include ecological sustainability challenges [@bugge2016WhatBioeconomyReview].
+Biological resources, not technologies, form the basis of a new economy.
+In essence, there are two ways in which bio-resources are poised to replace fossil resources: as substitution, e.g., biofuels replacing fossil fuels, and as high value added products, e.g., cross-laminated timber used as an alternative to concrete in construction [@vivien2019HijackingBioeconomy].
+To substitute fossil resources requires significant efforts in innovation.
+In contrast to the biotechnology vision, however, innovation is not confined to scientific advances from a narrow biochemical and pharmaceutical field.
+Instead, new and old actors are required to collaborate between and beyond established sectors [@bugge2016WhatBioeconomyReview].
+Although sustainability plays a more important role in this vision of the bioeconomy, this vision is considered an example of weak sustainability [@damato2017GreenCircularBio; @vivien2019HijackingBioeconomy].
+The reason is the focus of substitution and high biomass demands.
+The consumption of raw materials may even increase, not only for bio-based, but also for fossil-based resources [@asada2020BioeconomicTransitionProjecting].
+
+The exact role played by forests in a bioeconomy transition depends on the desired vision [@kleinschmit2014ShadesGreenSocial].
+There is no doubt that the Swedish vision places high demands on forests as primary providers of the required biomass.
+Or, as @belandlindahl2015SwedishForestryModel call it a forest sector that operates under a "more of everything" paradigm.
+Both in the strategy formulated in @research2012StrategyBiobasedEconomy and in public discourse, a small, well-connected group of actors center the bio-resource vision as legitimate and desirable [@holmgren2022WhoseTransformationThis].
+
+A paragraph here explaining the detailed findings of @holmgren2022WhoseTransformationThis discourse network here.
+
+Despite -- or perhaps because of -- bioeconomy lacking a strong definition, the term is used to refer to different concepts, with different aims and objectives, foci, normative assumptions and hence implications for society and environment [@bugge2016WhatBioeconomyReview].
+Common ground in ambitions to realize a bioeconomy is the need for diverse types of social, technological and organizational innovation [@el-chichakli2016PolicyFiveCornerstones].
+According to @jankovsky2021InnovationsForestBioeconomy there are four types of innovation crucial for a forest-based bioeconomy: (i) substitute products, (ii) new (bio-based) processes, (iii) new (bio-based) products, (iv) new behavior.
+The need for innovation also features large in Swedish bioeconomy discourse and narratives therein [@bauer2017TechnologicalInnovationSystems, @bennich2021BiobasedEconomy2030], as, "\[t\]echnological innovation is key to a greener future" [@holmgren2022WhoseTransformationThis 42].
+
+For the remainder of the paper the nomenclature of @bugge2016WhatBioeconomyReview is adopted.
+Vivien et al.'s [-@vivien2019HijackingBioeconomy] Type I bioeconomy is taken to correspond with the bioecology vision, Type II and Type III to correspond with the bioresource and biotechnology vision, respectively.
+The analysis of bioeconomy discourse which revealed different visions, highlights that the visions have emerged at different times and with different actors at their centers [@vivien2019HijackingBioeconomy, @bugge2016WhatBioeconomyReview].
+@vivien2019HijackingBioeconomy trace the introduction of the term bioeconomics to the 1920s and a work on fishery by the Russian biologist Baranoff.
+The term bioeoconomics which they label as Type I bioeconomy, was introduced by Georgescu-Roegen, in -@georgescu-roegen1975BioeconomicAspectsEntropy.
+@vivien2019HijackingBioeconomy argue that the term biotechnology gradually emerged between 1990 and 2000, in the wake of enthusiasm about biotechnological advances, such as the discovery of DNA.
+It gained popularity and international importance when the OECD began increasingly supporting and reshaping its meaning at the turn of the millennium [@vivien2019HijackingBioeconomy].
+According to them, the last concept to enter the discourse was the bioresource oriented bioeconomy.
+While biomass has been used throughout history in various ways, the renewed interest with the intention to replace fossil inputs, for example in European Commission communication since 2010.
+However, @vivien2019HijackingBioeconomy point to the long history of biorefinery as a core aspect of this vision, which was discussed as early as the 1930s.
+Hence, while the organization of the discourse into the labels biotechnology, bioecology and bioresource vision is a very recent research outcome, the underlying concepts have a richer history.
+
 # Data and Methods {#sec-data_methods}
 
-To capture the actual innovation output associated with the Swedish forest based bioeconomy the SWINNO database was searched for innovation associated with core forest sectors and keywords describing forest products which may be used in sectors along the bioeconomy value chain.
+To quantify transition directionality, data about innovation activity is needed.
+Previous research has frequently relied on input and throughput indicators of innovation activities, such as R&D expenditure, or  publications and patents [@wydra2020MeasuringInnovationBioeconomy].
+Such data, however, suffers from limitations when attempting to assess the directionality of socio-technical systems.
+While R&D expenditure is a key lever to exercise normative directionality for policy makers, the uncertainty of innovation requires portfolio diversification to prevent misallocation of public and private resources.
+Patents on the other hand, suffer from well documented measurement errors; not all innovations are patented, and not all patents become innovations [@arundel1998WhatPercentageInnovations].
+
+Literature-based innovation output measures remedy these issues, by capturing commercialized innovations [@vanderpanne2007IssuesMeasuringInnovation].
+Constructed with this method in Sweden, the SWINNO database contains `{python} n_swinno` innovations commercialized by Swedish companies between `{python} year_min` and `{python} year_max` [@sjoo2014DatabaseSwedishInnovations].
+SWINNO is constructed from editor-reviewed articles published in 15 independent trade journals covering technological developments for general and highly specialized audiences [@sjoo2014DatabaseSwedishInnovations].
+Since this method relies on the judgment of editiors and journalists, it can be described as significant innovations, omitting small reconfigurations of exisiting products and processes.
+The SWINNO database has significant benefits compared to other innovation measures.
+For one, it ensures that actual innovation, meaning commercialized products and processes, and thus that its positive directionality is captured.
+Second, because the source articles are preserved, it provides a rich source of data, giving frequently providing insights into the aims, motivations, advantages and challenges of the covered innovation, allowing to assess and quantify the vector of technological change.
+
+To capture the forest-based bioeconomy the database was searched for innovation associated with core forest sectors and keywords describing forest products which may be used in sectors along the bioeconomy value chain based on @wolfslehner2016ForestBioeconomyNew.
 The list of keywords and sectors is presented in the appendix.
+`{python} n_total_query` innovation were identified from the query. 
 Due to the query matching keywords whenever they appear as substrings in the description of the innovation, the initial search result is rather susceptible to false positives.
-In later manual review of each innovation for classification these false positives were removed from the search results.
-
-```{python}
-#| label: Total SWINNO Innovation
-#| output: asis
-
-n_swinno = swinno["sinno_id"].nunique()
-year_min = swinno["year"].min()
-year_max = swinno["year"].max()
-```
-
-The SWINNO database used here contains `{python} n_swinno` innovations commercialized between `{python} year_min` and `{python} year_max`.
-It is constructed according to  literature-based innovation output (LBIO) method using Swedish source material [@sjoo2014DatabaseSwedishInnovations].
-This method relies on the judgment of editiors and journalists at trade journals and hence captures what can be described as significant innovation.
-
-But, measuring innovation with LBIO has also substantial advantages.
-For one, it ensures that actual innovation, meaning commercialized products and processes, are captured.
-Second, because the source articles are preserved, it provides a rich source of data, giving frequently providing insights into the aims, motivations, advantages and challenges of the covered innovation.
+In later manual review of each innovation for classification these false positives were removed from the search results, yielding `{python} n_bioeconomy_innovation` innovations.
 
@@
 This means that one innovation can be aligned with more than one vision at a time, for example, if it uses biological processes such as bacteria (biotechnology) to refine a wood based material to replace a fossil input elsewhere in the economy (bioresource vision).
-Not all innovations fullfill at least one criteria, yet still originate or are used within the analytical boundaries of the forest based bioeconomy.
+Not all innovations fulfill at least one criteria, yet still originate or are used within the analytical boundaries of the forest based bioeconomy.
 I classify these innovations as Vision Neutral.
@@
 
-@tbl-vision-examples shows five examples of classified innovation.
 This selection also highlights that multiple labels can be attached to one innovation.
@@
 These innovations are labeled as Vision Neutral, which does not imply that the innovation is not used, or originates in the forest based bioeconomy as defined above.
-An illustrative example of such an innovation are improved work boots, which promote safety of forest workers but do not correspond with any of the criteria in @tbl-bugge or @tbl-vivien.
+An illustrative example of such an innovation are improved work boots, which promote safety of forest workers but do not correspond with any of the criteria in @tbl-decision-matrix.
 

### 2024-05-03 paper.qmd
update text

@@
 Findings reveal a decrease in the innovation output of the forest-based bioeconomy relative to Sweden's overall commercialized innovation.
-Notably, innovations associated with a bioresource vision surged during the energy crises of the 1970s, pre-dating their formalization in strategy documents, suggesting a significant role for path dependency in shaping shared visions.
+Notably, innovations associated with a bioresource vision surged during the energy crises of the 1970s, pre-dating their formalization in strategy documents, suggesting a significant role for path-dependency in shaping shared visions.
 Concurrently, the presence of innovations aligning with multiple visions indicates an open opportunity space for future development across various directions.
@@
 raw_data = here("data/raw")
+uncleaned_bioeconomy_query = pd.read_csv(raw_data / "uncleaned_bioeconomy_query.csv")
 bioeconomy = pd.read_csv(raw_data / "id_year_bioeconomy.csv")
@@
 ```{python}
-# | label: Total SWINNO Innovation
-
+# | label: Innovation Counts
 n_swinno = swinno["sinno_id"].nunique()
@@
 year_max = swinno["year"].max()
+
+n_uncleaned_bioeconomy_innovations = uncleaned_bioeconomy_query["sinno_id"].nunique()
+sum_bioeconomy_innovations = swinno["bioeconomy"].sum()
 ```
@@
 However, the strategy should not be seen as visionary, as the innovation system had been producing bioeconomy innovation and even bioresource aligned innovation from the 1970s onwards.
-This points to a strong interconnection between path dependence and the creation of directionality in the policy space.
+This points to a strong interconnection between path-dependence and the creation of directionality in the policy space.
 
@@
 
-<!-- From an empirical perspective, this is essentially about furthering the current understanding of how sociotechnical systems grow as a result of historical path dependencies, system-internal dynamics and external influence, towards a more elaborate understanding of how these mechanisms influence how system configurations change over time.
+<!-- From an empirical perspective, this is essentially about furthering the current understanding of how sociotechnical systems grow as a result of historical path-dependencies, system-internal dynamics and external influence, towards a more elaborate understanding of how these mechanisms influence how system configurations change over time.
 From a conceptual perspective, it is rather about enabling such investigations, by developing analytical frameworks that capture not only the pace, but also the directionality of innovation processes.
@@
 
+Path-dependency arises from events at one point in time affect later events, and may even lead to suboptimal outcomes [@david1985ClioEconomicsQWERTY].
+It arises due to positive feedback effects, such as increasing returns [@arthur1989CompetingTechnologiesIncreasing], or through extension of possibilities adjecent to exiting technology [@kauffman2000Investigations].
+Hence, technological change is contingent and influences technological change at previous or later points in time.
+Instead of more or less progress, each innovation implies a new direction for socio-technical systems [@stirling2011PluralisingProgressIntegrative].
+Andersson et al., [-@andersson2021OutcomesDirectionalityMorphology] term this quality of innovation possessing direction _positive_ directionality.
+
 As innovation policy took a normative turn under which innovation is needed to solve grand societal challenges [@schot2018ThreeFramesInnovation], the question of how to steer innovation into desired directions became a central issue in innovation research [@kohler2019AgendaSustainabilityTransitions].
-Innovation turned from a scalar concept, where the central questions where how to promote the right amount of innovation, into a question of influencing the vector in which society develops through innovation [@stirling2011PluralisingProgressIntegrative].
-But, as innovation systems are embedded in a spatial and temporal context they tend to develop along trajectories characterized by path dependency [@kemp2022DiversifyingDeepTransitions, @bergek2015TechnologicalInnovationSystems].
+Innovation policy now ought to promote the desired vector of societal development, instead of focusing on providing the right amount of progress [@schot2018ThreeFramesInnovation], lending innovation _normative_ directionality [@andersson2021OutcomesDirectionalityMorphology].
+As innovation systems are embedded in a spatial and temporal context, not all potentially desired pathways are available or feasible.
+Instead, socio-technical systems are constrained by path-dependency [@kemp2022DiversifyingDeepTransitions;@bergek2015TechnologicalInnovationSystems; @stirling2011PluralisingProgressIntegrative], sometimes to the point of being locked-in to specific system states [@unruh2000UnderstandingCarbonLockin].
 
-@hekkert2007FunctionsInnovationSystems provide a framework of intervention in innovation on a systemic level, which highlights the importance of guidance of search as one of seven functions innovation system must provide.
-They explain guidance of search as, "those activities within the innovation system that can positively affect the visibility and clarity of specific wants among technology users" (423).
-It is changing preferences in society, which through market and government influence and, "..
-an interactive and cumulative process of exchanging ideas between technology producers, technology users, and many other actors" (423) decides which direction of technological development is most promising.
-@bauer2017TechnologicalInnovationSystems adapts the innovation system functions to the specific case of biorefineries in Sweden and define the influence on the direction of search as "the incentives and/or pressures for organizations to enter the technological field.
-These may come in the form of visions, expectations of growth potential, regulation, policy targets, standards, articulation of demand from leading customers, crises in current business, etc." (4).
-
-<!-- 
-The most accepted view in sustainability transitions studies is that system change is driven by a combination of exogenous and endogenous driving forces.
-Moreover, the external shocks ‘do not mechanically impact niches and regimes, but need to be perceived and translated by actors to exert influence’ (Geels & Schot, 2007, p.
-404).
-This implies that the actual directions of change are shaped by strategic action (Grillitsch et al., 2018; Markard & Truffer, 2008; Pacheco et al., 2010; Smith et al., 2010; Yap & Truffer, 2019).
-Although there are many studies on the role of agency in shaping sustainability transitions, there is still limited understanding about how this agency shapes the directionality of the change process.
-@yang2022ShapingDirectionalitySustainability [752]
--->
-
-<!-- 
-The directionality battle is not about whether the new (niche actors) will win over the old (regime actors).
-In our research we do not want to tie regime actors upfront to a strategy of maintaining institutions (defending the regime) while niche actors do the creating (building niches) and disrupting work (destabilizing regimes).
-Rather, battles about the actual course of action may happen equally among regime actors within a prevailing regime as among actors supporting (potentially manifold) niches.
-Such a view accounts for a situation in which regime actors may operate in the niche and have an interest in promoting niches, while niche actors may not want to destroy the regime and prefer to operate at the niche level only.
-In the end, the unfolding directionality will be the result of interaction among actors with different degrees of incumbency (Jørgensen, 2012; Yap & Truffer, 2019).
-The fact that regime actors are not just defending the status quo has also been recognized in neo-institutional literature.
-The seminal work of Leblebici et al.
-(1991) emphasized that internal institutional contradictions may emerge as a starting point for dominant actors to engage with institutional change.
-In transition studies, Fünfschilling and Truffer (2014) elaborated how different institutional logics in a regime may create tensions within and among actors who are incumbents in the prevailing regime.
-We have therefore to account for a multiplicity of institutional work strategies of a multitude of actors, which are more or less tied to the prevailing regime structures.
-
-@yang2022ShapingDirectionalitySustainability [754] -->
-
-<!-- Our analysis supports that in the context of transformative industrial change, companies’ innovation activities change directionality.
-Companies’ central roles in the RIS still lie in exploiting, combining, and recombining knowledge and turning it into innovations (Autio, 1998; Isaksen & Trippl, 2016).
-The influence of the direction of search and legitimacy of change provided by pioneering organizations (Grillitsch & Sotarauta, 2020), such as Paper Province, additionally supports their identification with the environmental debate.
-
-@martin2023ReconsideringActorRoles 1643 -->
-
-@yang2022ShapingDirectionalitySustainability too highlight the importance of visions in directionality over transformations, stressing the importance of shared visions between niche and regime actors.
-They argue that a stretch-and-transform pattern is more likely if niche actors are given a leading role and core regime actors obtain new roles and identites.
-"The objective in the stretching and transforming form of niche empowerment is to convince the wider world that the rules of the game need to be changed.", @smith2012WhatProtectiveSpace [17].
-
-Visions as shared imaginaries, mental models and aspirations,
-narratives as their embodiment in discourse [@bauer2018NarrativesBiorefineryInnovation, @holmgren2022WhoseTransformationThis], technology as their material realization.
-
-@konnola2021TransformativeGovernanceInnovation highlight the importance of shared visions to provide a normative direction and purpose to a system.
-
-@grillitsch2019InnovationPolicySystemwidea illustrate that visions alone are not sufficent to provide clear directionality.
-They must be specific enough and with significant buyin from actors, which may be difficult to obtain with vision statements created elsewhere than within the system.
-
-<!-- Especially, this study indicates that the build-up of shared visions across niche and regime actors is key, and when these shared visions allow for a leading role of niche actors combined with openings for new roles and identities of core regime actors, the emergence of a stretch-and-transform pattern is more likely.
-
-@yang2022ShapingDirectionalitySustainability [766] -->
-
-@weber2012LegitimizingResearchTechnology synthesis policy insights from innovation systems and multilevel perspective research into a failures framework.
+Dealing with the question of how to provide and exercise normative transition directionality @weber2012LegitimizingResearchTechnology synthesis policy insights from innovation systems and multilevel perspective research into a failures framework.
 Therein, directionality failure is understood as the failure to provide direction to the process of search and unclear or missing setting of collective priorities.
@@
 
-Instead of linear view, all levels interact to provide directionality, i.e. influencing the direction of search.
-Other influences on the direction of search exists as well, such as landscape pressures, positive and negative pressures, windows of opportunties etc.
-Top down visions and bottom up pragmatism interact to shape transition dynamics [@bolton2019GrandVisionsPragmatic].
-Operationalizing directionality hence difficult.
-Few notable attempts such as
-@cappellano2022WhereInnovationMeets operationalize directionality as a measure of R&I expenditure earmarked for specific transition objectives.
-Other attempts to measure the directionality and progress towards a bioeconomy @jander2019MonitoringTransitionBioeconomy and @jander2020MonitoringBioeconomyTransitions, but lack the plurality of innovation.
+In the two central research streams of sustainability transitions, the innovation systems and the multi-level perspective literature, the importance of visions is highlighted.
+@hekkert2007FunctionsInnovationSystems provide a framework of intervention in innovation on a systemic level, which highlights the importance of guidance of search as one of seven functions innovation system must provide.
+They explain guidance of search as, "those activities within the innovation system that can positively affect the visibility and clarity of specific wants among technology users" [@hekkert2007FunctionsInnovationSystems, 423].
+It is changing preferences in society, which through market and government influence and, "..
+an interactive and cumulative process of exchanging ideas between technology producers, technology users, and many other actors" [@hekkert2007FunctionsInnovationSystems, 423] decides which direction of technological development is most promising.
+@bauer2017TechnologicalInnovationSystems adapts the innovation system functions to the specific case of biorefineries in Sweden and define the influence on the direction of search as "the incentives and/or pressures for organizations to enter the technological field.
+These may come in the form of visions, expectations of growth potential, regulation, policy targets, standards, articulation of demand from leading customers, crises in current business, etc." (p.
+4).
+Using the multi-level perspective, @yang2022ShapingDirectionalitySustainability too highlight the importance of visions in directionality over transformations, stressing the importance of shared visions between niche and regime actors.
+Similarly, @bolton2019GrandVisionsPragmatic illustrate how visions and pragmatic change combined in the development and evolution of a highly integrated European electricity system.
 
-<!-- As Stirling (2011) emphasizes, transition pathways are contingent.
-They involve contestations between different worldviews and interests.
-Instead of focusing entirely on the cascades of innovation that converge into clear-cut and singular regime shifts, transition analysis should remain attentive to the multiplicity of directions that a transition process may take.
-To be sure, the DT frameworkworks with an evolutionary logic that is far from deterministic.
-It addresses directionality explicitly.
-Schot & Kanger (2018, p.
-8) consider how the diversity of current socio-technical niches make for undercurrents that contain the seeds for alternative, sustainable and just directions of development.
-Kanger & Schot (2018, pp.
-15-16) describe a ‘portfolio of directionality’ that, throughout subsequent surges and notwithstanding the alignments across sectors, becomes broader over time.
-Along the formation of ‘regimes’ and ‘meta-regimes’, niche alternatives are not considered to just perish or become absorbed.
-Instead, these alternative directions may manifest after periods of hibernation, or they break through locally.
-@kemp2022DiversifyingDeepTransitions [112] -->
+@rotmans2001MoreEvolutionRevolution even go so far to argue that transition management is based on appealing and imaginative long-term visions and @loorbach2010TransitionManagementSustainable argues that shared visions need to be developed and disseminated between transition stakeholders.
+But, as @bolton2019GrandVisionsPragmatic show, visions alone are insufficient to provide directionality.
+In addition to being appealing, visions must also be specific and actors need to buy into them [@grillitsch2019InnovationPolicySystemwide].
+In summary, visions can be seen as shared imaginaries, mental models and aspirations of desired transitions which are expressed through narratives in discourse between actors [@bauer2018NarrativesBiorefineryInnovation; @holmgren2022WhoseTransformationThis].
 
-<!-- However, the DT framework displays a rather reductionist understanding of directionality.
-It reduces the matter to analytical openness, minding the evolutionary multiplicity of a transformation trajectory (e.g.
-towards, or away from, mass production).
-The normative reflection remains tied to a particular direction that a transition process should be taking (K ̈ ohler et al., 2019).
-Yet what is needed is a more fundamentalattending to the multiplicityof possible socio-technical configurations that may emerge,to the different process dynamics that may unfold, and to the different (competing) normative appraisals – by stakeholders and by analysts - of outcomes (Pel et al., 2020).
-Stirling (2009, 2011) underlined the need to unfold evolutionary diversity in a way that clarifies the normative implications of this diversity.To articulate directionality is to increaseevaluative diversity (Stirling, 2008: 268), opening up the wickedness and the ambiguities of transition issues.
-@kemp2022DiversifyingDeepTransitions [113] -->
-
-<!-- Do I need patents to actually show the entire funnel? -->
-
-Recent scholarship has called for advancing research fronts instead of remaining entrenched in contested discourse positions [@allain2022BioeconomicsVsBioeconomy].
-Here is how I propose to measure how the visions have provided directionality to the Swedish Transition to a bioeconomy.
-Directionality as influencing the direction of search.
-I then measure the effectivness of the directionality provided by different visions in actually producing innovation in the direction of search.
-The emergence of visions is at least partially grounded in development paths taken by system actors.
-These then serve as important vehicles to create legitimacy of the chosen paths and further provide directionality for the search.
+However, one transition can be characterized by multiple, competing visions, as is illustrated by the transition to a bioeconomy.
+The next section sharpens the conceptual understanding of what is meant with a bioeconomy and introduces the three visions.
 
@@
 Although the bioecononmy has gained importance as a research, policy and industry focus, it remains unclear what exactly it entails.
-Broadly speaking, the bioeconomy refers to an economy in which raw materials for energy and production processes are provided from biological instead of fossil sources.
-<!-- TODO: citation -->
+At a high level, the bioeconomy refers to an economy in which raw materials for energy and production processes are provided from biological instead of fossil sources [@bugge2016WhatBioeconomyReview].
 Hopes of increased resource efficiency and renewability as well as promised economic growth especially in regions rich in biological resources abound [@el-chichakli2016PolicyFiveCornerstones].
@@
 To quantify transition directionality, data about innovation activity is needed.
+Operationalizing directionality hence difficult.
+Few notable attempts such as
+@cappellano2022WhereInnovationMeets operationalize directionality as a measure of R&I expenditure earmarked for specific transition objectives.
+Other attempts to measure the directionality and progress towards a bioeconomy @jander2019MonitoringTransitionBioeconomy and @jander2020MonitoringBioeconomyTransitions, but lack the plurality of innovation.
 Previous research has frequently relied on input and throughput indicators of innovation activities, such as R&D expenditure, or  publications and patents [@wydra2020MeasuringInnovationBioeconomy].
@@
 The list of keywords and sectors is presented in the appendix.
-`{python} n_total_query` innovation were identified from the query. 
-Due to the query matching keywords whenever they appear as substrings in the description of the innovation, the initial search result is rather susceptible to false positives.
-In later manual review of each innovation for classification these false positives were removed from the search results, yielding `{python} n_bioeconomy_innovation` innovations.
 
@@
 Based on the analyses of competing bioeconomy visions in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy, I develop and apply a matrix of characterstics for each vision.
- Visions based on @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy, biotechnology definition based on @yeung2019CurrentResearchBiotechnology. 
-<!-- TODO: Matrix That matrix is presented in @tbl- -->
+While the criteria for the bioresource and bioecology vision are entirely based on the two previous publications, the biotechnology critieria are additionally based on a literature review of @yeung2019CurrentResearchBiotechnology, concerning biotechnology publications.
+
 Reading the source texts of innovations identified from SWINNO, I classify an innovation as being *aligned* with a vision if it fulfills at least one criteria in the matrix.
@@
 
-This selection also highlights that multiple labels can be attached to one innovation.
-The last innovation, describing a waterproof paper, is classified as a bioresource vision innovation because it describes a process by which biomass is converted and upgraded.
-It is also classified as a biotechnology innovation since, according to the innovation source article, the process mimics trees' microfiber formation.
+An innovation can align with multiple visions, for example, where biomass is converted and upgraded (bioresource vision) by application of biotechnology (biotechnology vision).
+Or the value chain decommodifying introduction of new production lines by a family-business which use wood from their own forests to produce furniture with high territorial identiy supporting explicit ecological conservation goals, that aligns with the bioresource and bioecology vision.
 Lastly, there is also the possibility that an innovation cannot be placed into at least one of the visions.
@@
 
-```{python}
-#| label: Number Bioeconomy Innovation
-#| output: asis
-
-sum_bioeconomy_innovations = swinno["bioeconomy"].sum()
-```
-
-Querying the SWINNO database for forest based bioeconomy innovations initially yields `{python} sum_bioeconomy_innovations` innovations.
+Querying the SWINNO database for forest based bioeconomy innovations initially yields `{python} n_uncleaned_bioeconomy_innovations` innovations.
 This number includes false positives from substrings matching keywords included in the search.
-While there is no indication that these false positives are non-randomly distributed across time and sectors, the following analysis for bioeconomy innovation trends at large needs to be viewed in the light of this being uncleaned data.
-Where the analysis turns to specific bioeconomy visions, only clean and certain cases are discussed.
+During classification, these false positives were removed from the dataset, yielding `{python} sum_bioeconomy_innovations` innovations which originate from, are used in or make use of products derived from the forest.
 
@@
 
+In conclusion, the technological artefacts produced in the transition to a forest-based bioeconomy in Sweden show a clear directionality aligned with the bioresource vision, which is most favored by central actors and policy makers.
+At the same time, there is evidence that the innovation system has produced knowledge aligned with alternative transition pathways towards an ecological bioeconomy, such as decommodification of value chains and efforts to increase the circularity of the system and reduce environmental pollution.
+The temporal distribution of innovation output aligns strongly with the Swedish tendency to produce consens oriented policy by aligning policy objectives with existing directionality.
+While this approach is pragmatic and has produced good outcomes in the past, it is at least doubtful how radical a shift can be expected for the bioeconomy transition, considering that innovation output since the 1970s has not meaningfully explored alternatives to the intensification of biomass resource extraction.
+
 {{< pagebreak >}}
-
 # References {.unnumbered}

### 2024-05-06 experiments/14-plots.qmd
add scripts to produce data and figures

```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(patchwork)
library(showtext)
library(paletteer)
library(rio)
library(here)
library(hrbrthemes)
library(stringr)

showtext_auto()
font_add_google("IBM Plex Sans", "IBMPlexSans")

theme_custom <- function() {
    ret <- hrbrthemes::theme_ipsum_ps(grid = "Y")
    ret <- ret + theme(axis.line.x = element_line(color = "black", linewidth = 0.5))
    ret
}
```

```{r}
bioeconomy_share <- import(here("data/modified/swinno_bioeconomy_share.csv"))
```

```{r}

plot_bioeconomy_stats <- function(df, y, yaxis_title) {
    ggplot(df, aes(x = year, y = .data[[y]])) +
        geom_area(color="black", fill = "gray40", alpha = 0.5, linewidth=1) +
        xlab("Year") +
        ylab(stringr::str_wrap(yaxis_title, width = 20)) +
        theme_custom()
}
```

```{r}
#| label: fig-n-bioeconomy
bioeconomy_plot <- plot_bioeconomy_stats(bioeconomy_share, "ma5_n_bioeconomy", "Forest Bioeconomy")
```

```{r}
#| label: fig-n-innovation

innovation_plot <- plot_bioeconomy_stats(bioeconomy_share, "ma5_n_innovation", "Total SWINNO")
```

```{r}
#| label: fig-share
share_plot <- plot_bioeconomy_stats(bioeconomy_share, "ma5_share_bioeconomy", "Forest Bioeconomy Share") + scale_y_continuous(label = percent)
share_plot
```

```{r}
#| label: fig-trends

innovation_trends_plot  <- (innovation_plot + bioeconomy_plot) / share_plot + plot_annotation(tag_levels = "a") & theme(plot.margin = margin(t = 0, 
                             r = 1,  # Right margin
                             b = 0,  # Bottom margin
                             l = 0))

innovation_trends_plot
ggsave(here("results/figures/innovation_trends.pdf"), width=7, height=5)
```

## Bioeconomy Visions Plot

```{r}
visions <- import(here("data/modified/count_certain_visions.csv"))

vision_summary <- visions |>
    group_by(bioeconomy_vision) |>
    summarize(total_count = sum(count))

visions <- merge(visions, vision_summary, by = "bioeconomy_vision")

```

```{r}
batlow_discrete <- , n = 5)
```

```{r}
paletteer::paletteer_c("scico::batlowW", n = 7)
```
```{r}
#| label: fig-vision-count
vision_count_plot <- ggplot(visions, aes(x = year, y = count, color = bioeconomy_vision, fill = bioeconomy_vision)) +
    geom_col(alpha = 0.9, linewidth = 0.75) +
    facet_wrap(~bioeconomy_vision) +
    scale_color_manual(values = batlow_discrete) +
    scale_fill_manual(values = batlow_discrete) +
    theme_custom() +
    theme(legend.position = "none") +
    scale_y_continuous(breaks = seq(0, max(visions$count), by = 2), expand = c(0, 0)) +
    xlab("Year") +
    ylab("Innovation Count") +
    geom_text(
        data = vision_summary, aes(label = paste("Total:", total_count), x = Inf, y = Inf),
        position = position_nudge(y = 0), hjust = 1, vjust = 1.3, size = 3.5
    )

ggsave(here("results/figures/vision_count.png"), vision_count_plot)
```

```{r}
#| label: fig-visions-stacked-area

# Reorder the levels of bioeconomy_vision based on the sum of counts
visions$bioeconomy_vision <- factor(visions$bioeconomy_vision,
    levels = vision_summary$bioeconomy_vision[order(vision_summary$total_count, decreasing = TRUE)]
)

p <- ggplot(visions, aes(x = year, y = count, fill = bioeconomy_vision, color = bioeconomy_vision)) +
    geom_area(color = "black", alpha = 0.8) +
    scale_fill_manual(values = batlow_discrete, name = "Bioeconomy Vision") + # Set legend title
    scale_color_manual(values = batlow_discrete, guide = "none") + # Hide color legend
    theme_custom() +
    theme(legend.position = c(0.9, 0.9)) +
    scale_y_continuous(expand = c(0, 0)) +
    labs(color = NULL)

p

```

### 2024-05-06 scripts/analysis/00-pull_data.py

@@ -1,0 +1,19 @@
+from src.data_handling import database
+from pyhere import here
+import pandas as pd
+
+if __name__ == "__main__":
+    conn = database.connect_swinno_db()
+    outpath = here("data/raw")
+
+    # Get sinno ids with year and bioeconomy indicator
+    sinno_ids = database.query_sinno_ids(conn)
+    sinno_ids.write_csv(here(outpath, "id_year_bioeconomy.csv"))
+
+    bioeconomy_visions = database.query_bioeconomy_visions(conn)
+    bioeconomy_visions.to_csv(here(outpath / "bioeconomy_visions.csv"), index=False)
+
+    uncleaned_bioeconomy = database.query_uncleaned_bioeconomy(conn)
+    uncleaned_bioeconomy.to_csv(
+        here(outpath / "uncleaned_bioeconomy_query.csv"), index=False
+    )

### 2024-05-06 scripts/analysis/01-create_bioeconomy_share.py

@@ -1,0 +1,26 @@
+from typing import List
+
+import polars as pl
+from pyhere import here
+
+
+def calculate_5yma(data: pl.DataFrame, col: str | List[str]) -> pl.DataFrame:
+    return data.with_columns(pl.col(col).rolling_mean(window_size=5).prefix("ma5_"))
+
+
+if __name__ == "__main__":
+    data_raw = here("data/raw")
+    data_modified = here("data/modified")
+    swinno_bioeconomy = pl.read_csv(data_raw / "id_year_bioeconomy.csv")
+
+    swinno_bioeconomy = (
+        swinno_bioeconomy.group_by("year")
+        .agg(
+            pl.col("bioeconomy").sum().alias("n_bioeconomy"),
+            pl.col("sinno_id").n_unique().alias("n_innovation"),
+        )
+        .with_columns(share_bioeconomy=pl.col("n_bioeconomy") / pl.col("n_innovation"))
+        .sort("year")
+        .pipe(calculate_5yma, col=["n_innovation", "n_bioeconomy", "share_bioeconomy"])
+    )
+    swinno_bioeconomy.write_csv(data_modified / "swinno_bioeconomy_share.csv")

### 2024-05-06 scripts/analysis/02_create_.py

@@ -1,0 +1,51 @@
+bioeconomy_visions = pd.read_sql(
+    """
+SELECT bv.sinno_id
+, codes.Category AS bioeconomy_vision
+, i.year_of_commercialization AS "year"
+FROM bioeconomy_visions AS bv
+JOIN [classification_codes] AS codes ON codes.code = bv.bioeconomy_vision
+JOIN innovation AS i ON i.sinno_id = bv.sinno_id
+WHERE bv.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
+UNION
+SELECT DISTINCT(ei.sinno_id)
+, 'Bioecology Vision' AS bioeconomy_vision
+, i.year_of_commercialization AS "i.year"
+FROM eco_innovations AS ei
+JOIN innovation AS i ON i.sinno_id = ei.sinno_id
+WHERE ei.innovation_type IN (206, 107)
+AND ei.sinno_id NOT IN (
+  SELECT ei2.sinno_id
+  FROM eco_innovations as ei2
+  WHERE ei2.innovation_type IN (999, '000')
+  );
+-- adds recycling innovation to bioecology
+""",
+    conn,
+)
+
+bioeconomy_visions.loc[
+    bioeconomy_visions["bioeconomy_vision"] == "Not Bioeconomy",
+    ["bioeconomy_vision"],
+] = "Vision Neutral"
+
+uncertain_bioeconomy_visions = bioeconomy_visions.query("bioeconomy_vision == 'Unsure'")
+
+
+certain_bioeconomy_visions = bioeconomy_visions.loc[
+    ~bioeconomy_visions["sinno_id"].isin(uncertain_bioeconomy_visions["sinno_id"]), :
+]
+
+count_certain_bioeconomy_visions = (
+    certain_bioeconomy_visions.groupby(["year", "bioeconomy_vision"])
+    .size()
+    .reset_index(name="count")
+)
+
+count_certain_bioeconomy_visions.to_csv("data/modified/count_certain_visions.csv")
+total_certain_bioeconomy_visions = (
+    count_certain_bioeconomy_visions.groupby("bioeconomy_vision")
+    .agg("sum", numeric_only=True)
+    .sort_values("count", ascending=False)
+    .reset_index()
+)

### 2024-05-06 src/data_handling/database.py

@@ -1,0 +1,485 @@
+from sqlalchemy import create_engine
+from pathlib import Path
+import polars as pl
+import pandas as pd
+from sqlalchemy.engine.base import Engine
+
+
+def connect_swinno_db() -> Engine:
+    path_to_db = (
+        Path(
+            "/Users/research/Library/CloudStorage/OneDrive-LundUniversity/research/swinno-db/data/swinno.db"
+        )
+        .absolute()
+        .as_posix()
+    )
+    database_uri = f"sqlite:///{path_to_db}"
+    engine = create_engine(database_uri)
+    return engine
+
+
+def query_innovation_experience(conn: Engine) -> pl.DataFrame:
+    query = """
+    SELECT
+    i.literal_name_of_innovating_firm_sinnos_common_name_sinno_id AS firm,
+    SUM(
+        CASE
+        WHEN bv.bioeconomy_vision = 1 THEN 1
+        ELSE 0
+        END
+    ) AS biotechnology,
+    SUM(
+        CASE
+        WHEN bv.bioeconomy_vision = 2 THEN 1
+        ELSE 0
+        END
+    ) AS bioresources,
+    SUM(
+        CASE
+        WHEN bv.bioeconomy_vision = 3 THEN 1
+        ELSE 0
+        END
+    ) AS bioecology,
+    SUM(
+        CASE
+        WHEN bv.bioeconomy_vision = 9 THEN 1
+        ELSE 0
+        END
+    ) AS not_bioeconomy,
+    SUM(
+        CASE
+        WHEN bv.bioeconomy_vision = 0 THEN 1
+        ELSE 0
+        END
+    ) AS unsure,
+    COUNT(distinct i.sinno_id) as total_innovation_count
+    FROM
+    bioeconomy_visions AS bv
+    JOIN innovation AS i ON i.sinno_id = bv.sinno_id
+    GROUP BY
+    i.literal_name_of_innovating_firm_sinnos_common_name_sinno_id;
+    """
+
+    return pl.read_database(query, conn)
+
+
+def query_isolated_innovation(conn: Engine):
+    conn = conn
+    query = """SELECT
+  DISTINCT(sinno_id),
+  year_of_commercialization AS year,
+  literal_name_of_innovating_firm_sinnos_common_name_sinno_id AS name
+FROM
+  innovation
+WHERE
+  developed_in_collaboration_with_other_actors IS NOT 1"""
+
[…]

### 2024-05-06 src/figures.py

@@ -5,6 +5,7 @@ import seaborn as sns
 import plotly.graph_objects as go
 import plotly.express as px
 import plotly.io as pio
+from cmcrameri import cm
 
 # Set global options for all Plotly plots
 
@@ -266,7 +267,10 @@ def plot_eco_innovation_by_bioeconomy_vision(data, label_lookup):
 
     data["innovation_type_label"] = data["innovation_type_code"].map(label_mapping)
 
-    palette = sns.color_palette("colorblind")
+    palette = cm.batlowS.colors
+
+    # Set font to IBM Plex Sans
+    plt.rcParams["font.family"] = "IBM Plex Sans"
 
     fig = sns.catplot(
         data=data,
@@ -285,6 +289,15 @@ def plot_eco_innovation_by_bioeconomy_vision(data, label_lookup):
             ax.bar_label(c, label_type="edge", padding=2, fontsize=8)
         ax.margins(x=0.1)
         ax.tick_params(axis="y", labelsize=10)
+
+        # Remove spines
+        ax.spines["top"].set_visible(False)
+        ax.spines["right"].set_visible(False)
+        ax.spines["bottom"].set_visible(False)
+        ax.spines["left"].set_visible(False)
+
+        # Add x-grid lines for every tick
+        ax.xaxis.grid(True, linestyle="-")
 
     fig.set_titles("{col_name}")
     fig.set_axis_labels("", "")

### 2024-05-06 paper.qmd
update paper for midway draft

@@
 number-sections: true
+number-depth: 3
 format:
@@
     sansfont: "Adobe Garamond Pro Semibold"
+    indent: true
 execute:
@@
 @andersson2021OutcomesDirectionalityMorphology term these two properties _positive_ and _normative_ directionality, respectively.
-Positive directionality of innovation arises from the fact that technological change is path-dependent [@arthur1989CompetingTechnologiesIncreasing, @david1985ClioEconomicsQWERTY], while normative directionality is rooted in the assumption that without public policy, private actors lack intrinsic incentives to pursue sustainability [@kohler2019AgendaSustainabilityTransitions].
+Positive directionality of innovation arises from the fact that technological change is path-dependent [@arthur1989CompetingTechnologiesIncreasing;@david1985ClioEconomicsQWERTY], while normative directionality is rooted in the assumption that without public policy, private actors lack intrinsic incentives to pursue sustainability [@kohler2019AgendaSustainabilityTransitions].
 
@@
 As one of the main sources for biomass in a bioeconomy [@priefer2017PathwaysShapeBioeconomy], forests have become a key arena for the bioeconomy transition, highlighting tradeoffs between material provision, carbon storage and biodiversity [@hogbom2021TrilemmaNordicBaltic].
-Hence, underneath the bioeconomy umbrella, the concept and viable transition pathways are heavily contested [@allain2022BioeconomicsVsBioeconomy].
-Three different visions of a bioeconomy compete in discourse to provide normative directionality of the transition; the first vision centers on the provision of biomass and renewable resources, the second on the scientific development and application of biotechnology, while the third emphasizes ecological considerations[@bugge2016WhatBioeconomyReview; @vivien2019HijackingBioeconomy].
-While the discourse is well understood in terms of content [e.g., @bugge2016WhatBioeconomyReview, @vivien2019HijackingBioeconomy] its narratives [@bauer2018NarrativesBiorefineryInnovation], and actors [@holmgren2022WhoseTransformationThis], the material effects of this transition, however, remain important research topics [@fischer2020SwedishForestsBioeconomy].
-How many innovations are commercialized in this new economy, especially considering the central role of innovation in transition discourse?
+Underneath the umbrella term "bioeconomy", the concept and viable transition pathways are heavily contested [@allain2022BioeconomicsVsBioeconomy].
+Three different visions of a bioeconomy compete in discourse to provide normative directionality of the transition: the first vision centers on the provision of biomass and renewable resources, the second on the scientific development and application of biotechnology, while the third emphasizes ecological considerations[@bugge2016WhatBioeconomyReview; @vivien2019HijackingBioeconomy].
+Although the discourse is well understood in terms of content [e.g., @bugge2016WhatBioeconomyReview;@vivien2019HijackingBioeconomy] its narratives [@bauer2018NarrativesBiorefineryInnovation], and actors [@holmgren2022WhoseTransformationThis], the material effects of this transition, however, remain important research topics [@fischer2020SwedishForestsBioeconomy].
+How many innovations are commercialized in this new economy, especially considering the central role of innovation in bioeconomy discourse?
 And what does commercialized innovation suggest for the positive directionality of the bioeconomy transition?
@@
 
-To unpack the plurality of possible socio-technical development paths, places high demands on the data quality of innovation measurement.
+To unpack the plurality of possible socio-technical development paths, requires high data quality when measuring innovation.
 Although progress towards monitoring the bioeconomy transition has been made [@jander2019MonitoringTransitionBioeconomy; @jander2020MonitoringBioeconomyTransitions], assessing directionality is limited by a lack of innovation output measures [@wydra2020MeasuringInnovationBioeconomy].
@@
 
-Path-dependency arises from events at one point in time affect later events, and may even lead to suboptimal outcomes [@david1985ClioEconomicsQWERTY].
-It arises due to positive feedback effects, such as increasing returns [@arthur1989CompetingTechnologiesIncreasing], or through extension of possibilities adjecent to exiting technology [@kauffman2000Investigations].
-Hence, technological change is contingent and influences technological change at previous or later points in time.
+Technology does not change randomly.
+Instead, technological change is contingent and influences technological change at previous or later points in time.
+It is path-dependent, meaning events at one point in time affect later events, which may even lead to suboptimal outcomes [@david1985ClioEconomicsQWERTY].
+Path-dependency arises due to positive feedback effects, such as increasing returns [@arthur1989CompetingTechnologiesIncreasing], or through extension of possibilities adjacent to exiting technology [@kauffman2000Investigations].
 Instead of more or less progress, each innovation implies a new direction for socio-technical systems [@stirling2011PluralisingProgressIntegrative].
-Andersson et al., [-@andersson2021OutcomesDirectionalityMorphology] term this quality of innovation possessing direction _positive_ directionality.
+@andersson2021OutcomesDirectionalityMorphology term this quality of innovation possessing direction _positive_ directionality.
 
@@
 
-Dealing with the question of how to provide and exercise normative transition directionality @weber2012LegitimizingResearchTechnology synthesis policy insights from innovation systems and multilevel perspective research into a failures framework.
+Dealing with the question of how to provide and exercise normative transition directionality  Weber and Rohracher [-@weber2012LegitimizingResearchTechnology] synthesize policy insights from innovation systems and multilevel perspective research into a failures framework.
 Therein, directionality failure is understood as the failure to provide direction to the process of search and unclear or missing setting of collective priorities.
 They caution that it should not be confused with shortsighted or flawed anticipation of market developments, as this myopia lacks a view on the direction of change.
-To overcome this failure, that is, to provide directionality, they argue, guiding orientation needs to be translated and intermediated on two levels.
-External requirements placed on the innovation system need to be absorbed after which they need to be interpreted in a process of negotiation between the different actors.
-Reaching this collective understanding of direction, @weber2012LegitimizingResearchTechnology [p. 1042] write, is in practice, "achieved by establishing shared future visions".
+To overcome this failure, that is, to provide directionality, they argue, external requirements need to be internalized by the innovation system in a interpretive process of negotiation between different system actors.
+Reaching this collective understanding of direction, Weber and Rohracher [-@weber2012LegitimizingResearchTechnology p.
+1042] write, is in practice, "...
+achieved by establishing shared future visions".
 
@@
 They explain guidance of search as, "those activities within the innovation system that can positively affect the visibility and clarity of specific wants among technology users" [@hekkert2007FunctionsInnovationSystems, 423].
-It is changing preferences in society, which through market and government influence and, "..
-an interactive and cumulative process of exchanging ideas between technology producers, technology users, and many other actors" [@hekkert2007FunctionsInnovationSystems, 423] decides which direction of technological development is most promising.
+Changing preferences in society, decides which direction of technological development is most promising through market and government influence and, "..
+an interactive and cumulative process of exchanging ideas between technology producers, technology users, and many other actors" [@hekkert2007FunctionsInnovationSystems, 423].
 @bauer2017TechnologicalInnovationSystems adapts the innovation system functions to the specific case of biorefineries in Sweden and define the influence on the direction of search as "the incentives and/or pressures for organizations to enter the technological field.
@@
 But, as @bolton2019GrandVisionsPragmatic show, visions alone are insufficient to provide directionality.
-In addition to being appealing, visions must also be specific and actors need to buy into them [@grillitsch2019InnovationPolicySystemwide].
+In addition to being appealing, visions must also be specific and actors need to buy into them [@grillitsch2019InnovationPolicySystemwidea].
 In summary, visions can be seen as shared imaginaries, mental models and aspirations of desired transitions which are expressed through narratives in discourse between actors [@bauer2018NarrativesBiorefineryInnovation; @holmgren2022WhoseTransformationThis].
@@
 
-Although the bioecononmy has gained importance as a research, policy and industry focus, it remains unclear what exactly it entails.
+Although the bioeconomy has gained importance as a research, policy and industry focus, it remains unclear what exactly it entails.
 At a high level, the bioeconomy refers to an economy in which raw materials for energy and production processes are provided from biological instead of fossil sources [@bugge2016WhatBioeconomyReview].
@@
 As one of the leading forestry nations in the world, Sweden sees substantial economic and political potential in advancing the bioeconomy .
-In a report commissioned by the Swedish Government, the @2012StrategyBiobasedEconomy lays out the plan for a Swedish bioeconomy based on two central pillars.
+In a report commissioned by the Swedish Government, @formas2012SwedishResearchInnovation lays out the plan for a Swedish bioeconomy based on two central pillars.
 It is to be an economy based on,
@@
 
+While the official Swedish bioeconomy strategy is still under development, the content of the quotes above echos across policy documents and the national forest program [@fischer2020SwedishForestsBioeconomy].
 These pillars reflect the two most influential visions of the bioeconomy concept.
@@
 The application of biotechnology to promote economic growth and provide jobs lends the biotechnology vision its name [@bugge2016WhatBioeconomyReview]. 
+Biotechnology is the application of cellular and / or molecular biological agents in the production or processing of materials in a scientific and engineering informed manner [@yeung2019CurrentResearchBiotechnology].
 Science driven innovation features prominently in this vision, with collaborations between research institutions in academia and industry taking on an important role [@bugge2016WhatBioeconomyReview].
@@
 Instead, new and old actors are required to collaborate between and beyond established sectors [@bugge2016WhatBioeconomyReview].
-Although sustainability plays a more important role in this vision of the bioeconomy, this vision is considered an example of weak sustainability [@damato2017GreenCircularBio; @vivien2019HijackingBioeconomy].
+Although sustainability plays a more important role in this vision of the bioeconomy, this vision is also considered an example of weak sustainability [@damato2017GreenCircularBio; @vivien2019HijackingBioeconomy].
 The reason is the focus of substitution and high biomass demands.
@@
 
-The exact role played by forests in a bioeconomy transition depends on the desired vision [@kleinschmit2014ShadesGreenSocial].
-There is no doubt that the Swedish vision places high demands on forests as primary providers of the required biomass.
-Or, as @belandlindahl2015SwedishForestryModel call it a forest sector that operates under a "more of everything" paradigm.
-Both in the strategy formulated in @research2012StrategyBiobasedEconomy and in public discourse, a small, well-connected group of actors center the bio-resource vision as legitimate and desirable [@holmgren2022WhoseTransformationThis].
+An ecology focused counter-vision exists, which @bugge2016WhatBioeconomyReview term the bioecology vision, while ecological economists preferentially use the term bioeconomics [@vivien2019HijackingBioeconomy; @allain2022BioeconomicsVsBioeconomy].
+Unlike the previous two visions, the bioecology vision builds on strong sustainability assumptions and emphasizes the need to develop an alternative economic system which respects the limits of the biosphere.
+Normative change towards material sobriety and degrowth to protect and restore ecosystems and biodiversity take precedent over technological and economic ambitions [@vivien2019HijackingBioeconomy;@bugge2016WhatBioeconomyReview].
 
-A paragraph here explaining the detailed findings of @holmgren2022WhoseTransformationThis discourse network here.
+#### Development of the Term Bioeconomy {#sec-bioeconomy-term-development}
 
-Despite -- or perhaps because of -- bioeconomy lacking a strong definition, the term is used to refer to different concepts, with different aims and objectives, foci, normative assumptions and hence implications for society and environment [@bugge2016WhatBioeconomyReview].
-Common ground in ambitions to realize a bioeconomy is the need for diverse types of social, technological and organizational innovation [@el-chichakli2016PolicyFiveCornerstones].
-According to @jankovsky2021InnovationsForestBioeconomy there are four types of innovation crucial for a forest-based bioeconomy: (i) substitute products, (ii) new (bio-based) processes, (iii) new (bio-based) products, (iv) new behavior.
-The need for innovation also features large in Swedish bioeconomy discourse and narratives therein [@bauer2017TechnologicalInnovationSystems, @bennich2021BiobasedEconomy2030], as, "\[t\]echnological innovation is key to a greener future" [@holmgren2022WhoseTransformationThis 42].
-
-For the remainder of the paper the nomenclature of @bugge2016WhatBioeconomyReview is adopted.
-Vivien et al.'s [-@vivien2019HijackingBioeconomy] Type I bioeconomy is taken to correspond with the bioecology vision, Type II and Type III to correspond with the bioresource and biotechnology vision, respectively.
-The analysis of bioeconomy discourse which revealed different visions, highlights that the visions have emerged at different times and with different actors at their centers [@vivien2019HijackingBioeconomy, @bugge2016WhatBioeconomyReview].
-@vivien2019HijackingBioeconomy trace the introduction of the term bioeconomics to the 1920s and a work on fishery by the Russian biologist Baranoff.
-The term bioeoconomics which they label as Type I bioeconomy, was introduced by Georgescu-Roegen, in -@georgescu-roegen1975BioeconomicAspectsEntropy.
+Analysis of the bioeconomy discourse, highlights that the visions have emerged at different times and with different actors at their centers [@vivien2019HijackingBioeconomy; @bugge2016WhatBioeconomyReview].
+And while the two hitherto mentioned visions dominate contemporary discourse towards a forest based bioeconomy [@holmgren2022WhoseTransformationThis; @bauer2018NarrativesBiorefineryInnovation], the bioeconomics concept itself was introduced in the 1920s in a work on fishery by the Russian biologist Baranoff [@vivien2019HijackingBioeconomy].
+Later, in 1975 ecological economist Georgescu-Roegen [-@georgescu-roegen1975BioeconomicAspectsEntropy] used the term to argue for an economy which minimizes its material and energy throughput to respect biological, physical and geochemical limits.
 @vivien2019HijackingBioeconomy argue that the term biotechnology gradually emerged between 1990 and 2000, in the wake of enthusiasm about biotechnological advances, such as the discovery of DNA.
@@
 However, @vivien2019HijackingBioeconomy point to the long history of biorefinery as a core aspect of this vision, which was discussed as early as the 1930s.
-Hence, while the organization of the discourse into the labels biotechnology, bioecology and bioresource vision is a very recent research outcome, the underlying concepts have a richer history.
+Hence, while the organization of the discourse into the labels biotechnology, bioecology and bioresource vision is a recent development, the underlying concepts have a richer history.
+Over time these visions have not only coexisted, but also influenced each other by borrowing from, contributing to, and distancing themselves from the other visions of a bioeconomy [@allain2022BioeconomicsVsBioeconomy].
+And although the exact role played by forests in a bioeconomy transition depends on the desired vision [@kleinschmit2014ShadesGreenSocial], there is no doubt that the Swedish vision places high demands on forests as primary providers of the required biomass.
+Or, as @belandlindahl2015SwedishForestryModel call it a forest sector that operates under a "more of everything" paradigm.
+Whether in the strategy formulated in @formas2012SwedishResearchInnovation, in Sweden's national forest program [@fischer2020SwedishForestsBioeconomy], or in public discourse, a small, well-connected group of actors center the bio-resource vision as legitimate and desirable [@holmgren2022WhoseTransformationThis].
+
+Common ground in ambitions to realize a bioeconomy is the need for diverse types of social, technological and organizational innovation [@el-chichakli2016PolicyFiveCornerstones].
+According to @jankovsky2021InnovationsForestBioeconomy there are four types of innovation crucial for a forest-based bioeconomy: (i) substitute products, (ii) new (bio-based) processes, (iii) new (bio-based) products, (iv) new behavior.
+The need for innovation also features large in Swedish bioeconomy discourse and narratives therein [@bauer2017TechnologicalInnovationSystems; @bennich2021BiobasedEconomy2030], as, "\[t\]echnological innovation is key to a greener future" [@holmgren2022WhoseTransformationThis 42].
+In a survey study sent to private organizations that received EU funding, @lovric2020FactorsDevelopmentInnovations find that the main type of European forest-based bioeconomy innovations are aimed at production processes, followed by goods and services, then business models.
+
+In summary, despite -- or perhaps because of -- bioeconomy lacking a strong definition, the term is used to refer to different concepts, with different aims and objectives, foci, normative assumptions and hence implications for society and environment [@bugge2016WhatBioeconomyReview].
+The summarized visions express themselves as narratives, actor coalitions and discourse artefacts in attempts to provide normative directionality to the transition.
+This normative directionality is expected to translate into positive directionality of technological change through innovation.
 
@@
 
-To quantify transition directionality, data about innovation activity is needed.
-Operationalizing directionality hence difficult.
-Few notable attempts such as
-@cappellano2022WhereInnovationMeets operationalize directionality as a measure of R&I expenditure earmarked for specific transition objectives.
-Other attempts to measure the directionality and progress towards a bioeconomy @jander2019MonitoringTransitionBioeconomy and @jander2020MonitoringBioeconomyTransitions, but lack the plurality of innovation.
-Previous research has frequently relied on input and throughput indicators of innovation activities, such as R&D expenditure, or  publications and patents [@wydra2020MeasuringInnovationBioeconomy].
-Such data, however, suffers from limitations when attempting to assess the directionality of socio-technical systems.
-While R&D expenditure is a key lever to exercise normative directionality for policy makers, the uncertainty of innovation requires portfolio diversification to prevent misallocation of public and private resources.
-Patents on the other hand, suffer from well documented measurement errors; not all innovations are patented, and not all patents become innovations [@arundel1998WhatPercentageInnovations].
+Operationalizing directionality suffers from frequent lack of differentiation of positive and normative directionality, in addition to inherent issues in the measurement of innovation.
+Notable exceptions are the work on socio-technical system morphology by @andersson2021OutcomesDirectionalityMorphology, who identify changes in the configuration of three different renewable energy systems, and by @yang2022ShapingDirectionalitySustainability who use interviews and secondary source analysis to understand changes to insitutions in solar photovoltaic production in two Chinese provinces.
+However, both studies deal more with the aggregate effects of directionality on socio-technical systems rather than with positive directionality of innovation directly.
+
+Normative directionality can and has been assessed through analysis of discourse [@] and narratives [@bauer2018NarrativesBiorefineryInnovation], visions and aims expressed in strategy documents [@parks2022DirectionalityTransformativeInnovation], and R&D expenditure [@cappellano2022WhereInnovationMeets; @wydra2020MeasuringInnovationBioeconomy].
+Where positive directionality of technological change has been assessed, scientific publications [@wydra2020MeasuringInnovationBioeconomy; @jankovsky2021InnovationsForestBioeconomy], and more often, patents are used [e.g., @aghion2016CarbonTaxesPath;@popp2005LessonsPatentsUsing;@moscona2022DoesDirectedInnovation].
+Such data, however, suffer from limitations when attempting to assess the directionality of socio-technical systems.
+While measures of normative directionality are equipped to identify the desired direction by different actors, visions need to first be accepted by innovators and reformulated into search.
+R&D funding can help in this task, but data on R&D funding is often too broad, targeting entire sectors or specific industries [@popp2005LessonsPatentsUsing].
+While patents are more granular and a closer proxy of technological change, they suffer from well documented measurement errors [@popp2005LessonsPatentsUsing]; not all innovations are patented, and not all patents become innovations [@arundel1998WhatPercentageInnovations].
 
 Literature-based innovation output measures remedy these issues, by capturing commercialized innovations [@vanderpanne2007IssuesMeasuringInnovation].
-Constructed with this method in Sweden, the SWINNO database contains `{python} n_swinno` innovations commercialized by Swedish companies between `{python} year_min` and `{python} year_max` [@sjoo2014DatabaseSwedishInnovations].
-SWINNO is constructed from editor-reviewed articles published in 15 independent trade journals covering technological developments for general and highly specialized audiences [@sjoo2014DatabaseSwedishInnovations].
+Applying this method to  editor-reviewed articles published in 15 independent trade journals covering technological developments for general and highly specialized audiences, the SWINNO database contains `{python} n_swinno` innovations commercialized by Swedish companies between `{python} year_min` and `{python} year_max` [@sjoo2014DatabaseSwedishInnovations].
 Since this method relies on the judgment of editiors and journalists, it can be described as significant innovations, omitting small reconfigurations of exisiting products and processes.
@@
 For one, it ensures that actual innovation, meaning commercialized products and processes, and thus that its positive directionality is captured.
-Second, because the source articles are preserved, it provides a rich source of data, giving frequently providing insights into the aims, motivations, advantages and challenges of the covered innovation, allowing to assess and quantify the vector of technological change.
+Second, because the source articles are preserved, it provides a rich source of data, frequently providing insights into the aims, motivations, advantages and challenges of the covered innovation, in addition to descriptions of what the innovation consists of, which allows to assess and quantify the direction of technological change.
 
 To capture the forest-based bioeconomy the database was searched for innovation associated with core forest sectors and keywords describing forest products which may be used in sectors along the bioeconomy value chain based on @wolfslehner2016ForestBioeconomyNew.
-The list of keywords and sectors is presented in the appendix.
+The list of keywords and sectors is presented in the appendix (@sec-appendix-data), additionally reproducible code to generate the figures and tables in this paper can be found at [`github.com/pjkreutzer/swinno-bioeconomy-directionality`](https://github.com/pjkreutzer/swinno-bioeconomy-directionality).
 
@@
 
-These source articles were carefully read and their associated unique identifiers in the database tagged with two types of labels.
+I carefully read the source articles of innovations identified in the database and tagged their associated unique identifiers with two types of labels.
 The first label is the main interest of this study: the bioeconomy vision an innovation can be associated with.
-The second label provides a deeper understanding if and how the innovation reduces environmental impacts.
-As @broring2020InnovationTypesBioeconomy highlight, not all bioeconomy innovations need to be sustainability oriented innovation.
+The second label provides a deeper understanding if and how the innovation reduces environmental impacts; not all bioeconomy innovations need to be sustainability oriented innovation [@broring2020InnovationTypesBioeconomy].
 
@@
 
-Based on the analyses of competing bioeconomy visions in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy, I develop and apply a matrix of characterstics for each vision.
-While the criteria for the bioresource and bioecology vision are entirely based on the two previous publications, the biotechnology critieria are additionally based on a literature review of @yeung2019CurrentResearchBiotechnology, concerning biotechnology publications.
+Based on the analyses of competing bioeconomy visions in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy, I developed and applied a matrix of characterstics for each vision.
+While the criteria for the bioresource and bioecology vision are entirely based on the two previous publications, the biotechnology criteria are additionally based on a literature review of @yeung2019CurrentResearchBiotechnology, concerning biotechnology publications.
 
-Reading the source texts of innovations identified from SWINNO, I classify an innovation as being *aligned* with a vision if it fulfills at least one criteria in the matrix.
-This means that one innovation can be aligned with more than one vision at a time, for example, if it uses biological processes such as bacteria (biotechnology) to refine a wood based material to replace a fossil input elsewhere in the economy (bioresource vision).
-Not all innovations fulfill at least one criteria, yet still originate or are used within the analytical boundaries of the forest based bioeconomy.
-I classify these innovations as Vision Neutral.
-Examples include innovations aimed at the health and saftey of employees 
-<!-- TODO enter source -->
-, or general improvements to the longevity of materials and tools used somewhere in the value chain of the forest bioeconomy.
-The majority of innovations in this category are of the latter nature.
+Reading the source texts of innovations identified from SWINNO, I classified an innovation as being *aligned* with a vision if it fulfills at least one criteria in the matrix.
+An innovation can align with multiple visions, for example, where biomass is converted and upgraded (bioresource vision) by application of biotechnology (biotechnology vision).
+Or the value chain decommodifying introduction of new production lines by a family-business which use wood from their own forests to produce houses with high territorial identity (*Svensk trävaru- & pappersmassetidning*, 1976:7, pp.
+539,541,544), that aligns with the bioresource and bioecology vision.
+Lastly, there is also the possibility that an innovation cannot be placed into at least one of the visions.
+This is the case for innovation which brings general technological improvements, purely advances economic performance, or relates to a general purpose technology such as appliance of computers.
+These innovations are labeled as Vision Neutral, which does not imply that the innovation is not used, or originates in the forest based bioeconomy as defined above.
+An illustrative example of such an innovation are improved work boots (*Sågverken*, 1974:2, p.
+121), which promote safety of forest workers but do not correspond with any of the criteria in @tbl-decision-matrix.
 
@@
 |                      | Provides method for research in the above areas                                                       |
+|||
 | Bioresource          | Biomass used to substitute other input                                                                |
@@
 |                      | Intensifies use of biomass, including from waste                                                      |
+|||
 | Bioecology           | Prioritizes biodiversity, conservation, ecosystem or soil health                                      |
@@
 
-An innovation can align with multiple visions, for example, where biomass is converted and upgraded (bioresource vision) by application of biotechnology (biotechnology vision).
-Or the value chain decommodifying introduction of new production lines by a family-business which use wood from their own forests to produce furniture with high territorial identiy supporting explicit ecological conservation goals, that aligns with the bioresource and bioecology vision.
-Lastly, there is also the possibility that an innovation cannot be placed into at least one of the visions.
-This is the case for innovation which brings general technological improvements, purely advances economic performance, or relates to a general purpose technology such as appliance of computers.
-These innovations are labeled as Vision Neutral, which does not imply that the innovation is not used, or originates in the forest based bioeconomy as defined above.
-An illustrative example of such an innovation are improved work boots, which promote safety of forest workers but do not correspond with any of the criteria in @tbl-decision-matrix.
-
 ### Eco-Innovation Types {#sec-eco_innovation_types}
@@
 The focus of this paper is understanding what bioeconomy visions dominate the Swedish innovation system's output.
-In order to gain a deeper understanding of the nature of innovation within each bioeconomy vision, however, the innovation were also tagged with labels of eco-innovation types.
+However, in order to gain a deeper understanding of the nature of innovation within each bioeconomy vision, I also tagged the innovation with labels of eco-innovation types.
 This classification uses eco-innovation types identified in a recent literature review by @garcia-granero2018EcoinnovationMeasurementReview.
@@
 While @tbl-full-codes shows all new added labels, in the following text those that do not clearly represent an eco-innovation -- following @garcia-granero2018EcoinnovationMeasurementReview "…fewer adverse effects on the environment and more efficient use of resources" (p.
-305) -- are excluded.
+305) -- were excluded.
 These excluded labels are Gen-manipulated plant, Increased plant yield, Gen-manipulated animal (incl new breed), Human Health and Safety, Automatization, Precision, and Biocide.
 
-The decision process for labelling eco-innovation types is also based on the source texts for each innovation.
-Based on the information contained therein, the innovation is labeled with one or more categories.
-In case of uncertainty, the uncertain label is added, in case where the text does not mention any eco-innovation aspect, the innovation is labeled as not an eco-innovation.
-Where the source text only mentions that the innovation is less harmful to the environment, the label "Environmental-friendly technologies" is used.
-
-<!-- Use examples of clear cases here -->
+The decision process for labelling eco-innovation types was also based on the source texts for each innovation.
+Based on the information contained therein, I labeled the innovation  with one or more categories.
+In case of uncertainty, I added the label uncertain. In case where the text does not mention any eco-innovation aspect, the innovation was labeled as not an eco-innovation.
+Where the source text only mentions that the innovation is less harmful to the environment, the label "Environmental-friendly technologies" was used.
 
 # Results {#sec-results}
-
-Querying the SWINNO database for forest based bioeconomy innovations initially yields `{python} n_uncleaned_bioeconomy_innovations` innovations.
-This number includes false positives from substrings matching keywords included in the search.
-During classification, these false positives were removed from the dataset, yielding `{python} sum_bioeconomy_innovations` innovations which originate from, are used in or make use of products derived from the forest.
 
@@
 ].values[0]
+
+pct_diff_max_latest = (share_max - share_latest) * 100
+total_max = bioeconomy_share["total_count"].max()
+total_max_year = bioeconomy_share.loc[
+    bioeconomy_share["total_count"] == total_max, "year"
+].values[0]
+
+bioeconomy_max = bioeconomy_share["bioeconomy_count"].max()
+bioeconomy_max_year = bioeconomy_share.loc[
+    bioeconomy_share["bioeconomy_count"] == bioeconomy_max, "year"
+].values[0]
+
+total_bioeconomy_share = sum_bioeconomy_innovations / n_swinno
+
 ```
+
+Querying the SWINNO database for forest based bioeconomy innovations initially yields `{python} n_uncleaned_bioeconomy_innovations` innovations.
+This number includes false positives.
+During classification, these false positives were removed from the dataset, resulting in `{python} sum_bioeconomy_innovations` innovations which originate from, are used in, or make use of products derived from the forest.
+
+## Innovation Output of the Swedish Forest-Based Bioeconomy {#sec-results-innovative}
+
+In total, the forest-based bioeconomy produced `{python} f"{total_bioeconomy_share:.0%}"` of innovation captured in the SWINNO database.
+While the total innovation output and bioeconomy innovations feature a period of high output from the 1970s to the 1990s, with a peak in `{python} total_max_year` for total output and `{python} bioeconomy_max_year` for the bioeconomy, the economic downturn and industrial restructurization of the 1990s was accompanied with a decline in commercialized innovation.
+Unlike the other sectors and socio-technical systems, however, it appears that the bioeconomy was not able to recover to the same extent.
+Its percentage of total innovation declined by `{python} f"{pct_diff_max_latest:.0f}"` percentage points from an all-time high of `{python} f"{share_max:.0%}"` in `{python} share_max_year` to `{python} f"{share_latest:.0%}"` in `{python} share_last_year`.
+@fig-innovation-trends illustrates this using the five year moving average for the number of innovations commercialized per year in the bioeconomy and in the entire dataset, including the bioeconomy.
+The decline in relative importance of bioeconomy innovations can be identified in the share of bioeconomy innovation of total innovation.
 
@@
 b.
-Forest based bioeconomy innovations registered in SWINNO database.
+Forest-based bioeconomy innovations registered in SWINNO database.
 c.
@@
 Shown are the 5 year moving averages for each time series.](results/figures/innovation_trends.pdf){#fig-innovation-trends}
-
-To find plausible explanations for the decline in innovation output, as well as explaining the variation in high and low output periods will be the focus of future versions of this paper.
-For now, it seems likely that the 1970s oil crisis and 1969 Environmental Protection Act provided strong pressures and windows of opportunity for the forest bioeconomy to innovate.
-Aspects of this have been discussed by many authors, for example @bergquist2011GreenInnovationSystems who illustrate how several government projects between 1960 and 1980 supported green innovation in Sweden's pulp and paper industry, and @taalbi2017WhatDrivesInnovation who argues that innovation in biomass harvest and use for energy was driven by the oil crisis and simultaneous wood shortage.
-Perhaps more interesting is to explain the slowdown in innovation output, despite the increasing role of the bioeconomy transition in political discourse.
 
@@
 
-For the following part, innovations which do not relate to the forest are excluded.
-Excluded are also the `{python} n_uncertain_bioeconomy_visions` innovations whose source text did not allow labeling without significant uncertaintiy.
-These innovations are planned to undergo a revision by subject matter experts at a later point.
-
 ```{python}
-#| label: Number of Results
-#| output: asis
+# | label: Number of Results
+# | output: asis
 
@@
 
-n_most_bioeconomy_vision = total_certain_bioeconomy_visions.iloc[0, 1]
+n_most_bioeconomy_vision = total_certain_bioeconomy_visions.iloc[0, 2]
 
-display(
-    Markdown(
-        f"This leaves {n_certain_bioeconomy_visions} innovations have been classified for the period between {certain_bioeconomy_visions_min_year} and {certain_bioeconomy_visions_max_year}.
-With {n_most_bioeconomy_vision} innovations, the {total_certain_bioeconomy_visions.iloc[0, 0]} has the most associated innovations.
-Next the {total_certain_bioeconomy_visions.iloc[1, 0]} was associated with {total_certain_bioeconomy_visions.iloc[1,1]}, while {total_certain_bioeconomy_visions.iloc[2,1]} and  {total_certain_bioeconomy_visions.iloc[3,1]} innovation were classified as {total_certain_bioeconomy_visions.iloc[2,0]} and {total_certain_bioeconomy_visions.iloc[3,0]}, respectively.
-Note that the sum by category will not yield the total number of classified innovations, as one innovation can be associated with more than one bioeconomy vision."
+def calculate_share(a: str):
+    return (
+        total_certain_bioeconomy_visions.loc[
+            total_certain_bioeconomy_visions["bioeconomy_vision"] == a, "count"
+        ].values[0]
+        / n_certain_bioeconomy_visions
     )
-)
+
+share_vision_neutral = calculate_share("Vision Neutral")
+share_bioecology = calculate_share("Bioecology Vision")
+share_bioresource = calculate_share("Bioresource Vision")
+share_biotechnology = calculate_share("Biotechnology Vision")
+
+vision_1_count = total_certain_bioeconomy_visions.iloc[0, 2] 
+vision_2_count = total_certain_bioeconomy_visions.iloc[1, 2] 
+vision_3_count = total_certain_bioeconomy_visions.iloc[2, 2] 
+vision_4_count = total_certain_bioeconomy_visions.iloc[3, 2]
+
+vision_1_count = total_certain_bioeconomy_visions.iloc[0, 2] 
+vision_2_count = total_certain_bioeconomy_visions.iloc[1, 2] 
+vision_3_count = total_certain_bioeconomy_visions.iloc[2, 2] 
+vision_4_count = total_certain_bioeconomy_visions.iloc[3, 2]
+
+vision_1 = total_certain_bioeconomy_visions.iloc[0, 0]
+vision_2 = total_certain_bioeconomy_visions.iloc[1, 0]
+vision_3 = total_certain_bioeconomy_visions.iloc[2, 0]
+vision_4 = total_certain_bioeconomy_visions.iloc[3, 0]
+
+n_bioresource_1970_1990 = certain_bioeconomy_visions.query("year.between(1970, 1990) & bioeconomy_vision == 'Bioresource Vision'")["sinno_id"].nunique()
 ```
-
-Perhaps surprisingly, the majority of innovation related to the forest based bioeconomy does not involve one of the visions discussed in policy, industry and research.
-Working with the source articles suggests that one possible reason is that many innovations used or produced by sectors in the forest bioeconomy broadly aim to improve tools.
-For example to increase the lifespan of a milling machine metal tip.
-<!-- SINNO ID: 10022001 -->
-The second surprising thing, is the relative absence of biotechnology vision related innovation.
-Considering that this vision was promoted heavily at an international level by the OECD, it could be expected to play a more significant role in the Swedish innovation system.
 
@@
 ].sum()
-
-display(
-    Markdown(
-        f'For the Bioecology Vision it is important to note that a majority, but not all, innovation associated with this vision earned their label through the mention of recycling in the source.
-Currently, this applies to {recycle_innovation_count} innovations involving either a product or process innovation related to recyling.
-Hence, {n_bioecology_vision - recycle_innovation_count} innovation are associated with the bioecology vision, without using or facilitating recycling.
-These innovations require special focus, as they illustrate that the innovation system does already possess some "[c]ounter-expertise rather than concrete technical solutions" [@vivien2019HijackingBioeconomy 194], to embedd economic activity within the biosphere.
-This focus, however, is not within the scope of the current paper.'
-    )
-)
 ```
 
-```{python}
-#| label: tbl-count_bioeconomy_visions
-#| tbl-cap: Count of Bioeconomy Visions
-#| output: asis
-#| include: false
+For the following part, `{python} n_uncertain_bioeconomy_visions` innovations whose source text did not allow labeling without significant uncertainty are excluded.
+These innovations are planned to undergo a revision by subject matter experts at a later point.
 
-display(
-    Markdown(
-        count_certain_bioeconomy_visions.groupby("bioeconomy_vision")
-        .agg("sum", numeric_only=True)
-        .sort_values("count", ascending=False)
-        .reset_index()
-        .rename(columns={"bioeconomy_vision": "Vision", "count": "Count"})
-        .to_markdown(index=False)
-    )
-)
-```
+Among the remaining `{python} n_certain_bioeconomy_visions` innovations,
+`{python} total_certain_bioeconomy_visions.iloc[0, 2]` (`{python} f"{share_vision_neutral:.0%}"` of all commercialized innovation) were not aligned with any vision.
+Next the `{python} vision_2.lower()` was associated with `{python} vision_2_count` (`{python} f"{calculate_share(vision_2):.0%}"`), while `{python} vision_3_count` (`{python} f"{calculate_share(vision_3):.0%}"`) and  `{python} vision_4_count` (`{python} f"{calculate_share(vision_4):.0%}"`) innovations were classified as `{python} vision_3.lower()` and `{python} vision_4.lower()`, respectively.
+Note that the sum of all categories does not equal the total number of classified innovations, as one innovation can be associated with more than one bioeconomy vision.
+@fig-bioeconomy-trends shows the count of innovation for each categorization per year.
+Output for the biotechnology and bioecology vision appears inconsistent, with years in which not a single commercialized innovation aligned with these visions.
+Additionally, due to the low number of innovations aligned with these visions no clear trends can be identified.
+In contrast, the two peaks of total innovation output of @fig-innovation-trends a, can be identified in the bioresource and vision neutral innovations.
+Unlike vision neutral innovation, bioresource innovation declined to under 4 commercialized innovations after 2010.
 
@@
 
-@fig-bioeconomy-trends illustrates that the peaks in innovation output, identified in @fig-bioeconomy-share and discussed above, were driven by innovation related to the Bioresource Vision as well as Vision Neutral.
-While the Bioresource Visions did strongly contribute to the first peak, the second appears to have been primarily driven by innovation which cannot be associated as belonging to any vision of a bioeconomy.
-This raises a couple of questions regarding the effectiveness of past and current policy.
-It does add, for example, additional heft to the argument put forth by @holmgren2022WhoseTransformationThis that industry interest dominate the transformation to a bioeconomy.
-Further, the persistent dominance of the bioresource vision over the full study period suggests that there might not be much effort, at least in terms of production of innovations, put into creating a transformation towards a bioeconomy.
+### A Deeper Look at Innovations in Each Vision {#sec-results-deeper-visions}
 
-Within the bioresource vision, two technological development blocks provide the basis for a substantial share of innovations: more efficient harvesting methods through technology and more efficient utilization of timber downstream.
-Write here about the pendelsag, about the forwarders, about the extent to which sustainability was incorporated but only to the extent that it allowed more profitable harvesting of wood.
+#### Bioresource Vision
 
-The role that the 1970s played in the development of core technologies for the bioresource vision can be illustrated in the fact that 7 of the 9 forwarder innovations recorded in SWINNO were commercialized between 1974 and 1982.
-Forwarders are machines used to retrieve felled timber from forest plots before loading them on heavier timber transport options.
-Forwarders replaced forest tractors or unmechanized timber retrival methods such as horse pulled sledges.
-@nordfjell2019TechnicalDevelopmentForwarders show that although they have become heavier, their core design has not changed dramatically.
+Although the bioresource vision claims the most aligned innovations, its most active period in terms of commercialized innovations was at the beginning of this study's time frame.
+`{python} n_bioresource_1970_1990` innovations were commercialized between 1970 and 1990.
+These can be grouped into two themes.
 
-Perhaps a little bit about the uncertainty and promise that it would be better / not harmful to forest soil.
+The first theme addressed environmental issues, especially greening the pulp and paper production process and has been extensively documented, for example by Bergquist and Söderholm [-@bergquist2011GreenInnovationSystems; -@bergquist2015TransitionGreenerPulp], and @bergquist2018GreeningPulpPaper.
+The second innovation group was a response to the energy crisis resulting from oil export restrictions placed by OPEC countries in the 1970s [@bergquist2016SustainableEnergyTransition].
+Once again a major player was the pulp and paper industry which commercialized numerous innovations to improve energy efficiency in drying (*Ny Teknik*, 1972:1, p.
+4), reduce cool down of furnaces (*Kemisk Tidskrift*, 1976:9 pp.
+96,99-100) and generate cheaper energy from waste, byproducts (*VVS*, 1980:7-8, p.69) and excess process heat (*Kemisk Tidskrift*, 1981:7, p.
+82).
+The second major player was the forestry sector.
+A cluster of technologies such as forwarders, tree harvesters, wood driers and processors emerged in response to a wood shortage and soaring energy prices from the oil shocks in the 1970s [@taalbi2014InnovationCreativeResponse].
+Key innovations in this technology cluster were the pendulum saw (*Sågverken*, 1977:2, pp.
+97-101,125), which could be mounted on vehicles to mechanically fell entire trees at once (*Sågverken*, 1974:10, pp.
+725-727), and vehicles that, despite being heavy, limited the damage to soft forest soil and importantly, the roots therein (e.g., *Sågverken*, 1974:10, p.
+725-727; *Sågverken*, 1975:10, p.
+731).
+
+The second period of increased activity starting in 2000 focused more on
+upgrading of biomass into new products such as construction timber (e.g., *NTT* 2003:27, p.
+11; *NTT*, 2003:8, p.6; *NTT* 2004:18, pp.16-17), various paper products (e.g., *Kemivärlden Biotech med Kemisk Tidskrift*, 2009:11, pp.
+60-61; *SPCI Svensk Papperstidning*, 2005:7, pp.10-11), but also biobased alternatives to plastics (e.g., *SPCI Svensk Papperstidning*, 2018:4, pp.
+28-33; *SPCI Svensk Papperstidning*, 2019:6, p.24; *Plastforum*,2019:5, p.31), or refined bio-fuel alternatives to fossil fuels, such as the LignoBoost Process (*SPCI Svensk Papperstidning*, 2006:7, pp.
+14-16), or gassification of black liquour -- a by-product of pulp digestion -- for electricity generation (*Ny Teknik*, 1990:16, p.5) .
+A second theme is rationalization of production processes, especially through the application of computers and software (e.g.,  *NTT*, 2007:14, p. 13; *Automation*, 2021:1, p. 24).
+
+Therefore, while the first period of high innovation output until the 1990s was characterized by attempts to modernize timber production, mainly through the mechanization of harvesting, and cleaning the pulp and paper process, the second period of active commercialization focused on finding new applications for biomass and improving the cost effectivness of exisiting processes.
+This is illustrated with innovations in forwarders -- specialized machines to aid in the removal of timber from harvest sides before loading them onto transport -- where of 9  innovations, 7 were commercialized before 1990.
+Moreover, the two innovations commercialized after focused on greening fuel consumption through hybrid engines (*Ny Teknik*, 2008:34, p.
+24) and weight reduction (*Ny Teknik*, 2008:34, p.
+24; *Automation*, 2012:40, p.
+26).
+Meanwhile, earlier innovations focused on establishing the product category by solving issues of soil damage from use of heavy machinery (*Sågverken*,1975:10, p.
+731; *Sågverken*, 1982:10, p.
+77) as well as integrating forwarders with other technologies for planting (*Sågverken*, 1979:3, p.
+237) and removing trees (*Ny Teknik*, 1976:43, p.20; *Sågverken*, 1975:12, pp.
+905,907).
+<!-- TODO make dynamic -->
 
@@
 
-Within bioecology, few innovations aim to decommodify value chains.
-Dominated by technological development block to green the pollution of pulp and paper industry.
+#### Biotechnology Vision
 
-In biotechnology, attempts to form a development block around new material such as tencel flaunder.
+Only 37 innovations were aligned with the biotechnology vision.
+Of those, the majority focused on cleaning waste water from, for example, fibers (e.g., *Kemisk Tidskrift*, 1972:4, p.
+19), or chemicals (e.g., *Ny Teknik*, 1970:4, p.
+8), with applications mostly in the pulp and paper industry.
+Other innovation applied biomass derived from the forest in new products, such as new fibers (*Ny Teknik*, 1976:9, pp.
+1,4-5), or applications of cellulose in heat isolation materials (*Ny Teknik*, 2002:9, p.
+14).
+Innovations in this vision frequently overlapped with bioresource innovations, such as the above mentioned LignoBoost process.
+
+#### Bioecology Vision
+
+Recycling accounted for `{python} recycle_innovation_count` unique innovations, or `{python} f"{(recycle_innovation_count / n_bioecology_vision):.0%}"` of innovation associated with the bioecology vision.
+The majority of innovations in aligned with this vision due to some explicitly stated environmental aim within existing production methods.
+An example is an instrument that measures the extent to which a tree is decayed from the inside (*NTT*, 2006:2, p. 20).
+While such trees are of little economic value, they play a big role in biodiversity conservation if left standing.
+This innovation aims to identify these trees before they are felled and thus to protect it from being felled.
+Some innovations stood out as attempts to generate social or ecological counter expertise.
+For example, in 1971 *Sågverken* (3, p. 195), covered an innovation in wooden floors made from wide, red pine planks.
+Notable about this innovation was that it countered the trend of economic rationalization of the time, in favor of tight integration between production and forest ownership.
+Another example of regional timber being treated not as a commodity, but as a desirable high quality product is the Masonite sytem for pre-fabricated wooden house (*Svensk trävaru- & pappersmassetidning*, 1976:7, pp.
+539,541,544).
+The specifc growing conditions of Northern Sweden are provided as a reason for denser wood, especially well suited for construction due to its strength properties and the "[t]otal integration makes the company feel optimistic: own forests, top-modern sawmill, board mill, newly built component factory are positive factors meaning advantages in all stages" (p.
+544).
+Similiar decommodification of value chains in favor of local product could even be found in the paper industry, where fresh wood from within a 100km radius was found to produce better quality and reduce chemical and energy usage (*SPCI Svensk Papperstidning*, 2015:8, pp.
+26-30).
 
@@
 How does each vision contribute to lessening the environmental impact of the forest bioeconomy?
-While a quantitative assessment of technological impact on the environment comes with a host of challenges, the labelling of eco-innovations at least allows a glimpse at what could be expected from each vision.
+While a quantitative assessment of technological impact on the environment comes with a host of challenges, the labelling of eco-innovations at least allowed a glimpse at what could be expected from each vision.
 @fig-eco-innovation-bioeconomy-vision shows the count of eco-innovation associated with each bioeconomy vision by type.
-As before, there is no weighting, which is why the total sum of all eco-innovation across all bioeconomy visions outnumbers the sum of innovations in the bioeconomy.
+As before, the total sum of all eco-innovation across all bioeconomy visions outnumbers the sum of innovations in the bioeconomy.
 A single innovation can be more than one eco-innovation type, for example, if it reduces the amount of raw material needed by recycling waste.
-The picture of innovation in the forest bioeconomy which @fig-eco-innovation-bioeconomy-vision suggests fits previous literature and what can be expected from LBIO data.
-Product and Process innovation are by far the most frequent across all visions.
-The Bioresource Vision is heavily involved with improving raw-material utilization, while the Bioecology Vision is most frequently associated with recycling.
-At the very least, this suggests that the definitions for each Vision and the classification scheme are in agreement.
+Product and Process innovation were by far the most frequent across all visions.
+The bioresource vision was heavily associated with improving raw-material utilization, while the Bioecology Vision is most frequently associated with recycling.
+At least, this suggests that the definitions for each Vision and the classification scheme are in agreement.
 
-A noteworthy feature of innovation belonging to No Bioresource Vision is that this category has by far the most Non-Eco innovations.
-The chi-squared test of independence reveals a statistically significant association between an innovation being aligned with a vision and it being an eco-innovation ($\chi^2 =$ `{python} round(chi2_stat, 3)` df = `{python} dof`, p < 0.05).
-For the sustainability transition of the forest bioeconomy this could imply that having some Vision of a bioeconomy creates more eco-innovation than pursuing other innovation paths.
+```{python}
+#| label: fig-eco-innovation-bioeconomy-vision
+#| fig-cap: Count of Eco-Innovation Types by Bioeconomy Vision
+p = plot_eco_innovation_by_bioeconomy_vision(
+    eco_innovation_by_bioeconomy_visions, codes_df
+)
+
+sns.move_legend(p, "lower center", ncol=5)
+# TODO: perhaps a per cent display by group size would be interesting too?
+```
+
+A noteworthy feature of innovation lacking alignment to a bioeconomy vision was that this category has by far the most non-eco-innovations.
+The chi-squared test of independence revealed a statistically significant association between an innovation being aligned with a vision and it being an eco-innovation ($\chi^2 =$ `{python} round(chi2_stat, 3)` df = `{python} dof`, p < 0.05).
 
@@
 
-```{python}
-#| label: fig-eco-innovation-bioeconomy-vision
-#| fig-cap: Count of Eco-Innovation Types by Bioeconomy Vision
-p = plot_eco_innovation_by_bioeconomy_vision(
-    eco_innovation_by_bioeconomy_visions, codes_df
-)
-
-sns.move_legend(p, "lower center", ncol=5)
-# TODO: perhaps a per cent display by group size would be interesting too?
-```
-
-<!-- To answer this, I need to decide what is an eco innovation in the first place.
-should take the original values from @garcia-granero2018EcoinnovationMeasurementReview and remove codes 
-501 -- health and saftey
-701 -- Biocide
-108 -- Gen manipulated plant
-109 -- increased plant yield
-111 -- Gen manipulated animal
-  -->
-
-<!-- From the review of innovation two clusters of highly related innovations emerge.
-One revolves around optimizing harvest of biomass from forest as a response to the 1970s oil crisis.
-The second is a response to the envrionmental pollution problems faced by the pulp and paper industry.
-
-While in the first cluster environmental considerations played only a secondary role, they were crucial in the cluster around recyclability and water pollution prevention of the pulp and paper industry.
-This is not to say that the environmental impacts of harvest intensification were not at least to some extent on the agenda of the innovation system.
-One innovation source cites possible concerns about the environemntal effects of a root harvester, pointing to the need of future studies to establish the extent to which this demages forests.
-<!-- TODO: find innovation that discusses forest floor issues -->
-<!-- One source, after describing a new root harvester from  -->
-<!-- TODO: find company -->
-<!-- , also included a hand drawn charicature of a hiker, who, taking a break on the  last stump on a hill and watching the harvester at work, exclaims "Thankfully they left at least one standing".
--->
-
 # Discussion {#sec-discussion}
 
-Within the bioeconomic innovation system, the different visions could reasonably be conceived of as niches themselves, vying not only for importance over each other, but even for providing sufficient direction to overcome general purpose improvements to and from a regime system.
-While the bioeconomy innovation system can be viewed as a niche in the transition to a economic system which does not rely on fossil based technologies, this analysis highlights the layered nature of contested transitions.
+The decline in innovations commercialized in the forest-based bioeconomy relative to all commercialized innovations suggests that Sweden's capacity to produce technological change towards a bioeconomy has diminished over the last 50 years.
+While innovation counts should be interpreted with caution, as the magnitude of technological change stemming from individual innovations can differ, it is surprising that a economic sector, which is expected to realize ambitious economic, social and ecological targets, is losing importance in relation to other Swedish innovations.
+While it appears that the crises in forestry and energy of the 1970s stimulated innovation directed towards economic rationalization of forestry, the bioeconomy could not benefit from similar stimulants when exploring conversion of biomass into new products after the financial crisis in 2008, and the shock to timber production after storm Gudrun in 2005.
+
+As for the positive directionality of the Swedish transition to a forest-bioeconomy, the dominant vision appears to be successful in so far as it is the vision most innovations align with.
+That a majority of innovation appear neutral in regards to vision based directionality is not to say that these do not posess positive directionality.
+These innovations should be regarded as general improvements to exisiting technology and hence are not neutral in regards to technological change, only in regards to not aligning with normatively directionality.
+The congruency between positive and normative directionality can be explained by a specific feature of Swedish socio-technical systems: a strong desire to create consensus [@belandlindahl2015SwedishForestryModel].
+@bergquist2011GreenInnovationSystems illuminate how strong collaboration networks between public and private actors shaped the greening of Sweden's pulp and paper industry between 1960 and 1980.
+The findings of this study support this view.
+Rather than providing directionality towards sustainability, preparations for a national forest strategy and the national forest program strongly align with the trajectories of technological change produced by the innovation system.
+While innovation systems literature highlights this as the mark of well implemented transition management, this consensus building also comes at a cost to transition paths.
+However, this consensus building comes at a cost, where dissenting views are marginalized [@fischer2020SwedishForestsBioeconomy].
+@holmgren2022WhoseTransformationThis show that industry interest dominate the transformation to a bioeconomy.
+@fischer2020SwedishForestsBioeconomy even go so far to liken it to consensus building through exclusion.
+
+Looking at the direction embedded in commercialized innovations suggests are more nuanced picture.
+While the bioresource vision is the most pronounced out of the three visions for a bioeconomy, innovations can and do align with multiple visions at once.
+In addition, those innovations that align with any of the bioeconomy visions are more often eco-innovations than vision neutral technological change.
+However, it must be said that whether this generates positive environmental impact depends on a number of factors such as technological diffusion and possible rebound effects.
 
@@
 
-An important limitation of LBIO measures, especially for research into forest sector innovation, is a bias against non-commercialized innovation.
-New processes, or new organizational approaches for example, will only be captured if they are sold to another company.
-This is especially sensitive for bioeconomy research as, for example, new methods or plans for managing forest stands are less likely to be included.
+In regards to the above, an important limitation is the use of innovation as a measure of socio-technical transitions.
+While the results of this study agree with the survey results of 
+@lovric2020FactorsDevelopmentInnovations and capture the production of technological change, it does not directly capture its adoption and use.
+In addition literature-based innovation output measures have data limitations might result in a bias against non-commercialized innovation, especially for research into forest sector innovation where organizational innovations such as silviculture management can be important levers to improve, for example biodiversity or cultural ecosystem services.
+Such new processes, or new organizational approaches, will only be captured if they are sold to another company.
 Relevant product innovation which facilitates these plans, for example products related to the planting of trees, on the other hand, are captured.
-A likely consequence of this limitation is that the number of innovation related to the bioecology vision are under counted, as this vision highlights "Counter-expertise rather than concrete technical solutions" [@vivien2019HijackingBioeconomy 194].
-Nonetheless, a few innovations associated with this vision were already identified in the database which will be discussed later.
+A likely consequence of this limitation is that the number of innovation related to the bioecology vision could be under counted, as this vision highlights "[c]ounter-expertise rather than concrete technical solutions" [@vivien2019HijackingBioeconomy 194].
+
+A second limitation arising from the data is the usage of innovation counts.
+While the literature-based innovation output captures significant innovation rather than mere reconfiguration, counting innovation thus captured can be misleading as to the impact of each individual innovation.
+Some technologies require few break-through innovations before meeting large market demand and diffusing quickly.
+Others might go through multiple rounds of more incremental improvements.
+Although literature-based innovation output measures assume editors at trade journales already discard incremental changes to include only important improvements, they cannot ex-ante anticipate which innovation will be most important in terms of impact [@kander2019InnovationTrendsIndustrial].
+Therefore the decline in innovation output this paper identifies should not be taken to necessarily imply a decline in the importance of the commercialized innovation.
+Similarily, the positive directionality measured in this paper should be seen more as a the vector of change embedded in the transition to a bioeconomy, rather than a scalar of its progress.
 
@@
 In conclusion, the technological artefacts produced in the transition to a forest-based bioeconomy in Sweden show a clear directionality aligned with the bioresource vision, which is most favored by central actors and policy makers.
-At the same time, there is evidence that the innovation system has produced knowledge aligned with alternative transition pathways towards an ecological bioeconomy, such as decommodification of value chains and efforts to increase the circularity of the system and reduce environmental pollution.
-The temporal distribution of innovation output aligns strongly with the Swedish tendency to produce consens oriented policy by aligning policy objectives with existing directionality.
+At the same time, there is evidence that the innovation system has produced knowledge aligned with alternative transition pathways towards an ecological bioeconomy, such as decommodification of value chains as well as efforts to increase the circularity of the system and to reduce environmental pollution.
+The temporal distribution of innovation output aligns strongly with the Swedish tendency to produce consens oriented policy by aligning policy objectives with existing technolgical trajectories.
 While this approach is pragmatic and has produced good outcomes in the past, it is at least doubtful how radical a shift can be expected for the bioeconomy transition, considering that innovation output since the 1970s has not meaningfully explored alternatives to the intensification of biomass resource extraction.
@@
 
+## Trade Journals
+
+\noindent
+*Automation* (1973-2007).
+Sundbyberg: Verkstadstidningen Förlags AB
+
+\noindent
+*Kemisk tidskrift: svensk kemisk tidskrift, kemiteknik.* (1969-1993). Stockholm: Ingenjörsförlaget.
+
+\noindent
+*Kemivärlden, Biotech, Kemisk tidskrift.* (2005-2018) Stockholm : Mentor Communications
+
+\noindent
+*Ny teknik.* (1970-2024). Stockholm: Teknisk tidskrifts förlag.
+
+\noindent
+*NTT såg & trä: Nordisk träteknik, Sågverken, Träindustrin.* (2003-2024).
+Lidingö: Mentor Communications
+
+\noindent
+*Svensk papperstidning: medlemstidning för Svenska pappers- och cellulosaingeniörsföreningen - SPCI.*(2003-2024). Stockholm: AB Svensk
+papperstidning.
+
+\noindent
+*Svensk trävaru- & pappersmassetidning. *(1970-1990). Stockholm: AB
+Svensk trävarutidning.
+
+\noindent
+*Sågverken: tidskrift för sågverksindustri, trävaruhandel och skogsbruk.* (1974 - 1998). Stockholm: Svensk trävarutidning.
+
+\noindent
+*VVS: tidskrift för värme-, ventilations- och sanitetsteknik : organ för Sven-
+ska värme- och sanitetstekniska föreningen.* (1970-1982). Stockholm:
+VVS.
+
 # Appendix {.appendix}
@@
 
+## Data Appendix {#sec-appendix-data}
 ```{python}
@@
 
-::::{.columns}
-
-:::{.column width="50%"}
-**Using producer / user codes
-**
-
-|SNI Code | Description                            |
-|--------:|:---------------------------------------|
-|02       | Forestry and related services
-|20       | Wood and wood product manufacturing except furniture
-|21       | Pulp, paper and paper product manufacturing
-|36       | Furniture manufacturing; other manufacturing
-:::
-
-:::{.column width="50%"}
-**Adding keywords of value chain**
-
-    - "virke",
-    - "cellulos",
-    - "lignin",
-    - "spån", 
-    - "bark",
-    - "levulinsyra" (Levulinic acid),
-    - "furfural" (Furfural),
-    - "svarttjära",
-    - "svartlut",
-    - "växtbas",
-    - "ved",
-    - "trä",
-    - "skog",
-    - "papper",
-    - "biobränsle",
-    - "biologiskt",
-    - "nedbrytbar",
-    - "papper",
-    - "pappret",
-    - "karton",
-    - "tencel",
-:::
-
-::::
+{{< include _appendix-data.qmd >}}

### 2024-05-07 rename_files.sh
clean up repo

@@ -1,15 +1,0 @@
-#!/bin/bash
-
-# specify the directory to perform the replacement in
-dir="scripts"
-
-# loop through all files and directories in the specified directory
-for file in "$dir"/*; do
-    # check if the file name contains "-"
-    if [[ "$file" == *-* ]]; then
-        # replace "-" with "_"
-        new_file="${file//-/_}"
-        # rename the file with the new name
-        mv "$file" "$new_file"
-    fi
-done

### 2024-05-07 scripts/calculate_bioeconomy_share_data.py

@@ -1,146 +1,0 @@
-# | label: Imports
-
-import pandas as pd
-import numpy as np
-from pathlib import Path
-from src.figures import *
-from src.utils import get_project_root, format_table
-from src.swinno_helpers import connect_swinno_db
-
-ROOT = get_project_root()
-conn = connect_swinno_db()
-
-# set all text on plots to be monospace with rcParams
-plt.rcParams["font.family"] = "Space Grotesk"
-
-sns.set_theme(style="white")
-
-# full code labels for merging
-
-
-codes_df = pd.read_sql("SELECT * FROM classification_codes", conn)
-
-codes_df["Category"] = codes_df["Category"].str.strip()
-uncertain_eco = {"Code": "000", "Category": "Unsure"}
-codes_df = codes_df.append(uncertain_eco, ignore_index=True)
-
-codes_df = codes_df[~codes_df["Code"].duplicated(keep="last")]
-
-codes_df = codes_df.set_index("Code")
-
-# | label: Query bioeconomy
-
-bioeconomy = pd.read_sql(
-    """
-select
-  i.sinno_id,
-  i.innovation_name_in_swedish AS name,
-  i.description_in_swedish AS description,
-  i.additional_information_if_origin_new_scientific_discovery || i.additional_information_if_origin_new_technologies_or_materials || i.additional_info_if_origin_official_regulation_legislation_and_standards || i.additional_information_if_origin_solution_for_a_problem || i.additional_information_if_origin_performance || i.additional_information_if_origin_other AS info,
-  i.year_of_commercialization AS year,
-  us.use_sector
-from
-  innovation i
-  join use_sectors us on i.sinno_id = us.sinno_id
-where
-  (
-    us.use_sector like '02%'
-    or us.use_sector like '20%'
-    or us.use_sector like '21%'
-    or us.use_sector like '36%'
-    or product_code like '02%'
-    or product_code like '20%'
-    or product_code like '21%'
-    or product_code like '36%'
-  )
-  or (
-    description like '%virke%'
-    or description like '%cellulos%'
-    or description like '%lignin%'
-    or description like '%spån%'
-    or description like '%bark%'
-    or description like '%levulinsyra%'
-    or description like '%furfural%'
-    or description like '%svarttjära%'
-    or description like '%svartlut%'
-    or description like '%växtbas%'
[…]

### 2024-05-07 experiments/14-plots.qmd
Delete unused plot

@@
     ggplot(df, aes(x = year, y = .data[[y]])) +
-        geom_area(color="black", fill = "gray40", alpha = 0.5, linewidth=1) +
+        geom_area(color = "black", fill = "gray40", alpha = 0.5, linewidth = 1) +
         xlab("Year") +
@@
 
-innovation_trends_plot  <- (innovation_plot + bioeconomy_plot) / share_plot + plot_annotation(tag_levels = "a") & theme(plot.margin = margin(t = 0, 
+innovation_trends_plot <- (innovation_plot + bioeconomy_plot) / share_plot + plot_annotation(tag_levels = "a") & theme(plot.margin = margin(
+    t = 0,
     r = 1, # Right margin
     b = 0, # Bottom margin
-                             l = 0))
+    l = 0
+))
 
 innovation_trends_plot
-ggsave(here("results/figures/innovation_trends.pdf"), width=7, height=5)
+ggsave(here("results/figures/innovation_trends.pdf"), width = 7, height = 5)
 ```
-
-## Bioeconomy Visions Plot
-
-```{r}
-visions <- import(here("data/modified/count_certain_visions.csv"))
-
-vision_summary <- visions |>
-    group_by(bioeconomy_vision) |>
-    summarize(total_count = sum(count))
-
-visions <- merge(visions, vision_summary, by = "bioeconomy_vision")
-
-```
-
-```{r}
-batlow_discrete <- , n = 5)
-```
-
-```{r}
-paletteer::paletteer_c("scico::batlowW", n = 7)
-```
-```{r}
-#| label: fig-vision-count
-vision_count_plot <- ggplot(visions, aes(x = year, y = count, color = bioeconomy_vision, fill = bioeconomy_vision)) +
-    geom_col(alpha = 0.9, linewidth = 0.75) +
-    facet_wrap(~bioeconomy_vision) +
-    scale_color_manual(values = batlow_discrete) +
-    scale_fill_manual(values = batlow_discrete) +
-    theme_custom() +
-    theme(legend.position = "none") +
-    scale_y_continuous(breaks = seq(0, max(visions$count), by = 2), expand = c(0, 0)) +
-    xlab("Year") +
-    ylab("Innovation Count") +
-    geom_text(
-        data = vision_summary, aes(label = paste("Total:", total_count), x = Inf, y = Inf),
-        position = position_nudge(y = 0), hjust = 1, vjust = 1.3, size = 3.5
-    )
-
-ggsave(here("results/figures/vision_count.png"), vision_count_plot)
-```
-
-```{r}
-#| label: fig-visions-stacked-area
-
-# Reorder the levels of bioeconomy_vision based on the sum of counts
-visions$bioeconomy_vision <- factor(visions$bioeconomy_vision,
-    levels = vision_summary$bioeconomy_vision[order(vision_summary$total_count, decreasing = TRUE)]
-)
-
-p <- ggplot(visions, aes(x = year, y = count, fill = bioeconomy_vision, color = bioeconomy_vision)) +
-    geom_area(color = "black", alpha = 0.8) +
-    scale_fill_manual(values = batlow_discrete, name = "Bioeconomy Vision") + # Set legend title
-    scale_color_manual(values = batlow_discrete, guide = "none") + # Hide color legend
-    theme_custom() +
-    theme(legend.position = c(0.9, 0.9)) +
-    scale_y_continuous(expand = c(0, 0)) +
-    labs(color = NULL)
-
-p
-
-```

### 2024-05-07 experiments/14-plots.qmd
add vision_counts.pdf plot

@@
 #| label: fig-n-innovation
-
 innovation_plot <- plot_bioeconomy_stats(bioeconomy_share, "ma5_n_innovation", "Total SWINNO")
@@
 ```
+
+## Bioeconomy Visions Plot
+
+```{r}
+visions <- import(here("data/modified/count_certain_visions.csv"))
+
+vision_summary <- visions |>
+    group_by(bioeconomy_vision) |>
+    summarize(total_count = sum(count))
+
+visions <- merge(visions, vision_summary, by = "bioeconomy_vision")
+
+```
+
+```{r}
+batlow_discrete <- paletteer::paletteer_c("scico::batlowW", n = 5)
+```
+
+```{r}
+#| label: fig-vision-count
+vision_count_plot <- ggplot(visions, aes(x = year, y = count, color = bioeconomy_vision, fill = bioeconomy_vision)) +
+    geom_col(alpha = 0.9, linewidth = 0.75) +
+    facet_wrap(~bioeconomy_vision) +
+    scale_color_manual(values = batlow_discrete) +
+    scale_fill_manual(values = batlow_discrete) +
+    theme_custom() +
+    theme(legend.position = "none") +
+    scale_y_continuous(breaks = seq(0, max(visions$count), by = 2), expand = c(0, 0)) +
+    xlab("Year") +
+    ylab("Innovation Count") +
+    geom_text(
+        data = vision_summary, aes(label = paste("Total:", total_count), x = Inf, y = Inf),
+        position = position_nudge(y = 0), hjust = 1, vjust = 1.3, size = 3.5
+    )
+
+ggsave(here("results/figures/vision_count.pdf"), vision_count_plot)
+```

### 2024-09-05 20240906-pjk-NEST.qmd
add NEST presentation

---
title: Quantifying Directionality and Innovation Output in Sweden's Transition to a Forest-Based Bioeconomy
author: Philipp Jonas Kreutzer
date: last-modified
bibliography: swinno-shades-green.bib
csl: 'https://zotero.org/styles/apa?source=1'
format: 
    pjk_theme-revealjs
---

## [Forest-Based Bioeconomy]{.text-white} {background-image="/assets/images/forwarder.jpeg" .center}

## RQ {.center}

[1.]{.text-gray-600} [How many innovations are commercialized in this new economy]{.text-royal-blue-600}, especially considering the central role of innovation in bioeconomy discourse?

[2.]{.text-gray-600} What does commercialized innovation suggest for the [positive directionality of the bioeconomy transition]{.text-royal-blue-600}?

# Data From SWINNO Database {.inverted}

## Significant Swedish Innovations

::: {.columns}
::: {.column}
~ 5000 unique innovations

Literature-Based Innovation Output method

from 15 independent trade journals [@sjoo2014DatabaseSwedishInnovations]

:::
::: {.column}
![Example Page From a Source Article](/assets/images/swinno-example-1.jpg)
:::
:::

## Defining the Bioeconomy {.smaller}

::: {.columns}
::: {.column}
|SNI Code | Description                            |
|--------:|:---------------------------------------|
|02       | Forestry and related services
|20       | Wood and wood product manufacturing except furniture
|21       | Pulp, paper and paper product manufacturing
|36       | Furniture manufacturing; other manufacturing
:Key Sectors Used in Query {#tbl-query}

:::
::: {.column}
| Swedish           | English                    |
|-------------------|----------------------------|
| virke             | timber                     |
| cellulos          | cellulose                  |
| lignin            | lignin                     |
| spån              | chip                       |
| bark              | bark                       |
| levulinsyra       | levulinic acid             |
| furfural          | furfural                   |
| svarttjära        | black tar                  |
| svartlut          | black liquor               |
| växtbas           | plant-based                |
| ved               | wood                       |
| trä               | timber                     |
| skog              | forest                     |
| biobränsle        | biofuel                    |
| biologiskt        | biological                 |
| nedbrytbar        | biodegradable              |
| papper            | paper                      |
| karton            | carton                     |
| lyocell            | lyocell |
: Swedish Keywords used in Query: `WHERE description LIKE %keyword% OR ...` {#tbl-keywords}
:::
:::

# Results {.inverted}

## Bioeconomy Innovation Has Declined After 1970s

![a. A total number of innovation registered in SWINNO database. b. Forest-based bioeconomy innovations registered in SWINNO database. c. Percentage of forest based bioeconomy innovations to total innovation registered in SWINNO database. Shown are the 5 year moving averages for each time series.](results/figures/innovation_trends.svg){#fig-innovation-trends}

## Three Visions for Normative Directionality

::: {.columns}
::: {.column width="33%"}

### [Bioresource]{.text-batlow-darkgreen}

![](/assets/images/bioresource.svg)
:::
::: {.column width="33%"}

### [Biotechnology]{.text-batlow-forestgreen}

![](/assets/images/biotechnology.svg)

:::
::: {.column width="33%"}

### [Bioecology]{.text-batlow-darkblue}

![](/assets/images/bioecology.svg)

:::
:::

Classification based on @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy

## Bioresource Was Biggest Vision Category

But most innovations did not align with any vision

![Counts of innovation classification by Bioeconomy Vision Category](/results/figures/vision_count.svg){#fig-bioeconomy-trends}

## [Bioresource Vision]{.text-batlow-darkgreen} {background-video="/assets/images/harvester.mp4" background-video-loop=true background-video-muted=true background-opacity=0.4}

1970-1990:

Intensification of Harvesting

Energy & Chemical Pollution Reduction in Pulp and Paper

. . .

2000-2021:

New Materials

## {.center}

::: {.columns}
::: {.column width="33%"}

### [Bioresource]{.text-batlow-darkgreen}

[1970-1990:]{.text-gray-600}

Intensification of Harvesting

Energy Usage & Chemical Pollution Reduction in Pulp and Paper

[2000-2021:]{.text-gray-600}

New Materials

:::
::: {.column width="33%"}

### [Biotechnology]{.text-batlow-forestgreen}

Mostly applications to clean water in pulp and paper

New materials from cellulose

:::
::: {.column width="33%"}

### [Bioecology]{.text-batlow-darkblue}

74% Recycling related

Decommodification of value chains through regional identity products

:::
:::

# Outlook & Aims {.inverted}

## Sharpen Analytical Focus on Directionality

To what extent is there a tension between the normative (desired) and positive (observed) directionality in the transition to a bioeconomy?

How does positive directionality shape normative directionality?

# Take Aways

1. [The bioeconomy is declining]{.text-royal-blue-600} relative to total innovation

2. [Positive directionality towards bioresource vision]{.text-royal-blue-600}, but overlap between visions and open development paths

# References

::: {#refs}
:::

### 2024-09-06 index.qmd
create gh-pages for NEST

--- 2024-09-05 20240906-pjk-NEST.qmd
@@
 format: 
-    pjk_theme-revealjs
+    pjk_theme-revealjs:
+        embedd-resources: true
 ---

### 2024-09-11 20240906-pjk-NEST.qmd
update nest presentation

@@
 
+::: {.columns}
+::: {.column width=66%}
+
 1. [The bioeconomy is declining]{.text-royal-blue-600} relative to total innovation
@@
 2. [Positive directionality towards bioresource vision]{.text-royal-blue-600}, but overlap between visions and open development paths
+:::
+::: {.column width=33%}
+
+**Get the Paper**
+
+{{< qrcode https://pjkreutzer.github.io/files/2024-05-05_pjk_draft_directionality.pdf width=100 height=100 colorDark='#0011bb' >}}
+:::
+:::
 

### 2024-09-12 notebooks/14-plots.qmd
rename experiments to notebooks

--- 2024-05-07 experiments/14-plots.qmd
@@
 ggsave(here("results/figures/innovation_trends.pdf"), width = 7, height = 5)
+ggsave(here("results/figures/innovation_trends.svg"), width = 7, height = 5)
 ```
@@
 batlow_discrete <- paletteer::paletteer_c("scico::batlowW", n = 5)
+batlow_discrete
 ```
@@
 ggsave(here("results/figures/vision_count.pdf"), vision_count_plot)
+ggsave(here("results/figures/vision_count.svg"), vision_count_plot)
 ```

### 2024-09-12 paper.qmd
refactor to pixi

@@
 
-from src.figures import *
-from src.utils import get_project_root, format_table
-from src.data_handling import database
+from swinno_bioeconomy_directionality.figures import *
+from swinno_bioeconomy_directionality.utils import get_project_root, format_table
+from swinno_bioeconomy_directionality.data_handling import database
 

### 2024-09-12 _extensions/jmbuhr/qrcode/assets/qrcode.js

@@ -1,0 +1,700 @@
+/**
+ * @fileoverview
+ * - Using the 'QRCode for Javascript library'
+ * - Fixed dataset of 'QRCode for Javascript library' for support full-spec.
+ * - this library has no dependencies.
+ * 
+ * @author davidshimjs
+ * @see <a href="http://www.d-project.com/" target="_blank">http://www.d-project.com/</a>
+ * @see <a href="http://jeromeetienne.github.com/jquery-qrcode/" target="_blank">http://jeromeetienne.github.com/jquery-qrcode/</a>
+ */
+var QRCode;
+
+(function () {
+	//---------------------------------------------------------------------
+	// QRCode for JavaScript
+	//
+	// Copyright (c) 2009 Kazuhiko Arase
+	//
+	// URL: http://www.d-project.com/
+	//
+	// Licensed under the MIT license:
+	//   http://www.opensource.org/licenses/mit-license.php
+	//
+	// The word "QR Code" is registered trademark of 
+	// DENSO WAVE INCORPORATED
+	//   http://www.denso-wave.com/qrcode/faqpatent-e.html
+	//
+	//---------------------------------------------------------------------
+	function QR8bitByte(data) {
+		this.mode = QRMode.MODE_8BIT_BYTE;
+		this.data = data;
+		this.parsedData = [];
+
+		// Added to support UTF-8 Characters
+		for (var i = 0, l = this.data.length; i < l; i++) {
+			var byteArray = [];
+			var code = this.data.charCodeAt(i);
+
+			if (code > 0x10000) {
+				byteArray[0] = 0xF0 | ((code & 0x1C0000) >>> 18);
+				byteArray[1] = 0x80 | ((code & 0x3F000) >>> 12);
+				byteArray[2] = 0x80 | ((code & 0xFC0) >>> 6);
+				byteArray[3] = 0x80 | (code & 0x3F);
+			} else if (code > 0x800) {
+				byteArray[0] = 0xE0 | ((code & 0xF000) >>> 12);
+				byteArray[1] = 0x80 | ((code & 0xFC0) >>> 6);
+				byteArray[2] = 0x80 | (code & 0x3F);
+			} else if (code > 0x80) {
+				byteArray[0] = 0xC0 | ((code & 0x7C0) >>> 6);
+				byteArray[1] = 0x80 | (code & 0x3F);
+			} else {
[…]

### 2024-09-12 scripts/analysis/00-pull_data.py

@@ -1,4 +1,4 @@
-from src.data_handling import database
+from swinno_bioeconomy_directionality.data_handling import database
 from pyhere import here
 import pandas as pd
 

### 2024-09-12 scripts/combine_categories.py

@@ -1,7 +1,7 @@
 import pandas as pd
-from src.utils import get_project_root
-from src.categorization_helpers import *
-from src.swinno_helpers import *
+from swinno_bioeconomy_directionality.utils import get_project_root
+from swinno_bioeconomy_directionality.categorization_helpers import *
+from swinno_bioeconomy_directionality.swinno_helpers import *
 
 ROOT = get_project_root()
 
@@ -12,7 +12,6 @@ list_visions_df = []
 list_notes_df = []
 
 for csv in Path(ROOT, "data", "raw-data", "tags").glob("*.csv"):
-
     file_name = Path(csv).stem
 
     print(f"****************** \n {file_name}")

### 2024-09-12 scripts/filter_bioeconomy.py

@@ -2,7 +2,10 @@ import pandas as pd
 import numpy as np
 from datetime import datetime
 from pathlib import Path
-from src.swinno_helpers import connect_swinno_db, get_project_root
+from swinno_bioeconomy_directionality.swinno_helpers import (
+    connect_swinno_db,
+    get_project_root,
+)
 
 conn = connect_swinno_db()
 

### 2024-09-12 scripts/find_sources.py

@@ -1,8 +1,13 @@
 import argparse
 from pathlib import Path
 
-from src.file_utils import copy_files, create_directory, get_input_ids, get_source_names
-from src.utils import get_project_root
+from swinno_bioeconomy_directionality.file_utils import (
+    copy_files,
+    create_directory,
+    get_input_ids,
+    get_source_names,
+)
+from swinno_bioeconomy_directionality.utils import get_project_root
 
 parser = argparse.ArgumentParser()
 parser.add_argument(

### 2024-09-12 src/swinno_helpers.py

@@ -1,85 +1,0 @@
-import pandas as pd
-import numpy as np
-from sqlalchemy import create_engine
-from pathlib import Path
-from src.utils import get_project_root
-
-
-def split_cols(df, col_to_split: str, sep: str):
-    """Splits columns that have joined information (e.g., SNI codes)
-
-    Args:
-        df (pd.DataFrame): The data frame in which the split should occur
-        col_to_split (str): The name of the column to split
-
-    Returns:
-        pd.DataFrame: A new data frame with the original column content split into new columns and the original column removed.
-    """
-    temp_df = df
-    split_cols = (
-        temp_df.loc[:, col_to_split]
-        .astype(str)
-        .str.strip()
-        .str.split(sep, expand=True)
-        .fillna(np.nan)
-    )
-
-    for col in split_cols:
-        temp_df.loc[:, f"{col_to_split}_{col}"] = split_cols.loc[:, col]
-
-    temp_df.drop(labels=col_to_split, axis=1, inplace=True)
-
-    return temp_df
-
-
-def melt_table(df, id_vars: str, col_start: str, value_name: str, **kwargs):
-    """Melts tables into lookup tables
-
-    Args:
-        df (pd.DataFrame): The dataframe to melt
-        id_vars (str): Variables to keep as ids
-        col_start (str): The start of the columns to select
-        value_name (str): The name of the value column in the melted table
-
-    Returns:
-        df (pd.DataFrame): Melted dataframe
-    """
-    temp_df = df.melt(
-        id_vars=id_vars,
-        value_vars=df.columns[df.columns.str.startswith(col_start)],
-        value_name=value_name,
-    )
-
-    temp_df.drop("variable", axis=1, inplace=True)
-    temp_df = temp_df[temp_df.iloc[:, 1] != ""]
-    temp_df.dropna(inplace=True)
-
-    temp_df = temp_df.sort_values(by=id_vars).reset_index(drop=True)
-
-    return temp_df
-
-
-def clean_column_names(df):
-    df.columns = df.columns.str.strip().str.lower().str.replace(" ", "_")
-    return df
-
-
-def clean_import(file_type, file_path, **kwargs):
-    if file_type == "csv":
-        df = pd.read_csv(file_path, **kwargs)
-
-    elif file_type == "xlsx":
-        df = pd.read_excel(file_path, **kwargs)
-
-    else:
-        raise ValueError("Unsupported file type")
-
-    df = clean_column_names(df)
[…]

### 2024-09-12 swinno_bioeconomy_directionality/swinno_helpers.py

@@ -1,0 +1,85 @@
+import pandas as pd
+import numpy as np
+from sqlalchemy import create_engine
+from pathlib import Path
+from swinno_bioeconomy_directionality.utils import get_project_root
+
+
+def split_cols(df, col_to_split: str, sep: str):
+    """Splits columns that have joined information (e.g., SNI codes)
+
+    Args:
+        df (pd.DataFrame): The data frame in which the split should occur
+        col_to_split (str): The name of the column to split
+
+    Returns:
+        pd.DataFrame: A new data frame with the original column content split into new columns and the original column removed.
+    """
+    temp_df = df
+    split_cols = (
+        temp_df.loc[:, col_to_split]
+        .astype(str)
+        .str.strip()
+        .str.split(sep, expand=True)
+        .fillna(np.nan)
+    )
+
+    for col in split_cols:
+        temp_df.loc[:, f"{col_to_split}_{col}"] = split_cols.loc[:, col]
+
+    temp_df.drop(labels=col_to_split, axis=1, inplace=True)
+
+    return temp_df
+
+
+def melt_table(df, id_vars: str, col_start: str, value_name: str, **kwargs):
+    """Melts tables into lookup tables
+
+    Args:
+        df (pd.DataFrame): The dataframe to melt
+        id_vars (str): Variables to keep as ids
+        col_start (str): The start of the columns to select
+        value_name (str): The name of the value column in the melted table
+
+    Returns:
+        df (pd.DataFrame): Melted dataframe
+    """
+    temp_df = df.melt(
+        id_vars=id_vars,
+        value_vars=df.columns[df.columns.str.startswith(col_start)],
+        value_name=value_name,
+    )
+
+    temp_df.drop("variable", axis=1, inplace=True)
+    temp_df = temp_df[temp_df.iloc[:, 1] != ""]
+    temp_df.dropna(inplace=True)
+
+    temp_df = temp_df.sort_values(by=id_vars).reset_index(drop=True)
+
+    return temp_df
+
+
+def clean_column_names(df):
+    df.columns = df.columns.str.strip().str.lower().str.replace(" ", "_")
+    return df
+
+
+def clean_import(file_type, file_path, **kwargs):
+    if file_type == "csv":
+        df = pd.read_csv(file_path, **kwargs)
+
+    elif file_type == "xlsx":
+        df = pd.read_excel(file_path, **kwargs)
+
+    else:
+        raise ValueError("Unsupported file type")
+
[…]

### 2024-09-12 tests/test_clean_codes.py

@@ -1,13 +1,13 @@
 import pandas as pd
 from pandas.testing import assert_frame_equal
-from  src.categorization_helpers import *
+from swinno_bioeconomy_directionality.categorization_helpers import *
 
 input_df = pd.DataFrame(
     {"id": [1, 2, 3, 4, 5], "codes": ["A 0", "   1,23", "A,1", "ABC", " 1, 2, 0"]}
 )
 
-def test_letter_to_code():
 
+def test_letter_to_code():
     expected_df = pd.DataFrame(
         {"id": [1, 2, 3, 4, 5], "codes": ["1 0", "   1,23", "1,1", "123", " 1, 2, 0"]}
     )
@@ -18,7 +18,6 @@ def test_letter_to_code():
 
 
 def test_clean_codes_1_digit():
-
     expected_df = pd.DataFrame(
         {"id": [1, 2, 3, 4, 5], "codes": ["1,0", "1,2,3", "1,1", "1,2,3", "1,2,0"]}
     )
@@ -30,9 +29,7 @@ def test_clean_codes_1_digit():
     assert expected_df["codes"].equals(output_df["codes"])
 
 
-
 def test_clean_codes_3_digit():
-
     input_df = pd.DataFrame(
         {"id": [1, 2, 3, 4], "codes": ["112322,0", "1230", "123123", "101201207"]}
     )
@@ -46,11 +43,9 @@ def test_clean_codes_3_digit():
 
     assert expected_df["codes"].equals(output_df["codes"])
 
-def test_clean_codes_int_handling():
 
-    input_df = pd.DataFrame(
-        {"id": [1, 2, 3], "codes": ["112322,0", "1230", 123123]}
-    )
+def test_clean_codes_int_handling():
+    input_df = pd.DataFrame({"id": [1, 2, 3], "codes": ["112322,0", "1230", 123123]})
     expected_df = pd.DataFrame(
         {"id": [1, 2, 3], "codes": ["112,322,0", "123,0", "123, 123"]}
     )
@@ -61,10 +56,11 @@ def test_clean_codes_int_handling():
 
     assert expected_df["codes"].equals(output_df["codes"])
 
+
 def test_white_space():
-    input_df = pd.DataFrame({"id":["6990001"], "codes":["101 201 207"]})
-    
-    expected_df = pd.DataFrame({"id":["6990001"], "codes":["101,201,207"]})
[…]

### 2024-09-12 tests/test_connect_db.py

@@ -2,7 +2,7 @@ import pytest
 from pathlib import Path
 from sqlalchemy import create_engine
 
-from src.swinno_helpers import connect_swinno_db
+from swinno_bioeconomy_directionality.swinno_helpers import connect_swinno_db
 
 
 def test_connect_swinno_db():

### 2024-09-12 tests/test_copy_files.py

@@ -3,7 +3,8 @@ import tempfile
 from pathlib import Path
 from shutil import copy
 from unittest.mock import patch, MagicMock
-from src.file_utils import copy_files
+from swinno_bioeconomy_directionality.file_utils import copy_files
+
 
 def test_copy_files():
     with tempfile.TemporaryDirectory() as temp_dir:
@@ -25,6 +26,7 @@ def test_copy_files():
         assert destination_file1.exists()
         assert destination_file2.exists()
 
+
 def test_copy_files_permission_error():
     with tempfile.TemporaryDirectory() as temp_dir:
         source_dir = Path(temp_dir, "source")
@@ -41,7 +43,10 @@ def test_copy_files_permission_error():
         with patch("shutil.copy", MagicMock(side_effect=PermissionError)):
             with patch("builtins.print") as mock_print:
                 copy_files("id1", ["file1"], source_dir, destination_dir)
-                mock_print.assert_called_with(f"Error: Permission denied when copying {source_file1} to {destination_file1}")
+                mock_print.assert_called_with(
+                    f"Error: Permission denied when copying {source_file1} to {destination_file1}"
+                )
+
 
 def test_copy_files_no_files():
     with tempfile.TemporaryDirectory() as temp_dir:

### 2024-09-12 tests/test_create_directory.py

@@ -1,7 +1,8 @@
 import shutil
 import tempfile
 from pathlib import Path
-from src.file_utils import create_directory
+from swinno_bioeconomy_directionality.file_utils import create_directory
+
 
 def test_create_directory():
     # Test that directory is created if it doesn't exist

### 2024-09-12 tests/test_get_input_ids.py

@@ -1,6 +1,7 @@
 import pytest
 from pathlib import Path
-from src.file_utils import get_input_ids
+from swinno_bioeconomy_directionality.file_utils import get_input_ids
+
 
 @pytest.fixture
 def input_file(tmp_path):
@@ -20,6 +21,7 @@ def input_file(tmp_path):
     with open(input_file, "w") as f:
         f.write("\n".join(str(id) for id in input_data))
     return input_file
+
 
 def test_get_input_ids(input_file):
     # Test that function returns expected IDs

### 2024-09-12 tests/test_get_project_root.py

@@ -1,4 +1,5 @@
-from src.utils import get_project_root
+from swinno_bioeconomy_directionality.utils import get_project_root
+
 
 def test_is_dir():
     assert get_project_root().is_dir()

### 2024-09-12 tests/test_get_source_names.py

@@ -2,7 +2,8 @@ import os
 import tempfile
 from collections import defaultdict
 import pytest
-from src.file_utils import get_source_names
+from swinno_bioeconomy_directionality.file_utils import get_source_names
+
 
 def test_get_source_names():
     with tempfile.TemporaryDirectory() as temp_dir:
@@ -17,6 +18,7 @@ def test_get_source_names():
         output = get_source_names(lookup_path, input_ids)
         assert output == expected_output
 
+
 def test_get_source_names_empty_input():
     with tempfile.TemporaryDirectory() as temp_dir:
         lookup_path = os.path.join(temp_dir, "lookup.txt")
@@ -29,6 +31,7 @@ def test_get_source_names_empty_input():
         expected_output = {}
         output = get_source_names(lookup_path, input_ids)
         assert output == expected_output
+
 
 def test_get_source_names_invalid_file():
     input_ids = ["1", "2", "3"]

### 2024-09-12 tests/test_melt.py

@@ -1,7 +1,7 @@
 import pandas as pd
 import numpy as np
 from pandas.testing import assert_frame_equal
-from src.swinno_helpers import melt_table
+from swinno_bioeconomy_directionality.swinno_helpers import melt_table
 
 
 test_input_df = pd.DataFrame(

### 2024-09-12 tests/test_split_cols.py

@@ -1,7 +1,7 @@
 import pandas as pd
 import numpy as np
 from pandas.testing import assert_frame_equal
-from src.swinno_helpers import split_cols
+from swinno_bioeconomy_directionality.swinno_helpers import split_cols
 
 test_input_df = pd.DataFrame({"A": ["1 2", np.nan, "0", "1 2"]})
 

### 2024-09-12 swinno_bioeconomy_directionality/config.py
add script to enter all classifications into db

@@ -1,0 +1,26 @@
+from pathlib import Path
+
+from loguru import logger
+
+# Paths
+PROJ_ROOT = Path(__file__).resolve().parents[1]
+logger.info(f"PROJ_ROOT path is: {PROJ_ROOT}")
+
+DATA_DIR = PROJ_ROOT / "data"
+RAW_DATA_DIR = DATA_DIR / "raw"
+INTERIM_DATA_DIR = DATA_DIR / "interim"
+PROCESSED_DATA_DIR = DATA_DIR / "processed"
+EXTERNAL_DATA_DIR = DATA_DIR / "external"
+
+RESULTS_DIR = PROJ_ROOT / "results"
+FIGURES_DIR = RESULTS_DIR / "figures"
+
+# If tqdm is installed, configure loguru with tqdm.write
+# https://github.com/Delgan/loguru/issues/135
+try:
+    from tqdm import tqdm
+
+    logger.remove(0)
+    logger.add(lambda msg: tqdm.write(msg, end=""), colorize=True)
+except ModuleNotFoundError:
+    pass

### 2024-11-28 paper.qmd
update website listings

--- 2024-05-06 paper.qmd
@@
 author:
-  name: Jonas Kreutzer
-  affiliation: 'Department of Economic History, Lund University'
+  name: Philipp Jonas Kreutzer
+  email: philipp_jonas.kreutzer@ekh.lu.se
+  orcid: 0000-0003-4234-9043
+  affiliations: 
+        - id: LU
+          name: Lund University
+          department: Department of Economic History
+          city: Lund
+          country: Sweden
 date: last-modified
@@
 This research underscores the dynamic interplay between historical contexts, shared visions, and innovation trajectories within the bioeconomy transition, and questions the extent to which visions can provide novel directionality to transitions.
+keywords: 
+  - Directionality
+  - Bioeconomy
+  - Forest
+  - Transformative Innovation Policy
 
-csl: 'https://zotero.org/styles/apa?source=1'
 bibliography:
-  - swinno-shades-green.bib
-link-citations: true
-link-bibliography: true
-colorlinks: true
-number-sections: true
-number-depth: 3
+  - swinno-bioeconomy-directionality.bib
+
 format:
-  pdf:
-    documentclass: scrartcl
-    papersize: a4
-    linestretch: 1.25
-    mainfont: "Adobe Garamond Pro"
-    sansfont: "Adobe Garamond Pro Semibold"
-    indent: true
+#   pdf:
+#     documentclass: scrartcl
+#     papersize: a4
+#     linestretch: 1.25
+#     mainfont: "Adobe Garamond Pro"
+#     sansfont: "Adobe Garamond Pro Semibold"
+#     indent: true
+#     csl: apa.csl
+    # link-citations: true
+    # link-bibliography: true
+    # colorlinks: true
+#       
+# number-sections: true
+# number-depth: 3
+#     output-file: "lund_style"
+  elsevier-pdf:
+    keep-tex: true  
+    journal:
+        name: Environmental Innovation and Societal Transitions
+        formatting: review
+        model: 3p
+        cite-style: authoryear
+        highlights:
+            - Highlight 1
+            - Highlight 2 
+            - Highlight 3
+
 execute:
@@
 import scipy.stats as stats
-from pyhere import here
 from IPython.display import display, Markdown
@@
 
-from src.figures import *
-from src.utils import get_project_root, format_table
-from src.data_handling import database
+from swinno_bioeconomy_directionality.utils import format_table
+from swinno_bioeconomy_directionality.data_handling import database
+from swinno_bioeconomy_directionality.config import (
+    PROJ_ROOT,
+    RAW_DATA_DIR,
+    PROCESSED_DATA_DIR,
+)
 
-ROOT = get_project_root()
-raw_data = here("data/raw")
-modified_data = here("data/modified")
 conn = database.connect_swinno_db()
-
-# set all text on plots to be monospace with rcParams
-plt.rcParams["font.family"] = "Space Grotesk"
 
@@
 #| label: Load Data
-raw_data = here("data/raw")
-uncleaned_bioeconomy_query = pd.read_csv(raw_data / "uncleaned_bioeconomy_query.csv")
-bioeconomy = pd.read_csv(raw_data / "id_year_bioeconomy.csv")
+uncleaned_bioeconomy_query = pd.read_csv(RAW_DATA_DIR / "uncleaned_bioeconomy_query.csv")
+bioeconomy = pd.read_csv(RAW_DATA_DIR / "id_year_bioeconomy.csv")
 bioeconomy = bioeconomy.loc[bioeconomy["bioeconomy"] == 1]
@@
 
-bioeconomy_visions = pd.read_csv(raw_data / "bioeconomy_visions.csv")
+bioeconomy_visions = pd.read_csv(RAW_DATA_DIR / "bioeconomy_visions.csv")
 ```
@@
 n_uncleaned_bioeconomy_innovations = uncleaned_bioeconomy_query["sinno_id"].nunique()
-sum_bioeconomy_innovations = swinno["bioeconomy"].sum()
+sum_bioeconomy_innovations = int(swinno["bioeconomy"].sum())
 ```
@@
 
-# Directionality and Visions of Bioeconomy {#sec-literature}
+This article has a dual purpose.
+On one hand, it examines the directionality of Sweden's bioeconomy transition using innovation output data matched to public funding applications.
+On the other hand, it examines the idea that visions are suitable to steer transitions in normatively desired directions.
+First it finds that only about half of significant commercialized innovation in Sweden corresponds to any bioeconomy vision at all, while the remainder can better be described as general improvements to exisiting technology.
+A point here that vision have possibly only recently been formulated coherently enough to develop guiding power.
+In addition, despite the bioresource vision dominating discourse the bioeconomy transition is still highly contested.
+Comparing innovations aligned with each vision shows that the bioresource vision does indeed account for a majority of vision aligned innovations.
+However, by extending the temporal horizon into the last century this paper further shows that innovation alinged with the bioresource vision has dominated other visions since the 1970s, therefore casting doubt on the external directionality of visions formulated at the policy level relative to an endogenous vision shared by key / a majority of innovation producers.
+For the Swedish case study this raises the question of who is horse and who is carrot in drafting and concentrating on a vision for normative directionality.
+For transitions at large this raises additional questions over the effectiveness of visions to provide directionality to transitions, such as to which extent do existing technology sets influence the formulation of vision?
+Can visions for transforming existing socio-technical systems be formulated without the buy-in of regime actors?
 
-Previous literature has investigated directionality of socio-technical systems and emphasized the importance of generating shared visions as tools to provide normative directionality.
-Three distinct visions of desirable transition pathways and outcomes compete for normative directionality of the bioeconomy transition.
-This section introduces key concepts from this literature.
+<!-- TODO: Read up on proponents of bottom up vision generation  -->
 
-## Directionality {#sec-directionality}
+# Literature Review
 
-Technology does not change randomly.
-Instead, technological change is contingent and influences technological change at previous or later points in time.
-It is path-dependent, meaning events at one point in time affect later events, which may even lead to suboptimal outcomes [@david1985ClioEconomicsQWERTY].
-Path-dependency arises due to positive feedback effects, such as increasing returns [@arthur1989CompetingTechnologiesIncreasing], or through extension of possibilities adjacent to exiting technology [@kauffman2000Investigations].
-Instead of more or less progress, each innovation implies a new direction for socio-technical systems [@stirling2011PluralisingProgressIntegrative].
-@andersson2021OutcomesDirectionalityMorphology term this quality of innovation possessing direction _positive_ directionality.
+## Transformative Innovation Policy
 
-As innovation policy took a normative turn under which innovation is needed to solve grand societal challenges [@schot2018ThreeFramesInnovation], the question of how to steer innovation into desired directions became a central issue in innovation research [@kohler2019AgendaSustainabilityTransitions].
-Innovation policy now ought to promote the desired vector of societal development, instead of focusing on providing the right amount of progress [@schot2018ThreeFramesInnovation], lending innovation _normative_ directionality [@andersson2021OutcomesDirectionalityMorphology].
-As innovation systems are embedded in a spatial and temporal context, not all potentially desired pathways are available or feasible.
-Instead, socio-technical systems are constrained by path-dependency [@kemp2022DiversifyingDeepTransitions;@bergek2015TechnologicalInnovationSystems; @stirling2011PluralisingProgressIntegrative], sometimes to the point of being locked-in to specific system states [@unruh2000UnderstandingCarbonLockin].
+### Directionality
 
-Dealing with the question of how to provide and exercise normative transition directionality  Weber and Rohracher [-@weber2012LegitimizingResearchTechnology] synthesize policy insights from innovation systems and multilevel perspective research into a failures framework.
-Therein, directionality failure is understood as the failure to provide direction to the process of search and unclear or missing setting of collective priorities.
-They caution that it should not be confused with shortsighted or flawed anticipation of market developments, as this myopia lacks a view on the direction of change.
-To overcome this failure, that is, to provide directionality, they argue, external requirements need to be internalized by the innovation system in a interpretive process of negotiation between different system actors.
-Reaching this collective understanding of direction, Weber and Rohracher [-@weber2012LegitimizingResearchTechnology p.
-1042] write, is in practice, "...
-achieved by establishing shared future visions".
+#### The Role of Visions
 
-In the two central research streams of sustainability transitions, the innovation systems and the multi-level perspective literature, the importance of visions is highlighted.
-@hekkert2007FunctionsInnovationSystems provide a framework of intervention in innovation on a systemic level, which highlights the importance of guidance of search as one of seven functions innovation system must provide.
-They explain guidance of search as, "those activities within the innovation system that can positively affect the visibility and clarity of specific wants among technology users" [@hekkert2007FunctionsInnovationSystems, 423].
-Changing preferences in society, decides which direction of technological development is most promising through market and government influence and, "..
-an interactive and cumulative process of exchanging ideas between technology producers, technology users, and many other actors" [@hekkert2007FunctionsInnovationSystems, 423].
-@bauer2017TechnologicalInnovationSystems adapts the innovation system functions to the specific case of biorefineries in Sweden and define the influence on the direction of search as "the incentives and/or pressures for organizations to enter the technological field.
-These may come in the form of visions, expectations of growth potential, regulation, policy targets, standards, articulation of demand from leading customers, crises in current business, etc." (p.
-4).
-Using the multi-level perspective, @yang2022ShapingDirectionalitySustainability too highlight the importance of visions in directionality over transformations, stressing the importance of shared visions between niche and regime actors.
-Similarly, @bolton2019GrandVisionsPragmatic illustrate how visions and pragmatic change combined in the development and evolution of a highly integrated European electricity system.
+## Contested Transition Towards Bioeconomy in Sweden
 
-@rotmans2001MoreEvolutionRevolution even go so far to argue that transition management is based on appealing and imaginative long-term visions and @loorbach2010TransitionManagementSustainable argues that shared visions need to be developed and disseminated between transition stakeholders.
-But, as @bolton2019GrandVisionsPragmatic show, visions alone are insufficient to provide directionality.
-In addition to being appealing, visions must also be specific and actors need to buy into them [@grillitsch2019InnovationPolicySystemwidea].
-In summary, visions can be seen as shared imaginaries, mental models and aspirations of desired transitions which are expressed through narratives in discourse between actors [@bauer2018NarrativesBiorefineryInnovation; @holmgren2022WhoseTransformationThis].
+### Differences in Visions
 
-However, one transition can be characterized by multiple, competing visions, as is illustrated by the transition to a bioeconomy.
-The next section sharpens the conceptual understanding of what is meant with a bioeconomy and introduces the three visions.
+###
 
-## Clarification of Bioeconomy Concept
-
-Although the bioeconomy has gained importance as a research, policy and industry focus, it remains unclear what exactly it entails.
-At a high level, the bioeconomy refers to an economy in which raw materials for energy and production processes are provided from biological instead of fossil sources [@bugge2016WhatBioeconomyReview].
-Hopes of increased resource efficiency and renewability as well as promised economic growth especially in regions rich in biological resources abound [@el-chichakli2016PolicyFiveCornerstones].
-However, a such a profound change to economic systems will not be possible without balancing significant tradeoffs between the needs and desires of different human and non-human stakeholders [@dieken2021MultitudesBioeconomiesSystematic].
-@hogbom2021TrilemmaNordicBaltic summarize these tradeoffs as a trilemma between forests ability to provide sufficent wood to replace fossil based materials, carbon sequestration and storage, and the delivery of other eco-system services such as forest biodiversity.
-Resolving these tradeoffs is challenging, as goal conflicts between different development pathways make simulatneous progress on social, ecological and economic goals difficult [@bennich2021BiobasedEconomy2030].
-
-As one of the leading forestry nations in the world, Sweden sees substantial economic and political potential in advancing the bioeconomy .
-In a report commissioned by the Swedish Government, @formas2012SwedishResearchInnovation lays out the plan for a Swedish bioeconomy based on two central pillars.
-It is to be an economy based on,
-
-> \[a\] sustainable production of biomass to enable increased use within a number of different sectors of society. The objective is to reduce climate effects and the use of fossil-based raw materials.
-
-With,
-
-> \[a\]n increased added value for biomass materials, concomitant with a reduction in energy consumption and recovery of nutrients and energy as additional end products.
-The objective is to optimize the value and contribution of ecosystem services to the economy.
-(p.
-9).
-
-While the official Swedish bioeconomy strategy is still under development, the content of the quotes above echos across policy documents and the national forest program [@fischer2020SwedishForestsBioeconomy].
-These pillars reflect the two most influential visions of the bioeconomy concept.
-Notions of a biotechnology and bio-resource oriented bioeconomy are widespread, dominate discourse at various levels [@dieken2021MultitudesBioeconomiesSystematic] and form the main industrial visions [@befort2020GoingDefinitionsUnderstand].
-A third vision, centered around ecology and an economy compatible with the biosphere acts as a counter point.
-
-### Bioeconomic Visions {#sec-bioeconomy-visions}
-
-The application of biotechnology to promote economic growth and provide jobs lends the biotechnology vision its name [@bugge2016WhatBioeconomyReview]. 
-Biotechnology is the application of cellular and / or molecular biological agents in the production or processing of materials in a scientific and engineering informed manner [@yeung2019CurrentResearchBiotechnology].
-Science driven innovation features prominently in this vision, with collaborations between research institutions in academia and industry taking on an important role [@bugge2016WhatBioeconomyReview].
-@vivien2019HijackingBioeconomy present a review of the concept showing that the focus on science as an absolute motor of change also shaped policy recommendations, driven especially by the OECD, towards increasing the speed and diffusion of innovation through partnerships between new start-ups and establish pharmaceutical companies.
-As a policy narrative this vision was particularly potent and widespread at the turn of the 21st century.
-According to @vivien2019HijackingBioeconomy, the intellectual heritage of Schumpeter and Kontradieff play an important role in this vision.
-Technological breakthroughs are poised to solve economic and ecological challenges, for example through genetic engineering.
-Consequently, this vision has a weak sustainability notion, viewing sustainability challenges as temporary, and nature as substitutable given sufficient research and knowledge creation [@vivien2019HijackingBioeconomy].
-
-In more recent years, visions of a bio-resource bioeconomy have taken over as the leading narrative [@allain2022BioeconomicsVsBioeconomy].
-Unlike the biotechnology vision, this conceptualization does not only focus on economic growth and job creation, but attempts to include ecological sustainability challenges [@bugge2016WhatBioeconomyReview].
-Biological resources, not technologies, form the basis of a new economy.
-In essence, there are two ways in which bio-resources are poised to replace fossil resources: as substitution, e.g., biofuels replacing fossil fuels, and as high value added products, e.g., cross-laminated timber used as an alternative to concrete in construction [@vivien2019HijackingBioeconomy].
-To substitute fossil resources requires significant efforts in innovation.
-In contrast to the biotechnology vision, however, innovation is not confined to scientific advances from a narrow biochemical and pharmaceutical field.
-Instead, new and old actors are required to collaborate between and beyond established sectors [@bugge2016WhatBioeconomyReview].
-Although sustainability plays a more important role in this vision of the bioeconomy, this vision is also considered an example of weak sustainability [@damato2017GreenCircularBio; @vivien2019HijackingBioeconomy].
-The reason is the focus of substitution and high biomass demands.
-The consumption of raw materials may even increase, not only for bio-based, but also for fossil-based resources [@asada2020BioeconomicTransitionProjecting].
-
-An ecology focused counter-vision exists, which @bugge2016WhatBioeconomyReview term the bioecology vision, while ecological economists preferentially use the term bioeconomics [@vivien2019HijackingBioeconomy; @allain2022BioeconomicsVsBioeconomy].
-Unlike the previous two visions, the bioecology vision builds on strong sustainability assumptions and emphasizes the need to develop an alternative economic system which respects the limits of the biosphere.
-Normative change towards material sobriety and degrowth to protect and restore ecosystems and biodiversity take precedent over technological and economic ambitions [@vivien2019HijackingBioeconomy;@bugge2016WhatBioeconomyReview].
-
-#### Development of the Term Bioeconomy {#sec-bioeconomy-term-development}
-
-Analysis of the bioeconomy discourse, highlights that the visions have emerged at different times and with different actors at their centers [@vivien2019HijackingBioeconomy; @bugge2016WhatBioeconomyReview].
-And while the two hitherto mentioned visions dominate contemporary discourse towards a forest based bioeconomy [@holmgren2022WhoseTransformationThis; @bauer2018NarrativesBiorefineryInnovation], the bioeconomics concept itself was introduced in the 1920s in a work on fishery by the Russian biologist Baranoff [@vivien2019HijackingBioeconomy].
-Later, in 1975 ecological economist Georgescu-Roegen [-@georgescu-roegen1975BioeconomicAspectsEntropy] used the term to argue for an economy which minimizes its material and energy throughput to respect biological, physical and geochemical limits.
-@vivien2019HijackingBioeconomy argue that the term biotechnology gradually emerged between 1990 and 2000, in the wake of enthusiasm about biotechnological advances, such as the discovery of DNA.
-It gained popularity and international importance when the OECD began increasingly supporting and reshaping its meaning at the turn of the millennium [@vivien2019HijackingBioeconomy].
-According to them, the last concept to enter the discourse was the bioresource oriented bioeconomy.
-While biomass has been used throughout history in various ways, the renewed interest with the intention to replace fossil inputs, for example in European Commission communication since 2010.
-However, @vivien2019HijackingBioeconomy point to the long history of biorefinery as a core aspect of this vision, which was discussed as early as the 1930s.
-Hence, while the organization of the discourse into the labels biotechnology, bioecology and bioresource vision is a recent development, the underlying concepts have a richer history.
-Over time these visions have not only coexisted, but also influenced each other by borrowing from, contributing to, and distancing themselves from the other visions of a bioeconomy [@allain2022BioeconomicsVsBioeconomy].
-And although the exact role played by forests in a bioeconomy transition depends on the desired vision [@kleinschmit2014ShadesGreenSocial], there is no doubt that the Swedish vision places high demands on forests as primary providers of the required biomass.
-Or, as @belandlindahl2015SwedishForestryModel call it a forest sector that operates under a "more of everything" paradigm.
-Whether in the strategy formulated in @formas2012SwedishResearchInnovation, in Sweden's national forest program [@fischer2020SwedishForestsBioeconomy], or in public discourse, a small, well-connected group of actors center the bio-resource vision as legitimate and desirable [@holmgren2022WhoseTransformationThis].
-
-Common ground in ambitions to realize a bioeconomy is the need for diverse types of social, technological and organizational innovation [@el-chichakli2016PolicyFiveCornerstones].
-According to @jankovsky2021InnovationsForestBioeconomy there are four types of innovation crucial for a forest-based bioeconomy: (i) substitute products, (ii) new (bio-based) processes, (iii) new (bio-based) products, (iv) new behavior.
-The need for innovation also features large in Swedish bioeconomy discourse and narratives therein [@bauer2017TechnologicalInnovationSystems; @bennich2021BiobasedEconomy2030], as, "\[t\]echnological innovation is key to a greener future" [@holmgren2022WhoseTransformationThis 42].
-In a survey study sent to private organizations that received EU funding, @lovric2020FactorsDevelopmentInnovations find that the main type of European forest-based bioeconomy innovations are aimed at production processes, followed by goods and services, then business models.
-
-In summary, despite -- or perhaps because of -- bioeconomy lacking a strong definition, the term is used to refer to different concepts, with different aims and objectives, foci, normative assumptions and hence implications for society and environment [@bugge2016WhatBioeconomyReview].
-The summarized visions express themselves as narratives, actor coalitions and discourse artefacts in attempts to provide normative directionality to the transition.
-This normative directionality is expected to translate into positive directionality of technological change through innovation.
-
-# Data and Methods {#sec-data_methods}
+# Data and Methods {#sec-data-methods}
 
@@
 
-### Eco-Innovation Types {#sec-eco_innovation_types}
-
-The focus of this paper is understanding what bioeconomy visions dominate the Swedish innovation system's output.
-However, in order to gain a deeper understanding of the nature of innovation within each bioeconomy vision, I also tagged the innovation with labels of eco-innovation types.
-This classification uses eco-innovation types identified in a recent literature review by @garcia-granero2018EcoinnovationMeasurementReview.
-They review 104 publications and produce a list 30 key Eco-Innovation Performance Indicators of product, process, organizational and marketing innovation.
-
-This initial list was tested on a subset of innovation in SWINNO and augmented with 15 categories.
-Because the list of eco-innovation types was relatively exhaustive from the beginning, few adjustments had to be made.
-These include: expanding the labeling for other sustainability criteria, such as human health and safety, and specific forest related aspects / tasks to gain a more detailed understanding of an innovations aim.
-For example, some innovations aim to protect harvested timber, either at harvest, during transport or storage.
-While @tbl-full-codes shows all new added labels, in the following text those that do not clearly represent an eco-innovation -- following @garcia-granero2018EcoinnovationMeasurementReview "…fewer adverse effects on the environment and more efficient use of resources" (p.
-305) -- were excluded.
-These excluded labels are Gen-manipulated plant, Increased plant yield, Gen-manipulated animal (incl new breed), Human Health and Safety, Automatization, Precision, and Biocide.
-
-The decision process for labelling eco-innovation types was also based on the source texts for each innovation.
-Based on the information contained therein, I labeled the innovation  with one or more categories.
-In case of uncertainty, I added the label uncertain. In case where the text does not mention any eco-innovation aspect, the innovation was labeled as not an eco-innovation.
-Where the source text only mentions that the innovation is less harmful to the environment, the label "Environmental-friendly technologies" was used.
-
 # Results {#sec-results}
@@
 ```{python}
-#| label: Bioeconomy Trend Description
-#| output: asis
+# | label: Bioeconomy Trend Description
+# | output: asis
 share_max = bioeconomy_share["bioeconomy_share"].max()
@@
 Querying the SWINNO database for forest based bioeconomy innovations initially yields `{python} n_uncleaned_bioeconomy_innovations` innovations.
-This number includes false positives.
-During classification, these false positives were removed from the dataset, resulting in `{python} sum_bioeconomy_innovations` innovations which originate from, are used in, or make use of products derived from the forest.
+False positives were removed from the dataset during classification, resulting in `{python} sum_bioeconomy_innovations` innovations which originate from, are used in, or make use of products derived from the forest.
 
@@
 Percentage of forest based bioeconomy innovations to total innovation registered in SWINNO database.
-Shown are the 5 year moving averages for each time series.](results/figures/innovation_trends.pdf){#fig-innovation-trends}
+Shown are the 5 year moving averages for each time series.](results/figures/share_test.tex){#fig-innovation-trends width=100%}
 
@@
 
-certain_bioeconomy_visions_min_year = certain_bioeconomy_visions["year"].min()
+certain_bioeconomy_visions_min_year = certain_bioeconomy_visions["year"].min().item()
 
-certain_bioeconomy_visions_max_year = certain_bioeconomy_visions["year"].max()
+certain_bioeconomy_visions_max_year = certain_bioeconomy_visions["year"].max().item()
 
-n_most_bioeconomy_vision = total_certain_bioeconomy_visions.iloc[0, 2]
+n_most_bioeconomy_vision = total_certain_bioeconomy_visions.iloc[0, 2].item()
 
@@
 
-vision_1_count = total_certain_bioeconomy_visions.iloc[0, 2] 
-vision_2_count = total_certain_bioeconomy_visions.iloc[1, 2] 
-vision_3_count = total_certain_bioeconomy_visions.iloc[2, 2] 
-vision_4_count = total_certain_bioeconomy_visions.iloc[3, 2]
+vision_1_count = total_certain_bioeconomy_visions.iloc[0, 2].item()
+vision_2_count = total_certain_bioeconomy_visions.iloc[1, 2].item()
+vision_3_count = total_certain_bioeconomy_visions.iloc[2, 2].item()
+vision_4_count = total_certain_bioeconomy_visions.iloc[3, 2].item()
 
-vision_1_count = total_certain_bioeconomy_visions.iloc[0, 2] 
-vision_2_count = total_certain_bioeconomy_visions.iloc[1, 2] 
-vision_3_count = total_certain_bioeconomy_visions.iloc[2, 2] 
-vision_4_count = total_certain_bioeconomy_visions.iloc[3, 2]
+vision_1_count = total_certain_bioeconomy_visions.iloc[0, 2].item()
+vision_2_count = total_certain_bioeconomy_visions.iloc[1, 2].item()
+vision_3_count = total_certain_bioeconomy_visions.iloc[2, 2].item()
+vision_4_count = total_certain_bioeconomy_visions.iloc[3, 2].item()
 
@@
 
-n_bioresource_1970_1990 = certain_bioeconomy_visions.query("year.between(1970, 1990) & bioeconomy_vision == 'Bioresource Vision'")["sinno_id"].nunique()
+n_bioresource_1970_1990 = certain_bioeconomy_visions.query(
+    "year.between(1970, 1990) & bioeconomy_vision == 'Bioresource Vision'"
+)["sinno_id"].nunique()
 ```
@@
 ```{python}
-#| label: Count Recycling Innovation
+# | label: Count Recycling Innovation
 
@@
 Innovations can be classified into multiple visions simulatenously.
-The sum of all vision counts does not match the sum of all bioeconomy innovations.](results/figures/vision_count.pdf){#fig-bioeconomy-trends}
+The sum of all vision counts does not match the sum of all bioeconomy innovations.](results/figures/vision_plot.tex){#fig-bioeconomy-trends width=100%}
 
@@
 
-### How many eco-innovations are there?
+# Discussion {#sec-discussion}
 
-```{python}
-# | label: merge eco-innovation with bioeconomy visions
+This discussion first turns towards the directionality of the Swedish forest bioeconomy before extracting insights from the case on the interplay between visions and directionality in sustainability transitions.
 
-eco_innovation_merged_bv = certain_bioeconomy_visions.merge(
-    eco_innovations, on="sinno_id", validate="many_to_many"
-)
-
-eco_innovation_merged_bv["innovation_type_group"] = np.where(
-    eco_innovation_merged_bv["innovation_type_code"].str[0] == "1",
-    "Product",
-    np.where(
-        eco_innovation_merged_bv["innovation_type_code"].str[0] == "2",
-        "Process",
-        np.where(
-            eco_innovation_merged_bv["innovation_type_code"].str[0] == "3",
-            "Organizational",
-            np.where(
-                eco_innovation_merged_bv["innovation_type_code"].str[0] == "4",
-                "Marketing",
-                "No Eco-Innovation",
-            ),
-        ),
-    ),
-)
-
-eco_innovation_by_bioeconomy_visions = (
-    eco_innovation_merged_bv.groupby(["bioeconomy_vision", "innovation_type_code"])
-    .size()
-    .to_frame("count")
-    .reset_index()
-)
-
-eco_innovation_by_bioeconomy_visions["innovation_type_group"] = np.where(
-    eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0] == "1",
-    "Product",
-    np.where(
-        eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0] == "2",
-        "Process",
-        np.where(
-            eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0] == "3",
-            "Organizational",
-            np.where(
-                eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0]
-                == "4",
-                "Marketing",
-                "No Eco-Innovation",
-            ),
-        ),
-    ),
-)
-```
-
-```{python}
-# | label: chi-square-vision-eco
-other_sustainability = {501, 701, 108, 109, 111}
-eco_innovation_filtered = eco_innovation_merged_bv[
-    ~(
-        eco_innovation_merged_bv["innovation_type_code"]
-        .astype(int)
-        .isin(other_sustainability)
-    )
-]
-# Create the groups for the cross tab
-eco_innovation_filtered["eco_innovation"] = (
-    eco_innovation_filtered["innovation_type_group"] != "No Eco-Innovation"
-)
-eco_innovation_filtered["vision_neutral"] = (
-    eco_innovation_filtered["bioeconomy_vision"] == "Vision Neutral"
-)
-
-# Calculate the chi squared test
-vision_eco_cross_tab = (
-    eco_innovation_filtered.drop_duplicates(subset="sinno_id")
-    .groupby(["vision_neutral", "eco_innovation"])
-    .agg(count=("sinno_id", "nunique"))
-    .unstack()
-)
-chi2_stat, p_val, dof, expected = stats.chi2_contingency(vision_eco_cross_tab)
-```
-
-How does each vision contribute to lessening the environmental impact of the forest bioeconomy?
-While a quantitative assessment of technological impact on the environment comes with a host of challenges, the labelling of eco-innovations at least allowed a glimpse at what could be expected from each vision.
-@fig-eco-innovation-bioeconomy-vision shows the count of eco-innovation associated with each bioeconomy vision by type.
-As before, the total sum of all eco-innovation across all bioeconomy visions outnumbers the sum of innovations in the bioeconomy.
-A single innovation can be more than one eco-innovation type, for example, if it reduces the amount of raw material needed by recycling waste.
-Product and Process innovation were by far the most frequent across all visions.
-The bioresource vision was heavily associated with improving raw-material utilization, while the Bioecology Vision is most frequently associated with recycling.
-At least, this suggests that the definitions for each Vision and the classification scheme are in agreement.
-
-```{python}
-#| label: fig-eco-innovation-bioeconomy-vision
-#| fig-cap: Count of Eco-Innovation Types by Bioeconomy Vision
-p = plot_eco_innovation_by_bioeconomy_vision(
-    eco_innovation_by_bioeconomy_visions, codes_df
-)
-
-sns.move_legend(p, "lower center", ncol=5)
-# TODO: perhaps a per cent display by group size would be interesting too?
-```
-
-A noteworthy feature of innovation lacking alignment to a bioeconomy vision was that this category has by far the most non-eco-innovations.
-The chi-squared test of independence revealed a statistically significant association between an innovation being aligned with a vision and it being an eco-innovation ($\chi^2 =$ `{python} round(chi2_stat, 3)` df = `{python} dof`, p < 0.05).
-
-```{python}
-# | label: tbl-cross-table-eco-visions
-# | tbl-cap: Cross Table of Unique Innovations Classified as Vision Neutral And Eco-Innovation
-# Headers
-headers = ["Non-Eco-Innovation", "Eco-Innovation"]
-index = ["Bioeconomy Vision", "Vision Neutral"]
-
-# Generate the table
-cross_table = tabulate(
-    vision_eco_cross_tab.values, headers=headers, showindex=index, tablefmt="pipe"
-)
-
-display(Markdown(cross_table))
-```
-
-# Discussion {#sec-discussion}
+## Positive Directionality of the Swedish Forest-Based Bioeconomy and Possible Future Transition Paths
 
@@
 
+## Visions as Instruments of
+
+While this paper cannot proof conclusively that the visions of future bioeconomy pathways are reformulations of available technologies, the innovation output does suggest that at least in this case the vision dominating bioeconomy transition discours is also the vision most innovation output aligns with.
+Considering that @holmgren2022WhoseTransformationThis showed that the discourse is captured by
+
+The fact that innovations aligned with the bioresource vision have been commercialized since the
+
 ## Limitations {#sec-limitations}
@@
 Similarily, the positive directionality measured in this paper should be seen more as a the vector of change embedded in the transition to a bioeconomy, rather than a scalar of its progress.
+
+# Policy Implications {#sec-policy}
+
+Some recommendations that follow from this study directly for Swedish policy makers include:
+
+1.
+To align innovation output with key aspects of the Swedish vision for a forest-based bioeconomy, there needs to be more focus on high-tech conversion of biomass to high-value added products including, but not limited to, biorefinery.
+2. Technological innovation whitin forestry has mainly focused on mechanization of harvesting. More proactive policy interventions can help spur innovation facilitating other silvioculture practices.
 
@@
 ```{python}
-#| label: tbl-full-codes
-#| tbl-cap: List of Eco-Innovation Labels
+# | label: tbl-full-codes
+# | tbl-cap: List of Eco-Innovation Labels
 format_table(codes_df.iloc[:5, :])

### 2025-04-25 swinno_bioeconomy_directionality/analysis/create_bioeconomy_share.py
fix bioeconomy data

@@ -7,6 +7,8 @@ from swinno_bioeconomy_directionality.config import (
     PROCESSED_DATA_DIR,
 )
 
+from swinno_bioeconomy_directionality.data_handling.io import DataLoader
+
 
 def calculate_5yma(data: pl.DataFrame, col: str | List[str]) -> pl.DataFrame:
     return data.with_columns(
@@ -15,7 +17,7 @@ def calculate_5yma(data: pl.DataFrame, col: str | List[str]) -> pl.DataFrame:
 
 
 if __name__ == "__main__":
-    swinno_bioeconomy = pl.read_csv(RAW_DATA_DIR / "id_year_bioeconomy.csv")
+    swinno_bioeconomy = DataLoader(RAW_DATA_DIR).load_id_year_bioeconomy()
 
     swinno_bioeconomy = (
         swinno_bioeconomy.group_by("year")
@@ -23,8 +25,8 @@ if __name__ == "__main__":
             pl.col("bioeconomy").sum().alias("n_bioeconomy"),
             pl.col("sinno_id").n_unique().alias("n_innovation"),
         )
-        .with_columns(share_bioeconomy=pl.col("n_bioeconomy") / pl.col("n_innovation"))
+        .with_columns(pct_bioeconomy=pl.col("n_bioeconomy") / pl.col("n_innovation"))
         .sort("year")
-        .pipe(calculate_5yma, col=["n_innovation", "n_bioeconomy", "share_bioeconomy"])
+        .pipe(calculate_5yma, col=["n_innovation", "n_bioeconomy", "pct_bioeconomy"])
     )
     swinno_bioeconomy.write_csv(PROCESSED_DATA_DIR / "swinno_bioeconomy_share.csv")

### 2025-04-25 swinno_bioeconomy_directionality/analysis/pull_data.py

@@ -1,19 +1,45 @@
 from swinno_bioeconomy_directionality.data_handling import database
-import pandas as pd
-
+import polars as pl
 from swinno_bioeconomy_directionality.config import RAW_DATA_DIR
 
-if __name__ == "__main__":
+
+def pull_data():
     conn = database.connect_swinno_db()
 
     # Get sinno ids with year and bioeconomy indicator
     sinno_ids = database.query_sinno_ids(conn)
     sinno_ids.write_csv(RAW_DATA_DIR / "id_year_bioeconomy.csv")
 
+    swinno = pl.read_database(
+        """
+        select sinno_id, year_of_commercialization as year, innovation_name_in_swedish as name
+        from innovation;
+        """,
+        conn,
+    )
+    swinno.write_csv(RAW_DATA_DIR / "swinno_data.csv")
+
     bioeconomy_visions = database.query_bioeconomy_visions(conn)
-    bioeconomy_visions.to_csv(RAW_DATA_DIR / "bioeconomy_visions.csv", index=False)
+    bioeconomy_visions.write_csv(RAW_DATA_DIR / "bioeconomy_visions.csv")
+
+    recycle_innovation = pl.read_database(
+        """
+        SELECT *
+        FROM classification_codes
+        LEFT JOIN eco_innovations ON classification_codes.Code = eco_innovations.innovation_type
+        WHERE eco_innovations.innovation_type = 107 OR eco_innovations.innovation_type = 206 
+        AND sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes LIKE '%not forest%')
+        """,
+        conn,
+    )
+    recycle_innovation.write_csv(RAW_DATA_DIR / "recycling_innovation.csv")
 
     uncleaned_bioeconomy = database.query_uncleaned_bioeconomy(conn)
-    uncleaned_bioeconomy.to_csv(
-        RAW_DATA_DIR / "uncleaned_bioeconomy_query.csv", index=False
-    )
+    uncleaned_bioeconomy.write_csv(RAW_DATA_DIR / "uncleaned_bioeconomy_query.csv")
+
+    cleaned_bioeconomy = database.query_clean_bioeconomy(conn)
+    cleaned_bioeconomy.write_csv(RAW_DATA_DIR / "id_year_bioeconomy.csv")
+
+
+if __name__ == "__main__":
+    pull_data()

### 2025-04-25 swinno_bioeconomy_directionality/categorization_helpers.py

@@ -1,51 +1,0 @@
-import pandas as pd
-from pathlib import Path
-import re
-
-
-def clean_codes(df, column, code_digits, remove_whitespace=True, drop_na=True):
-    if drop_na:
-        df = df.dropna(subset=[column]).copy()
-
-    df.loc[:, column] = df[column].astype(str)
-
-    df.loc[:, column] = df[column].apply(lambda x: re.sub(r";", r",", x))
-
-    df.loc[:, column] = df[column].apply(
-        lambda x: re.sub(r"\b(0)\b", r"0" * code_digits, x)
-    )
-    df.loc[:, column] = df[column].apply(
-        lambda x: re.sub(r"\b(9)\b", r"9" * code_digits, x)
-    )
-
-    df.loc[:, column] = df[column].apply(
-        lambda x: re.sub(rf"(\d{ {code_digits} })(?!$)", r"\1,", str(x))
-    )
-
-    df.loc[:, column] = df[column].apply(lambda x: re.sub(r",,", r",", str(x)))
-
-    if remove_whitespace:
-        df.loc[:, column] = df[column].apply(lambda x: re.sub(r"\s", "", x))
-        df.loc[:, column] = df[column].str.strip()
-
-    return df
-
-
-def replace_letters_codes(input_df, col):
-    d = {"A": "1", "B": "2", "C": "3"}
-    input_df[col] = input_df[col].str.replace(
-        "(A|B|C)", lambda x: d[x.group()], regex=True
-    )
-    return input_df
-
-
-def check_duplicates(df, output_name, output_path, subset="sinno_id"):
-    duplicates = df.loc[df.duplicated(subset=subset, keep=False), :]
-    duplicates = duplicates.sort_values(by=subset)
-
-    if duplicates.empty:
-        print("No duplicates found.")
-    else:
-        print(f"{len(duplicates)} duplicates found.")
-        df.to_csv(output_path)
-        print(f"{output_name} written to csv.")

### 2025-04-25 swinno_bioeconomy_directionality/data_handling/combine_categories.py

@@ -1,88 +1,226 @@
-import pandas as pd
-
-from swinno_bioeconomy_directionality.categorization_helpers import *
-from swinno_bioeconomy_directionality.config import RAW_DATA_DIR, INTERIM_DATA_DIR
-from swinno_bioeconomy_directionality.swinno_helpers import *
+import polars as pl
+from pathlib import Path
 
 
-engine = connect_swinno_db()
+from swinno_bioeconomy_directionality.config import RAW_DATA_DIR, PROCESSED_DATA_DIR
+from swinno_bioeconomy_directionality.swinno_helpers import connect_swinno_db
+
+
+def load_tag_data(tags_path: Path) -> pl.DataFrame:
+    """Load manually tagged innovation data."""
+    tag_files = tags_path.glob("*.csv")
+    tag_dfs = [
+        pl.read_csv(
+            p, columns=["sinno_id", "bioeconomy_vision", "innovation_type", "notes"]
+        )
+        for p in tag_files
+    ]
+    return pl.concat(tag_dfs)
+
+
+def load_final_classifications(path: Path) -> pl.DataFrame:
+    """Load finalized innovation classifications."""
+    files = path.glob("*innovations-to-check.csv")
+    dfs = [
+        pl.read_csv(
+            f,
+            columns=[
+                "sinno_id",
+                "bioeconomy_vision",
+                "innovation_type",
+                "article_checked",
+                "notes",
+            ],
+        )
+        for f in files
+    ]
+    return pl.concat(dfs)
+
+
+def resolve_classifications(
+    tags_df: pl.DataFrame, final_df: pl.DataFrame
+) -> pl.DataFrame:
+    """Merge uncertain and final classifications, giving preference to checked entries."""
+    # Include manually tagged entries that aren't in the final classification
+    missing = tags_df.filter(~pl.col("sinno_id").is_in(final_df["sinno_id"]))
+    merged = pl.concat([final_df, missing], how="diagonal")
+
+    # Prefer duplicates with checked articles
+    duplicates = merged.filter(pl.col("sinno_id").is_duplicated()).sort("sinno_id")
+    preferred = duplicates.filter(pl.col("article_checked").is_not_null())
+
+    # Combine non-duplicates and preferred duplicates
+    unique_entries = merged.filter(~pl.col("sinno_id").is_duplicated())
+    return pl.concat([unique_entries, preferred])
+
+
+def clean_codes(
+    df: pl.DataFrame,
[…]

### 2025-04-25 swinno_bioeconomy_directionality/data_handling/database.py

@@ -395,28 +395,45 @@ def query_total_innovation_counts(conn: Engine):
 
 def query_bioeconomy_visions(conn):
     query = """
-  SELECT bv.sinno_id
-  , codes.Category AS bioeconomy_vision
-  , i.year_of_commercialization AS "year"
+  WITH all_visions AS (
+  SELECT 
+    bv.sinno_id,
+    codes.Category AS bioeconomy_vision,
+    i.year_of_commercialization AS "year"
   FROM bioeconomy_visions AS bv
   JOIN [classification_codes] AS codes ON codes.code = bv.bioeconomy_vision
   JOIN innovation AS i ON i.sinno_id = bv.sinno_id
-  WHERE bv.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
+  
   UNION
-  SELECT DISTINCT(ei.sinno_id)
-  , 'Bioecology Vision' AS bioeconomy_vision
-  , i.year_of_commercialization AS "i.year"
+  
+  SELECT DISTINCT(ei.sinno_id),
+    'Bioecology Vision' AS bioeconomy_vision,
+    i.year_of_commercialization AS "year" 
   FROM eco_innovations AS ei
   JOIN innovation AS i ON i.sinno_id = ei.sinno_id
   WHERE ei.innovation_type IN (206, 107)
   AND ei.sinno_id NOT IN (
     SELECT ei2.sinno_id
     FROM eco_innovations as ei2
-    WHERE ei2.innovation_type IN (999, '000')
-    );
+    WHERE ei2.innovation_type IN (999, '000')  -- Consistent quoting of string values
+  )
   -- adds recycling innovation to bioecology
+)
+
+SELECT * 
+FROM all_visions 
+WHERE sinno_id NOT IN (
+  SELECT DISTINCT(sinno_id) 
+  FROM categorization_notes 
+  WHERE notes NOT LIKE '%not forest%' 
+) 
+AND sinno_id NOT IN (
+  SELECT DISTINCT(sinno_id) 
+  FROM bioeconomy_visions 
+  WHERE bioeconomy_vision = 0 
+)
   """
-    return pd.read_sql(query, conn)
+    return pl.read_database(query, conn)
 
 
 def query_eco_innovations(conn):
@@ -432,7 +449,7 @@ WHERE
 innovation_type NOT IN (701, 602, 601, 501, 111, 109, 108)
 ;
 """
[…]

### 2025-04-25 swinno_bioeconomy_directionality/data_handling/io.py

@@ -1,0 +1,95 @@
+from pathlib import Path
+import polars as pl
+
+from swinno_bioeconomy_directionality.config import (
+    RAW_DATA_DIR,
+)
+from swinno_bioeconomy_directionality.data_handling import process_data
+
+
+def ensure_directory_exists(directory: Path):
+    directory.mkdir(parents=True, exist_ok=True)
+
+
+class DataLoader:
+    def __init__(self, data_dir: Path):
+        self.data_dir = Path(data_dir)
+
+    def _validate_file(self, file_path):
+        """
+        Validate if a file exists.
+
+        Args:
+            file_path (Path): Path to the file.
+
+        Raises:
+            FileNotFoundError: If the file does not exist.
+        """
+        if not file_path.exists():
+            raise FileNotFoundError(f"File '{file_path}' does not exist.")
+
+    def load_aggregated_collaborations(self, clean: bool = True) -> pl.DataFrame:
+        file_path = self.data_dir / "aggregated_collaborations.csv"
+
+        self._validate_file(file_path)
+
+        aggregated_collaborations = pl.read_csv(
+            file_path, schema_overrides={"sinno_id": pl.String}
+        )
+
+        if clean:
+            aggregated_collaborations = aggregated_collaborations.pipe(
+                process_data.clean_collaboration_data
+            )
+
+        return aggregated_collaborations
+
+    def load_swinno(
+        self, file_name: str = "swinno_data.csv", clean: bool = True, **kwargs
+    ) -> pl.DataFrame:
+        file_path = self.data_dir / file_name
+
+        self._validate_file(file_path)
+
+        return pl.read_csv(
+            file_path,
+            schema_overrides={
+                "sinno_id": pl.String,
+                "year": pl.Int64,
+                "name": pl.String,
+            },
+        )
+
+    def load_id_year_bioeconomy(self) -> pl.DataFrame:
+        file_path = RAW_DATA_DIR / "cleaned_bioeconomy_query.csv"
+
+        self._validate_file(file_path)
+
+        id_year_bioeconomy = pl.read_csv(
+            file_path,
+            schema_overrides={
+                "sinno_id": pl.String,
+                "year": pl.Int64,
+                "bioeconomy": pl.Int64,
+            },
+        )
+
+        return id_year_bioeconomy
+
+    def load_visions(self, file_name="bioeconomy_visions.csv"):
+        file_path = self.data_dir / file_name
+
+        self._validate_file(file_path)
+
[…]

### 2025-04-25 swinno_bioeconomy_directionality/data_handling/process_data.py

@@ -1,0 +1,104 @@
+from pathlib import Path
+from typing import Dict, List, Optional, Tuple, Union
+
+import polars as pl
+import polars.selectors as cs
+
+from swinno_bioeconomy_directionality.config import INTERIM_DATA_DIR
+
+
+def clean_collaboration_data(data: pl.DataFrame) -> pl.DataFrame:
+    return data.unique("sinno_id").select(
+        [
+            "sinno_id",
+            "firm",
+            cs.starts_with("collaborator"),
+        ],
+    )
+
+
+def clean_taalbi_column_names(data: pl.DataFrame) -> pl.DataFrame:
+    return data.rename({"SINNO_ID": "sinno_id", "Firm level 2": "firm"}).rename(
+        lambda c: c.lower().strip().replace(" ", "_")
+    )
+
+
+def clean_taalbi_aggregation(data: pl.DataFrame) -> pl.DataFrame:
+    return (
+        data.pipe(clean_taalbi_column_names)
+        .with_columns(pl.all().cast(pl.String))
+        .with_columns(pl.col("*").replace("0", None))
+        .pipe(clean_collaboration_data)
+        .with_columns(cs.string().str.strip_chars())
+    )
+
+
+def split_columns(data: pl.DataFrame) -> pl.DataFrame:
+    return (
+        data.with_columns(
+            cs.string()
+            .exclude("sinno_id")
+            .str.split(";")
+            .list.eval(pl.element().str.strip_chars())
+        ).pipe(_pad_node_list)
+        # for each list columns apply the pad_node_list function
+    )
+
+
+def _pad_node_list(data: pl.DataFrame) -> pl.DataFrame:
+    """
+    Pad all string list columns with None if they have exactly 2 elements.
+
+    Args:
+        data (pl.DataFrame): Input DataFrame with list columns
+
+    Returns:
+        pl.DataFrame: DataFrame with padded list columns
+    """
+    # Get list of column names that are string lists
+    list_cols = data.select(cs.by_dtype(pl.List(pl.String))).columns
+
+    # Create expressions for each list column
+    expressions = [
+        pl.when(pl.col(col).list.len() == 2)
+        .then(pl.col(col).list.concat([None]))
+        .otherwise(pl.col(col))
+        .alias(col)
+        for col in list_cols
+    ]
+
+    return data.with_columns(expressions)
+
+
[…]

### 2025-04-25 20250409-pjk-CIRCLE.qmd
update for talk at circle

---
title: "Unpacking Directionality: How Aligned Are Swedish Innovation with Bioeconomy Visions?"
author: Philipp Jonas Kreutzer
date: last-modified
bibliography: swinno-bioeconomy-directionality.bib
csl: 'https://zotero.org/styles/apa?source=1'
format:
    pjk_theme-revealjs
---

## Forest-Based Bioeconomy {background-opacity=0.25 background-image="/assets/images/forwarder.jpeg" .center}

Transition from fossil-based economy to an economy built on bio-based material and fuel resources, but highly contested transition arena with competing visions

[@holmgren2022WhoseTransformationThis]{.slide-cite}

## Directionality in Innovation Systems

Directionality as **possibility** [@stirling2023TransformingImaginationsMultiple], **normative governance** [@degraaff2025PoliticsDirectionalityInnovation] and **systemic outcomes** [@andersson2021OutcomesDirectionalityMorphology].

:::{.fragment}
#### Achieved Through Visions 
Reaching a collective understanding of direction in practice, [**"... achieved by establishing shared future visions"**.]{.text-royal-blue-900} [@weber2012LegitimizingResearchTechnology p. 1042]
:::

## Three Visions for Normative Directionality

::: {.columns}
::: {.column width="33%"}

### [Bioresource]{.text-batlow-darkgreen}

![](/assets/images/bioresource.svg)
:::
::: {.column width="33%"}

### [Biotechnology]{.text-batlow-forestgreen}

![](/assets/images/biotechnology.svg)

:::
::: {.column width="33%"}

### [Bioecology]{.text-batlow-darkblue}

![](/assets/images/bioecology.svg)

:::
:::

Classification based on @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy

## RQ {.center}

<!-- The aim of this article is to provide a deeper understanding of the emerging Swedish bioeconomy by investigating (1) who are the key actors, and (2) how they have shaped the direction of the Swedish bioeconomy to date. Understanding by whom and how the Swedish bieconomy is shaped and institutionalized will fill an important knowledge gap and can help to open up the (forest) bioeconomy debate for discussion, contestation and remaking. -->

[1.]{.text-gray-600} Beyond discourse, [what visions do commercialized innovations in Sweden's forest-based bioeconomy align with?]{.text-royal-blue-600}

[2.]{.text-gray-600} What does Sweden's transition to a forest-based bioeconomy suggest for [the interplay of normative visions and commercialized innovation?]{.text-royal-blue-600}

# Data From SWINNO Database {.inverted}

## Significant Swedish Innovations

::: {.columns}
::: {.column}
~ 5000 unique innovations

Literature-Based Innovation Output method

from 15 independent trade journals [@sjoo2014DatabaseSwedishInnovations]

:::
::: {.column}
![Example Page From a Source Article](/assets/images/swinno-example-1.jpg)
:::
:::

## Defining the Bioeconomy {.smaller}

::: {.columns}
::: {.column}
|SNI Code | Description                            |
|--------:|:---------------------------------------|
|02       | Forestry and related services
|20       | Wood and wood product manufacturing except furniture
|21       | Pulp, paper and paper product manufacturing
|36       | Furniture manufacturing; other manufacturing
:Key Sectors Used in Query {#tbl-query}

:::
::: {.column}
| Swedish           | English                    |
|-------------------|----------------------------|
| virke             | timber                     |
| cellulos          | cellulose                  |
| lignin            | lignin                     |
| spån              | chip                       |
| bark              | bark                       |
| levulinsyra       | levulinic acid             |
| furfural          | furfural                   |
| svarttjära        | black tar                  |
| svartlut          | black liquor               |
| växtbas           | plant-based                |
| ved               | wood                       |
| trä               | timber                     |
| skog              | forest                     |
| biobränsle        | biofuel                    |
| biologiskt        | biological                 |
| nedbrytbar        | biodegradable              |
| papper            | paper                      |
| karton            | carton                     |
| lyocell            | lyocell |
: Swedish Keywords used in Query: `WHERE description LIKE %keyword% OR ...` {#tbl-keywords}
:::
:::

# Results {.inverted}

## Bioeconomy Innovation Has Declined After 1970s

![a. A total number of innovation registered in SWINNO database. b. Forest-based bioeconomy innovations registered in SWINNO database. c. Percentage of forest based bioeconomy innovations to total innovation registered in SWINNO database. Shown are the 5 year moving averages for each time series.](results/figures/fig_innovation_trends.svg){#fig-innovation-trends}

:::notes
Querying the SWINNO database for forest based bioeconomy innovations **initially yields 817 innovations**.False positives were removed from the dataset during classification, resulting in **811 innovations** which originate from, are used in, or make use of products derived from the forest.
:::

## Bioresource Was Biggest Vision Category

But most innovations did not align with any vision

![Counts of innovation classification by Bioeconomy Vision Category](/results/figures/fig_bioeconomy_vision_counts.svg){#fig-bioeconomy-trends}

## Bioecology and Biotechnology Vision Innovations Singular Events for Producers

![Counts of Producing Firms by Number of Innovation Aligned With Bioeconomy Visions.](results/figures/fig_vision_producer_counts.svg){#fig-bioeconomy-vision-producer-counts width=100%}

:::notes
| Rank | Bioresource Vision | Not Bioeconomy |
|------|-------------------|----------------|
| 1    | Sunds (7)         | Atlas Copco (10) |
| 2    | MoDo (5)          | Sandvik (5)    |
| 3    | Billerud (5)      | LKAB (5)       |
| 4    | Kamyr (5)         | Asea/ABB (4)   |
:::

## [Bioresource Vision]{.text-batlow-darkgreen} {background-video="/assets/images/harvester.mp4" background-video-loop=true background-video-muted=true background-opacity=0.4}

1970-1990:

Intensification of Harvesting **Development Block** [@dahmen1988DevelopmentBlocksIndustrial]

Energy & Chemical Pollution Reduction in Pulp and Paper

. . .

2000-2021:

New Materials

## {.center}

::: {.columns}
::: {.column width="33%"}

### [Bioresource]{.text-batlow-darkgreen}

[1970-1990:]{.text-gray-600}

Intensification of Harvesting **Development Block** [@dahmen1988DevelopmentBlocksIndustrial]

Energy Usage & Chemical Pollution Reduction in Pulp and Paper

[2000-2021:]{.text-gray-600}

New Materials

:::
::: {.column width="33%"}

### [Biotechnology]{.text-batlow-forestgreen}

Mostly applications to clean water in pulp and paper

New materials from cellulose

:::
::: {.column width="33%"}

### [Bioecology]{.text-batlow-darkblue}

74% Recycling related

Decommodification of value chains through regional identity products

:::
:::

# Take Aways

1. [The bioeconomy is declining]{.text-royal-blue-600} relative to total innovation, casting doubts on the overall transition to a bioeconomy.

1. [Bioresource Vision strongest among commercialized innovation]{.text-royal-blue-600}, but overlap between visions and open development paths.

1. [Discourse dominance and innovation output likely strongly intertwined]{.text-royal-blue-600}, but other drivers of innovation matter at least as much.

# References

::: {#refs}
:::

# Appendix

## Classification Matrix {.scrollable}

| Corresponding Vision | Source article explicitly mentions that innovation:                          |
|----------------------|-------------------------------------------------------------------------------------------------------|
| Biotechnology    | Develops new, or modifies existing, organism, biological system, or process              |
|           | Uses cellular or molecular process to solve problem or in production process             |
|           | Applies biological agent in scientific or engineering based principle to produce product or service  |
|           | Based on molecular biology, biochemistry, cell biology, embryology, genetics or microbiology research |
|           | Provides method for research in the above areas                            |
|||
| Bioresource     | Uses biomass to substitute other input                                |
|           | Refines biomass through mechanical, chemical or other process                     |
|           | Converts and upgrades bioresource                                   |
|           | Optimizes land use                                          |
|           | Includes degraded land in production of biofuels                           |
|           | Intensifies use of biomass, including from waste                           |
|||
| Bioecology      | Prioritizes biodiversity, conservation, ecosystem or soil health                   |
|           | Integrates production systems                                     |
|           | Develops high quality product with territorial identity                        |
|           | Reuses or recycles waste                                       |
|           | States compatibility with biosphere as motivation or aim                       |
|           | States need for social or ecological counter-expertise relative to other sociotechnical innovations |
|           | Promotes or uses a circular production mode                              |
|           | Promotes or uses a self-sustained production mode                           |

: Decision matrix used to classify innovation. {#tbl-decision-matrix tbl-colwidths="[20,80]"}

## Top Five Producers

| Rank | Bioresource Vision | Vision Neutral |
|------|-------------------|----------------|
| 1    | Sunds (7)         | Atlas Copco (10) |
| 2    | MoDo (5)          | Sandvik (5)    |
| 3    | Billerud (5)      | LKAB (5)       |
| 4    | Kamyr (5)         | Asea/ABB (4)   |
: Top Five Producers and their produced innovation by vision category {#tbl-top-producers}

### 2025-04-25 _code_setup.qmd
update for talk at circle

```{python}
# | label: Imports

import polars as pl
import numpy as np
import scipy.stats as stats
from IPython.display import display, Markdown
import matplotlib.pyplot as plt
import matplotlib.ticker as mtick
import seaborn as sns
from tabulate import tabulate

from swinno_bioeconomy_directionality.utils import format_table

from swinno_bioeconomy_directionality.data_handling.io import DataLoader
from swinno_bioeconomy_directionality.data_handling import database
from swinno_bioeconomy_directionality.config import (
    PROJ_ROOT,
    RAW_DATA_DIR,
    PROCESSED_DATA_DIR,
)

conn = database.connect_swinno_db()

sns.set_theme(style="white")
```

```{python}
# | label: Load Data
from_raw = DataLoader(RAW_DATA_DIR)
from_processed = DataLoader(PROCESSED_DATA_DIR)

uncleaned_bioeconomy_query = pl.read_csv(
    RAW_DATA_DIR / "uncleaned_bioeconomy_query.csv", ignore_errors=True
)

id_year_bioeconomy = from_raw.load_id_year_bioeconomy()

# bioeconomy_ids = id_year_bioeconomy.filter(pl.col("bioeconomy") ==  "1")

swinno = from_raw.load_swinno()

bioeconomy_visions = from_processed.load_visions()
```

```{python}
# | label: Innovation Counts
n_swinno = swinno.select("sinno_id").n_unique()
year_min = swinno["year"].min()
year_max = swinno["year"].max()

n_uncleaned_bioeconomy_innovations = uncleaned_bioeconomy_query["sinno_id"].n_unique()
sum_bioeconomy_innovations = int(id_year_bioeconomy["bioeconomy"].sum())
```

```{python}
# | label: Calculate Bioeconomy Share

def calculate_pct_bioeconomy(swinno):
    return (
        id_year_bioeconomy.group_by("year")
        .agg(
            pl.col("sinno_id").n_unique().alias("n_innovation"),
            pl.col("bioeconomy").sum().alias("n_bioeconomy"),
        )
        .with_columns(pct_bioeconomy=pl.col("n_bioeconomy") / pl.col("n_innovation"))
    ).sort("year")

pct_bioeconomy = calculate_pct_bioeconomy(swinno)
```

```{python}
# | label: Bioeconomy Trend Description
# | output: asis
pct_max = pct_bioeconomy.select("pct_bioeconomy").max().item()
pct_max_year = (
    pct_bioeconomy.filter(pl.col("pct_bioeconomy") == pct_max).select("year").item()
)
pct_last_year = pct_bioeconomy.select("year").max().item()
pct_latest = (
    pct_bioeconomy.filter(pl.col("year") == pct_last_year)
    .select("pct_bioeconomy")
    .item()
)
pct_diff_max_latest = (pct_max - pct_latest) * 100

total_max = pct_bioeconomy.select("n_innovation").max().item()
total_max_year = (
    pct_bioeconomy.filter(pl.col("n_innovation") == total_max).select("year").item()
)
bioeconomy_max = pct_bioeconomy.select("n_bioeconomy").max().item()
bioeconomy_max_year = (
    pct_bioeconomy.filter(pl.col("n_bioeconomy") == bioeconomy_max)
    .select("year")
    .item()
)

total_pct_bioeconomy = sum_bioeconomy_innovations / n_swinno

```

```{python}
# | label: Number of Results
# | output: asis

count_bioeconomy_visions = bioeconomy_visions.group_by(
    ["year", "bioeconomy_vision"]
).agg(pl.col("sinno_id").count().alias("count"))

total_bioeconomy_visions = (
    count_bioeconomy_visions.group_by("bioeconomy_vision")
    .agg(pl.sum("count").alias("count"))
    .sort("count", descending=True)
)

n_bioeconomy_visions = bioeconomy_visions.select("sinno_id").n_unique()
bioeconomy_visions_min_year = bioeconomy_visions.select(pl.min("year")).item()
bioeconomy_visions_max_year = bioeconomy_visions.select(pl.max("year")).item()
n_most_bioeconomy_vision = total_bioeconomy_visions.row(0)[1]

def calculate_share(vision: str):
    return (
        total_bioeconomy_visions.filter(pl.col("bioeconomy_vision") == vision)
        .select("count")
        .item()
        / n_bioeconomy_visions
    )

pct_vision_neutral = calculate_share("Vision Neutral")
pct_bioecology = calculate_share("Bioecology Vision")
pct_bioresource = calculate_share("Bioresource Vision")
pct_biotechnology = calculate_share("Biotechnology Vision")

n_bioresource_1970_1990 = (
    bioeconomy_visions.filter(
        (pl.col("year").is_between(1970, 1990))
        & (pl.col("bioeconomy_vision") == "Bioresource Vision")
    )
    .select("sinno_id")
    .n_unique()
)

n_bioresource = (
    bioeconomy_visions.filter(pl.col("bioeconomy_vision") == "Bioresource Vision")
    .select("sinno_id")
    .n_unique()
)

n_biotechnology = (
    bioeconomy_visions.filter(pl.col("bioeconomy_vision") == "Biotechnology Vision")
    .select("sinno_id")
    .n_unique()
)

n_bioecology = (
    bioeconomy_visions.filter(pl.col("bioeconomy_vision") == "Bioecology Vision")
    .select("sinno_id")
    .n_unique()
)

n_vision_neutral = (
    bioeconomy_visions.filter(pl.col("bioeconomy_vision") == "Vision Neutral")
    .select("sinno_id")
    .n_unique()
)
```

```{python}
# | label: Count Recycling Innovation

recycle_innovation = pl.read_csv(RAW_DATA_DIR / "recycling_innovation.csv")

recycle_innovation_count = recycle_innovation["sinno_id"].n_unique()

```

### 2025-04-25 paper.qmd
update for talk at circle

--- 2024-05-06 paper.qmd
@@
 author:
-  name: Jonas Kreutzer
-  affiliation: 'Department of Economic History, Lund University'
+  name: Philipp Jonas Kreutzer
+  email: philipp_jonas.kreutzer@ekh.lu.se
+  orcid: 0000-0003-4234-9043
+acknowledgements: |
+  I am thankful for helpful comments from Josef Taalbi, Astrid Kander, Frederic Bauer as well as participants of the 9th NEST Conference and the Lund Economic History Conference Seminar Series.
+affiliations: 
+  - id: LU
+    department: Department of Economic History
+    name: Lund University
+    city: Lund
+    country: Sweden
+funding: This work has been funded by Vinnova (2020-01963).
 date: last-modified
@@
   The transition to a bioeconomy is high on the agenda for research, industry and policy.
-Like other sustainability transition, the bioeconomy transition is characterized by contestation over desired outcomes and transition pathways.
+Like other sustainability transitions, the bioeconomy transition is characterized by contestation over desired outcomes and transition pathways.
 While shared visions are seen as pivotal in guiding transitions, the bioeconomy discourse is marked by the contention of three competing visions.
@@
 The use of detailed innovation data from a Swedish literature-based innovation output database, SWINNO, allows to capture the actual innovation output of the Swedish forest-based bioeconomy innovation system.
-Qualitative coding of the source material links innovation from 1970 to 2021 with visions of a bioresource, biotechnology and / or bioecology bioeconomy.
+Qualitative coding of the source material links innovation from 1970 to 2021 with visions of a bioresource, biotechnology and/or bioecology bioeconomy.
 Findings reveal a decrease in the innovation output of the forest-based bioeconomy relative to Sweden's overall commercialized innovation.
@@
 This research underscores the dynamic interplay between historical contexts, shared visions, and innovation trajectories within the bioeconomy transition, and questions the extent to which visions can provide novel directionality to transitions.
-
-csl: 'https://zotero.org/styles/apa?source=1'
+keywords: 
+  - Directionality
+  - Bioeconomy
+  - Forest
+  - Transformative Innovation Policy
 bibliography:
-  - swinno-shades-green.bib
-link-citations: true
-link-bibliography: true
-colorlinks: true
-number-sections: true
-number-depth: 3
+  - swinno-bioeconomy-directionality.bib
 format:
-  pdf:
-    documentclass: scrartcl
-    papersize: a4
-    linestretch: 1.25
-    mainfont: "Adobe Garamond Pro"
-    sansfont: "Adobe Garamond Pro Semibold"
-    indent: true
+  elsevier-pdf:
+    keep-tex: true
+    journal:
+      name: Environmental Innovation and Societal Transitions
+      formatting: review
+      model: 3p
+      cite-style: authoryear
+      highlights:
+        - Highlight 1
+        - Highlight 2
+        - Highlight 3
 execute:
@@
 
-```{python}
-# | label: Imports
+{{< include _code_setup.qmd >}}
 
-import pandas as pd
-import numpy as np
-import scipy.stats as stats
-from pyhere import here
-from IPython.display import display, Markdown
-import matplotlib.pyplot as plt
-import matplotlib.ticker as mtick
-import seaborn as sns
-from tabulate import tabulate
-
-from src.figures import *
-from src.utils import get_project_root, format_table
-from src.data_handling import database
-
-ROOT = get_project_root()
-raw_data = here("data/raw")
-modified_data = here("data/modified")
-conn = database.connect_swinno_db()
-
-# set all text on plots to be monospace with rcParams
-plt.rcParams["font.family"] = "Space Grotesk"
-
-sns.set_theme(style="white")
-```
-
-```{python}
-#| label: Load Data
-raw_data = here("data/raw")
-uncleaned_bioeconomy_query = pd.read_csv(raw_data / "uncleaned_bioeconomy_query.csv")
-bioeconomy = pd.read_csv(raw_data / "id_year_bioeconomy.csv")
-bioeconomy = bioeconomy.loc[bioeconomy["bioeconomy"] == 1]
-
-sni_codes = pd.read_sql_query(
-    """
-select
-*
-from 
-sni_codes
-""",
-    conn,
-)
-
-sni_codes = sni_codes.rename(columns={"code": "use_sector"})
-
-swinno = pd.read_sql_query(
-    """
-select sinno_id, year_of_commercialization as year, innovation_name_in_swedish as name
-from innovation;
-""",
-    conn,
-)
-
-swinno["bioeconomy"] = (
-    swinno["sinno_id"].astype(str).isin(bioeconomy["sinno_id"].astype(str).unique())
-)
-
-# full code labels for merging
-codes_df = pd.read_sql("SELECT * FROM classification_codes", conn)
-
-codes_df["Category"] = codes_df["Category"].str.strip()
-uncertain_eco = {"Code": ["000"], "Category": ["Unsure"]}
-codes_df = pd.concat([codes_df, pd.DataFrame(uncertain_eco)], ignore_index=True)
-
-codes_df = codes_df[~codes_df["Code"].duplicated(keep="last")]
-
-codes_df = codes_df.set_index("Code")
-
-eco_innovations = database.query_eco_innovations(conn)
-eco_innovations["innovation_type"] = eco_innovations["innovation_type"].str.strip()
-
-bioeconomy_visions = pd.read_csv(raw_data / "bioeconomy_visions.csv")
-```
-
-```{python}
-# | label: Innovation Counts
-n_swinno = swinno["sinno_id"].nunique()
-year_min = swinno["year"].min()
-year_max = swinno["year"].max()
-
-n_uncleaned_bioeconomy_innovations = uncleaned_bioeconomy_query["sinno_id"].nunique()
-sum_bioeconomy_innovations = swinno["bioeconomy"].sum()
-```
-
-# Introduction {#sec-introduction}
+<!-- TODO there are 6 unsure bioeconomy innovations remaining -->
 
@@
 -->
-At the heart of sustainability transitions lies the question how societal functions can be fulfilled within ecological boundaries [@kohler2019AgendaSustainabilityTransitions].
-Societal functions are provided through socio-technical systems, which consist of the interplay between the production, distribution, and use of technology [@geels2004SectoralSystemsInnovation].
-Transitions are complex, long-term processes of shifts between different systems, involving diverse actors who have to chose between multiple, open-ended pathways, uncertain of their technological, social and political success [@kohler2019AgendaSustainabilityTransitions].
-Further, success in the outcome and chosen pathway of sustainability transitions is contested based on the values, norms, and goals of  different actors and social groups.
-Hence, innovation cannot be seen as a scalar, but must rather be understood as a vector, as possessing and providing direction to change rather than being simply progress forward [@stirling2011PluralisingProgressIntegrative].
-@andersson2021OutcomesDirectionalityMorphology term these two properties _positive_ and _normative_ directionality, respectively.
-Positive directionality of innovation arises from the fact that technological change is path-dependent [@arthur1989CompetingTechnologiesIncreasing;@david1985ClioEconomicsQWERTY], while normative directionality is rooted in the assumption that without public policy, private actors lack intrinsic incentives to pursue sustainability [@kohler2019AgendaSustainabilityTransitions].
+At the heart of sustainability transitions is the idea that some configurations of sociotechnical systems are preferable to others.
+At least since the normative turn of innovation policy, this has gone hand in hand with the idea that policymakers ought to steer sociotechnical transitions into preferred directions.
+A key instrument literature proposes for providing this normative directionality are shared visions among actors.
+A central question is how these visions should emerge, should they be imposed top-down, or develop from the bottom-up?
 
-A sustainability transition which has been gaining relevance for researchers, policy makers, and producers alike is the bioeconomy.
-Broadly speaking, it aims to radically shift economic systems from relying on fossil fuels for energy and material resource, to an economic system built on biobased materials [@bennich2018BioBasedEconomyDynamics], circular and renewable value chains [@damato2017GreenCircularBio] and higher environmental integrity through innovation [@bugge2016WhatBioeconomyReview].
-As one of the main sources for biomass in a bioeconomy [@priefer2017PathwaysShapeBioeconomy], forests have become a key arena for the bioeconomy transition, highlighting tradeoffs between material provision, carbon storage and biodiversity [@hogbom2021TrilemmaNordicBaltic].
-Underneath the umbrella term "bioeconomy", the concept and viable transition pathways are heavily contested [@allain2022BioeconomicsVsBioeconomy].
-Three different visions of a bioeconomy compete in discourse to provide normative directionality of the transition: the first vision centers on the provision of biomass and renewable resources, the second on the scientific development and application of biotechnology, while the third emphasizes ecological considerations[@bugge2016WhatBioeconomyReview; @vivien2019HijackingBioeconomy].
-Although the discourse is well understood in terms of content [e.g., @bugge2016WhatBioeconomyReview;@vivien2019HijackingBioeconomy] its narratives [@bauer2018NarrativesBiorefineryInnovation], and actors [@holmgren2022WhoseTransformationThis], the material effects of this transition, however, remain important research topics [@fischer2020SwedishForestsBioeconomy].
-How many innovations are commercialized in this new economy, especially considering the central role of innovation in bioeconomy discourse?
-And what does commercialized innovation suggest for the positive directionality of the bioeconomy transition?
+A sustainability transition which has been high on the agenda of policymakers and researchers alike is the transition to a bioeconomy.
+The fundamental idea of this transition is to move from an economic system which depends on fossil resources for fuels and raw material input, to one that builds on renewable biomass instead.
+However, beyond this very broad understanding of the bioeconomy, the what, how, who, when and why of the bioeconomy transition is heavily contested.
+Different actor coalitions have proposed, diffused and countered three competing visions of bioeconomies.
+Expectations on future bioeconomy, and hence research on bioeconomy transition, are especially pronounced in countries endowed in biomass sources.
+As a country whose land area is covered to 70% with forests and a technology leader, Sweden places high expectations on the transition to a forest-based bioeconomy and hence is well explored as a case study on the bioeconomy transition.
+The central aims of this paper are to understand the extent to which commercialized innovation aligns with the contested bioeconomy visions.
+Based on this case study, the paper also aims to provide insights on how innovation output and directionality through visions interact.
 
-<!-- @fischer2020SwedishForestsBioeconomy also note the consensus builidng of Swedish bioeconomy discourse marginalizes non-mainstream opinions.
-I can show that this is rooted, at least in part, in the path-dependence on technology output.
+<!-- Paragraphs 3-4: Challenges.
+These paragraphs explain why your research question has not already been answered, i.e., what are the central challenges a researcher must tackle to answer this question.
 -->
 
-<!-- • Paragraphs 3-4: Challenges.
-These paragraphs explain why your research question has not already been answered, i.e., what are the central challenges a researcher must tackle to answer this question.
--->
-<!-- • Paragraph 5: This Paper. This paragraph states in a nutshell what the paper accomplishes and how. -->
-<!-- • Paragraphs 6-7: Model. Summarize the key formal assumptions you will maintain in your analysis. -->
-<!-- • Paragraphs 8-9: Data. Explain where you obtain your data and how you measure the concepts that are central to your study. -->
-<!-- • Paragraphs 10-11: Methods. Explain how you take your model to the data and how you overcome the challenges you raised in paragraphs 3-4. -->
+Previous research on directionality in bioeconomy transitions has relied heavily on discourse analysis as it is challenging to identify transition outcomes of an ongoing, or expected transition.
+Assessing the extent to which visions are formed bottom-up or top-down is hindered by the requirement for clearly delineated visions for each transition.
 
-To unpack the plurality of possible socio-technical development paths, requires high data quality when measuring innovation.
-Although progress towards monitoring the bioeconomy transition has been made [@jander2019MonitoringTransitionBioeconomy; @jander2020MonitoringBioeconomyTransitions], assessing directionality is limited by a lack of innovation output measures [@wydra2020MeasuringInnovationBioeconomy].
-In this paper, I use SWINNO, a literature-based innovation output database capturing all significant commercialized innovation in Sweden [@sjoo2014DatabaseSwedishInnovations], to analyze the innovation output of the Swedish forest-based bioeconomy.
-I develop and apply a classification scheme to identify which bioeconomy vision an innovation produced or used in forestry or the downstream value chain aligns with.
+<!-- Paragraph 5: This Paper. This paragraph states in a nutshell what the paper accomplishes and how. -->
 
-<!-- • Paragraphs 12-13: Findings. Describe the key findings. Make sure they connect clearly to the motiva-
-tion in paragraphs 1-2. -->
-Results show that Swedish bioeconomy innovation output has declined over the last 50 years.
-Of the three different visions the bioresource vision is the most pronounced, in line with expectations from discourse analysis and the Swedish bioeconomy strategy.
-However, the strategy should not be seen as visionary, as the innovation system had been producing bioeconomy innovation and even bioresource aligned innovation from the 1970s onwards.
-This points to a strong interconnection between path-dependence and the creation of directionality in the policy space.
+This paper overcomes these issues by turning to innovation output data, obtained from a literature-based innovation output method database of significant Swedish innovation, and classifying innovation commercialized between 1970 and 2021 into corresponding visions.
+Thus, it becomes possible to observe the extent to which innovation producers have pursued different directions.
+From the results it becomes evident that the vision dominating discourse about future bioeconomy configurations by key actors is also the vision which most innovations were aligned with.
+However, considering that the discourse around the bioeconomy transition has only in recent years gained substantial attention resulting in the definition of what each vision entails, it seems more likely that the dominating vision formed around innovations already produced by key actors, rather than this vision providing top-down directionality.
+Qualitative analysis of the innovations further supports the idea that economic and political incentives had more substantial impacts on innovation output than collectively shared understandings about ideal innovation outcomes.
+This raises questions about the effectiveness of vision as instruments for translating normative directionality into innovation outcomes and echos recent work on the intertwined nature of directionality policy [@degraaff2025PoliticsDirectionalityInnovation].
+By analyzing the vehciles of technological change, commercialized innovation, it also fills a gap in our understanding of directionality in sociotechnical systems [@yang2022ShapingDirectionalitySustainability].
 
-<!-- • Paragraphs 14-15: Literature.
+<!-- Paragraphs 6-7: Model. Summarize the key formal assumptions you will maintain in your analysis. -->
+
+<!-- Paragraphs 8-9: Data. Explain where you obtain your data and how you measure the concepts that are central to your study. -->
+
+<!-- Paragraphs 10-11: Methods. Explain how you take your model to the data and how you overcome the challenges you raised in paragraphs 3-4. -->
+
+<!-- Paragraphs 12-13: Findings. Describe the key findings. Make sure they connect clearly to the motivation in paragraphs 1-2. -->
+
+<!-- Paragraphs 14-15: Literature.
 Lay out the two main ways your paper contributes to the literature.
@@
 
-<!-- By analyzing the innovations produced by the forest-bioeconomy system in Sweden, this paper adds to existing literature on directionality in sustainability transitions and bioeconomy literature in two main ways.
-On one hand, using literature covering innovation output can be used to assess the directionality a sociotechnical system  -->
+# Previous Literature {#sec-literature}
 
-<!-- From an empirical perspective, this is essentially about furthering the current understanding of how sociotechnical systems grow as a result of historical path-dependencies, system-internal dynamics and external influence, towards a more elaborate understanding of how these mechanisms influence how system configurations change over time.
-From a conceptual perspective, it is rather about enabling such investigations, by developing analytical frameworks that capture not only the pace, but also the directionality of innovation processes.
-For example, there may be a need to advancethe commonly used technological innovation systems framework, since the typology of functional processes is geared towards investigating system growth in a way that hides the potential for vastly different technical, social and spatial development trajectories (Andersson, 2020).
-@andersson2021OutcomesDirectionalityMorphology [127-128] -->
+Previous literature has investigated directionality of sociotechnical systems and emphasized the importance of generating shared visions as tools to provide normative directionality.
+This section introduces key concepts and open questions from this literature.
 
-# Directionality and Visions of Bioeconomy {#sec-literature}
+## Key concepts: 
+### Directionality
 
-Previous literature has investigated directionality of socio-technical systems and emphasized the importance of generating shared visions as tools to provide normative directionality.
+Increasingly scholars have paid attention to opening up innovation from notions of one directional progress or advancement of society to acknowledging that
+
+#### Positive Directionality
+#### Normative Directionality
+
+## Open Questions
+How directionality is:
+        a. set, 
+        b. translated,
+        c. if it is (or ought to be) a top-down or bottom-up process
+
+Directionality in the (forest) bioeconomy.
+
+1. Concept of the bioeconomy
+2. Contested Visions
+    As is often the case in sustainability transitions [@wanzenbock2020FrameworkMissionorientedInnovation], the directionality of the bioeconomy is highly contested [@vivien2019HijackingBioeconomy; @bugge2016WhatBioeconomyReview].
 Three distinct visions of desirable transition pathways and outcomes compete for normative directionality of the bioeconomy transition.
-This section introduces key concepts from this literature.
 
 ## Directionality {#sec-directionality}
+<!-- “Through this co-evolution, transitions unfold as complex, open-ended processes.
+Importantly, this implies that transitions can take more and less desirable courses (Rotmans, 2005; Grin et al., 2010).
+In recent years, transitions scholarship has become more sensitive to this directionality of transitions, i.e.
+the diversity of possible transition pathways (Stirling, 2011, 2019; Røpke, 2012; Pel et al., 2020b; Andersson et al., 2021).” ([Pel et al., 2023, p.
+7](zotero://select/library/items/22YIZBA8)) ([pdf](zotero://open-pdf/library/items/2UYJXF8N?page=7&annotation=YQRI3EPP)) -->
 
-Technology does not change randomly.
-Instead, technological change is contingent and influences technological change at previous or later points in time.
-It is path-dependent, meaning events at one point in time affect later events, which may even lead to suboptimal outcomes [@david1985ClioEconomicsQWERTY].
+First I need to show that technological change has a direction.
+This explains that fundamentally innovation does not simply bring progress or advance society, but that it takes us in different directions, or different configurations of sociotechnical systems [@andersson2021OutcomesDirectionalityMorphology;@yang2022ShapingDirectionalitySustainability].
+
+Directed technological change @aghion2016CarbonTaxesPath, @moscona2022DoesDirectedInnovation. 
+It is path-dependent, meaning events at one point in time effect later events, which may even lead to suboptimal outcomes [@david1985ClioEconomicsQWERTY].
 Path-dependency arises due to positive feedback effects, such as increasing returns [@arthur1989CompetingTechnologiesIncreasing], or through extension of possibilities adjacent to exiting technology [@kauffman2000Investigations].
-Instead of more or less progress, each innovation implies a new direction for socio-technical systems [@stirling2011PluralisingProgressIntegrative].
-@andersson2021OutcomesDirectionalityMorphology term this quality of innovation possessing direction _positive_ directionality.
+Instead of more or less progress, each innovation implies a new direction for sociotechnical systems [@stirling2011PluralisingProgressIntegrative].
+
+@andersson2021OutcomesDirectionalityMorphology term the fact that some directions for technological change are more plausible than others _positive_ directionality.
+
+, lending innovation _normative_ directionality [@andersson2021OutcomesDirectionalityMorphology].
+As innovation systems are embedded in a spatial and temporal context, not all potentially desired pathways are available or feasible.
+Instead, sociotechnical systems are constrained by path-dependency [@kemp2022DiversifyingDeepTransitions;@bergek2015TechnologicalInnovationSystems; @stirling2011PluralisingProgressIntegrative], sometimes to the point of being locked-in to specific system states [@unruh2000UnderstandingCarbonLockin].
+
+### Directing Technological Change
+
+Then I will explain that with the normative turn of innovation policy the idea that policymakers should provide direction rather than just stimulating more innovation has spread.
 
 As innovation policy took a normative turn under which innovation is needed to solve grand societal challenges [@schot2018ThreeFramesInnovation], the question of how to steer innovation into desired directions became a central issue in innovation research [@kohler2019AgendaSustainabilityTransitions].
-Innovation policy now ought to promote the desired vector of societal development, instead of focusing on providing the right amount of progress [@schot2018ThreeFramesInnovation], lending innovation _normative_ directionality [@andersson2021OutcomesDirectionalityMorphology].
-As innovation systems are embedded in a spatial and temporal context, not all potentially desired pathways are available or feasible.
-Instead, socio-technical systems are constrained by path-dependency [@kemp2022DiversifyingDeepTransitions;@bergek2015TechnologicalInnovationSystems; @stirling2011PluralisingProgressIntegrative], sometimes to the point of being locked-in to specific system states [@unruh2000UnderstandingCarbonLockin].
-
-Dealing with the question of how to provide and exercise normative transition directionality  Weber and Rohracher [-@weber2012LegitimizingResearchTechnology] synthesize policy insights from innovation systems and multilevel perspective research into a failures framework.
+Innovation policy now ought to promote the desired vector of societal development, instead of focusing on providing the right amount of progress [@schot2018ThreeFramesInnovation].
+Dealing with the question of how to provide and exercise normative transition directionality @weber2012LegitimizingResearchTechnology synthesize policy insights from innovation systems and multilevel perspective research into a failures' framework.
 Therein, directionality failure is understood as the failure to provide direction to the process of search and unclear or missing setting of collective priorities.
 They caution that it should not be confused with shortsighted or flawed anticipation of market developments, as this myopia lacks a view on the direction of change.
-To overcome this failure, that is, to provide directionality, they argue, external requirements need to be internalized by the innovation system in a interpretive process of negotiation between different system actors.
-Reaching this collective understanding of direction, Weber and Rohracher [-@weber2012LegitimizingResearchTechnology p.
-1042] write, is in practice, "...
-achieved by establishing shared future visions".
+To overcome this failure, that is, to provide directionality, they argue, external requirements need to be internalized by the innovation system in an interpretive process of negotiation between different system actors.
+Reaching this collective understanding of direction, @weber2012LegitimizingResearchTechnology [p. 1042] write, is in practice, "... achieved by establishing shared future visions".
+In their failures' framework, @weber2012LegitimizingResearchTechnology identify that the lack of (sufficiently clear) direction can represent a failure of the innovation system, which they termed directionality failure
+<!-- of or in? -->.
+To address this failure they proposed the need to generate collective visions for development outcomes.
+<!-- Does it also include the means of achieving said outcomes? -->
+
+Next I should explain who should provide directionality / decide.
+Key authors are @mazzucato2018MissionorientedInnovationPolicies, @schot2018ThreeFramesInnovation, @parks2022DirectionalityTransformativeInnovation and @salasgirones2020RolePolicyEntrepreneurs.
+Two central issues of normative directionality, or direction setting, are how desires might result in observable outcomes of new technological directions and, second, how to arrive at normative directionality in the first place.
+On the theoretical side, @TOP-DOWN argue for the setting of directionality through a top-down process, while @BOTTOM-UP argue it should emerge from the bottom up.
+Moving beyond the question of how normative directionality should emerge, @salasgirones2020RolePolicyEntrepreneurs studied how policy entrepreneurs attempt to influence the policymaking process to actively set directions for change in Dutch transport towards automated driving.
+<!-- Empirically, the analysis has consisted of a multi-level case study that included the Swedish Innovation Agency, a strategic innovation programme, and a project organized by regional and urban actors.
+The analysis has shown how directionality was avoided in favour of demand articulation, allowing urban infrastructure companies alone to ‘give directions’ and articulate societal challenges.
+By excluding users and civil society, this process left little room for alternative ideas or conflicts that could encourage higher-order learning or experiment with radical solutions to societal challenges.
+The article’s findings indicate that policy layering has both advantages and disadvantages in transformative innovation policy.
+A disadvantage was how it resulted in an innovation programme led by actors from the supply side of innovation processes.
+These actors are not necessarily well-suited to organize innovation processes that address societal challenges, where it is necessary to open-up innovation to include stakeholders from the consumption and end-use of socio-technical networks such as urban infrastructure.
+Neither do these actors have the capacity to provide directionality to innovation processes.
+Yet at the urban and regional level, policy layering had the advantage of engaging with actors responsible for regional economic development.
+Their capacity to initiate innovation projects facilitated the participation of actors from the demand side, who might otherwise lack the capacity and networks to initiative innovations projects on their own.
+In this sense, existing institutions in innovation systems can contribute to societal challenges by helping to open-up innovation processes to actors from the demand side.
+Perhaps, if provided with models of innovation other than procurement, these existing institutions might facilitate more radical innovation that could encourage more transformative change.
+@parks2022DirectionalityTransformativeInnovation [11] -->
+
+Lastly I turn to the question of how normative directionality is supposed to provide direction.
+For this part I will use @stirling2024ResponsibilityHiddenPolitics.
+I somewhere need to explain that what Striling [-@stirling2008OpeningClosingPower;-@stirling2010MulticriteriaDiversityAnalysis; -@stirling2011PluralisingProgressIntegrative] means with directionality is different from what other authors understand [c.f.
+@andersson2021OutcomesDirectionalityMorphology].
+
+#### Shared Visions are Central Instruments to Provide Direction to Sustainability Transitions
+Dealing with the question of how to provide and exercise normative transition directionality @weber2012LegitimizingResearchTechnology synthesize policy insights from innovation systems and multilevel perspective research into a failures' framework.
+Therein, directionality failure is understood as the failure to provide direction to the process of search and unclear or missing setting of collective priorities.
+They caution that it should not be confused with shortsighted or flawed anticipation of market developments, as this myopia lacks a view on the direction of change.
+To overcome this failure, that is, to provide directionality, they argue, external requirements need to be internalized by the innovation system in an interpretive process of negotiation between different system actors.
+Reaching this collective understanding of direction, @weber2012LegitimizingResearchTechnology [p. 1042] write, is in practice, "... achieved by establishing shared future visions".
 
@@
 They explain guidance of search as, "those activities within the innovation system that can positively affect the visibility and clarity of specific wants among technology users" [@hekkert2007FunctionsInnovationSystems, 423].
-Changing preferences in society, decides which direction of technological development is most promising through market and government influence and, "..
-an interactive and cumulative process of exchanging ideas between technology producers, technology users, and many other actors" [@hekkert2007FunctionsInnovationSystems, 423].
+Changing preferences in society, decides which direction of technological development is most promising through market and government influence and, "… an interactive and cumulative process of exchanging ideas between technology producers, technology users, and many other actors" [@hekkert2007FunctionsInnovationSystems, 423].
 @bauer2017TechnologicalInnovationSystems adapts the innovation system functions to the specific case of biorefineries in Sweden and define the influence on the direction of search as "the incentives and/or pressures for organizations to enter the technological field.
@@
 4).
-Using the multi-level perspective, @yang2022ShapingDirectionalitySustainability too highlight the importance of visions in directionality over transformations, stressing the importance of shared visions between niche and regime actors.
+Using the multi-level perspective, @yang2022ShapingDirectionalitySustainability to highlight the importance of visions in directionality over transformations, stressing the importance of shared visions between niche and regime actors.
 Similarly, @bolton2019GrandVisionsPragmatic illustrate how visions and pragmatic change combined in the development and evolution of a highly integrated European electricity system.
@@
 In summary, visions can be seen as shared imaginaries, mental models and aspirations of desired transitions which are expressed through narratives in discourse between actors [@bauer2018NarrativesBiorefineryInnovation; @holmgren2022WhoseTransformationThis].
+However, one transition can be characterized by multiple, competing visions, as is illustrated by the transition to a bioeconomy.
 
-However, one transition can be characterized by multiple, competing visions, as is illustrated by the transition to a bioeconomy.
-The next section sharpens the conceptual understanding of what is meant with a bioeconomy and introduces the three visions.
+### Sweden's Forest Bioeconomy
 
-## Clarification of Bioeconomy Concept
-
-Although the bioeconomy has gained importance as a research, policy and industry focus, it remains unclear what exactly it entails.
 At a high level, the bioeconomy refers to an economy in which raw materials for energy and production processes are provided from biological instead of fossil sources [@bugge2016WhatBioeconomyReview].
@@
 However, a such a profound change to economic systems will not be possible without balancing significant tradeoffs between the needs and desires of different human and non-human stakeholders [@dieken2021MultitudesBioeconomiesSystematic].
-@hogbom2021TrilemmaNordicBaltic summarize these tradeoffs as a trilemma between forests ability to provide sufficent wood to replace fossil based materials, carbon sequestration and storage, and the delivery of other eco-system services such as forest biodiversity.
-Resolving these tradeoffs is challenging, as goal conflicts between different development pathways make simulatneous progress on social, ecological and economic goals difficult [@bennich2021BiobasedEconomy2030].
+@hogbom2021TrilemmaNordicBaltic summarize these tradeoffs as a trilemma between forests ability to provide sufficient wood to replace fossil based materials, carbon sequestration and storage, and the delivery of other ecosystem services such as forest biodiversity.
+Resolving these tradeoffs is challenging, as goal conflicts between different development pathways make simultaneous progress on social, ecological and economic goals difficult [@bennich2021BiobasedEconomy2030].
 
-As one of the leading forestry nations in the world, Sweden sees substantial economic and political potential in advancing the bioeconomy .
+As one of the leading forestry nations in the world, Sweden perceives substantial economic and political potential in advancing the bioeconomy and hence aims to promote a transition towards it.
 In a report commissioned by the Swedish Government, @formas2012SwedishResearchInnovation lays out the plan for a Swedish bioeconomy based on two central pillars.
-It is to be an economy based on,
+It foresees an economy based on,
 
@@
 Notions of a biotechnology and bio-resource oriented bioeconomy are widespread, dominate discourse at various levels [@dieken2021MultitudesBioeconomiesSystematic] and form the main industrial visions [@befort2020GoingDefinitionsUnderstand].
-A third vision, centered around ecology and an economy compatible with the biosphere acts as a counter point.
+A third vision, centered around ecology and an economy compatible with the biosphere acts as a counterpoint.
 
@@
 In summary, despite -- or perhaps because of -- bioeconomy lacking a strong definition, the term is used to refer to different concepts, with different aims and objectives, foci, normative assumptions and hence implications for society and environment [@bugge2016WhatBioeconomyReview].
-The summarized visions express themselves as narratives, actor coalitions and discourse artefacts in attempts to provide normative directionality to the transition.
+The summarized visions express themselves as narratives, actor coalitions and discourse artifacts in attempts to provide normative directionality to the transition.
 This normative directionality is expected to translate into positive directionality of technological change through innovation.
 
-# Data and Methods {#sec-data_methods}
+# Data and Methods {#sec-data-methods}
 
 Operationalizing directionality suffers from frequent lack of differentiation of positive and normative directionality, in addition to inherent issues in the measurement of innovation.
-Notable exceptions are the work on socio-technical system morphology by @andersson2021OutcomesDirectionalityMorphology, who identify changes in the configuration of three different renewable energy systems, and by @yang2022ShapingDirectionalitySustainability who use interviews and secondary source analysis to understand changes to insitutions in solar photovoltaic production in two Chinese provinces.
-However, both studies deal more with the aggregate effects of directionality on socio-technical systems rather than with positive directionality of innovation directly.
+Notable exceptions are the work on sociotechnical system morphology by @andersson2021OutcomesDirectionalityMorphology, who identify changes in the configuration of three different renewable energy systems, and by @yang2022ShapingDirectionalitySustainability who use interviews and secondary source analysis to understand changes to institutions in solar photovoltaic production in two Chinese provinces.
+However, both studies deal more with the aggregate effects of directionality on sociotechnical systems rather than with positive directionality of innovation directly.
 
@@
 Where positive directionality of technological change has been assessed, scientific publications [@wydra2020MeasuringInnovationBioeconomy; @jankovsky2021InnovationsForestBioeconomy], and more often, patents are used [e.g., @aghion2016CarbonTaxesPath;@popp2005LessonsPatentsUsing;@moscona2022DoesDirectedInnovation].
-Such data, however, suffer from limitations when attempting to assess the directionality of socio-technical systems.
+Such data, however, suffer from limitations when attempting to assess the directionality of sociotechnical systems.
 While measures of normative directionality are equipped to identify the desired direction by different actors, visions need to first be accepted by innovators and reformulated into search.
 R&D funding can help in this task, but data on R&D funding is often too broad, targeting entire sectors or specific industries [@popp2005LessonsPatentsUsing].
-While patents are more granular and a closer proxy of technological change, they suffer from well documented measurement errors [@popp2005LessonsPatentsUsing]; not all innovations are patented, and not all patents become innovations [@arundel1998WhatPercentageInnovations].
+While patents are more granular and a closer proxy of technological change, they suffer from well documented measurement errors [@popp2005LessonsPatentsUsing]: not all innovations are patented, and not all patents become innovations [@arundel1998WhatPercentageInnovations].
 
@@
 Applying this method to editor-reviewed articles published in 15 independent trade journals covering technological developments for general and highly specialized audiences, the SWINNO database contains `{python} n_swinno` innovations commercialized by Swedish companies between `{python} year_min` and `{python} year_max` [@sjoo2014DatabaseSwedishInnovations].
-Since this method relies on the judgment of editiors and journalists, it can be described as significant innovations, omitting small reconfigurations of exisiting products and processes.
+Since this method relies on the judgment of editions and journalists, it can be described as significant innovations, omitting small reconfigurations of exisiting products and processes.
 The SWINNO database has significant benefits compared to other innovation measures.
 For one, it ensures that actual innovation, meaning commercialized products and processes, and thus that its positive directionality is captured.
-Second, because the source articles are preserved, it provides a rich source of data, frequently providing insights into the aims, motivations, advantages and challenges of the covered innovation, in addition to descriptions of what the innovation consists of, which allows to assess and quantify the direction of technological change.
+Second, because the source articles are preserved, it provides a rich source of data, frequently providing insights into the aims, motivations, advantages and challenges of the covered innovation, in addition to descriptions of what the innovation consists of, which allows assessing and quantify the direction of technological change.
 
@@
 
-I carefully read the source articles of innovations identified in the database and tagged their associated unique identifiers with two types of labels.
-The first label is the main interest of this study: the bioeconomy vision an innovation can be associated with.
-The second label provides a deeper understanding if and how the innovation reduces environmental impacts; not all bioeconomy innovations need to be sustainability oriented innovation [@broring2020InnovationTypesBioeconomy].
+Based on the analyses of competing bioeconomy visions in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy, I developed and applied a matrix of characteristics for each vision.
+While the criteria for the bioresource and bioecology vision are entirely based on the two previous publications, the biotechnology criteria are additionally based on a literature review of @yeung2019CurrentResearchBiotechnology concerning biotechnology publications.
 
-### Bioeconomy Visions {#sec-bioeconomy_classification_method}
-
-Based on the analyses of competing bioeconomy visions in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy, I developed and applied a matrix of characterstics for each vision.
-While the criteria for the bioresource and bioecology vision are entirely based on the two previous publications, the biotechnology criteria are additionally based on a literature review of @yeung2019CurrentResearchBiotechnology, concerning biotechnology publications.
-
-Reading the source texts of innovations identified from SWINNO, I classified an innovation as being *aligned* with a vision if it fulfills at least one criteria in the matrix.
-An innovation can align with multiple visions, for example, where biomass is converted and upgraded (bioresource vision) by application of biotechnology (biotechnology vision).
+Reading the source texts of innovations identified from SWINNO, I classified an innovation as being *aligned* with a vision if it fulfilled at least one criterion in the matrix.
+An innovation can align with multiple visions. 
+For example, the conversion and upgrading of biomass (bioresource vision) by application of a biological agent or cellular or molecular processes (biotechnology vision).
 Or the value chain decommodifying introduction of new production lines by a family-business which use wood from their own forests to produce houses with high territorial identity (*Svensk trävaru- & pappersmassetidning*, 1976:7, pp.
-539,541,544), that aligns with the bioresource and bioecology vision.
+539,541,544), which aligns with the bioresource and bioecology vision.
 Lastly, there is also the possibility that an innovation cannot be placed into at least one of the visions.
-This is the case for innovation which brings general technological improvements, purely advances economic performance, or relates to a general purpose technology such as appliance of computers.
+This is the case for innovation which brings general technological improvements, purely advances economic performance, or relates to a general purpose technology such as the application of computers.
 These innovations are labeled as Vision Neutral, which does not imply that the innovation is not used, or originates in the forest based bioeconomy as defined above.
@@
 |||
-| Bioresource          | Biomass used to substitute other input                                                                |
+| Bioresource     | Uses biomass to substitute other input                                |
 |           | Refines biomass through mechanical, chemical or other process                     |
 |           | Converts and upgrades bioresource                                   |
-|                      | Optimises land use                                                                                    |
+|           | Optimizes land use                                          |
 |           | Includes degraded land in production of biofuels                           |
@@
 |           | States compatibility with biosphere as motivation or aim                       |
-|                      | States need for social or ecological counter-expertise relative to other socio-technical innovations  |
+|           | States need for social or ecological counter-expertise relative to other sociotechnical innovations |
 |           | Promotes or uses a circular production mode                              |
@@
 
-### Eco-Innovation Types {#sec-eco_innovation_types}
-
-The focus of this paper is understanding what bioeconomy visions dominate the Swedish innovation system's output.
-However, in order to gain a deeper understanding of the nature of innovation within each bioeconomy vision, I also tagged the innovation with labels of eco-innovation types.
-This classification uses eco-innovation types identified in a recent literature review by @garcia-granero2018EcoinnovationMeasurementReview.
-They review 104 publications and produce a list 30 key Eco-Innovation Performance Indicators of product, process, organizational and marketing innovation.
-
-This initial list was tested on a subset of innovation in SWINNO and augmented with 15 categories.
-Because the list of eco-innovation types was relatively exhaustive from the beginning, few adjustments had to be made.
-These include: expanding the labeling for other sustainability criteria, such as human health and safety, and specific forest related aspects / tasks to gain a more detailed understanding of an innovations aim.
-For example, some innovations aim to protect harvested timber, either at harvest, during transport or storage.
-While @tbl-full-codes shows all new added labels, in the following text those that do not clearly represent an eco-innovation -- following @garcia-granero2018EcoinnovationMeasurementReview "…fewer adverse effects on the environment and more efficient use of resources" (p.
-305) -- were excluded.
-These excluded labels are Gen-manipulated plant, Increased plant yield, Gen-manipulated animal (incl new breed), Human Health and Safety, Automatization, Precision, and Biocide.
-
-The decision process for labelling eco-innovation types was also based on the source texts for each innovation.
-Based on the information contained therein, I labeled the innovation  with one or more categories.
-In case of uncertainty, I added the label uncertain. In case where the text does not mention any eco-innovation aspect, the innovation was labeled as not an eco-innovation.
-Where the source text only mentions that the innovation is less harmful to the environment, the label "Environmental-friendly technologies" was used.
-
 # Results {#sec-results}
 
-```{python}
-#| label: Calculate Bioeconomy Share
-
-def calculate_bioeconomy_share(swinno):
-    df = swinno.copy()
-    df = df[["sinno_id", "year", "bioeconomy"]]
-    df = (
-        df.groupby("year")
-        .agg(
-            total_count=("sinno_id", "nunique"), bioeconomy_count=("bioeconomy", "sum")
-        )
-        .reset_index()
-    )
-    df["bioeconomy_share"] = df["bioeconomy_count"] / df["total_count"]
-
-    return df
-
-bioeconomy_share = calculate_bioeconomy_share(swinno)
-```
-
-```{python}
-#| label: Bioeconomy Trend Description
-#| output: asis
-share_max = bioeconomy_share["bioeconomy_share"].max()
-share_max_year = bioeconomy_share.loc[
-    bioeconomy_share["bioeconomy_share"] == share_max, "year"
-].values[0]
-share_last_year = bioeconomy_share["year"].max()
-share_latest = bioeconomy_share.loc[
-    bioeconomy_share["year"] == share_last_year, "bioeconomy_share"
-].values[0]
-
-pct_diff_max_latest = (share_max - share_latest) * 100
-total_max = bioeconomy_share["total_count"].max()
-total_max_year = bioeconomy_share.loc[
-    bioeconomy_share["total_count"] == total_max, "year"
-].values[0]
-
-bioeconomy_max = bioeconomy_share["bioeconomy_count"].max()
-bioeconomy_max_year = bioeconomy_share.loc[
-    bioeconomy_share["bioeconomy_count"] == bioeconomy_max, "year"
-].values[0]
-
-total_bioeconomy_share = sum_bioeconomy_innovations / n_swinno
-
-```
-
 Querying the SWINNO database for forest based bioeconomy innovations initially yields `{python} n_uncleaned_bioeconomy_innovations` innovations.
-This number includes false positives.
-During classification, these false positives were removed from the dataset, resulting in `{python} sum_bioeconomy_innovations` innovations which originate from, are used in, or make use of products derived from the forest.
+False positives were removed from the dataset during classification, resulting in `{python} sum_bioeconomy_innovations` innovations which originate from, are used in, or make use of products derived from the forest.
 
@@
 
-In total, the forest-based bioeconomy produced `{python} f"{total_bioeconomy_share:.0%}"` of innovation captured in the SWINNO database.
-While the total innovation output and bioeconomy innovations feature a period of high output from the 1970s to the 1990s, with a peak in `{python} total_max_year` for total output and `{python} bioeconomy_max_year` for the bioeconomy, the economic downturn and industrial restructurization of the 1990s was accompanied with a decline in commercialized innovation.
-Unlike the other sectors and socio-technical systems, however, it appears that the bioeconomy was not able to recover to the same extent.
-Its percentage of total innovation declined by `{python} f"{pct_diff_max_latest:.0f}"` percentage points from an all-time high of `{python} f"{share_max:.0%}"` in `{python} share_max_year` to `{python} f"{share_latest:.0%}"` in `{python} share_last_year`.
-@fig-innovation-trends illustrates this using the five year moving average for the number of innovations commercialized per year in the bioeconomy and in the entire dataset, including the bioeconomy.
+In total, the forest-based bioeconomy produced `{python} f"{total_pct_bioeconomy:.0%}"` of innovation captured in the SWINNO database.
+While the total innovation output and bioeconomy innovations feature a period of high output from the 1970s to the 1990s, with a peak in `{python} total_max_year` for total output and `{python} bioeconomy_max_year` for the bioeconomy, the economic downturn and industrial restructurization of the 1990s was accompanied by a decline in commercialized innovation.
+Unlike other sectors and sociotechnical systems, however, it appears that the bioeconomy was not able to recover to the same extent.
+Its percentage of total innovation declined by `{python} f"{pct_diff_max_latest:.0f}"` percentage points from an all-time high of `{python} f"{pct_max:.0%}"` in `{python} pct_max_year` to `{python} f"{pct_latest:.0%}"` in `{python} pct_last_year`.
+@fig-innovation-trends illustrates this using the five-year moving average for the number of innovations commercialized per year in the bioeconomy and in the entire dataset, including the bioeconomy.
 The decline in relative importance of bioeconomy innovations can be identified in the share of bioeconomy innovation of total innovation.
@@
 Percentage of forest based bioeconomy innovations to total innovation registered in SWINNO database.
-Shown are the 5 year moving averages for each time series.](results/figures/innovation_trends.pdf){#fig-innovation-trends}
+Shown are the 5-year moving averages for each time series.](results/figures/fig_innovation_trends.tex){#fig-innovation-trends width=100%}
 
@@
 
-```{python}
-#| label: Uncertain Visions
-
-bioeconomy_visions.loc[
-    bioeconomy_visions["bioeconomy_vision"] == "Not Bioeconomy",
-    ["bioeconomy_vision"],
-] = "Vision Neutral"
-
-uncertain_bioeconomy_visions = bioeconomy_visions.query("bioeconomy_vision == 'Unsure'")
-
-certain_bioeconomy_visions = bioeconomy_visions.loc[
-    ~bioeconomy_visions["sinno_id"].isin(uncertain_bioeconomy_visions["sinno_id"]), :
-]
-
-count_certain_bioeconomy_visions = (
-    certain_bioeconomy_visions.groupby(["year", "bioeconomy_vision"])
-    .size()
-    .reset_index(name="count")
-)
-
-total_certain_bioeconomy_visions = (
-    count_certain_bioeconomy_visions.groupby("bioeconomy_vision")
-    .agg("sum", numeric_only=True)
-    .sort_values("count", ascending=False)
-    .reset_index()
-)
-```
-
-```{python}
-#| label: Number of Uncertains
-#| output: asis
-
-n_uncertain_bioeconomy_visions = uncertain_bioeconomy_visions["sinno_id"].nunique()
-```
-
-```{python}
-# | label: Number of Results
-# | output: asis
-
-n_certain_bioeconomy_visions = certain_bioeconomy_visions["sinno_id"].nunique()
-
-certain_bioeconomy_visions_min_year = certain_bioeconomy_visions["year"].min()
-
-certain_bioeconomy_visions_max_year = certain_bioeconomy_visions["year"].max()
-
-n_most_bioeconomy_vision = total_certain_bioeconomy_visions.iloc[0, 2]
-
-def calculate_share(a: str):
-    return (
-        total_certain_bioeconomy_visions.loc[
-            total_certain_bioeconomy_visions["bioeconomy_vision"] == a, "count"
-        ].values[0]
-        / n_certain_bioeconomy_visions
-    )
-
-share_vision_neutral = calculate_share("Vision Neutral")
-share_bioecology = calculate_share("Bioecology Vision")
-share_bioresource = calculate_share("Bioresource Vision")
-share_biotechnology = calculate_share("Biotechnology Vision")
-
-vision_1_count = total_certain_bioeconomy_visions.iloc[0, 2] 
-vision_2_count = total_certain_bioeconomy_visions.iloc[1, 2] 
-vision_3_count = total_certain_bioeconomy_visions.iloc[2, 2] 
-vision_4_count = total_certain_bioeconomy_visions.iloc[3, 2]
-
-vision_1_count = total_certain_bioeconomy_visions.iloc[0, 2] 
-vision_2_count = total_certain_bioeconomy_visions.iloc[1, 2] 
-vision_3_count = total_certain_bioeconomy_visions.iloc[2, 2] 
-vision_4_count = total_certain_bioeconomy_visions.iloc[3, 2]
-
-vision_1 = total_certain_bioeconomy_visions.iloc[0, 0]
-vision_2 = total_certain_bioeconomy_visions.iloc[1, 0]
-vision_3 = total_certain_bioeconomy_visions.iloc[2, 0]
-vision_4 = total_certain_bioeconomy_visions.iloc[3, 0]
-
-n_bioresource_1970_1990 = certain_bioeconomy_visions.query("year.between(1970, 1990) & bioeconomy_vision == 'Bioresource Vision'")["sinno_id"].nunique()
-```
-
-```{python}
-#| label: Count Recycling Innovation
-
-recycle_innovation = pd.read_sql(
-    """
-SELECT *
-FROM classification_codes
-LEFT JOIN eco_innovations ON classification_codes.Code = eco_innovations.innovation_type
-WHERE eco_innovations.innovation_type = 107 OR eco_innovations.innovation_type = 206 AND sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes LIKE '%not forest%')
-""",
-    conn,
-)
-
-recycle_innovation_count = recycle_innovation["sinno_id"].nunique()
-
-n_bioecology_vision = count_certain_bioeconomy_visions.loc[
-    count_certain_bioeconomy_visions["bioeconomy_vision"] == "Bioecology Vision",
-    "count",
-].sum()
-```
-
-For the following part, `{python} n_uncertain_bioeconomy_visions` innovations whose source text did not allow labeling without significant uncertainty are excluded.
-These innovations are planned to undergo a revision by subject matter experts at a later point.
-
-Among the remaining `{python} n_certain_bioeconomy_visions` innovations,
-`{python} total_certain_bioeconomy_visions.iloc[0, 2]` (`{python} f"{share_vision_neutral:.0%}"` of all commercialized innovation) were not aligned with any vision.
-Next the `{python} vision_2.lower()` was associated with `{python} vision_2_count` (`{python} f"{calculate_share(vision_2):.0%}"`), while `{python} vision_3_count` (`{python} f"{calculate_share(vision_3):.0%}"`) and  `{python} vision_4_count` (`{python} f"{calculate_share(vision_4):.0%}"`) innovations were classified as `{python} vision_3.lower()` and `{python} vision_4.lower()`, respectively.
+Among the remaining `{python} n_bioeconomy_visions` innovations,
+`{python} n_vision_neutral` (`{python} f"{pct_vision_neutral:.0%}"` of all commercialized innovation) were not aligned with any vision.
+Next the Bioresource Vision was associated with `{python} n_bioresource` (`{python} f"{pct_bioresource:.0%}"`), while `{python} n_bioecology` (`{python} f"{pct_bioecology:.0%}"`) and `{python} n_biotechnology` (`{python} f"{pct_biotechnology:.0%}"`) innovations were classified as Bioecology and Biotechnology Vision, respectively.
 Note that the sum of all categories does not equal the total number of classified innovations, as one innovation can be associated with more than one bioeconomy vision.
-@fig-bioeconomy-trends shows the count of innovation for each categorization per year.
+@fig-bioeconomy-vision-counts shows the count of innovation for each categorization per year.
 Output for the biotechnology and bioecology vision appears inconsistent, with years in which not a single commercialized innovation aligned with these visions.
@@
 Innovations can be classified into multiple visions simulatenously.
-The sum of all vision counts does not match the sum of all bioeconomy innovations.](results/figures/vision_count.pdf){#fig-bioeconomy-trends}
+The sum of all vision counts does not match the sum of all bioeconomy innovations.](results/figures/fig_bioeconomy_vision_counts.tex){#fig-bioeconomy-vision-counts width=100%}
+
+![Counts of Producing Firms by Number of Innovation Aligned With Bioeconomy Visions.](results/figures/fig_vision_producer_counts.tex){#fig-bioeconomy-vision-producer-counts width=100%}
 
@@
 11; *NTT*, 2003:8, p.6; *NTT* 2004:18, pp.16-17), various paper products (e.g., *Kemivärlden Biotech med Kemisk Tidskrift*, 2009:11, pp.
-60-61; *SPCI Svensk Papperstidning*, 2005:7, pp.10-11), but also biobased alternatives to plastics (e.g., *SPCI Svensk Papperstidning*, 2018:4, pp.
-28-33; *SPCI Svensk Papperstidning*, 2019:6, p.24; *Plastforum*,2019:5, p.31), or refined bio-fuel alternatives to fossil fuels, such as the LignoBoost Process (*SPCI Svensk Papperstidning*, 2006:7, pp.
-14-16), or gassification of black liquour -- a by-product of pulp digestion -- for electricity generation (*Ny Teknik*, 1990:16, p.5) .
+60-61; *SPCI Svensk Papperstidning*, 2005:7, pp.10-11), but also bio-based alternatives to plastics (e.g., *SPCI Svensk Papperstidning*, 2018:4, pp.
+28-33; *SPCI Svensk Papperstidning*, 2019:6, p.24; *Plastforum*,2g019:5, p.31), or refined biofuel alternatives to fossil fuels, such as the LignoBoost Process (*SPCI Svensk Papperstidning*, 2006:7, pp.
+14-16), or gassification of black liquor -- a by-product of pulp digestion -- for electricity generation (*Ny Teknik*, 1990:16, p.5).
 A second theme is rationalization of production processes, especially through the application of computers and software (e.g., *NTT*, 2007:14, p. 13; *Automation*, 2021:1, p. 24).
 
-Therefore, while the first period of high innovation output until the 1990s was characterized by attempts to modernize timber production, mainly through the mechanization of harvesting, and cleaning the pulp and paper process, the second period of active commercialization focused on finding new applications for biomass and improving the cost effectivness of exisiting processes.
+Therefore, while the first period of high innovation output until the 1990s was characterized by attempts to modernize timber production, mainly through the mechanization of harvesting, and cleaning the pulp and paper process, the second period of active commercialization focused on finding new applications for biomass and improving the cost-effectiveness of existing processes.
 This is illustrated with innovations in forwarders -- specialized machines to aid in the removal of timber from harvest sides before loading them onto transport -- where of 9 innovations, 7 were commercialized before 1990.
@@
 
-```{python}
-#| label: forwarder
-#| include: false
-forwarder_innovation = pd.read_sql(
-    "SELECT sinno_id, year_of_commercialization from innovation where description_in_swedish LIKE '%skotare%'",
-    conn,
-)
-
-forwarder_innovation
-```
-
 #### Biotechnology Vision
 
-Only 37 innovations were aligned with the biotechnology vision.
+Only `{python} n_biotechnology` innovations were aligned with the biotechnology vision.
 Of those, the majority focused on cleaning waste water from, for example, fibers (e.g., *Kemisk Tidskrift*, 1972:4, p.
@@
 14).
-Innovations in this vision frequently overlapped with bioresource innovations, such as the above mentioned LignoBoost process.
+Innovations in this vision frequently overlapped with bioresource innovations, such as the above-mentioned LignoBoost process.
 
@@
 
-Recycling accounted for `{python} recycle_innovation_count` unique innovations, or `{python} f"{(recycle_innovation_count / n_bioecology_vision):.0%}"` of innovation associated with the bioecology vision.
+Recycling accounted for `{python} recycle_innovation_count` unique innovations, or `{python} f"{(recycle_innovation_count / n_bioecology):.0%}"` of innovation associated with the bioecology vision.
 The majority of innovations in aligned with this vision due to some explicitly stated environmental aim within existing production methods.
@@
 539,541,544).
-The specifc growing conditions of Northern Sweden are provided as a reason for denser wood, especially well suited for construction due to its strength properties and the "[t]otal integration makes the company feel optimistic: own forests, top-modern sawmill, board mill, newly built component factory are positive factors meaning advantages in all stages" (p.
+The specific growing conditions of Northern Sweden are provided as a reason for denser wood, especially well suited for construction due to its strength properties and the "[t]otal integration makes the company feel optimistic: own forests, top-modern sawmill, board mill, newly built component factory are positive factors meaning advantages in all stages" (p.
 544).
-Similiar decommodification of value chains in favor of local product could even be found in the paper industry, where fresh wood from within a 100km radius was found to produce better quality and reduce chemical and energy usage (*SPCI Svensk Papperstidning*, 2015:8, pp.
+Similar decommodification of value chains in favor of local product could even be found in the paper industry, where fresh wood from within a 100km radius was found to produce better quality and reduce chemical and energy usage (*SPCI Svensk Papperstidning*, 2015:8, pp.
 26-30).
-
-### How many eco-innovations are there?
-
-```{python}
-# | label: merge eco-innovation with bioeconomy visions
-
-eco_innovation_merged_bv = certain_bioeconomy_visions.merge(
-    eco_innovations, on="sinno_id", validate="many_to_many"
-)
-
-eco_innovation_merged_bv["innovation_type_group"] = np.where(
-    eco_innovation_merged_bv["innovation_type_code"].str[0] == "1",
-    "Product",
-    np.where(
-        eco_innovation_merged_bv["innovation_type_code"].str[0] == "2",
-        "Process",
-        np.where(
-            eco_innovation_merged_bv["innovation_type_code"].str[0] == "3",
-            "Organizational",
-            np.where(
-                eco_innovation_merged_bv["innovation_type_code"].str[0] == "4",
-                "Marketing",
-                "No Eco-Innovation",
-            ),
-        ),
-    ),
-)
-
-eco_innovation_by_bioeconomy_visions = (
-    eco_innovation_merged_bv.groupby(["bioeconomy_vision", "innovation_type_code"])
-    .size()
-    .to_frame("count")
-    .reset_index()
-)
-
-eco_innovation_by_bioeconomy_visions["innovation_type_group"] = np.where(
-    eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0] == "1",
-    "Product",
-    np.where(
-        eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0] == "2",
-        "Process",
-        np.where(
-            eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0] == "3",
-            "Organizational",
-            np.where(
-                eco_innovation_by_bioeconomy_visions["innovation_type_code"].str[0]
-                == "4",
-                "Marketing",
-                "No Eco-Innovation",
-            ),
-        ),
-    ),
-)
-```
-
-```{python}
-# | label: chi-square-vision-eco
-other_sustainability = {501, 701, 108, 109, 111}
-eco_innovation_filtered = eco_innovation_merged_bv[
-    ~(
-        eco_innovation_merged_bv["innovation_type_code"]
-        .astype(int)
-        .isin(other_sustainability)
-    )
-]
-# Create the groups for the cross tab
-eco_innovation_filtered["eco_innovation"] = (
-    eco_innovation_filtered["innovation_type_group"] != "No Eco-Innovation"
-)
-eco_innovation_filtered["vision_neutral"] = (
-    eco_innovation_filtered["bioeconomy_vision"] == "Vision Neutral"
-)
-
-# Calculate the chi squared test
-vision_eco_cross_tab = (
-    eco_innovation_filtered.drop_duplicates(subset="sinno_id")
-    .groupby(["vision_neutral", "eco_innovation"])
-    .agg(count=("sinno_id", "nunique"))
-    .unstack()
-)
-chi2_stat, p_val, dof, expected = stats.chi2_contingency(vision_eco_cross_tab)
-```
-
-How does each vision contribute to lessening the environmental impact of the forest bioeconomy?
-While a quantitative assessment of technological impact on the environment comes with a host of challenges, the labelling of eco-innovations at least allowed a glimpse at what could be expected from each vision.
-@fig-eco-innovation-bioeconomy-vision shows the count of eco-innovation associated with each bioeconomy vision by type.
-As before, the total sum of all eco-innovation across all bioeconomy visions outnumbers the sum of innovations in the bioeconomy.
-A single innovation can be more than one eco-innovation type, for example, if it reduces the amount of raw material needed by recycling waste.
-Product and Process innovation were by far the most frequent across all visions.
-The bioresource vision was heavily associated with improving raw-material utilization, while the Bioecology Vision is most frequently associated with recycling.
-At least, this suggests that the definitions for each Vision and the classification scheme are in agreement.
-
-```{python}
-#| label: fig-eco-innovation-bioeconomy-vision
-#| fig-cap: Count of Eco-Innovation Types by Bioeconomy Vision
-p = plot_eco_innovation_by_bioeconomy_vision(
-    eco_innovation_by_bioeconomy_visions, codes_df
-)
-
-sns.move_legend(p, "lower center", ncol=5)
-# TODO: perhaps a per cent display by group size would be interesting too?
-```
-
-A noteworthy feature of innovation lacking alignment to a bioeconomy vision was that this category has by far the most non-eco-innovations.
-The chi-squared test of independence revealed a statistically significant association between an innovation being aligned with a vision and it being an eco-innovation ($\chi^2 =$ `{python} round(chi2_stat, 3)` df = `{python} dof`, p < 0.05).
-
-```{python}
-# | label: tbl-cross-table-eco-visions
-# | tbl-cap: Cross Table of Unique Innovations Classified as Vision Neutral And Eco-Innovation
-# Headers
-headers = ["Non-Eco-Innovation", "Eco-Innovation"]
-index = ["Bioeconomy Vision", "Vision Neutral"]
-
-# Generate the table
-cross_table = tabulate(
-    vision_eco_cross_tab.values, headers=headers, showindex=index, tablefmt="pipe"
-)
-
-display(Markdown(cross_table))
-```
 
@@
 
+This research aimed at understanding the direction of technological change embedded in innovations produced within the Swedish forest-based bioeconomy and how these correspond with overarching visions of a future bioeconomy.
+
+## Positive Directionality of the Swedish Forest-Based Bioeconomy and Possible Future Transition Paths
+
 The decline in innovations commercialized in the forest-based bioeconomy relative to all commercialized innovations suggests that Sweden's capacity to produce technological change towards a bioeconomy has diminished over the last 50 years.
-While innovation counts should be interpreted with caution, as the magnitude of technological change stemming from individual innovations can differ, it is surprising that a economic sector, which is expected to realize ambitious economic, social and ecological targets, is losing importance in relation to other Swedish innovations.
+While innovation counts should be interpreted with caution, as the magnitude of technological change stemming from individual innovations can differ, it is surprising that an economic sector, which is expected to realize ambitious economic, social and ecological targets, is losing importance in relation to other Swedish innovations.
 While it appears that the crises in forestry and energy of the 1970s stimulated innovation directed towards economic rationalization of forestry, the bioeconomy could not benefit from similar stimulants when exploring conversion of biomass into new products after the financial crisis in 2008, and the shock to timber production after storm Gudrun in 2005.
 
+<!-- TODO change this paragraph to be about the importance of bottom-up innovation rather than top-down.
+Tie this to the studies of @holmgren2022WhoseTransformationThis to show that innovation producers may favor bioresource innovation at least partially due to path-dependence of previous technological change -->
 As for the positive directionality of the Swedish transition to a forest-bioeconomy, the dominant vision appears to be successful in so far as it is the vision most innovations align with.
-That a majority of innovation appear neutral in regards to vision based directionality is not to say that these do not posess positive directionality.
-These innovations should be regarded as general improvements to exisiting technology and hence are not neutral in regards to technological change, only in regards to not aligning with normatively directionality.
-The congruency between positive and normative directionality can be explained by a specific feature of Swedish socio-technical systems: a strong desire to create consensus [@belandlindahl2015SwedishForestryModel].
+That a majority of innovation appear neutral with regard to vision based directionality is not to say that these do not possess positive directionality.
+These innovations should be regarded as general improvements to existing technology and hence are not neutral with regard to technological change, only with regard to not aligning with normatively directionality.
+The congruence between positive and normative directionality can be explained by a specific feature of Swedish sociotechnical systems: a strong desire to create consensus [@belandlindahl2015SwedishForestryModel].
 @bergquist2011GreenInnovationSystems illuminate how strong collaboration networks between public and private actors shaped the greening of Sweden's pulp and paper industry between 1960 and 1980.
@@
 
+The history of sustainability transitions in the Swedish forest-bioeconomy, specifically in the pulp and paper sector suggests, that directing technological change is not only possible, but can be highly successful.
+@bergquist2011GreenInnovationSystems and @bergquist2015TransitionGreenerPulp show how social pressure, together with government led regulations and innovation efforts by private actors resulted in harmful chlorine bleaching of paper to be replaced with more environmentally friendly alternatives in the 1970s and 1980s.
+Similarly, the 1990s and 2000s saw a proliferation of greener pulping methods in the sector as response to renewable and efficient energy requirements stimulated by policy aiming to reduce the energy intensity of Swedish industry [@ottosson2013SociotechnicalRegimesHeterogeneous].
+The resulting innovations were aligned mostly with bioresource or bioecology visions, but unlike these broader visions followed concrete and well articulated aims.
+This further underlines the importance of translating visions into measurable and well-defined targets.
+
+<!-- “This case study suggests that without specific routes (e.g.
+in terms of technologies used societal applications), the policy can be easily captured by interests and result in disappointing outcomes such as in our case study.
+For this reason, a more active role of public authorities is required to define, together with societal actors, how a particular technology can contribute to the societal challenges that it intends to address.
+It is striking how this initiative, even though it has a societaldemand orientation, barely engaged with users to define its directions.” ([Salas Gironés et al., 2020, p.
+9](zotero://select/library/items/CIC4FYAW)) ([pdf](zotero://open-pdf/library/items/4TCN7VPB?page=9&annotation=6KS3DMAK)) -->
+
 ## Limitations {#sec-limitations}
 
-In regards to the above, an important limitation is the use of innovation as a measure of socio-technical transitions.
+Regarding the above, an important limitation is the use of innovation as a measure of sociotechnical transitions.
 While the results of this study agree with the survey results of 
 @lovric2020FactorsDevelopmentInnovations and capture the production of technological change, it does not directly capture its adoption and use.
-In addition literature-based innovation output measures have data limitations might result in a bias against non-commercialized innovation, especially for research into forest sector innovation where organizational innovations such as silviculture management can be important levers to improve, for example biodiversity or cultural ecosystem services.
+In addition, literature-based innovation output measures have data limitations might result in a bias against non-commercialized innovation, especially for research into forest sector innovation where organizational innovations such as silviculture management can be important levers to improve, for example biodiversity or cultural ecosystem services.
 Such new processes, or new organizational approaches, will only be captured if they are sold to another company.
 Relevant product innovation which facilitates these plans, for example products related to the planting of trees, on the other hand, are captured.
-A likely consequence of this limitation is that the number of innovation related to the bioecology vision could be under counted, as this vision highlights "[c]ounter-expertise rather than concrete technical solutions" [@vivien2019HijackingBioeconomy 194].
+A likely consequence of this limitation is that the number of innovation related to the bioecology vision could be undercounted, as this vision highlights "[c]ounter-expertise rather than concrete technical solutions" [@vivien2019HijackingBioeconomy 194].
 
@@
 Others might go through multiple rounds of more incremental improvements.
-Although literature-based innovation output measures assume editors at trade journales already discard incremental changes to include only important improvements, they cannot ex-ante anticipate which innovation will be most important in terms of impact [@kander2019InnovationTrendsIndustrial].
-Therefore the decline in innovation output this paper identifies should not be taken to necessarily imply a decline in the importance of the commercialized innovation.
-Similarily, the positive directionality measured in this paper should be seen more as a the vector of change embedded in the transition to a bioeconomy, rather than a scalar of its progress.
+Although literature-based innovation output measures assume editors at trade journals already discard incremental changes to include only important improvements, they cannot ex-ante anticipate which innovation will be most important in terms of impact [@kander2019InnovationTrendsIndustrial].
+Therefore, the decline in innovation output this paper identifies should not be taken to necessarily imply a decline in the importance of the commercialized innovation.
+Similarly, the positive directionality measured in this paper should be seen more as a vector of change embedded in the transition to a bioeconomy, rather than a scalar of its progress.
+
+# Policy Implications {#sec-policy}
+
+Some recommendations that follow from this study directly for Swedish policymakers include:
+
+1.
+To align innovation output with key aspects of the Swedish vision for a forest-based bioeconomy, there needs to be more focus on high-tech conversion of biomass to high-value added products including, but not limited to, biorefinery.
+2. Technological innovation within forestry has mainly focused on mechanization of harvesting. More proactive policy interventions can help spur innovation facilitating other silviculture practices.
 
@@
 
-In conclusion, the technological artefacts produced in the transition to a forest-based bioeconomy in Sweden show a clear directionality aligned with the bioresource vision, which is most favored by central actors and policy makers.
+In conclusion, the technological artifacts produced in the transition to a forest-based bioeconomy in Sweden show a clear directionality aligned with the bioresource vision, which is most favored by central actors and policymakers.
 At the same time, there is evidence that the innovation system has produced knowledge aligned with alternative transition pathways towards an ecological bioeconomy, such as decommodification of value chains as well as efforts to increase the circularity of the system and to reduce environmental pollution.
-The temporal distribution of innovation output aligns strongly with the Swedish tendency to produce consens oriented policy by aligning policy objectives with existing technolgical trajectories.
+The temporal distribution of innovation output aligns strongly with the Swedish tendency to produce consensus oriented policy by aligning policy objectives with existing technological trajectories.
 While this approach is pragmatic and has produced good outcomes in the past, it is at least doubtful how radical a shift can be expected for the bioeconomy transition, considering that innovation output since the 1970s has not meaningfully explored alternatives to the intensification of biomass resource extraction.
 
+# Funding {.unnumbered}
+This work has been funded by Vinnova (2020-01963).
+
 {{< pagebreak >}}
+
 # References {.unnumbered}
@@
 \noindent
-*VVS: tidskrift för värme-, ventilations- och sanitetsteknik : organ för Sven-
-ska värme- och sanitetstekniska föreningen.* (1970-1982). Stockholm:
-VVS.
+*VVS: tidskrift för värme-, ventilations- och sanitetsteknik : organ för Svenska värme- och sanitetstekniska föreningen.* (1970-1982). Stockholm: VVS.
 
@@
 ## Data Appendix {#sec-appendix-data}
-```{python}
-#| label: tbl-full-codes
-#| tbl-cap: List of Eco-Innovation Labels
-format_table(codes_df.iloc[:5, :])
-```
 
 {{< include _appendix-data.qmd >}}
+
+# Notes

### 2025-04-25 swinno_bioeconomy_directionality/analysis/figures.py

@@ -1,350 +1,368 @@
-import pandas as pd
+import string
+import polars as pl
 import matplotlib.pyplot as plt
-import matplotlib.ticker as mtick
-import seaborn as sns
-import plotly.graph_objects as go
-import plotly.express as px
-import plotly.io as pio
-import cmcrameri as cm
-from swinno_bioeconomy_directionality.config import FIGURES_DIR
+import matplotlib.ticker as mtick  # Added missing import
+import seaborn.objects as so
+from seaborn import axes_style
 
-# Set global options for all Plotly plots
+from cmcrameri import cm
+from swinno_bioeconomy_directionality.config import (
+    FIGURES_DIR,
+    PROCESSED_DATA_DIR,
+    RAW_DATA_DIR,
+    EXTERNAL_DATA_DIR,
+)
+
+from swinno_bioeconomy_directionality.data_handling.io import DataLoader
+
+WIDTH = 12
+HEIGHT = 8
+
+
+def create_figures():
+    theme_dict = {
+        **axes_style("whitegrid"),
+        "grid.linestyle": "-",
+        "axes.grid.axis": "y",
+        "axes.spines.top": False,
+        "axes.spines.right": False,
+        "axes.spines.left": False,
+    }
+
+    so.Plot().theme(theme_dict)
+
+    bioeconomy_share = pl.read_csv(PROCESSED_DATA_DIR / "swinno_bioeconomy_share.csv")
+
+    bioeconomy_vision_data = DataLoader(RAW_DATA_DIR).load_visions()
+
+    vision_summary = bioeconomy_vision_data.group_by("bioeconomy_vision").agg(
+        pl.col("sinno_id").n_unique().alias("sum")
+    )
+
+    vision_colormap = create_vision_colormap()
+
+    # Create and save figures
+    fig_innovation_trends = plot_innovation_trends(bioeconomy_share)
+    save_fig(fig_innovation_trends, "fig_innovation_trends")
+
+    # Pass both required arguments
+    fig_bioeconomy_vision_counts = plot_vision_counts(
+        bioeconomy_vision_data, vision_colormap
+    )
+    save_fig(fig_bioeconomy_vision_counts, "fig_bioeconomy_vision_counts")
+
+    fig_vision_producer_counts = plot_firm_counts(
+        bioeconomy_vision_data, vision_colormap
+    )
+    save_fig(fig_vision_producer_counts, "fig_vision_producer_counts")
 
 
[…]
